### 2015-10-03 manuscript.Rmd
first commit

---
#layout: preprint
layout: review, 11pt
linenumbers: true
title: "rgbif: a package for working with species occurrence data in R"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: scott(at)ropensci.org
    footnote: Corresponding author
address:
  - code: cstar
    address: |
      University of California, Berkeley, CA, USA
abstract: |
      1. xxx

      2. xxx

      3. xxx

      4. xxxx

bibliography: components/references.bib
csl: components/methods-in-ecology-and-evolution.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE,
               cache = 1, comment = NA, verbose = TRUE)
basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

```

# Introduction

Users of the popular statistical and mathematical computing platform
R [@R] enjoy a wealth of readily installable comparative phylogenetic
methods and tools [@taskview]. Exploiting the opportunities arising from
this wealth for complex and integrative comparative research questions
relies on the ability to reuse and integrate previously generated or
published data and metadata. The expanding data exchange needs of the
evolutionary research community are rapidly outpacing the capabilities of
most current and widely used data exchange standards [@Vos_2012], which
were all developed a decade or more ago. This has resulted in a radiation
of different data representations and exchange standard "flavors" that are
no longer interoperable at the very time when the growth of available data
and methods has made that interoperability most valuable. In response to
the unmet needs for standardized data exchange in phylogenetics, a modern
XML-based exchange standard, called NeXML, has recently been developed
[@Vos_2012]. NeXML comprehensively supports current data exchange needs,
is predictably machine-readable, and is forward compatible.

The exchange problem for phylogenetic data is particularly acute in light
of the challenges in finding and sharing phylogenetic data without the
otherwise common loss of most data and metadata semantics [@Drew_2013;
@Stoltzfus_2012; @Cranston_2014]. For example, the still popular NEXUS
file format [@Maddison_1997] cannot consistently represent horizontal
gene transfer or ambiguity in reading a character (such as a DNA sequence
base pair).  This and other limitations have led to modifications of NEXUS
in different ways for different needs, with the unfortunate result that
NEXUS files generated by one program can be incompatible with another
[@Vos_2012]. Without a formal grammar, software based on NEXUS files
may also make inconsistent assumptions about tokens, quoting, or element
lengths.  @Vos_2012 estimates that as many as 15% of the NEXUS files in
the CIPRES portal contain unrecoverable but hard to diagnose errors.

A detailed account of how the NeXML standard addresses these and
other relevant challenges can be found in @Vos_2012. In brief,
NeXML was designed with the following important properties. First,
NeXML is defined by a precise grammar that can be programmatically
**validated**; i.e., it can be verified whether a file precisely
follows this grammar, and therefore whether it can be read (parsed)
without errors by software that uses the NeXML grammar (e.g. RNeXML)
is predictable.
Second, NeXML is **extensible**: a user
can define representations of new, previously unanticipated information
(as we will illustrate) without violating its defining grammar. Third
and most importantly, NeXML is rich in **computable semantics**: it is
designed for expressing metadata such that machines can understand
their meaning and make inferences from it. For example, OTUs in a tree
or character matrix for frog species can be linked to concepts in a
formally defined hierarchy of taxonomic concepts such as the
Vertebrate Taxonomy Ontology [@Midford2013], which enables a machine
to infer that a query for amphibia is to include the frog data in what
is returned. (For a more broader discussion of the value of such
capabilities for evolutionary and biodiversity science we refer the
reader to @Parr2011.)

To make the capabilities of NeXML available to R users in an easy-to-use
form, and to lower the hurdles to adoption of the standard, we present
RNeXML, an R package that aims to provide easy programmatic access to
reading and writing NeXML documents, tailored for the kinds of use-cases
that will be common for users and developers of the wealth of evolutionary
analysis methods within the R ecosystem.

```{r echo=FALSE, results='hide', eval=FALSE}
install.packages("rgbif", dependencies=TRUE)
```

```{r echo=FALSE, results='hide'}
library("rgbif")
```

# The rgbif package

The `rgbif` package ...

# Conclusions and future directions

`rgbif` ...

## Acknowledgements

This project was supported in part by the Alfred P Sloan
Foundation (Grant 2013-6-22).

## Data Accessibility

All software, scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif).

# References

### 2015-10-03 manuscript.tex
first commit

\documentclass[author-year, review, 11pt]{components/elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
  \linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rgbif: a package for working with species occurrence data in R},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{rgbif: a package for working with species occurrence data in R}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{scott(at)ropensci.org} 
   \cortext[c1]{Corresponding author}
      \address[cstar]{University of California, Berkeley, CA, USA}

  \begin{abstract}
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \item
    xxx
  \item
    xxx
  \item
    xxx
  \item
    xxxx
  \end{enumerate}
  \end{abstract}

 \end{frontmatter}

\section{Introduction}\label{introduction}

Users of the popular statistical and mathematical computing platform R
(R Core Team 2014) enjoy a wealth of readily installable comparative
phylogenetic methods and tools (O'Meara 2014). Exploiting the
opportunities arising from this wealth for complex and integrative
comparative research questions relies on the ability to reuse and
integrate previously generated or published data and metadata. The
expanding data exchange needs of the evolutionary research community are
rapidly outpacing the capabilities of most current and widely used data
exchange standards (Vos \emph{et al.} 2012), which were all developed a
decade or more ago. This has resulted in a radiation of different data
representations and exchange standard ``flavors'' that are no longer
interoperable at the very time when the growth of available data and
methods has made that interoperability most valuable. In response to the
unmet needs for standardized data exchange in phylogenetics, a modern
XML-based exchange standard, called NeXML, has recently been developed
(Vos \emph{et al.} 2012). NeXML comprehensively supports current data
exchange needs, is predictably machine-readable, and is forward
compatible.

The exchange problem for phylogenetic data is particularly acute in
light of the challenges in finding and sharing phylogenetic data without
the otherwise common loss of most data and metadata semantics (Stoltzfus
\emph{et al.} 2012; Drew \emph{et al.} 2013; Cranston \emph{et al.}
2014). For example, the still popular NEXUS file format (Maddison
\emph{et al.} 1997) cannot consistently represent horizontal gene
transfer or ambiguity in reading a character (such as a DNA sequence
base pair). This and other limitations have led to modifications of
NEXUS in different ways for different needs, with the unfortunate result
that NEXUS files generated by one program can be incompatible with
another (Vos \emph{et al.} 2012). Without a formal grammar, software
based on NEXUS files may also make inconsistent assumptions about
tokens, quoting, or element lengths. Vos et al. (2012) estimates that as
many as 15\% of the NEXUS files in the CIPRES portal contain
unrecoverable but hard to diagnose errors.

A detailed account of how the NeXML standard addresses these and other
relevant challenges can be found in Vos et al. (2012). In brief, NeXML
was designed with the following important properties. First, NeXML is
defined by a precise grammar that can be programmatically
\textbf{validated}; i.e., it can be verified whether a file precisely
follows this grammar, and therefore whether it can be read (parsed)
without errors by software that uses the NeXML grammar (e.g.~RNeXML) is
predictable. Second, NeXML is \textbf{extensible}: a user can define
representations of new, previously unanticipated information (as we will
illustrate) without violating its defining grammar. Third and most
importantly, NeXML is rich in \textbf{computable semantics}: it is
designed for expressing metadata such that machines can understand their
meaning and make inferences from it. For example, OTUs in a tree or
character matrix for frog species can be linked to concepts in a
formally defined hierarchy of taxonomic concepts such as the Vertebrate
Taxonomy Ontology (Midford \emph{et al.} 2013), which enables a machine
to infer that a query for amphibia is to include the frog data in what
is returned. (For a more broader discussion of the value of such
capabilities for evolutionary and biodiversity science we refer the
reader to Parr et al. (2011).)

To make the capabilities of NeXML available to R users in an easy-to-use
form, and to lower the hurdles to adoption of the standard, we present
RNeXML, an R package that aims to provide easy programmatic access to
reading and writing NeXML documents, tailored for the kinds of use-cases
that will be common for users and developers of the wealth of
evolutionary analysis methods within the R ecosystem.

\section{The rgbif package}\label{the-rgbif-package}

The \texttt{rgbif} package \ldots{}

\section{Conclusions and future
directions}\label{conclusions-and-future-directions}

\texttt{rgbif} \ldots{}

\subsection{Acknowledgements}\label{acknowledgements}

This project was supported in part by the Alfred P Sloan Foundation
(Grant 2013-6-22).

\subsection{Data Accessibility}\label{data-accessibility}

All software, scripts and data used in this paper can be found in the
permanent data archive Zenodo under the digital object identifier (DOI).
This DOI corresponds to a snapshot of the GitHub repository at
\href{https://github.com/ropensci/rgbif}{github.com/ropensci/rgbif}.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

Cranston, K., Harmon, L.J., O'Leary, M.A. \& Lisle, C. (2014). Best
practices for data sharing in phylogenetic research. \emph{PLoS Curr}.
Retrieved from
\url{http://dx.doi.org/10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645}

Drew, B.T., Gazis, R., Cabezas, P., Swithers, K.S., Deng, J., Rodriguez,
R., Katz, L.A., Crandall, K.A., Hibbett, D.S. \& Soltis, D.E. (2013).
Lost branches on the tree of life. \emph{PLoS Biol}, \textbf{11},
e1001636. Retrieved from
\url{http://dx.doi.org/10.1371/journal.pbio.1001636}

Maddison, D., Swofford, D. \& Maddison, W. (1997). NEXUS: An extensible
file format for systematic information. \emph{Syst. Biol.}, \textbf{46},
590--621. Retrieved from
\url{http://www.ncbi.nlm.nih.gov/pubmed/11975335}

Midford, P., Dececchi, T., Balhoff, J., Dahdul, W., Ibrahim, N., Lapp,
H., Lundberg, J., Mabee, P., Sereno, P., Westerfield, M., Vision, T. \&
Blackburn, D. (2013). The vertebrate taxonomy ontology: A framework for
reasoning across model organism and species phenotypes. \emph{J. Biomed.
Semantics}, \textbf{4}, 34. Retrieved from
\url{http://dx.doi.org/10.1186/2041-1480-4-34}

O'Meara, B. (2014). CRAN task view: Phylogenetics, especially
comparative methods. Retrieved from
\url{http://cran.r-project.org/web/views/Phylogenetics.html}

Parr, C.S., Guralnick, R., Cellinese, N. \& Page, R.D.M. (2011).
Evolutionary informatics: unifying knowledge about the diversity of
life. \emph{Trends in ecology \& evolution}, \textbf{27}, 94--103.
Retrieved from \url{http://www.ncbi.nlm.nih.gov/pubmed/22154516}

R Core Team. (2014). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing, Vienna, Austria.
Retrieved from \url{http://www.R-project.org/}

Stoltzfus, A., O'Meara, B., Whitacre, J., Mounce, R., Gillespie, E.L.,
Kumar, S., Rosauer, D.F. \& Vos, R.A. (2012). Sharing and re-use of
phylogenetic trees (and associated data) to facilitate synthesis.
\emph{BMC Research Notes}, \textbf{5}, 574. Retrieved from
\url{http://dx.doi.org/10.1186/1756-0500-5-574}

Vos, R.A., Balhoff, J.P., Caravas, J.A., Holder, M.T., Lapp, H.,
Maddison, W.P., Midford, P.E., Priyam, A., Sukumaran, J., Xia, X. \&
Stoltzfus, A. (2012). NeXML: Rich, extensible, and verifiable
representation of comparative data and metadata. \emph{Systematic
Biology}, \textbf{61}, 675--689. Retrieved from
\url{http://dx.doi.org/10.1093/sysbio/sys025}

\end{document}

### 2015-10-03 manuscript.Rmd
remove some things, remove not needed refs, updated circleci badge

@@
                cache.path = paste("components/cache/", basename, "/", sep=""))
-
 ```
@@
 R [@R] enjoy a wealth of readily installable comparative phylogenetic
-methods and tools [@taskview]. Exploiting the opportunities arising from
-this wealth for complex and integrative comparative research questions
-relies on the ability to reuse and integrate previously generated or
-published data and metadata. The expanding data exchange needs of the
-evolutionary research community are rapidly outpacing the capabilities of
-most current and widely used data exchange standards [@Vos_2012], which
-were all developed a decade or more ago. This has resulted in a radiation
-of different data representations and exchange standard "flavors" that are
-no longer interoperable at the very time when the growth of available data
-and methods has made that interoperability most valuable. In response to
-the unmet needs for standardized data exchange in phylogenetics, a modern
-XML-based exchange standard, called NeXML, has recently been developed
-[@Vos_2012]. NeXML comprehensively supports current data exchange needs,
-is predictably machine-readable, and is forward compatible.
-
-The exchange problem for phylogenetic data is particularly acute in light
-of the challenges in finding and sharing phylogenetic data without the
-otherwise common loss of most data and metadata semantics [@Drew_2013;
-@Stoltzfus_2012; @Cranston_2014]. For example, the still popular NEXUS
-file format [@Maddison_1997] cannot consistently represent horizontal
-gene transfer or ambiguity in reading a character (such as a DNA sequence
-base pair).  This and other limitations have led to modifications of NEXUS
-in different ways for different needs, with the unfortunate result that
-NEXUS files generated by one program can be incompatible with another
-[@Vos_2012]. Without a formal grammar, software based on NEXUS files
-may also make inconsistent assumptions about tokens, quoting, or element
-lengths.  @Vos_2012 estimates that as many as 15% of the NEXUS files in
-the CIPRES portal contain unrecoverable but hard to diagnose errors.
-
-A detailed account of how the NeXML standard addresses these and
-other relevant challenges can be found in @Vos_2012. In brief,
-NeXML was designed with the following important properties. First,
-NeXML is defined by a precise grammar that can be programmatically
-**validated**; i.e., it can be verified whether a file precisely
-follows this grammar, and therefore whether it can be read (parsed)
-without errors by software that uses the NeXML grammar (e.g. RNeXML)
-is predictable.
-Second, NeXML is **extensible**: a user
-can define representations of new, previously unanticipated information
-(as we will illustrate) without violating its defining grammar. Third
-and most importantly, NeXML is rich in **computable semantics**: it is
-designed for expressing metadata such that machines can understand
-their meaning and make inferences from it. For example, OTUs in a tree
-or character matrix for frog species can be linked to concepts in a
-formally defined hierarchy of taxonomic concepts such as the
-Vertebrate Taxonomy Ontology [@Midford2013], which enables a machine
-to infer that a query for amphibia is to include the frog data in what
-is returned. (For a more broader discussion of the value of such
-capabilities for evolutionary and biodiversity science we refer the
-reader to @Parr2011.)
+methods and tools.
 
@@
 All software, scripts and data used in this paper can be found in the permanent
-data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif).
+data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/sckott/msrgbif](https://github.com/sckott/msrgbif).
 

### 2015-10-03 manuscript.tex
update pdf, trying circleci artifacts

@@
 (R Core Team 2014) enjoy a wealth of readily installable comparative
-phylogenetic methods and tools (O'Meara 2014). Exploiting the
-opportunities arising from this wealth for complex and integrative
-comparative research questions relies on the ability to reuse and
-integrate previously generated or published data and metadata. The
-expanding data exchange needs of the evolutionary research community are
-rapidly outpacing the capabilities of most current and widely used data
-exchange standards (Vos \emph{et al.} 2012), which were all developed a
-decade or more ago. This has resulted in a radiation of different data
-representations and exchange standard ``flavors'' that are no longer
-interoperable at the very time when the growth of available data and
-methods has made that interoperability most valuable. In response to the
-unmet needs for standardized data exchange in phylogenetics, a modern
-XML-based exchange standard, called NeXML, has recently been developed
-(Vos \emph{et al.} 2012). NeXML comprehensively supports current data
-exchange needs, is predictably machine-readable, and is forward
-compatible.
-
-The exchange problem for phylogenetic data is particularly acute in
-light of the challenges in finding and sharing phylogenetic data without
-the otherwise common loss of most data and metadata semantics (Stoltzfus
-\emph{et al.} 2012; Drew \emph{et al.} 2013; Cranston \emph{et al.}
-2014). For example, the still popular NEXUS file format (Maddison
-\emph{et al.} 1997) cannot consistently represent horizontal gene
-transfer or ambiguity in reading a character (such as a DNA sequence
-base pair). This and other limitations have led to modifications of
-NEXUS in different ways for different needs, with the unfortunate result
-that NEXUS files generated by one program can be incompatible with
-another (Vos \emph{et al.} 2012). Without a formal grammar, software
-based on NEXUS files may also make inconsistent assumptions about
-tokens, quoting, or element lengths. Vos et al. (2012) estimates that as
-many as 15\% of the NEXUS files in the CIPRES portal contain
-unrecoverable but hard to diagnose errors.
-
-A detailed account of how the NeXML standard addresses these and other
-relevant challenges can be found in Vos et al. (2012). In brief, NeXML
-was designed with the following important properties. First, NeXML is
-defined by a precise grammar that can be programmatically
-\textbf{validated}; i.e., it can be verified whether a file precisely
-follows this grammar, and therefore whether it can be read (parsed)
-without errors by software that uses the NeXML grammar (e.g.~RNeXML) is
-predictable. Second, NeXML is \textbf{extensible}: a user can define
-representations of new, previously unanticipated information (as we will
-illustrate) without violating its defining grammar. Third and most
-importantly, NeXML is rich in \textbf{computable semantics}: it is
-designed for expressing metadata such that machines can understand their
-meaning and make inferences from it. For example, OTUs in a tree or
-character matrix for frog species can be linked to concepts in a
-formally defined hierarchy of taxonomic concepts such as the Vertebrate
-Taxonomy Ontology (Midford \emph{et al.} 2013), which enables a machine
-to infer that a query for amphibia is to include the frog data in what
-is returned. (For a more broader discussion of the value of such
-capabilities for evolutionary and biodiversity science we refer the
-reader to Parr et al. (2011).)
+phylogenetic methods and tools.
 
@@
 This DOI corresponds to a snapshot of the GitHub repository at
-\href{https://github.com/ropensci/rgbif}{github.com/ropensci/rgbif}.
+\href{https://github.com/sckott/msrgbif}{github.com/sckott/msrgbif}.
 
@@
 \addcontentsline{toc}{section}{References}
-
-Cranston, K., Harmon, L.J., O'Leary, M.A. \& Lisle, C. (2014). Best
-practices for data sharing in phylogenetic research. \emph{PLoS Curr}.
-Retrieved from
-\url{http://dx.doi.org/10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645}
-
-Drew, B.T., Gazis, R., Cabezas, P., Swithers, K.S., Deng, J., Rodriguez,
-R., Katz, L.A., Crandall, K.A., Hibbett, D.S. \& Soltis, D.E. (2013).
-Lost branches on the tree of life. \emph{PLoS Biol}, \textbf{11},
-e1001636. Retrieved from
-\url{http://dx.doi.org/10.1371/journal.pbio.1001636}
-
-Maddison, D., Swofford, D. \& Maddison, W. (1997). NEXUS: An extensible
-file format for systematic information. \emph{Syst. Biol.}, \textbf{46},
-590--621. Retrieved from
-\url{http://www.ncbi.nlm.nih.gov/pubmed/11975335}
-
-Midford, P., Dececchi, T., Balhoff, J., Dahdul, W., Ibrahim, N., Lapp,
-H., Lundberg, J., Mabee, P., Sereno, P., Westerfield, M., Vision, T. \&
-Blackburn, D. (2013). The vertebrate taxonomy ontology: A framework for
-reasoning across model organism and species phenotypes. \emph{J. Biomed.
-Semantics}, \textbf{4}, 34. Retrieved from
-\url{http://dx.doi.org/10.1186/2041-1480-4-34}
-
-O'Meara, B. (2014). CRAN task view: Phylogenetics, especially
-comparative methods. Retrieved from
-\url{http://cran.r-project.org/web/views/Phylogenetics.html}
-
-Parr, C.S., Guralnick, R., Cellinese, N. \& Page, R.D.M. (2011).
-Evolutionary informatics: unifying knowledge about the diversity of
-life. \emph{Trends in ecology \& evolution}, \textbf{27}, 94--103.
-Retrieved from \url{http://www.ncbi.nlm.nih.gov/pubmed/22154516}
 
@@
 
-Stoltzfus, A., O'Meara, B., Whitacre, J., Mounce, R., Gillespie, E.L.,
-Kumar, S., Rosauer, D.F. \& Vos, R.A. (2012). Sharing and re-use of
-phylogenetic trees (and associated data) to facilitate synthesis.
-\emph{BMC Research Notes}, \textbf{5}, 574. Retrieved from
-\url{http://dx.doi.org/10.1186/1756-0500-5-574}
-
-Vos, R.A., Balhoff, J.P., Caravas, J.A., Holder, M.T., Lapp, H.,
-Maddison, W.P., Midford, P.E., Priyam, A., Sukumaran, J., Xia, X. \&
-Stoltzfus, A. (2012). NeXML: Rich, extensible, and verifiable
-representation of comparative data and metadata. \emph{Systematic
-Biology}, \textbf{61}, 675--689. Retrieved from
-\url{http://dx.doi.org/10.1093/sysbio/sys025}
-
 \end{document}

### 2015-10-03 manuscript.Rmd
added gitignore, manuscript edits, added refs

@@
 
-Users of the popular statistical and mathematical computing platform
-R [@R] enjoy a wealth of readily installable comparative phylogenetic
-methods and tools.
+Biodiversity loss is one of the greatest challenges of our
+time [@Pimm_2014]. Some have called this the sixth great mass
+extinction [@Ceballos_2015]. Given this challenge there is a
+great need for, and much research is done on, biodiversity
+science. The pool of questions about biodiversity we can answer
+effectively becomes much larger with more and better data.
 
-To make the capabilities of NeXML available to R users in an easy-to-use
-form, and to lower the hurdles to adoption of the standard, we present
-RNeXML, an R package that aims to provide easy programmatic access to
-reading and writing NeXML documents, tailored for the kinds of use-cases
-that will be common for users and developers of the wealth of evolutionary
-analysis methods within the R ecosystem.
+The Global Biodiversity Information Facility (hereafter, GBIF,
+http://www.gbif.org/) is the largest collection of biodiversity
+records globally, currently with 580 million records, 1.6 million
+taxa, 15,000 datasets from 770 publishers. In addition, large
+biodiversity warehouses such as iNaturalist, VertNet, and USGS's
+BISON all feed into GBIF.
+
+Herein, we describe a library for working with GBIF data in
+the R programming environment [@R]. R is an extremely widely used
+language in academia, non-profits, and the private sector.
+Importantly, R make it easy to do all of the steps of the
+research process, including data management, data manipulation
+and cleaning, statistics, and vizualization. An R client
+for getting GBIF is a powerful tool for reproducible research.
+
+# Install rgbif
 
@@
 The `rgbif` package ...
+
+## Registry
+
+```{r}
+"xxx"
+```
+
+## Species
+
+## Occurrences
+
+### Download API
+
+### Search API
 

### 2015-10-04 manuscript.Rmd
tweaks
manuscript edits, ignore git_deploy scritpt for now

@@
 
-Herein, we describe a library for working with GBIF data in
-the R programming environment [@R]. R is an extremely widely used
-language in academia, non-profits, and the private sector.
+Herein, we describe a library (rgbif [@rgbif]) for working with
+GBIF data in the R programming environment [@R]. R is an extremely
+widely used language in academia, non-profits, and the private sector.
 Importantly, R make it easy to do all of the steps of the
@@
 
+The GBIF reqistry API services are spread across four major sets
+of functions:
+
+* Datasets
+* Installations
+* Networks
+* Nodes
+* Organizations
+
+### Datasets
+
+Search for datasets
+
 ```{r}
-"xxx"
+dataset_search(keyword="france", return = "data")
+```
+
+Get dataset metrics
+
+```{r}
+res <- dataset_metrics(uuid='66dd0960-2d7d-46ee-a491-87b9adcfe7b1')
+data.frame(names(res$countByRank), unname(unlist(res$countByRank)))
+```
+
+### Networks, nodes, and installations
+
+Here, we search for the first give GBIF networks, returning just the
+key and title fields.
+
+```{r}
+networks(limit=5)$data[ , c('key', 'title')]
 ```
@@
 
+GBIF provides two ways to get occurrence data: through the
+/occurrence/search route (see `occ_search`), or via the
+/occurrence/download route (many functions, see below).
+`occ_search` is more appropriate for smaller data, while
+occ_download*() functions are more appropriate for larger
+data requests.
+
 ### Download API
+
+The download API syntax is similar to the occurrence search API
+in that the same parameters are used, but the syntas is quite different.
+For example, in the download API you can do greater than searches,
+whereas you can not do that in the occurrence search API.
+
+```{r}
+occ_download_get("0000066-140928181241064", overwrite = TRUE)
+```
 
@@
 # Conclusions and future directions
-
-`rgbif` ...
 

### 2015-10-04 manuscript.Rmd
manuscript additions

@@
 
-Biodiversity loss is one of the greatest challenges of our
-time [@Pimm_2014]. Some have called this the sixth great mass
-extinction [@Ceballos_2015]. Given this challenge there is a
-great need for, and much research is done on, biodiversity
-science. The pool of questions about biodiversity we can answer
-effectively becomes much larger with more and better data.
+Perhaps the most fundamental element in many fields of ecology is
+the inividual. How many individuals of each species in a given
+location forms the basis for many sub-fields of ecology and evolution.
+Some research questions necessitate collecting new data, while
+others can easily take advantage of existing data. In fact, some
+ecology fields are built largely on existing data, e.g., macro-ecology
+[@Brown_1995; @Beck_2012].
 
-The Global Biodiversity Information Facility (hereafter, GBIF,
+Data on individuals, including which species, and where they're found,
+can be used for a large number of research questions. In addition,
+the pool of questions we can answer becomes much larger with more
+and better data. In addition to wide utility, this data is important
+for conservation. Biodiversity loss is one of the greatest challenges
+of our time [@Pimm_2014]. Some have called this the sixth great mass
+extinction [@Ceballos_2015]. Given this challenge there is a great
+need for data on specimen records, whether collected from live sightings
+in the field or specimens in museums.
+
+There are many online services that collect and maintain specimen records.
+However, Global Biodiversity Information Facility (hereafter, GBIF,
 http://www.gbif.org/) is the largest collection of biodiversity
 records globally, currently with 580 million records, 1.6 million
-taxa, 15,000 datasets from 770 publishers. In addition, large
-biodiversity warehouses such as iNaturalist, VertNet, and USGS's
+taxa, 15,000 datasets from 770 publishers (figures collected on 2015-10-04).
+Many large biodiversity warehouses such as iNaturalist, VertNet, and USGS's
 BISON all feed into GBIF.
@@
 GBIF data in the R programming environment [@R]. R is an extremely
-widely used language in academia, non-profits, and the private sector.
-Importantly, R make it easy to do all of the steps of the
+widely used language in academia, and in non-profit and private sectors.
+Importantly, R makes it easy to do all of the steps of the
 research process, including data management, data manipulation
-and cleaning, statistics, and vizualization. An R client
-for getting GBIF is a powerful tool for reproducible research.
-
-# Install rgbif
+and cleaning, statistics, and vizualization. Thus, an R client
+for getting GBIF data is a powerful tool for reproducible research.
 
@@
 
-The `rgbif` package ...
+The `rgbif` package is completely written in R, uses an [MIT license][mit]
+to maximize use everywhere. `rgbif` is developed publicly on GitHub
+at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
+of the package can be installed, and bugs and feature requests reported.
+Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
+distribution network for R packages. `rgbif` is part of the rOpenSci project,
+a developer network making R software to facilitate reproducible research.
+
+## Package interface
+
+`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
+or API. The GBIF API has four major components: registry, species names,
+occurrence data, and maps. We ignore maps in `rgbif` as it is concerned with
+generating maps for web applications. `rgbif` has a suite of functions dealing
+with each of registry, species names, and occurrence data - we'll go through
+each in turn describing design and example usage.
 
@@
 
+The search API is very similar to the download API, but is meant for smaller
+data acquisition jobs.
+
 # Conclusions and future directions
@@
 # References
+
+[mit]: http://choosealicense.com/licenses/mit/
+[ghrgbif]: https://github.com/ropensci/rgbif
+[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
+[gbifdocs]: http://www.gbif.org/developer/summary

### 2015-10-04 manuscript.Rmd
code eg edits for simplicity

@@
 ```{r}
-dataset_search(keyword="france", return = "data")
+res <- dataset_search(query = "oregon")
+res$data$datasetTitle[1:10]
 ```
@@
 res <- dataset_metrics(uuid='66dd0960-2d7d-46ee-a491-87b9adcfe7b1')
-data.frame(names(res$countByRank), unname(unlist(res$countByRank)))
+df <- data.frame(rank = names(res$countByRank),
+                 count = unname(unlist(res$countByRank)))
+knitr::kable(df)
 ```
@@
 ```{r}
-networks(limit=5)$data[ , c('key', 'title')]
+networks(limit=10)$data$title
 ```
@@
 ```{r}
-occ_download_get("0000066-140928181241064", overwrite = TRUE)
+occ_download_get("0000066-140928181241064")
 ```

### 2015-10-05 manuscript.tex
built pdf

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
+\newcommand{\NormalTok}[1]{{#1}}
+\usepackage{longtable}
 \ifxetex
@@
 
-Users of the popular statistical and mathematical computing platform R
-(R Core Team 2014) enjoy a wealth of readily installable comparative
-phylogenetic methods and tools.
+Perhaps the most fundamental element in many fields of ecology is the
+inividual. How many individuals of each species in a given location
+forms the basis for many sub-fields of ecology and evolution. Some
+research questions necessitate collecting new data, while others can
+easily take advantage of existing data. In fact, some ecology fields are
+built largely on existing data, e.g., macro-ecology (Brown 1995; Beck
+\emph{et al.} 2012).
 
-To make the capabilities of NeXML available to R users in an easy-to-use
-form, and to lower the hurdles to adoption of the standard, we present
-RNeXML, an R package that aims to provide easy programmatic access to
-reading and writing NeXML documents, tailored for the kinds of use-cases
-that will be common for users and developers of the wealth of
-evolutionary analysis methods within the R ecosystem.
+Data on individuals, including which species, and where they're found,
+can be used for a large number of research questions. In addition, the
+pool of questions we can answer becomes much larger with more and better
+data. In addition to wide utility, this data is important for
+conservation. Biodiversity loss is one of the greatest challenges of our
+time (Pimm \emph{et al.} 2014). Some have called this the sixth great
+mass extinction (Ceballos \emph{et al.} 2015). Given this challenge
+there is a great need for data on specimen records, whether collected
+from live sightings in the field or specimens in museums.
+
+There are many online services that collect and maintain specimen
+records. However, Global Biodiversity Information Facility (hereafter,
+GBIF, \url{http://www.gbif.org/}) is the largest collection of
+biodiversity records globally, currently with 580 million records, 1.6
+million taxa, 15,000 datasets from 770 publishers (figures collected on
+2015-10-04). Many large biodiversity warehouses such as iNaturalist,
+VertNet, and USGS's BISON all feed into GBIF.
+
+Herein, we describe a library (rgbif (Chamberlain \emph{et al.})) for
+working with GBIF data in the R programming environment (R Core Team
+2014). R is an extremely widely used language in academia, and in
+non-profit and private sectors. Importantly, R makes it easy to do all
+of the steps of the research process, including data management, data
+manipulation and cleaning, statistics, and vizualization. Thus, an R
+client for getting GBIF data is a powerful tool for reproducible
+research.
 
@@
 
-The \texttt{rgbif} package \ldots{}
+The \texttt{rgbif} package is completely written in R, uses an
+\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
+use everywhere. \texttt{rgbif} is developed publicly on GitHub at
+\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
+where development versions of the package can be installed, and bugs and
+feature requests reported. Stable versions of \texttt{rgbif} can be
+installed from
+\href{https://cran.rstudio.com/web/packages/rgbif/}{CRAN}, the
+distribution network for R packages. \texttt{rgbif} is part of the
+rOpenSci project, a developer network making R software to facilitate
+reproducible research.
+
+\subsection{Package interface}\label{package-interface}
+
+\texttt{rgbif} is designed following the
+\href{http://www.gbif.org/developer/summary}{GBIF Application
+Programming Interface}, or API. The GBIF API has four major components:
+registry, species names, occurrence data, and maps. We ignore maps in
+\texttt{rgbif} as it is concerned with generating maps for web
+applications. \texttt{rgbif} has a suite of functions dealing with each
+of registry, species names, and occurrence data - we'll go through each
+in turn describing design and example usage.
+
+\subsection{Registry}\label{registry}
+
+The GBIF reqistry API services are spread across four major sets of
+functions:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Datasets
+\item
+  Installations
+\item
+  Networks
+\item
+  Nodes
+\item
+  Organizations
+\end{itemize}
+
+\subsubsection{Datasets}\label{datasets}
+
+Search for datasets
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"oregon"}\NormalTok{)}
+\NormalTok{res$data$datasetTitle[}\DecValTok{1}\NormalTok{:}\DecValTok{10}\NormalTok{]}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+ [1] "SDNHM Birds Collection"                                                
+ [2] "CM Birds Collection"                                                   
+ [3] "condoncollection"                                                      
+ [4] "Taxonomy in Flux Checklist"                                            
+ [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"     
+ [6] "Bryophyte Collection - University of Washington Herbarium (WTU)"       
+ [7] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"
+ [8] "UWFC Ichthyology Collection"                                           
+ [9] "Lichen Collection - University of Washington Herbarium (WTU)"          
+[10] "UWBM Mammalogy Collection"                                             
+\end{verbatim}
+
+Get dataset metrics
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'66dd0960-2d7d-46ee-a491-87b9adcfe7b1'}\NormalTok{)}
+\NormalTok{df <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
+                 \DataTypeTok{count =} \KeywordTok{unname}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(res$countByRank)))}
+\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(df)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[c]{@{}lr@{}}
+\toprule
+rank & count\tabularnewline
+\midrule
+\endhead
+SPECIES & 52452\tabularnewline
+GENUS & 12930\tabularnewline
+VARIETY & 4806\tabularnewline
+SUBSPECIES & 4440\tabularnewline
+SERIES & 1079\tabularnewline
+TRIBE & 844\tabularnewline
+FAMILY & 509\tabularnewline
+SUBTRIBE & 327\tabularnewline
+SUBFAMILY & 303\tabularnewline
+SUBGENUS & 241\tabularnewline
+FORM & 239\tabularnewline
+SECTION & 82\tabularnewline
+SUBVARIETY & 4\tabularnewline
+KINGDOM & 1\tabularnewline
+\bottomrule
+\end{longtable}
+
+\subsubsection{Networks, nodes, and
+installations}\label{networks-nodes-and-installations}
+
+Here, we search for the first give GBIF networks, returning just the key
+and title fields.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit=}\DecValTok{10}\NormalTok{)$data$title}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+ [1] "GBIF Backbone Sources"                                      
+ [2] "Canadensys"                                                 
+ [3] "Southwest Collections of Arthropods Network (SCAN)"         
+ [4] "VertNet"                                                    
+ [5] "Dryad"                                                      
+ [6] "GBIF Network"                                               
+ [7] "The Knowledge Network for Biocomplexity (KNB) "             
+ [8] "Online Zoological Collections of Australian Museums (OZCAM)"
+ [9] "Catalogue of Life"                                          
+[10] "Ocean Biogeographic Information System (OBIS)"              
+\end{verbatim}
+
+\subsection{Species}\label{species}
+
+\subsection{Occurrences}\label{occurrences}
+
+GBIF provides two ways to get occurrence data: through the
+/occurrence/search route (see \texttt{occ\_search}), or via the
+/occurrence/download route (many functions, see below).
+\texttt{occ\_search} is more appropriate for smaller data, while
+occ\_download*() functions are more appropriate for larger data
+requests.
+
+\subsubsection{Download API}\label{download-api}
+
+The download API syntax is similar to the occurrence search API in that
+the same parameters are used, but the syntas is quite different. For
+example, in the download API you can do greater than searches, whereas
+you can not do that in the occurrence search API.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000066-140928181241064"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+<<gbif downloaded get>>
+  Path: ./0000066-140928181241064.zip
+  File size: 0.14 MB
+\end{verbatim}
+
+\subsubsection{Search API}\label{search-api}
+
+The search API is very similar to the download API, but is meant for
+smaller data acquisition jobs.
 
@@
 directions}\label{conclusions-and-future-directions}
-
-\texttt{rgbif} \ldots{}
 
@@
 
+Beck, J., Ballesteros-Mejia, L., Buchmann, C.M., Dengler, J., Fritz,
+S.A., Gruber, B., Hof, C., Jansen, F., Knapp, S., Kreft, H., Schneider,
+A.-K., Winter, M. \& Dormann, C.F. (2012). Whats on the horizon for
+macroecology? \emph{Ecography}, \textbf{35}, 673--683. Retrieved from
+\url{http://dx.doi.org/10.1111/j.1600-0587.2012.07364.x}
+
+Brown, J.H. (1995). \emph{Macroecology}. University of Chicago Press.
+
+Ceballos, G., Ehrlich, P.R., Barnosky, A.D., Garcia, A., Pringle, R.M.
+\& Palmer, T.M. (2015). Accelerated modern human-induced species losses:
+Entering the sixth mass extinction. \emph{Science Advances}, \textbf{1},
+e1400253--e1400253. Retrieved from
+\url{http://dx.doi.org/10.1126/sciadv.1400253}
+
+Chamberlain, S., Ram, K., Barve, V. \& Mcglinn, D. \emph{Rgbif:
+Interface to the global 'biodiversity' information facility 'aPI'}.
+Retrieved from \url{https://github.com/ropensci/rgbif}
+
+Pimm, S.L., Jenkins, C.N., Abell, R., Brooks, T.M., Gittleman, J.L.,
+Joppa, L.N., Raven, P.H., Roberts, C.M. \& Sexton, J.O. (2014). The
+biodiversity of species and their rates of extinction, distribution, and
+protection. \emph{Science}, \textbf{344}, 1246752--1246752. Retrieved
+from \url{http://dx.doi.org/10.1126/science.1246752}
+
 R Core Team. (2014). \emph{R: A language and environment for statistical

### 2015-10-05 manuscript.Rmd
deleted cache stuff, gitignoring now, updates to occurence sectin

@@
 
-The GBIF reqistry API services are spread across four major sets
+The GBIF reqistry API services are spread across four sets
 of functions:
@@
 GBIF provides two ways to get occurrence data: through the
-/occurrence/search route (see `occ_search`), or via the
-/occurrence/download route (many functions, see below).
-`occ_search` is more appropriate for smaller data, while
-occ_download*() functions are more appropriate for larger
+`/occurrence/search` route (see `occ_search`), or via the
+`/occurrence/download` route (many functions, see below).
+`occ_search()` is the main funtion for the search route,
+and is more appropriate for smaller data, while
+`occ_download*()` functions are more appropriate for larger
 data requests.
+
+Large is of course a subjective term. When you hit a "large dataset"
+will depend primarily on the size of the your data request. GBIF imposes
+for any given search a limit of 200,000 records in the search service,
+after which point you can't download any more records for that search.
+However, you can download more records for different searches.
+
+We think the search service is still quite useful for many people
+even given the 200,000 limit. For those that need more data, we have
+created a similar interface in the `download_*()` functions, that should
+be easy to use. Users should take note that using the download service
+has a few extra steps to get data into R, but is straight-forward.
 
@@
 The download API syntax is similar to the occurrence search API
-in that the same parameters are used, but the syntas is quite different.
-For example, in the download API you can do greater than searches,
-whereas you can not do that in the occurrence search API.
+in that the same parameters are used, but the way in which the query
+is defined is different. For example, in the download API you can
+do greater than searches, whereas you can not do that in the
+occurrence search API. Thus, we can't make the query interace exactly
+the same for both search and download functions.
+
+Using the download service can be as few as three steps.
+
+Request data download given a query. Here, we xxxx
 
 ```{r}
-occ_download_get("0000066-140928181241064")
+"xxx"
+```
+
+You can check on when the download is ready using the functions
+`occ_download_list()` and `occ_download_meta()`. When it's ready
+use `occ_download_get()` to download the dataset to your computer.
+
+```{r}
+(res <- occ_download_get("0000066-140928181241064", overwrite = TRUE))
+```
+
+What's printed out above is a very brief summary of what was downloaded,
+the path to the file, and its size (in human readable form).
+
+Next, read the data in to R using the function `occ_download_import()`.
+
+```{r}
+library("dplyr")
+dat <- occ_download_import(res)
+dat %>%
+  select(gbifID, decimalLatitude, decimalLongitude)
 ```

### 2015-10-05 manuscript.tex
deleted cache stuff, gitignoring now, updates to occurence sectin

@@
 
-The GBIF reqistry API services are spread across four major sets of
-functions:
+The GBIF reqistry API services are spread across four sets of functions:
 
@@
 GBIF provides two ways to get occurrence data: through the
-/occurrence/search route (see \texttt{occ\_search}), or via the
-/occurrence/download route (many functions, see below).
-\texttt{occ\_search} is more appropriate for smaller data, while
-occ\_download*() functions are more appropriate for larger data
-requests.
+\texttt{/occurrence/search} route (see \texttt{occ\_search}), or via the
+\texttt{/occurrence/download} route (many functions, see below).
+\texttt{occ\_search()} is the main funtion for the search route, and is
+more appropriate for smaller data, while \texttt{occ\_download*()}
+functions are more appropriate for larger data requests.
+
+Large is of course a subjective term. When you hit a ``large dataset''
+will depend primarily on the size of the your data request. GBIF imposes
+for any given search a limit of 200,000 records in the search service,
+after which point you can't download any more records for that search.
+However, you can download more records for different searches.
+
+We think the search service is still quite useful for many people even
+given the 200,000 limit. For those that need more data, we have created
+a similar interface in the \texttt{download\_*()} functions, that should
+be easy to use. Users should take note that using the download service
+has a few extra steps to get data into R, but is straight-forward.
 
@@
 The download API syntax is similar to the occurrence search API in that
-the same parameters are used, but the syntas is quite different. For
-example, in the download API you can do greater than searches, whereas
-you can not do that in the occurrence search API.
+the same parameters are used, but the way in which the query is defined
+is different. For example, in the download API you can do greater than
+searches, whereas you can not do that in the occurrence search API.
+Thus, we can't make the query interace exactly the same for both search
+and download functions.
+
+Using the download service can be as few as three steps.
+
+Request data download given a query. Here, we xxxx
 
@@
 \begin{Highlighting}[]
-\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000066-140928181241064"}\NormalTok{)}
+\StringTok{"xxx"}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+[1] "xxx"
+\end{verbatim}
+
+You can check on when the download is ready using the functions
+\texttt{occ\_download\_list()} and \texttt{occ\_download\_meta()}. When
+it's ready use \texttt{occ\_download\_get()} to download the dataset to
+your computer.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(res <-}\StringTok{ }\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000066-140928181241064"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
 \end{Highlighting}
@@
   File size: 0.14 MB
+\end{verbatim}
+
+What's printed out above is a very brief summary of what was downloaded,
+the path to the file, and its size (in human readable form).
+
+Next, read the data in to R using the function
+\texttt{occ\_download\_import()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_download_import}\NormalTok{(res)}
+\NormalTok{dat %>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+      gbifID decimalLatitude decimalLongitude
+1  657590544              NA               NA
+2  657679551              NA               NA
+3  657791316        37.70805        -118.4162
+4  658180562              NA               NA
+5  441881672              NA               NA
+6  911596181              NA               NA
+7   56454601              NA               NA
+8  657848913              NA               NA
+9  658187373              NA               NA
+10 658279212        38.95917        -106.9892
+..       ...             ...              ...
 \end{verbatim}

### 2015-10-05 manuscript.Rmd
more writing ✏ ✏

@@
 ---
-#layout: preprint
 layout: review, 11pt
 linenumbers: true
-title: "rgbif: a package for working with species occurrence data in R"
+title: "rgbif: R client for working with GBIF species occurrence data"
 author:
@@
 bibliography: components/references.bib
-csl: components/methods-in-ecology-and-evolution.csl
+csl: components/peerj.csl
 documentclass: components/elsarticle
@@
 is defined is different. For example, in the download API you can
-do greater than searches, whereas you can not do that in the
-occurrence search API. Thus, we can't make the query interace exactly
-the same for both search and download functions.
+do greater than searches (i.e., `latitude > 50`), whereas you can not
+do that in the occurrence search API. Thus, we can't make the query
+interace exactly the same for both search and download functions.
 
-Using the download service can be as few as three steps.
+Using the download service can be as few as three steps: 1) Request data
+via a search; 2) Download data; 3) Import data into R.
 
-Request data download given a query. Here, we xxxx
+Request data download given a query. Here, we search for the taxon key
+`3119195`, which is the key for _Helianthus annuus_
+(http://www.gbif.org/species/3119195).
 
 ```{r}
-"xxx"
+occ_download('taxonKey = 3119195')
 ```
@@
 ```{r}
-(res <- occ_download_get("0000066-140928181241064", overwrite = TRUE))
+(res <- occ_download_get("0000840-150615163101818", overwrite = TRUE))
 ```
@@
 
+#### Downloaded data format
+
+The downloaded dataset from GBIF is actually a Darwin Core Archive (DwC-A), an
+internationally recognized biodiversity informatics standard
+(http://rs.tdwg.org/dwc/). The DwC-A downloaded is a compressed folder with
+a number of files, including metadata, citations for each of the datasets included
+in the download, and the data itself, in separate files for each dataset as well
+as one single `.txt` file. In `occ_download_import()`, we simply fetch data from the
+`.txt` file. If you want to dig into the metadata, citations, etc., it is easily
+accessible from the folder on your computer.
+
 ### Search API
 
-The search API is very similar to the download API, but is meant for smaller
-data acquisition jobs.
+The search API follows the GBIF API and is broken down into the following functions:
+
+* `occ_count()`
+* `occ_search()`
+* `occ_get()`
+* `occ_metadata()`
+
+The main search work-horse is `occ_search()`. This function allows very flexible
+search definitions. In addition, this function does paging internally, making it
+such that the user does not have worry about the 300 records per request limit -
+but of course we can't go over the 200,000 maximum limit.
+
+...
+
+#### Cleaning data
+
+GBIF provides optional data issues with each occurrence record. These issues fall
+into many different pre-defined classes, covering issues with taxonomic names,
+geographic data, and more (see `occ_issues_lookup()` to find out more information
+on GBIF issues; and the same data on [GBIF's development site][gbifissues]).
+
+`occ_issues()` provides a way to easily filter data downloaded via `occ_search()`
+based on GBIF issues.
+
+```{r}
+out <- occ_search(issue='DEPTH_UNLIKELY', limit = 500)
+NROW(out)
+out %>% occ_issues(-cudc) %>% .$data %>% NROW
+```
+
+## Use cases
+
+...
 
@@
 
-All software, scripts and data used in this paper can be found in the permanent
-data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/sckott/msrgbif](https://github.com/sckott/msrgbif).
+All scripts and data used in this paper can be found in the permanent
+data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
+can be found at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif), under the
+open and permissive MIT license.
 
@@
 [gbifdocs]: http://www.gbif.org/developer/summary
+[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html

### 2015-10-05 manuscript.tex
more writing ✏ ✏

@@
             pdfauthor={},
-            pdftitle={rgbif: a package for working with species occurrence data in R},
+            pdftitle={rgbif: R client for working with GBIF species occurrence data},
             colorlinks=true,
@@
 
-  \title{rgbif: a package for working with species occurrence data in R}
+  \title{rgbif: R client for working with GBIF species occurrence data}
     \author[cstar]{Scott Chamberlain\corref{c1}}
@@
 easily take advantage of existing data. In fact, some ecology fields are
-built largely on existing data, e.g., macro-ecology (Brown 1995; Beck
-\emph{et al.} 2012).
+built largely on existing data, e.g., macro-ecology (Brown, 1995; Beck
+et al., 2012).
 
@@
 conservation. Biodiversity loss is one of the greatest challenges of our
-time (Pimm \emph{et al.} 2014). Some have called this the sixth great
-mass extinction (Ceballos \emph{et al.} 2015). Given this challenge
-there is a great need for data on specimen records, whether collected
-from live sightings in the field or specimens in museums.
+time (Pimm et al., 2014). Some have called this the sixth great mass
+extinction (Ceballos et al., 2015). Given this challenge there is a
+great need for data on specimen records, whether collected from live
+sightings in the field or specimens in museums.
 
@@
 
-Herein, we describe a library (rgbif (Chamberlain \emph{et al.})) for
-working with GBIF data in the R programming environment (R Core Team
-2014). R is an extremely widely used language in academia, and in
-non-profit and private sectors. Importantly, R makes it easy to do all
-of the steps of the research process, including data management, data
-manipulation and cleaning, statistics, and vizualization. Thus, an R
-client for getting GBIF data is a powerful tool for reproducible
-research.
+Herein, we describe a library (rgbif (Chamberlain et al.)) for working
+with GBIF data in the R programming environment (R Core Team, 2014). R
+is an extremely widely used language in academia, and in non-profit and
+private sectors. Importantly, R makes it easy to do all of the steps of
+the research process, including data management, data manipulation and
+cleaning, statistics, and vizualization. Thus, an R client for getting
+GBIF data is a powerful tool for reproducible research.
 
@@
 is different. For example, in the download API you can do greater than
-searches, whereas you can not do that in the occurrence search API.
-Thus, we can't make the query interace exactly the same for both search
-and download functions.
+searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
+not do that in the occurrence search API. Thus, we can't make the query
+interace exactly the same for both search and download functions.
 
-Using the download service can be as few as three steps.
+Using the download service can be as few as three steps: 1) Request data
+via a search; 2) Download data; 3) Import data into R.
 
-Request data download given a query. Here, we xxxx
+Request data download given a query. Here, we search for the taxon key
+\texttt{3119195}, which is the key for \emph{Helianthus annuus}
+(\url{http://www.gbif.org/species/3119195}).
 
@@
 \begin{Highlighting}[]
-\StringTok{"xxx"}
+\KeywordTok{occ_download}\NormalTok{(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{verbatim}
-[1] "xxx"
+<<gbif download>>
+  Username: sckott
+  E-mail: myrmecocystus@gmail.com
+  Download key: 0004098-150922153815467
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\NormalTok{(res <-}\StringTok{ }\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000066-140928181241064"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
+\NormalTok{(res <-}\StringTok{ }\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000840-150615163101818"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
 \end{Highlighting}
@@
 <<gbif downloaded get>>
-  Path: ./0000066-140928181241064.zip
-  File size: 0.14 MB
+  Path: ./0000840-150615163101818.zip
+  File size: 3.19 MB
 \end{verbatim}
@@
 
+\paragraph{Downloaded data format}\label{downloaded-data-format}
+
+The downloaded dataset from GBIF is actually a Darwin Core Archive
+(DwC-A), an internationally recognized biodiversity informatics standard
+(\url{http://rs.tdwg.org/dwc/}). The DwC-A downloaded is a compressed
+folder with a number of files, including metadata, citations for each of
+the datasets included in the download, and the data itself, in separate
+files for each dataset as well as one single \texttt{.txt} file. In
+\texttt{occ\_download\_import()}, we simply fetch data from the
+\texttt{.txt} file. If you want to dig into the metadata, citations,
+etc., it is easily accessible from the folder on your computer.
+
 \subsubsection{Search API}\label{search-api}
 
-The search API is very similar to the download API, but is meant for
-smaller data acquisition jobs.
+The search API follows the GBIF API and is broken down into the
+following functions:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \texttt{occ\_count()}
+\item
+  \texttt{occ\_search()}
+\item
+  \texttt{occ\_get()}
+\item
+  \texttt{occ\_metadata()}
+\end{itemize}
+
+The main search work-horse is \texttt{occ\_search()}. This function
+allows very flexible search definitions. In addition, this function does
+paging internally, making it such that the user does not have worry
+about the 300 records per request limit - but of course we can't go over
+the 200,000 maximum limit.
+
+\ldots{}
+
+\paragraph{Cleaning data}\label{cleaning-data}
+
+GBIF provides optional data issues with each occurrence record. These
+issues fall into many different pre-defined classes, covering issues
+with taxonomic names, geographic data, and more (see
+\texttt{occ\_issues\_lookup()} to find out more information on GBIF
+issues; and the same data on
+\href{http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html}{GBIF's
+development site}).
+
+\texttt{occ\_issues()} provides a way to easily filter data downloaded
+via \texttt{occ\_search()} based on GBIF issues.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{issue=}\StringTok{'DEPTH_UNLIKELY'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
+\KeywordTok{NROW}\NormalTok{(out)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+[1] 4
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+[1] 2
+\end{verbatim}
+
+\subsection{Use cases}\label{use-cases}
+
+\ldots{}
 
@@
 
-All software, scripts and data used in this paper can be found in the
-permanent data archive Zenodo under the digital object identifier (DOI).
-This DOI corresponds to a snapshot of the GitHub repository at
+All scripts and data used in this paper can be found in the permanent
+data archive Zenodo under the digital object identifier (DOI). This DOI
+corresponds to a snapshot of the GitHub repository at
 \href{https://github.com/sckott/msrgbif}{github.com/sckott/msrgbif}.
+Software can be found at
+\href{https://github.com/ropensci/rgbif}{github.com/ropensci/rgbif},
+under the open and permissive MIT license.
 
@@
 
-Beck, J., Ballesteros-Mejia, L., Buchmann, C.M., Dengler, J., Fritz,
-S.A., Gruber, B., Hof, C., Jansen, F., Knapp, S., Kreft, H., Schneider,
-A.-K., Winter, M. \& Dormann, C.F. (2012). Whats on the horizon for
-macroecology? \emph{Ecography}, \textbf{35}, 673--683. Retrieved from
-\url{http://dx.doi.org/10.1111/j.1600-0587.2012.07364.x}
+Beck J., Ballesteros-Mejia L., Buchmann CM., Dengler J., Fritz SA.,
+Gruber B., Hof C., Jansen F., Knapp S., Kreft H., Schneider A-K., Winter
+M., Dormann CF. 2012. Whats on the horizon for macroecology?
+\emph{Ecography} 35:673--683.
 
-Brown, J.H. (1995). \emph{Macroecology}. University of Chicago Press.
+Brown JH. 1995. \emph{Macroecology}. University of Chicago Press.
 
-Ceballos, G., Ehrlich, P.R., Barnosky, A.D., Garcia, A., Pringle, R.M.
-\& Palmer, T.M. (2015). Accelerated modern human-induced species losses:
-Entering the sixth mass extinction. \emph{Science Advances}, \textbf{1},
-e1400253--e1400253. Retrieved from
-\url{http://dx.doi.org/10.1126/sciadv.1400253}
+Ceballos G., Ehrlich PR., Barnosky AD., Garcia A., Pringle RM., Palmer
+TM. 2015. Accelerated modern human-induced species losses: Entering the
+sixth mass extinction. \emph{Science Advances} 1:e1400253--e1400253.
 
-Chamberlain, S., Ram, K., Barve, V. \& Mcglinn, D. \emph{Rgbif:
-Interface to the global 'biodiversity' information facility 'aPI'}.
-Retrieved from \url{https://github.com/ropensci/rgbif}
+Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{Rgbif: Interface to
+the global 'biodiversity' information facility 'aPI'}.
 
-Pimm, S.L., Jenkins, C.N., Abell, R., Brooks, T.M., Gittleman, J.L.,
-Joppa, L.N., Raven, P.H., Roberts, C.M. \& Sexton, J.O. (2014). The
-biodiversity of species and their rates of extinction, distribution, and
-protection. \emph{Science}, \textbf{344}, 1246752--1246752. Retrieved
-from \url{http://dx.doi.org/10.1126/science.1246752}
+Pimm SL., Jenkins CN., Abell R., Brooks TM., Gittleman JL., Joppa LN.,
+Raven PH., Roberts CM., Sexton JO. 2014. The biodiversity of species and
+their rates of extinction, distribution, and protection. \emph{Science}
+344:1246752--1246752.
 
-R Core Team. (2014). \emph{R: A language and environment for statistical
-computing}. R Foundation for Statistical Computing, Vienna, Austria.
-Retrieved from \url{http://www.R-project.org/}
+R Core Team. 2014. \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing.
 

### 2015-10-06 manuscript.Rmd
dont run occ_download
edits, rendered ms

@@
 library("knitr")
-opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE,
-               cache = 1, comment = NA, verbose = TRUE)
+opts_chunk$set(
+  tidy = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  cache = 1,
+  comment = '#>',
+  collapse = TRUE,
+  verbose = TRUE
+)
 basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
@@
 
-```{r}
+```{r eval = FALSE}
 occ_download('taxonKey = 3119195')
+#> <<gbif download>>
+#>   Username: xxxx
+#>   E-mail: xxxx
+#>   Download key: 0000840-150615163101818
 ```
@@
 
-...
+### A
+
+### B
+
+### C
 
 # Conclusions and future directions
+
+* pt 1
+* pt 2
+* pt 3
+* pt 4
 

### 2015-10-06 manuscript.tex
edits, rendered ms

@@
 \NormalTok{res$data$datasetTitle[}\DecValTok{1}\NormalTok{:}\DecValTok{10}\NormalTok{]}
+\CommentTok{#>  [1] "SDNHM Birds Collection"                                                }
+\CommentTok{#>  [2] "CM Birds Collection"                                                   }
+\CommentTok{#>  [3] "condoncollection"                                                      }
+\CommentTok{#>  [4] "Taxonomy in Flux Checklist"                                            }
+\CommentTok{#>  [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"     }
+\CommentTok{#>  [6] "Bryophyte Collection - University of Washington Herbarium (WTU)"       }
+\CommentTok{#>  [7] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"}
+\CommentTok{#>  [8] "UWFC Ichthyology Collection"                                           }
+\CommentTok{#>  [9] "Lichen Collection - University of Washington Herbarium (WTU)"          }
+\CommentTok{#> [10] "UWBM Mammalogy Collection"}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
- [1] "SDNHM Birds Collection"                                                
- [2] "CM Birds Collection"                                                   
- [3] "condoncollection"                                                      
- [4] "Taxonomy in Flux Checklist"                                            
- [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"     
- [6] "Bryophyte Collection - University of Washington Herbarium (WTU)"       
- [7] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"
- [8] "UWFC Ichthyology Collection"                                           
- [9] "Lichen Collection - University of Washington Herbarium (WTU)"          
-[10] "UWBM Mammalogy Collection"                                             
-\end{verbatim}
 
@@
 \KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit=}\DecValTok{10}\NormalTok{)$data$title}
+\CommentTok{#>  [1] "GBIF Backbone Sources"                                      }
+\CommentTok{#>  [2] "Canadensys"                                                 }
+\CommentTok{#>  [3] "Southwest Collections of Arthropods Network (SCAN)"         }
+\CommentTok{#>  [4] "VertNet"                                                    }
+\CommentTok{#>  [5] "Dryad"                                                      }
+\CommentTok{#>  [6] "GBIF Network"                                               }
+\CommentTok{#>  [7] "The Knowledge Network for Biocomplexity (KNB) "             }
+\CommentTok{#>  [8] "Online Zoological Collections of Australian Museums (OZCAM)"}
+\CommentTok{#>  [9] "Catalogue of Life"                                          }
+\CommentTok{#> [10] "Ocean Biogeographic Information System (OBIS)"}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
- [1] "GBIF Backbone Sources"                                      
- [2] "Canadensys"                                                 
- [3] "Southwest Collections of Arthropods Network (SCAN)"         
- [4] "VertNet"                                                    
- [5] "Dryad"                                                      
- [6] "GBIF Network"                                               
- [7] "The Knowledge Network for Biocomplexity (KNB) "             
- [8] "Online Zoological Collections of Australian Museums (OZCAM)"
- [9] "Catalogue of Life"                                          
-[10] "Ocean Biogeographic Information System (OBIS)"              
-\end{verbatim}
 
@@
 \KeywordTok{occ_download}\NormalTok{(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
+\CommentTok{#> <<gbif download>>}
+\CommentTok{#>   Username: xxxx}
+\CommentTok{#>   E-mail: xxxx}
+\CommentTok{#>   Download key: 0000840-150615163101818}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-<<gbif download>>
-  Username: sckott
-  E-mail: myrmecocystus@gmail.com
-  Download key: 0004098-150922153815467
-\end{verbatim}
 
@@
 \NormalTok{(res <-}\StringTok{ }\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000840-150615163101818"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
+\CommentTok{#> <<gbif downloaded get>>}
+\CommentTok{#>   Path: ./0000840-150615163101818.zip}
+\CommentTok{#>   File size: 3.19 MB}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-<<gbif downloaded get>>
-  Path: ./0000840-150615163101818.zip
-  File size: 3.19 MB
-\end{verbatim}
 
@@
 \StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
+\CommentTok{#>       gbifID decimalLatitude decimalLongitude}
+\CommentTok{#> 1  657590544              NA               NA}
+\CommentTok{#> 2  657679551              NA               NA}
+\CommentTok{#> 3  657791316        37.70805        -118.4162}
+\CommentTok{#> 4  658180562              NA               NA}
+\CommentTok{#> 5  441881672              NA               NA}
+\CommentTok{#> 6  911596181              NA               NA}
+\CommentTok{#> 7   56454601              NA               NA}
+\CommentTok{#> 8  657848913              NA               NA}
+\CommentTok{#> 9  658187373              NA               NA}
+\CommentTok{#> 10 658279212        38.95917        -106.9892}
+\CommentTok{#> ..       ...             ...              ...}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-      gbifID decimalLatitude decimalLongitude
-1  657590544              NA               NA
-2  657679551              NA               NA
-3  657791316        37.70805        -118.4162
-4  658180562              NA               NA
-5  441881672              NA               NA
-6  911596181              NA               NA
-7   56454601              NA               NA
-8  657848913              NA               NA
-9  658187373              NA               NA
-10 658279212        38.95917        -106.9892
-..       ...             ...              ...
-\end{verbatim}
 
@@
 \KeywordTok{NROW}\NormalTok{(out)}
+\CommentTok{#> [1] 4}
+\NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
+\CommentTok{#> [1] 2}
 \end{Highlighting}
@@
 
-\begin{verbatim}
-[1] 4
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-[1] 2
-\end{verbatim}
-
 \subsection{Use cases}\label{use-cases}
 
-\ldots{}
+\subsubsection{A}\label{a}
+
+\subsubsection{B}\label{b}
+
+\subsubsection{C}\label{c}
 
@@
 directions}\label{conclusions-and-future-directions}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  pt 1
+\item
+  pt 2
+\item
+  pt 3
+\item
+  pt 4
+\end{itemize}
 

### 2015-10-06 manuscript.Rmd
more intro text, rebuilt pdf

@@
 Data on individuals, including which species, and where they're found,
-can be used for a large number of research questions. In addition,
-the pool of questions we can answer becomes much larger with more
-and better data. In addition to wide utility, this data is important
-for conservation. Biodiversity loss is one of the greatest challenges
-of our time [@Pimm_2014]. Some have called this the sixth great mass
+can be used for a large number of research questions. Biodiversity
+records have been used for a suite of other use cases:
+validating habitat suitability models with real occurrence data
+[@Ficetola_2014]; ancestral range reconstruction [@Ferretti_2015;
+@Mendoza_2015]; development of invasive species watch lists
+[@Faulkner_2014]; evaluate risk of invasive species spread
+[@Di_Febbraro_2013]; and effects of climate change on future
+biodiversity [@Brown_2015].
+
+In addition to wide utility, this data is important for conservation.
+Biodiversity loss is one of the greatest challenges of our time
+[@Pimm_2014]. Some have called this the sixth great mass
 extinction [@Ceballos_2015]. Given this challenge there is a great
@@
 taxa, 15,000 datasets from 770 publishers (figures collected on 2015-10-04).
-Many large biodiversity warehouses such as iNaturalist, VertNet, and USGS's
-BISON all feed into GBIF.
+Many large biodiversity warehouses such as iNaturalist
+(http://www.inaturalist.org/), VertNet (http://vertnet.org/), and USGS's
+Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov/)
+all feed into GBIF.
 
-Herein, we describe a library (rgbif [@rgbif]) for working with
-GBIF data in the R programming environment [@R]. R is an extremely
-widely used language in academia, and in non-profit and private sectors.
+Herein, we describe the `rgbif` library [@rgbif] for working with
+GBIF data in the R programming environment [@R]. R is a widely used
+language in academia, and in non-profit and private sectors.
 Importantly, R makes it easy to do all of the steps of the
@@
 and cleaning, statistics, and vizualization. Thus, an R client
-for getting GBIF data is a powerful tool for reproducible research.
+for getting GBIF data is a powerful tool to facilitate
+reproducible research.
 

### 2015-10-06 manuscript.tex
more intro text, rebuilt pdf

@@
 Data on individuals, including which species, and where they're found,
-can be used for a large number of research questions. In addition, the
-pool of questions we can answer becomes much larger with more and better
-data. In addition to wide utility, this data is important for
-conservation. Biodiversity loss is one of the greatest challenges of our
-time (Pimm et al., 2014). Some have called this the sixth great mass
-extinction (Ceballos et al., 2015). Given this challenge there is a
-great need for data on specimen records, whether collected from live
-sightings in the field or specimens in museums.
+can be used for a large number of research questions. Biodiversity
+records have been used for a suite of other use cases: validating
+habitat suitability models with real occurrence data (Ficetola et al.,
+2014); ancestral range reconstruction (Ferretti et al., 2015; Mar{í}a
+Mendoza et al., 2015); development of invasive species watch lists
+(Faulkner et al., 2014); evaluate risk of invasive species spread
+(Febbraro et al., 2013); and effects of climate change on future
+biodiversity (Brown et al., 2015).
+
+In addition to wide utility, this data is important for conservation.
+Biodiversity loss is one of the greatest challenges of our time (Pimm et
+al., 2014). Some have called this the sixth great mass extinction
+(Ceballos et al., 2015). Given this challenge there is a great need for
+data on specimen records, whether collected from live sightings in the
+field or specimens in museums.
 
@@
 million taxa, 15,000 datasets from 770 publishers (figures collected on
-2015-10-04). Many large biodiversity warehouses such as iNaturalist,
-VertNet, and USGS's BISON all feed into GBIF.
+2015-10-04). Many large biodiversity warehouses such as iNaturalist
+(\url{http://www.inaturalist.org/}), VertNet
+(\url{http://vertnet.org/}), and USGS's Biodiversity Information Serving
+Our Nation (BISON; \url{http://bison.usgs.ornl.gov/}) all feed into
+GBIF.
 
-Herein, we describe a library (rgbif (Chamberlain et al.)) for working
-with GBIF data in the R programming environment (R Core Team, 2014). R
-is an extremely widely used language in academia, and in non-profit and
+Herein, we describe the \texttt{rgbif} library (Chamberlain et al.) for
+working with GBIF data in the R programming environment (R Core Team,
+2014). R is a widely used language in academia, and in non-profit and
 private sectors. Importantly, R makes it easy to do all of the steps of
@@
 cleaning, statistics, and vizualization. Thus, an R client for getting
-GBIF data is a powerful tool for reproducible research.
+GBIF data is a powerful tool to facilitate reproducible research.
 
@@
 
+Brown KA., Parks KE., Bethell CA., Johnson SE., Mulligan M. 2015.
+Predicting plant diversity patterns in madagascar: Understanding the
+effects of climate and land cover change in a biodiversity hotspot.
+\emph{PLOS ONE} 10:e0122721.
+
 Ceballos G., Ehrlich PR., Barnosky AD., Garcia A., Pringle RM., Palmer
@@
 the global 'biodiversity' information facility 'aPI'}.
+
+Faulkner KT., Robertson MP., Rouget M., Wilson JR. 2014. A simple, rapid
+methodology for developing invasive species watch lists.
+\emph{Biological Conservation} 179:25--32.
+
+Febbraro MD., Lurz PWW., Genovesi P., Maiorano L., Girardello M.,
+Bertolino S. 2013. The use of climatic niches in screening procedures
+for introduced species to evaluate risk of spread: A case with the
+american eastern grey squirrel. \emph{PLoS ONE} 8:e66559.
+
+Ferretti F., Verd GM., Seret B., {Š}prem JS., Micheli F. 2015. Falling
+through the cracks: The fading history of a large iconic predator.
+\emph{Fish and Fisheries}:n/a--n/a.
+
+Ficetola GF., Rondinini C., Bonardi A., Baisero D., Padoa-Schioppa E.
+2014. Habitat availability for amphibians and extinction threat: A
+global analysis. \emph{Diversity and Distributions} 21:302--311.
+
+Mar{í}a Mendoza., Ospina OE., C{á}rdenas-Henao H., Garc{í}a-R JC. 2015.
+A likelihood inference of historical biogeography in the world's most
+diverse terrestrial vertebrate genus: Diversification of
+direct-developing frogs (craugastoridae: Pristimantis) across the
+neotropics. \emph{Molecular Phylogenetics and Evolution} 85:50--58.
 

### 2015-10-27 manuscript.Rmd
updated registry section

@@
 or API. The GBIF API has four major components: registry, species names,
-occurrence data, and maps. We ignore maps in `rgbif` as it is concerned with
-generating maps for web applications. `rgbif` has a suite of functions dealing
-with each of registry, species names, and occurrence data - we'll go through
+occurrences, and maps. We ignore maps in `rgbif` as it is concerned with
+generating maps primarily for web applications. `rgbif` has a suite of functions dealing
+with each of registry, species names, and occurrences - we'll go through
 each in turn describing design and example usage.
@@
 
+The relationship between these five things is XXXXXX.
+
 ### Datasets
 
-Search for datasets
+Dataset functions include search, dataset metadata retrieval, and dataset metrics.
+Searching for datasets is an important part of the discovery process. One can
+search for datasets on the GBIF web portal. However, programmatic searching using
+this package is much more powerful. Identifying datasets appropriate for a
+research question is helpful as you can get metadata for each dataset, and track
+down dataset specific problems, if any.
+
+The `dataset_search()` function is one way to search for datasets. Here, we search
+for the query term "oregon", which finds any datasets that have terms matching
+that term.
 
@@
 
-Get dataset metrics
+Also, check out `datasets()` and `dataset_suggest()` for searching for datasets.
+
+#### Dataset metrics
+
+Dataset metrics are another useful way of digging in to figure out what
+datasets you may want to use. One drawback is that these data are only
+available for datasets of type _checklist_, but there are quite a lot of
+them (`r dataset_search(type = "checklist")$meta$count`).
+
+Here, we seaerch for dataset metrics for a single dataset, with uuid
+`ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
+Formicidae) of the World.
 
 ```{r}
-res <- dataset_metrics(uuid='66dd0960-2d7d-46ee-a491-87b9adcfe7b1')
+res <- dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
 df <- data.frame(rank = names(res$countByRank),
@@
 
-Here, we search for the first give GBIF networks, returning just the
-key and title fields.
+Networks, nodes and installations are at a higher level of organization
+above datasets, but can be useful if you want to explore data from given
+organizations. Here, we search for the first 10 GBIF networks, returning
+just the title field.
 

### 2015-10-27 manuscript.tex
updated registry section

@@
 Programming Interface}, or API. The GBIF API has four major components:
-registry, species names, occurrence data, and maps. We ignore maps in
-\texttt{rgbif} as it is concerned with generating maps for web
+registry, species names, occurrences, and maps. We ignore maps in
+\texttt{rgbif} as it is concerned with generating maps primarily for web
 applications. \texttt{rgbif} has a suite of functions dealing with each
-of registry, species names, and occurrence data - we'll go through each
-in turn describing design and example usage.
+of registry, species names, and occurrences - we'll go through each in
+turn describing design and example usage.
 
@@
 
+The relationship between these five things is XXXXXX.
+
 \subsubsection{Datasets}\label{datasets}
 
-Search for datasets
+Dataset functions include search, dataset metadata retrieval, and
+dataset metrics. Searching for datasets is an important part of the
+discovery process. One can search for datasets on the GBIF web portal.
+However, programmatic searching using this package is much more
+powerful. Identifying datasets appropriate for a research question is
+helpful as you can get metadata for each dataset, and track down dataset
+specific problems, if any.
+
+The \texttt{dataset\_search()} function is one way to search for
+datasets. Here, we search for the query term ``oregon'', which finds any
+datasets that have terms matching that term.
 
@@
 
-Get dataset metrics
+Also, check out \texttt{datasets()} and \texttt{dataset\_suggest()} for
+searching for datasets.
+
+\paragraph{Dataset metrics}\label{dataset-metrics}
+
+Dataset metrics are another useful way of digging in to figure out what
+datasets you may want to use. One drawback is that these data are only
+available for datasets of type \emph{checklist}, but there are quite a
+lot of them (2472).
+
+Here, we seaerch for dataset metrics for a single dataset, with uuid
+\texttt{ec93a739-1681-4b04-b62f-3a687127a17f}, a checklist of the ants
+(Hymenoptera: Formicidae) of the World.
 
@@
 \begin{Highlighting}[]
-\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'66dd0960-2d7d-46ee-a491-87b9adcfe7b1'}\NormalTok{)}
+\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
 \NormalTok{df <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
@@
 \endhead
-SPECIES & 52452\tabularnewline
-GENUS & 12930\tabularnewline
-VARIETY & 4806\tabularnewline
-SUBSPECIES & 4440\tabularnewline
-SERIES & 1079\tabularnewline
-TRIBE & 844\tabularnewline
-FAMILY & 509\tabularnewline
-SUBTRIBE & 327\tabularnewline
-SUBFAMILY & 303\tabularnewline
-SUBGENUS & 241\tabularnewline
-FORM & 239\tabularnewline
-SECTION & 82\tabularnewline
-SUBVARIETY & 4\tabularnewline
+SPECIES & 13710\tabularnewline
+SUBSPECIES & 3234\tabularnewline
+GENUS & 726\tabularnewline
+TRIBE & 53\tabularnewline
+SUBFAMILY & 20\tabularnewline
+FAMILY & 2\tabularnewline
 KINGDOM & 1\tabularnewline
+PHYLUM & 1\tabularnewline
+CLASS & 1\tabularnewline
+ORDER & 1\tabularnewline
 \bottomrule
@@
 
-Here, we search for the first give GBIF networks, returning just the key
-and title fields.
+Networks, nodes and installations are at a higher level of organization
+above datasets, but can be useful if you want to explore data from given
+organizations. Here, we search for the first 10 GBIF networks, returning
+just the title field.
 

### 2015-10-28 manuscript.Rmd
updated names section

@@
 `rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
-or API. The GBIF API has four major components: registry, species names,
+or API. The GBIF API has four major components: registry, taxonomic names,
 occurrences, and maps. We ignore maps in `rgbif` as it is concerned with
 generating maps primarily for web applications. `rgbif` has a suite of functions dealing
-with each of registry, species names, and occurrences - we'll go through
+with each of registry, taxonomic names, and occurrences - we'll go through
 each in turn describing design and example usage.
@@
 
-```{r}
+```{r eval=FALSE}
+res <- dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
+data.frame(rank = names(res$countByRank),
+           count = unname(unlist(res$countByRank)))
+```
+
+```{r echo=FALSE}
 res <- dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
@@
 
-## Species
+## Taxonomic names
+
+The GBIF taxonomic names API services are spread across five functions:
+
+* Search GBIF name backbone - `name_backbone()`
+* Search across all checklists - `name_lookup()`
+* Quick name lookup - `name_suggest()`
+* Name usage of a name according to a checklist - `name_usage()`
+* GBIF name parser - `parsenames()`
+
+The goal of these name functions is often to settle on a taxonomic name
+known to GBIF's database. This serves two purposes: 1) when referring to
+a taxonomic name, you can point to a URI on the internet, and 2) you can
+search for metadata on that species, and occurrences of that species in
+GBIF.
+
+Taxonomic names are particularly tricky. Many different organizations have
+their own unique codes for the same taxonomic names, and some taxonomic
+groups have preferred sources for the definitive names for that group.
+
+When searching for occurrences (see below) you can search by taxonomic
+name (and other filters, e.g., taxonomic rank), but you're probably better off
+figuring out the taxonomic key in the GBIF backbone taxonomy, and using
+that to search for occurrences. The `taxonkey` parameter in the GBIF occurrences
+API expects a GBIF backbone taxon key.
+
+### GBIF Backbone
+
+The GBIF backbone taxonomy is used in GBIF to have a consistent way to refer
+to taxonomic names throughout their services. The backbone has `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$distinctNamesCount` unique
+names and `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$countByRank$SPECIES` species names. The backbone taxonomy is also a dataset with key
+`d7dddbf4-2cf0-4f39-9b2a-bb099caae36c`
+(http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c).
+
+We can search the backbone taxonomy with the function `name_backbone()`. Here,
+we're searching for the name _Poa_, restricting to genera, and the family
+_Poaceae_.
+
+```{r}
+res <- name_backbone(name='Poa', rank='genus', family='Poaceae')
+res[c('usageKey', 'kingdom')]
+```
+
+### Name searching
+
+One of the quickest ways to search for names is using `name_suggest()`, which
+does a very quick search and returns minimal data. Here, we're searching for
+the query tem _Pum_, and we get back a bunch of names:
+
+```{r eval=FALSE}
+name_suggest(q='Pum', limit = 6)
+```
+
+```{r echo=FALSE}
+df <- name_suggest(q='Pum', limit = 6)
+knitr::kable(df)
+```
 
@@
 
-* `occ_count()`
-* `occ_search()`
-* `occ_get()`
-* `occ_metadata()`
+* Get a single numeric count of occurrenes - `occ_count()`
+* Search for occurrences - `occ_search()`
+* Get occurrences by occurrence identifier - `occ_get()`
+* Get occurrence metadata - `occ_metadata()`
 

### 2015-10-28 manuscript.tex
updated names section

@@
 Programming Interface}, or API. The GBIF API has four major components:
-registry, species names, occurrences, and maps. We ignore maps in
+registry, taxonomic names, occurrences, and maps. We ignore maps in
 \texttt{rgbif} as it is concerned with generating maps primarily for web
 applications. \texttt{rgbif} has a suite of functions dealing with each
-of registry, species names, and occurrences - we'll go through each in
+of registry, taxonomic names, and occurrences - we'll go through each in
 turn describing design and example usage.
@@
 \NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
-\NormalTok{df <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
+\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
            \DataTypeTok{count =} \KeywordTok{unname}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(res$countByRank)))}
-\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(df)}
 \end{Highlighting}
@@
 
-\subsection{Species}\label{species}
+\subsection{Taxonomic names}\label{taxonomic-names}
+
+The GBIF taxonomic names API services are spread across five functions:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Search GBIF name backbone - \texttt{name\_backbone()}
+\item
+  Search across all checklists - \texttt{name\_lookup()}
+\item
+  Quick name lookup - \texttt{name\_suggest()}
+\item
+  Name usage of a name according to a checklist - \texttt{name\_usage()}
+\item
+  GBIF name parser - \texttt{parsenames()}
+\end{itemize}
+
+The goal of these name functions is often to settle on a taxonomic name
+known to GBIF's database. This serves two purposes: 1) when referring to
+a taxonomic name, you can point to a URI on the internet, and 2) you can
+search for metadata on that species, and occurrences of that species in
+GBIF.
+
+Taxonomic names are particularly tricky. Many different organizations
+have their own unique codes for the same taxonomic names, and some
+taxonomic groups have preferred sources for the definitive names for
+that group.
+
+When searching for occurrences (see below) you can search by taxonomic
+name (and other filters, e.g., taxonomic rank), but you're probably
+better off figuring out the taxonomic key in the GBIF backbone taxonomy,
+and using that to search for occurrences. The \texttt{taxonkey}
+parameter in the GBIF occurrences API expects a GBIF backbone taxon key.
+
+\subsubsection{GBIF Backbone}\label{gbif-backbone}
+
+The GBIF backbone taxonomy is used in GBIF to have a consistent way to
+refer to taxonomic names throughout their services. The backbone has
+4410899 unique names and 2497114 species names. The backbone taxonomy is
+also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
+(\url{http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).
+
+We can search the backbone taxonomy with the function
+\texttt{name\_backbone()}. Here, we're searching for the name
+\emph{Poa}, restricting to genera, and the family \emph{Poaceae}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Poa'}\NormalTok{, }\DataTypeTok{rank=}\StringTok{'genus'}\NormalTok{, }\DataTypeTok{family=}\StringTok{'Poaceae'}\NormalTok{)}
+\NormalTok{res[}\KeywordTok{c}\NormalTok{(}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{)]}
+\CommentTok{#> $usageKey}
+\CommentTok{#> [1] 2704173}
+\CommentTok{#> }
+\CommentTok{#> $kingdom}
+\CommentTok{#> [1] "Plantae"}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{Name searching}\label{name-searching}
+
+One of the quickest ways to search for names is using
+\texttt{name\_suggest()}, which does a very quick search and returns
+minimal data. Here, we're searching for the query tem \emph{Pum}, and we
+get back a bunch of names:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q=}\StringTok{'Pum'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{6}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[c]{@{}rll@{}}
+\toprule
+key & canonicalName & rank\tabularnewline
+\midrule
+\endhead
+3269133 & Pumilus & GENUS\tabularnewline
+4407849 & Pumilinura & GENUS\tabularnewline
+4323990 & Pumiliopsis & GENUS\tabularnewline
+4324083 & Pumiliopes & GENUS\tabularnewline
+4161281 & Pumilea & GENUS\tabularnewline
+4312370 & Pumilopagurus & GENUS\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 \CommentTok{#>       gbifID decimalLatitude decimalLongitude}
-\CommentTok{#> 1  657590544              NA               NA}
-\CommentTok{#> 2  657679551              NA               NA}
-\CommentTok{#> 3  657791316        37.70805        -118.4162}
-\CommentTok{#> 4  658180562              NA               NA}
-\CommentTok{#> 5  441881672              NA               NA}
-\CommentTok{#> 6  911596181              NA               NA}
-\CommentTok{#> 7   56454601              NA               NA}
-\CommentTok{#> 8  657848913              NA               NA}
-\CommentTok{#> 9  658187373              NA               NA}
-\CommentTok{#> 10 658279212        38.95917        -106.9892}
+\CommentTok{#> 1  725767384        61.01005         24.41740}
+\CommentTok{#> 2  725767447        59.82923         23.13550}
+\CommentTok{#> 3  725767450        60.38505         25.17449}
+\CommentTok{#> 4  725767513        68.37648         23.51963}
+\CommentTok{#> 5  725767546        67.19203         24.85820}
+\CommentTok{#> 6  725767579        60.21607         24.67412}
+\CommentTok{#> 7  725767609        66.49260         25.70471}
+\CommentTok{#> 8  725767645        61.36634         24.76218}
+\CommentTok{#> 9  725767678        62.29174         27.96500}
+\CommentTok{#> 10 725767681        60.28615         22.38489}
 \CommentTok{#> ..       ...             ...              ...}
@@
 \item
-  \texttt{occ\_count()}
+  Get a single numeric count of occurrenes - \texttt{occ\_count()}
 \item
-  \texttt{occ\_search()}
+  Search for occurrences - \texttt{occ\_search()}
 \item
-  \texttt{occ\_get()}
+  Get occurrences by occurrence identifier - \texttt{occ\_get()}
 \item
-  \texttt{occ\_metadata()}
+  Get occurrence metadata - \texttt{occ\_metadata()}
 \end{itemize}

### 2015-10-28 manuscript.tex
update description about datasets, orgs, netowrks, inst, etc., thanks tim

@@
 
-The relationship between these five things is XXXXXX.
+Datasets are owned by organizations. Organizations are endorsed by nodes
+to share datasets with GBIF. Datasets are published through
+institutions, which may be hosted at another organization. A network is
+a group of datasets (managed by GBIF).
 
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2472).
+lot of them (2474).
 

### 2015-11-14 manuscript.Rmd
update description about datasets, orgs, netowrks, inst, etc., thanks tim
added three use cases

@@
 
+## GBIF feedback loop
+
+With each request `rgbif` makes to GBIF's API, we send request headers that
+tell GBIF that the request is coming from `rgbif`, including what version of
+the package. This will help GBIF know what proportion of requests are coming from
+this package, and therefore R, as most requests likely will come from `rgbif`.
+
 ## Registry
@@
 
-The relationship between these five things is XXXXXX.
+Datasets are owned by organizations. Organizations are endorsed by nodes to share datasets with GBIF.
+Datasets are published through institutions, which may be hosted at another organization. A network
+is a group of datasets (managed by GBIF).
 
@@
 
-## Use cases
+# Use cases
 
-### A
+### Ecological niche modelling
 
-### B
+```{r niche-modelling, child="use-cases/niche-modelling.Rmd"}
+```
 
-### C
+### Biodiversity in big cities
+
+```{r city-diversity, child="use-cases/city-diversity.Rmd"}
+```
+
+### Valley oak occurrence data comparison
+
+```{r oak-occurrences, child="use-cases/oak-occurrences.Rmd"}
+```
 
@@
 
-## Acknowledgements
+# Acknowledgements
 
@@
 
-## Data Accessibility
+# Data Accessibility
 
 All scripts and data used in this paper can be found in the permanent
-data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
-can be found at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif), under the
-open and permissive MIT license.
+data archive Zenodo under the digital object identifier (DOI). This DOI
+corresponds to a snapshot of the GitHub repository at
+[github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
+can be found at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
+under the open and permissive MIT license.
 

### 2015-11-14 manuscript.tex
added three use cases

@@
 \usepackage{longtable}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 
+\subsection{GBIF feedback loop}\label{gbif-feedback-loop}
+
+With each request \texttt{rgbif} makes to GBIF's API, we send request
+headers that tell GBIF that the request is coming from \texttt{rgbif},
+including what version of the package. This will help GBIF know what
+proportion of requests are coming from this package, and therefore R, as
+most requests likely will come from \texttt{rgbif}.
+
 \subsection{Registry}\label{registry}
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2474).
+lot of them (2483).
 
@@
 
-\subsection{Use cases}\label{use-cases}
+\section{Use cases}\label{use-cases}
 
-\subsubsection{A}\label{a}
+\subsubsection{Ecological niche
+modelling}\label{ecological-niche-modelling}
 
-\subsubsection{B}\label{b}
+In this example, we plot actual occurrence data for \emph{Bradypus}
+species against a single predictor variable, BIO1 (annual mean
+temperature). This is only ont step in a species distribution modelling
+nworkflow.
 
-\subsubsection{C}\label{c}
+This example can be done using BISON data as well with our rbison
+package.
+
+\emph{Load libraries}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"dismo"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"maptools"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Raster files}
+
+Make a list of files that are installed with the dismo package, then
+create a rasterStack from these
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{files <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\KeywordTok{system.file}\NormalTok{(}\DataTypeTok{package =} \StringTok{"dismo"}\NormalTok{), }\StringTok{"/ex"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{),}
+                    \StringTok{"grd"}\NormalTok{, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{predictors <-}\StringTok{ }\KeywordTok{stack}\NormalTok{(files)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get world boundaries}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{data}\NormalTok{(wrld_simpl)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get GBIF data using the rOpenSci package rgbif}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nn <-}\StringTok{ }\KeywordTok{name_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
+\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn$data$nubKey))}
+\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
+\NormalTok{df <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{class}\NormalTok{(x$data)) %in%}\StringTok{ "data.frame"} \NormalTok{]}
+\NormalTok{df <-}\StringTok{ }\KeywordTok{ldply}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
+\NormalTok{df2 <-}\StringTok{ }\NormalTok{df[,}\KeywordTok{c}\NormalTok{(}\StringTok{'decimalLongitude'}\NormalTok{,}\StringTok{'decimalLatitude'}\NormalTok{)]}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Plot}
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Add raster data, (2) Add political boundaries, (3) Add the points
+  (occurrences)
+\end{enumerate}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{plot}\NormalTok{(predictors, }\DecValTok{1}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(wrld_simpl, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{)}
+\KeywordTok{points}\NormalTok{(df2, }\DataTypeTok{col =} \StringTok{"blue"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-unnamed-chunk-19-1.pdf}
+\caption{}
+\end{figure}
+
+\subsubsection{Biodiversity in big
+cities}\label{biodiversity-in-big-cities}
+
+In this example, we collect specimen records across different cities
+using GBIF data from the \texttt{rgbif} package.
+
+\emph{Load libraries}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"RCurl"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"RColorBrewer"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get bounding boxes for some cites}
+
+Bounding lat/long data is from
+\href{https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt}{here}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{getURL}\NormalTok{(}\StringTok{'https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt'}\NormalTok{)}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}\DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{sep=}\StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{))}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{dat$city, }\DataTypeTok{minlon=}\NormalTok{dat$minlon, }\DataTypeTok{minlat=}\NormalTok{dat$minlat, }\DataTypeTok{maxlon=}\NormalTok{dat$maxlon, }\DataTypeTok{maxlat=}\NormalTok{dat$maxlat)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{getdata <-}\StringTok{ }\NormalTok{function(x)\{}
+  \NormalTok{coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
+  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords)$meta$count}
+  \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{x[}\StringTok{'city'}\NormalTok{], }\DataTypeTok{richness=}\NormalTok{num, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out <-}\StringTok{ }\KeywordTok{apply}\NormalTok{(dat, }\DecValTok{1}\NormalTok{, getdata)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Merge to original table}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by=}\StringTok{"city"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Add centroids from bounding boxes}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat+maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon+maxlon)/}\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Plot data}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mapp <-}\StringTok{ }\KeywordTok{map_data}\NormalTok{(}\StringTok{'world'}\NormalTok{)}
+\KeywordTok{ggplot}\NormalTok{(mapp, }\KeywordTok{aes}\NormalTok{(long, lat)) +}
+\StringTok{  }\KeywordTok{geom_polygon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group=}\NormalTok{group), }\DataTypeTok{fill=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{alpha=}\DecValTok{0}\NormalTok{, }\DataTypeTok{color=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.4}\NormalTok{) +}
+\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{out, }\KeywordTok{aes}\NormalTok{(lon, lat, }\DataTypeTok{color=}\NormalTok{richness), }\DataTypeTok{size=}\DecValTok{5}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) +}
+\StringTok{  }\KeywordTok{scale_color_continuous}\NormalTok{(}\DataTypeTok{low =} \StringTok{"#60E1EE"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#0404C8"}\NormalTok{) +}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{""}\NormalTok{, }\DataTypeTok{y=}\StringTok{""}\NormalTok{) +}
+\StringTok{  }\KeywordTok{theme_grey}\NormalTok{(}\DataTypeTok{base_size=}\DecValTok{14}\NormalTok{) +}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\DataTypeTok{legend.key =} \KeywordTok{element_blank}\NormalTok{()) +}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{color =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{keywidth =} \DecValTok{2}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-unnamed-chunk-24-1.pdf}
+\caption{}
+\end{figure}
+
+\subsubsection{Valley oak occurrence data
+comparison}\label{valley-oak-occurrence-data-comparison}
+
+This example comes from \href{https://twitter.com/ajpelu}{Antonio J.
+Perez-Luque} who
+\href{https://twitter.com/ajpelu/status/473951167567757312}{shared his
+plot on Twitter}. Antonio compared the occurrences of Valley Oak
+(\emph{Quercus lobata}) from \href{http://www.gbif.org/}{GBIF} to the
+distribution of the same species from the
+\href{http://esp.cr.usgs.gov/data/little/}{Atlas of US Trees}.
+
+\emph{Load libraries}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{'rgbif'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{'raster'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{'sp'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{'maptools'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{'rgeos'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{'scales'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get GBIF Data for Quercus lobata}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{keyQl <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Quercus lobata'}\NormalTok{, }\DataTypeTok{kingdom=}\StringTok{'plants'}\NormalTok{)$speciesKey}
+\NormalTok{dat.Ql <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey=}\NormalTok{keyQl, }\DataTypeTok{return=}\StringTok{'data'}\NormalTok{, }\DataTypeTok{limit=}\DecValTok{50000}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)}
+
+From
+\href{http://esp.cr.usgs.gov/data/little/}{\url{http://esp.cr.usgs.gov/data/little/}}.
+And save shapefile in same directory
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{url <-}\StringTok{ 'http://esp.cr.usgs.gov/data/little/querloba.zip'}
+\NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
+\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"~/querloba.zip"}\NormalTok{)}
+\KeywordTok{unzip}\NormalTok{(}\StringTok{"~/querloba.zip"}\NormalTok{, }\DataTypeTok{exdir =} \NormalTok{tmp)}
+\NormalTok{ql <-}\StringTok{ }\KeywordTok{readShapePoly}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"querloba.shp"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get Elevation data of US}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{alt.USA <-}\StringTok{ }\KeywordTok{getData}\NormalTok{(}\StringTok{'alt'}\NormalTok{, }\DataTypeTok{country =} \StringTok{'USA'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Create Hillshade of US}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{alt.USA <-}\StringTok{ }\NormalTok{alt.USA[[}\DecValTok{1}\NormalTok{]]}
+\NormalTok{slope.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'slope'}\NormalTok{)}
+\NormalTok{aspect.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'aspect'}\NormalTok{)}
+\NormalTok{hill.USA <-}\StringTok{ }\KeywordTok{hillShade}\NormalTok{(slope.USA, aspect.USA, }\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{direction =} \DecValTok{315}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Plot map}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{), }\DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
+\CommentTok{# add shape from Atlas of US Trees}
+\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
+\CommentTok{# add Gbif presence points}
+\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{, }\StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E.
+1971)"}\NormalTok{, }\DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
+\caption{}
+\end{figure}
 
@@
 
-\subsection{Acknowledgements}\label{acknowledgements}
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 
-\subsection{Data Accessibility}\label{data-accessibility}
+\section{Data Accessibility}\label{data-accessibility}
 

### 2015-11-14 use-cases/city-diversity.Rmd
added three use cases

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  fig.width=8,
  fig.height=6,
  cache.path = "cache/"
)
```

In this example, we collect specimen records across different cities using GBIF data from the `rgbif` package.

_Load libraries_

```{r load}
library("rgbif")
library("ggplot2")
library("plyr")
library("RCurl")
library("RColorBrewer")
```

_Get bounding boxes for some cites_

Bounding lat/long data is from [here](https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt).

```{r box}
rawdat <- getURL('https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt')
dat <- read.table(text = rawdat, header = FALSE, sep="\t", col.names=c("city","minlat","maxlon","maxlat","minlon"))
dat <- data.frame(city=dat$city, minlon=dat$minlon, minlat=dat$minlat, maxlon=dat$maxlon, maxlat=dat$maxlat)
```

```{r newsearch}
getdata <- function(x){
  coords <- as.numeric(x[c('minlon','minlat','maxlon','maxlat')])
  num <- occ_search(geometry = coords)$meta$count
  data.frame(city=x['city'], richness=num, stringsAsFactors = FALSE)
}
```

```{r cache=TRUE}
out <- apply(dat, 1, getdata)
```

_Merge to original table_

```{r}
out <- merge(dat, ldply(out), by="city")
```

_Add centroids from bounding boxes_

```{r}
out <- transform(out, lat = (minlat+maxlat)/2, lon = (minlon+maxlon)/2)
```

_Plot data_

```{r}
mapp <- map_data('world')
ggplot(mapp, aes(long, lat)) +
  geom_polygon(aes(group=group), fill="white", alpha=0, color="black", size=0.4) +
  geom_point(data=out, aes(lon, lat, color=richness), size=5, alpha=0.8) +
  scale_color_continuous(low = "#60E1EE", high = "#0404C8") +
  labs(x="", y="") +
  theme_grey(base_size=14) +
  theme(legend.position = "bottom", legend.key = element_blank()) +
  guides(color = guide_legend(keywidth = 2))
```

### 2015-11-14 use-cases/niche-modelling.Rmd
added three use cases

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

In this example, we plot actual occurrence data for _Bradypus_ species against a single predictor variable, BIO1 (annual mean temperature). This is only ont step in a species distribution modelling nworkflow.

This example can be done using BISON data as well with our rbison package.

_Load libraries_

```{r}
library("rgbif")
library("dismo")
library("maptools")
library("plyr")
```

_Raster files_

Make a list of files that are installed with the dismo package, then create a rasterStack from these

```{r}
files <- list.files(paste(system.file(package = "dismo"), "/ex", sep = ""),
                    "grd", full.names = TRUE)
predictors <- stack(files)
```

_Get world boundaries_

```{r}
data(wrld_simpl)
```

_Get GBIF data using the rOpenSci package rgbif_

```{r cache=TRUE}
nn <- name_lookup("bradypus*", rank = "species")
nn <- na.omit(unique(nn$data$nubKey))
df <- occ_search(taxonKey = nn, hasCoordinate = TRUE, limit = 500)
df <- df[ sapply(df, function(x) class(x$data)) %in% "data.frame" ]
df <- ldply(lapply(df, "[[", "data"))
df2 <- df[,c('decimalLongitude','decimalLatitude')]
```

_Plot_

(1) Add raster data, (2) Add political boundaries, (3) Add the points (occurrences)

```{r fig.keep='last'}
plot(predictors, 1)
plot(wrld_simpl, add = TRUE)
points(df2, col = "blue")
```

### 2015-11-14 use-cases/oak-occurrences.Rmd
added three use cases

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

This example comes from [Antonio J. Perez-Luque](https://twitter.com/ajpelu) who [shared his plot on Twitter][tweet]. Antonio compared the occurrences of Valley Oak (_Quercus lobata_) from [GBIF][gbif] to the distribution of the same species from the [Atlas of US Trees][atlas].

_Load libraries_

```{r}
library('rgbif')
library('raster')
library('sp')
library('maptools')
library('rgeos')
library('scales')
```

_Get GBIF Data for Quercus lobata_

```{r cache=TRUE}
keyQl <- name_backbone(name='Quercus lobata', kingdom='plants')$speciesKey
dat.Ql <- occ_search(taxonKey=keyQl, return='data', limit=50000)
```

_Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)_

From [http://esp.cr.usgs.gov/data/little/](http://esp.cr.usgs.gov/data/little/). And save shapefile in same directory

```{r}
url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
tmp <- tempdir()
download.file(url, destfile = "~/querloba.zip")
unzip("~/querloba.zip", exdir = tmp)
ql <- readShapePoly(file.path(tmp, "querloba.shp"))
```

_Get Elevation data of US_

```{r}
alt.USA <- getData('alt', country = 'USA')
```

_Create Hillshade of US_

```{r}
alt.USA <- alt.USA[[1]]
slope.USA <- terrain(alt.USA, opt = 'slope')
aspect.USA <- terrain(alt.USA, opt = 'aspect')
hill.USA <- hillShade(slope.USA, aspect.USA, angle = 45, direction = 315)
```

_Plot map_

```{r fig.keep='last'}
plot(hill.USA, col = grey(0:100/100), legend = FALSE, xlim = c(-125, -116), ylim = c(32, 42), main = 'Distribution of Quercus lobata', xlab = "Long", ylab = 'Lat')
# add shape from Atlas of US Trees
plot(ql, add = TRUE, col = alpha("white", 0.6), border = FALSE)
# add Gbif presence points
points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = -121, y = 40.5, "GBIF Data", pch = 19, col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
legend(x = -121, y = 41.5, "Atlas of United States Trees \n (Little, E. 1971)", pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
```

[tweet]: https://twitter.com/ajpelu/status/473951167567757312
[gbif]: http://www.gbif.org/
[atlas]: http://esp.cr.usgs.gov/data/little/

### 2015-11-25 manuscript.Rmd
manuscript improvements on occurrences search section, conclusion fleshed out

@@
 Perhaps the most fundamental element in many fields of ecology is
-the inividual. How many individuals of each species in a given
+the individual. The number of individuals of each species in a given
 location forms the basis for many sub-fields of ecology and evolution.
@@
 @Mendoza_2015]; development of invasive species watch lists
-[@Faulkner_2014]; evaluate risk of invasive species spread
+[@Faulkner_2014]; evaluating risk of invasive species spread
 [@Di_Febbraro_2013]; and effects of climate change on future
@@
 Biodiversity loss is one of the greatest challenges of our time
-[@Pimm_2014]. Some have called this the sixth great mass
+[@Pimm_2014], and some have called this the sixth great mass
 extinction [@Ceballos_2015]. Given this challenge there is a great
@@
 However, Global Biodiversity Information Facility (hereafter, GBIF,
-http://www.gbif.org/) is the largest collection of biodiversity
-records globally, currently with 580 million records, 1.6 million
-taxa, 15,000 datasets from 770 publishers (figures collected on 2015-10-04).
+http://www.gbif.org) is the largest collection of biodiversity
+records globally, currently with 640 million records, 1.6 million
+taxa, 15,000 datasets from 780 publishers (current as of 2015-11-25).
 Many large biodiversity warehouses such as iNaturalist
-(http://www.inaturalist.org/), VertNet (http://vertnet.org/), and USGS's
-Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov/)
+(http://www.inaturalist.org), VertNet (http://vertnet.org), and USGS's
+Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov)
 all feed into GBIF.
 
-Herein, we describe the `rgbif` library [@rgbif] for working with
+Herein, we describe the `rgbif` software library [@rgbif] for working with
 GBIF data in the R programming environment [@R]. R is a widely used
@@
 
-The `rgbif` package is completely written in R, uses an [MIT license][mit]
+The `rgbif` package is nearly completely written in R (a small Javascript
+library is included for reading well known text), uses an [MIT license][mit]
 to maximize use everywhere. `rgbif` is developed publicly on GitHub
@@
 Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
-distribution network for R packages. `rgbif` is part of the rOpenSci project,
-a developer network making R software to facilitate reproducible research.
+distribution network for R packages. `rgbif` is part of the rOpenSci project
+(http://ropensci.org), a developer network making R software to facilitate
+reproducible research.
 
@@
 or API. The GBIF API has four major components: registry, taxonomic names,
-occurrences, and maps. We ignore maps in `rgbif` as it is concerned with
-generating maps primarily for web applications. `rgbif` has a suite of functions dealing
-with each of registry, taxonomic names, and occurrences - we'll go through
-each in turn describing design and example usage.
+occurrences, and maps. We also include functions to interface with the OAI-PMH
+GBIF service; only dataset information is available vis this service, however. We
+ignore maps in `rgbif` as it is concerned with generating maps primarily for web
+applications. `rgbif` has a suite of functions dealing with each of registry, taxonomic
+names, and occurrences - we'll go through each in turn describing design and example usage.
 
@@
 tell GBIF that the request is coming from `rgbif`, including what version of
-the package. This will help GBIF know what proportion of requests are coming from
-this package, and therefore R, as most requests likely will come from `rgbif`.
+the package. This helps GBIF know what proportion of requests are coming from
+this package, and therefore R, as most requests likely will come from `rgbif`;
+this information is helpful for them in thinking about how people are using
+GBIF data.
 
@@
 
-The GBIF reqistry API services are spread across four sets
-of functions:
+The GBIF reqistry API services are spread across five sets of functions
+via the main API:
 
@@
 
-Datasets are owned by organizations. Organizations are endorsed by nodes to share datasets with GBIF.
-Datasets are published through institutions, which may be hosted at another organization. A network
-is a group of datasets (managed by GBIF).
+And dataset information in general is available via the OAI-PMH service,
+functions in `rgbif` prefixed with `gbif_oai_`.
+
+Datasets are owned by organizations. Organizations are endorsed by nodes to share
+datasets with GBIF. Datasets are published through institutions, which may be
+hosted at another organization. A network is a group of datasets (managed by GBIF).
+Datasets are the units that matter the most with respect to registry information,
+while installations, networks, nodes, and organizations are simply higher level
+organizational structure.
 
@@
 
-Also, check out `datasets()` and `dataset_suggest()` for searching for datasets.
+See also `datasets()` and `dataset_suggest()` for searching for datasets.
 
@@
 
-Dataset metrics are another useful way of digging in to figure out what
-datasets you may want to use. One drawback is that these data are only
+Dataset metrics are another useful way of figuring out what
+datasets you may want to use. One drawback is that these metrics data are only
 available for datasets of type _checklist_, but there are quite a lot of
@@
 a taxonomic name, you can point to a URI on the internet, and 2) you can
-search for metadata on that species, and occurrences of that species in
+search for metadata on a taxon, and occurrences of that taxon in
 GBIF.
@@
 their own unique codes for the same taxonomic names, and some taxonomic
-groups have preferred sources for the definitive names for that group.
+groups have preferred sources for the definitive names for that group. That's
+why it's best to determine what name GBIF uses, and its associated identifier,
+for the taxon of interest instead of simply searching for occurrences with a
+taxonomic name.
 
@@
 does a very quick search and returns minimal data. Here, we're searching for
-the query tem _Pum_, and we get back a bunch of names:
+the query tem _Pum_, and we get back many names:
 
@@
 ```
+
+With these results, you can then proceed to search for occurrences with the
+taxon key(s), or drill down further with other name searching functions to
+get the exact taxon of interest.
 
@@
 
+The download service, like the occurrence search service, is rate-limited.
+That is, you can only have one to three downloads running simultaneously for
+your user credentials. However, simply check when a download job is complete,
+then you should be able to start a new download request.
+
 ### Download API
@@
 do greater than searches (i.e., `latitude > 50`), whereas you can not
-do that in the occurrence search API. Thus, we can't make the query
-interace exactly the same for both search and download functions.
+do that in the occurrence search API. Thus, unfortnately, we couldn't
+make the query interace exactly the same for both search and
+download functions.
 
-Using the download service can be as few as three steps: 1) Request data
-via a search; 2) Download data; 3) Import data into R.
+Using the download service can consist of as few as three steps: 1) Request
+data via a search; 2) Download data; 3) Import data into R.
 
@@
 
+#### Search for occurrences
+
 The main search work-horse is `occ_search()`. This function allows very flexible
@@
 
-...
+The output of `occ_search()` borrows the tidy `data.frame` idea from the `dplyr`
+R package, so that no matter how large the `data.frame`, the output is easily
+assessed because only a few of the records (rows) are shown, only a few columns
+are shown (with others shown in name only), and metadata is shown on top of
+the `data.frame` to indicate data found and returned, media
+records found, unique taxonomic hierarchies returned, and the query executed.
+
+The output of these examples, except one, aren't shown, but all run correctly.
+
+Search by species name, using `name_backbone()` first to get key
+
+```{r}
+(key <- name_suggest(q = 'Helianthus annuus', rank = 'species')$key[1])
+occ_search(taxonKey = key, limit = 2)
+```
+
+Instead of getting a taxon key first, you can search for a name directly
+
+```{r eval=FALSE}
+occ_search(scientificName = 'Ursus americanus')
+```
+
+Search for many species
+
+```{r eval=FALSE}
+splist <- c('Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa')
+keys <- sapply(splist, function(x) name_suggest(x)$key[1], USE.NAMES = FALSE)
+occ_search(taxonKey = keys, limit = 5, return = 'data')
+```
+
+Spatial search, based on well known text format, or a bounding box set of
+four coordinates
+
+```{r eval=FALSE}
+# well known text
+occ_search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
+# bounding box
+occ_search(geometry = c(-125.0,38.4,-121.8,40.9), limit = 20)
+```
+
+Get only occurrences with lat/long data
+
+```{r eval=FALSE}
+occ_search(hasCoordinate = TRUE, limit = 5)
+```
+
+Get only those occurrences with spatial issues
+
+```{r eval=FALSE}
+occ_search(hasGeospatialIssue = TRUE, limit = 5)
+```
 
@@
 ```{r}
-out <- occ_search(issue='DEPTH_UNLIKELY', limit = 500)
+out <- occ_search(issue = 'DEPTH_UNLIKELY', limit = 500)
 NROW(out)
@@
 
-* pt 1
-* pt 2
-* pt 3
-* pt 4
+The `rgbif` R package provides a programmatic interface to GBIF's
+application programming interface (API) - a powerful tool for making
+research using species occurrence data reproducible. In fact,
+the `rgbif` package has already been used in 22 scholarly publications
+(as of 2015-11-14).
+
+The `rgbif` package is relatively stable now, and should not have
+many breaking changes unless necessitated due to changes in the
+GBIF API.
+
+One area of focus in the future is to attempt to solve many use
+cases that have been brought up with respect to GBIF data. For
+example, some specimens are included in GBIF that are located
+in botanical gardens. For many research questions, researchers are
+interested in "wild" type occurrences, not those in human curated
+scenarios. Making removal of these occurrences easy would be
+very useful, but is actually quite a hard problem.
 

### 2015-11-25 manuscript.tex
manuscript improvements on occurrences search section, conclusion fleshed out

@@
 Perhaps the most fundamental element in many fields of ecology is the
-inividual. How many individuals of each species in a given location
-forms the basis for many sub-fields of ecology and evolution. Some
-research questions necessitate collecting new data, while others can
-easily take advantage of existing data. In fact, some ecology fields are
-built largely on existing data, e.g., macro-ecology (Brown, 1995; Beck
-et al., 2012).
+individual. The number of individuals of each species in a given
+location forms the basis for many sub-fields of ecology and evolution.
+Some research questions necessitate collecting new data, while others
+can easily take advantage of existing data. In fact, some ecology fields
+are built largely on existing data, e.g., macro-ecology (Brown, 1995;
+Beck et al., 2012).
 
@@
 Mendoza et al., 2015); development of invasive species watch lists
-(Faulkner et al., 2014); evaluate risk of invasive species spread
+(Faulkner et al., 2014); evaluating risk of invasive species spread
 (Febbraro et al., 2013); and effects of climate change on future
@@
 Biodiversity loss is one of the greatest challenges of our time (Pimm et
-al., 2014). Some have called this the sixth great mass extinction
+al., 2014), and some have called this the sixth great mass extinction
 (Ceballos et al., 2015). Given this challenge there is a great need for
@@
 records. However, Global Biodiversity Information Facility (hereafter,
-GBIF, \url{http://www.gbif.org/}) is the largest collection of
-biodiversity records globally, currently with 580 million records, 1.6
-million taxa, 15,000 datasets from 770 publishers (figures collected on
-2015-10-04). Many large biodiversity warehouses such as iNaturalist
-(\url{http://www.inaturalist.org/}), VertNet
-(\url{http://vertnet.org/}), and USGS's Biodiversity Information Serving
-Our Nation (BISON; \url{http://bison.usgs.ornl.gov/}) all feed into
-GBIF.
+GBIF, \url{http://www.gbif.org}) is the largest collection of
+biodiversity records globally, currently with 640 million records, 1.6
+million taxa, 15,000 datasets from 780 publishers (current as of
+2015-11-25). Many large biodiversity warehouses such as iNaturalist
+(\url{http://www.inaturalist.org}), VertNet (\url{http://vertnet.org}),
+and USGS's Biodiversity Information Serving Our Nation (BISON;
+\url{http://bison.usgs.ornl.gov}) all feed into GBIF.
 
-Herein, we describe the \texttt{rgbif} library (Chamberlain et al.) for
-working with GBIF data in the R programming environment (R Core Team,
-2014). R is a widely used language in academia, and in non-profit and
-private sectors. Importantly, R makes it easy to do all of the steps of
-the research process, including data management, data manipulation and
-cleaning, statistics, and vizualization. Thus, an R client for getting
-GBIF data is a powerful tool to facilitate reproducible research.
+Herein, we describe the \texttt{rgbif} software library (Chamberlain et
+al.) for working with GBIF data in the R programming environment (R Core
+Team, 2014). R is a widely used language in academia, and in non-profit
+and private sectors. Importantly, R makes it easy to do all of the steps
+of the research process, including data management, data manipulation
+and cleaning, statistics, and vizualization. Thus, an R client for
+getting GBIF data is a powerful tool to facilitate reproducible
+research.
 
@@
 
-The \texttt{rgbif} package is completely written in R, uses an
+The \texttt{rgbif} package is nearly completely written in R (a small
+Javascript library is included for reading well known text), uses an
 \href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
@@
 distribution network for R packages. \texttt{rgbif} is part of the
-rOpenSci project, a developer network making R software to facilitate
-reproducible research.
+rOpenSci project (\url{http://ropensci.org}), a developer network making
+R software to facilitate reproducible research.
 
@@
 Programming Interface}, or API. The GBIF API has four major components:
-registry, taxonomic names, occurrences, and maps. We ignore maps in
+registry, taxonomic names, occurrences, and maps. We also include
+functions to interface with the OAI-PMH GBIF service; only dataset
+information is available vis this service, however. We ignore maps in
 \texttt{rgbif} as it is concerned with generating maps primarily for web
@@
 headers that tell GBIF that the request is coming from \texttt{rgbif},
-including what version of the package. This will help GBIF know what
+including what version of the package. This helps GBIF know what
 proportion of requests are coming from this package, and therefore R, as
-most requests likely will come from \texttt{rgbif}.
+most requests likely will come from \texttt{rgbif}; this information is
+helpful for them in thinking about how people are using GBIF data.
 
@@
 
-The GBIF reqistry API services are spread across four sets of functions:
+The GBIF reqistry API services are spread across five sets of functions
+via the main API:
 
@@
 
+And dataset information in general is available via the OAI-PMH service,
+functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}.
+
 Datasets are owned by organizations. Organizations are endorsed by nodes
@@
 institutions, which may be hosted at another organization. A network is
-a group of datasets (managed by GBIF).
+a group of datasets (managed by GBIF). Datasets are the units that
+matter the most with respect to registry information, while
+installations, networks, nodes, and organizations are simply higher
+level organizational structure.
 
@@
 
-Also, check out \texttt{datasets()} and \texttt{dataset\_suggest()} for
+See also \texttt{datasets()} and \texttt{dataset\_suggest()} for
 searching for datasets.
@@
 
-Dataset metrics are another useful way of digging in to figure out what
-datasets you may want to use. One drawback is that these data are only
+Dataset metrics are another useful way of figuring out what datasets you
+may want to use. One drawback is that these metrics data are only
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2483).
+lot of them (2490).
 
@@
 a taxonomic name, you can point to a URI on the internet, and 2) you can
-search for metadata on that species, and occurrences of that species in
-GBIF.
+search for metadata on a taxon, and occurrences of that taxon in GBIF.
 
@@
 taxonomic groups have preferred sources for the definitive names for
-that group.
+that group. That's why it's best to determine what name GBIF uses, and
+its associated identifier, for the taxon of interest instead of simply
+searching for occurrences with a taxonomic name.
 
@@
 minimal data. Here, we're searching for the query tem \emph{Pum}, and we
-get back a bunch of names:
+get back many names:
 
@@
 \end{longtable}
+
+With these results, you can then proceed to search for occurrences with
+the taxon key(s), or drill down further with other name searching
+functions to get the exact taxon of interest.
 
@@
 
+The download service, like the occurrence search service, is
+rate-limited. That is, you can only have one to three downloads running
+simultaneously for your user credentials. However, simply check when a
+download job is complete, then you should be able to start a new
+download request.
+
 \subsubsection{Download API}\label{download-api}
@@
 searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
-not do that in the occurrence search API. Thus, we can't make the query
-interace exactly the same for both search and download functions.
+not do that in the occurrence search API. Thus, unfortnately, we
+couldn't make the query interace exactly the same for both search and
+download functions.
 
-Using the download service can be as few as three steps: 1) Request data
-via a search; 2) Download data; 3) Import data into R.
+Using the download service can consist of as few as three steps: 1)
+Request data via a search; 2) Download data; 3) Import data into R.
 
@@
 
+\paragraph{Search for occurrences}\label{search-for-occurrences}
+
 The main search work-horse is \texttt{occ\_search()}. This function
@@
 
-\ldots{}
+The output of \texttt{occ\_search()} borrows the tidy
+\texttt{data.frame} idea from the \texttt{dplyr} R package, so that no
+matter how large the \texttt{data.frame}, the output is easily assessed
+because only a few of the records (rows) are shown, only a few columns
+are shown (with others shown in name only), and metadata is shown on top
+of the \texttt{data.frame} to indicate data found and returned, media
+records found, unique taxonomic hierarchies returned, and the query
+executed.
+
+The output of these examples, except one, aren't shown, but all run
+correctly.
+
+Search by species name, using \texttt{name\_backbone()} first to get key
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(key <-}\StringTok{ }\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q =} \StringTok{'Helianthus annuus'}\NormalTok{, }\DataTypeTok{rank =} \StringTok{'species'}\NormalTok{)$key[}\DecValTok{1}\NormalTok{])}
+\CommentTok{#> [1] 3119195}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
+\CommentTok{#> Records found [21577] }
+\CommentTok{#> Records returned [2] }
+\CommentTok{#> No. unique hierarchies [1] }
+\CommentTok{#> No. media records [2] }
+\CommentTok{#> Args [taxonKey=3119195, limit=2, offset=0, fields=all] }
+\CommentTok{#> First 10 rows of data}
+\CommentTok{#> }
+\CommentTok{#>                name        key decimalLatitude decimalLongitude}
+\CommentTok{#> 1 Helianthus annuus 1143516596        35.42767        -105.0688}
+\CommentTok{#> 2 Helianthus annuus 1095851641         0.00000           0.0000}
+\CommentTok{#> Variables not shown: issues (chr), datasetKey (chr), publishingOrgKey}
+\CommentTok{#>      (chr), publishingCountry (chr), protocol (chr), lastCrawled (chr),}
+\CommentTok{#>      lastParsed (chr), extensions (chr), basisOfRecord (chr), taxonKey}
+\CommentTok{#>      (int), kingdomKey (int), phylumKey (int), classKey (int), orderKey}
+\CommentTok{#>      (int), familyKey (int), genusKey (int), speciesKey (int),}
+\CommentTok{#>      scientificName (chr), kingdom (chr), phylum (chr), order (chr),}
+\CommentTok{#>      family (chr), genus (chr), species (chr), genericName (chr),}
+\CommentTok{#>      specificEpithet (chr), taxonRank (chr), dateIdentified (chr), year}
+\CommentTok{#>      (int), month (int), day (int), eventDate (chr), modified (chr),}
+\CommentTok{#>      lastInterpreted (chr), references (chr), identifiers (chr), facts}
+\CommentTok{#>      (chr), relations (chr), geodeticDatum (chr), class (chr), countryCode}
+\CommentTok{#>      (chr), country (chr), rightsHolder (chr), identifier (chr),}
+\CommentTok{#>      verbatimEventDate (chr), datasetName (chr), gbifID (chr),}
+\CommentTok{#>      verbatimLocality (chr), collectionCode (chr), occurrenceID (chr),}
+\CommentTok{#>      taxonID (chr), recordedBy (chr), catalogNumber (chr),}
+\CommentTok{#>      http...unknown.org.occurrenceDetails (chr), institutionCode (chr),}
+\CommentTok{#>      rights (chr), occurrenceRemarks (chr), identificationID (chr),}
+\CommentTok{#>      elevation (dbl), elevationAccuracy (dbl), stateProvince (chr),}
+\CommentTok{#>      recordNumber (chr), locality (chr), municipality (chr), language}
+\CommentTok{#>      (chr), type (chr), ownerInstitutionCode (chr), identifiedBy (chr)}
+\end{Highlighting}
+\end{Shaded}
+
+Instead of getting a taxon key first, you can search for a name directly
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{scientificName =} \StringTok{'Ursus americanus'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Search for many species
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{splist <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{)}
+\NormalTok{keys <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(splist, function(x) }\KeywordTok{name_suggest}\NormalTok{(x)$key[}\DecValTok{1}\NormalTok{], }\DataTypeTok{USE.NAMES =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{keys, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Spatial search, based on well known text format, or a bounding box set
+of four coordinates
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# well known text}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\CommentTok{# bounding box}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{125.0}\NormalTok{,}\FloatTok{38.4}\NormalTok{,-}\FloatTok{121.8}\NormalTok{,}\FloatTok{40.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Get only occurrences with lat/long data
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Get only those occurrences with spatial issues
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasGeospatialIssue =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{issue=}\StringTok{'DEPTH_UNLIKELY'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
+\NormalTok{out <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{issue =} \StringTok{'DEPTH_UNLIKELY'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
 \KeywordTok{NROW}\NormalTok{(out)}
@@
 \centering
-\includegraphics{components/figure/manuscript-unnamed-chunk-19-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-25-1.pdf}
 \caption{}
@@
 \centering
-\includegraphics{components/figure/manuscript-unnamed-chunk-24-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-30-1.pdf}
 \caption{}
@@
 \centering
-\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-37-1.pdf}
 \caption{}
@@
 \item
-  pt 3
+  The \texttt{rgbif} package has already been used in 22 scholarly
+  publications (as of 2015-11-14).
 \item

### 2015-11-25 manuscript.tex
pdf updated

@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  pt 1
-\item
-  pt 2
-\item
-  The \texttt{rgbif} package has already been used in 22 scholarly
+The \texttt{rgbif} R package provides a programmatic interface to GBIF's
+application programming interface (API) - a powerful tool for making
+research using species occurrence data reproducible. In fact, the
+\texttt{rgbif} package has already been used in 22 scholarly
 publications (as of 2015-11-14).
-\item
-  pt 4
-\end{itemize}
+
+The \texttt{rgbif} package is relatively stable now, and should not have
+many breaking changes unless necessitated due to changes in the GBIF
+API.
+
+One area of focus in the future is to attempt to solve many use cases
+that have been brought up with respect to GBIF data. For example, some
+specimens are included in GBIF that are located in botanical gardens.
+For many research questions, researchers are interested in ``wild'' type
+occurrences, not those in human curated scenarios. Making removal of
+these occurrences easy would be very useful, but is actually quite a
+hard problem.
 

### 2016-01-07 manuscript.Rmd
manuscript updates, more text around GBIF uses in other related pkgs, map eg

@@
 * Search for occurrences - `occ_search()`
+* A simplified and optimized version of `occ_search()` - `occ_data()`
 * Get occurrences by occurrence identifier - `occ_get()`
@@
 
-#### Cleaning data
+#### Data cleaning
 
@@
 
+## Mapping
+
+An obvious downstream use case for species occurrence data is to map the
+data. `rgbif` per se is largely not concerned with making this easier,
+although we do have a simple wrappar around `ggplot2` to make it easy to
+get a quick plot of occurrence data. For example, here we plot 100
+occurrences for _Puma concolor_.
+
+```{r fig.width=8, fig.height=4}
+key <- name_backbone(name='Puma concolor')$speciesKey
+dat <- occ_search(taxonKey = key, limit = 100, hasCoordinate = TRUE)
+gbifmap(dat$data)
+```
+
+Another package, [spoccutils][spoccutils]
+
+## GBIF data in other R packages
+
+We discuss usage of GBIF data in other R packages throughout the
+manuscript, but provide a synopsis here for clarity.
+
+### taxize
+
+Some of the GBIF taxonomic services are also avialable in [taxize][taxize],
+an R package that focuses on getting data from taxonomic data sources on
+the web. For example, with `get_gbifid()` one can get GBIF IDs used for
+a set of taxonomic names - then use those IDs in other functions in `taxize`
+to get additional information, like taxonomically downstream children.
+
+### spocc
+
+GBIF occurrence data is available in the R package [spocc][spocc] via `rgbif`.
+`spocc` is a unified interface for fetching species occurrence data from
+many sources on the web. For example, a user can collect occurrence data
+from GBIF, iDigBio, and iNaturalist, and easily combine them, then use other
+pacakages to clean and vizualize the data.
+
 # Use cases
+
+The following are three use cases for `rgbif`: niche modelling,
+spatial change in biodiversity, and distribution mapping.
 
@@
 
-The `rgbif` package is relatively stable now, and should not have
+The `rgbif` package is relatively stable, and should not have
 many breaking changes unless necessitated due to changes in the
@@
 This project was supported in part by the Alfred P Sloan
-Foundation (Grant 2013-6-22).
+Foundation (Grant 2013-6-22), and the Helmsley Foundation
+(Grant XXXXXX).
 
@@
 [github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
-can be found at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
-under the open and permissive MIT license.
+can be found at [https://github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
+under an MIT license.
 
@@
 [gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
+[taxize]: https://github.com/ropensci/taxize
+[spocc]: https://github.com/ropensci/spocc
+[spoccutils]: https://github.com/ropensci/spocc

### 2016-01-07 manuscript.tex
manuscript updates, more text around GBIF uses in other related pkgs, map eg

@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2490).
+lot of them (2580).
 
@@
   Search for occurrences - \texttt{occ\_search()}
+\item
+  A simplified and optimized version of \texttt{occ\_search()} -
+  \texttt{occ\_data()}
 \item
@@
 
-\paragraph{Cleaning data}\label{cleaning-data}
+\paragraph{Data cleaning}\label{data-cleaning}
 
@@
 
+\subsection{Mapping}\label{mapping}
+
+An obvious downstream use case for species occurrence data is to map the
+data. \texttt{rgbif} per se is largely not concerned with making this
+easier, although we do have a simple wrappar around \texttt{ggplot2} to
+make it easy to get a quick plot of occurrence data. For example, here
+we plot 100 occurrences for \emph{Puma concolor}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{key <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Puma concolor'}\NormalTok{)$speciesKey}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{100}\NormalTok{, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{)}
+\KeywordTok{gbifmap}\NormalTok{(dat$data)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-unnamed-chunk-20-1.pdf}
+\caption{}
+\end{figure}
+
+Another package, \href{https://github.com/ropensci/spocc}{spoccutils}
+
+\subsection{GBIF data in other R
+packages}\label{gbif-data-in-other-r-packages}
+
+We discuss usage of GBIF data in other R packages throughout the
+manuscript, but provide a synopsis here for clarity.
+
+\subsubsection{taxize}\label{taxize}
+
+Some of the GBIF taxonomic services are also avialable in
+\href{https://github.com/ropensci/taxize}{taxize}, an R package that
+focuses on getting data from taxonomic data sources on the web. For
+example, with \texttt{get\_gbifid()} one can get GBIF IDs used for a set
+of taxonomic names - then use those IDs in other functions in
+\texttt{taxize} to get additional information, like taxonomically
+downstream children.
+
+\subsubsection{spocc}\label{spocc}
+
+GBIF occurrence data is available in the R package
+\href{https://github.com/ropensci/spocc}{spocc} via \texttt{rgbif}.
+\texttt{spocc} is a unified interface for fetching species occurrence
+data from many sources on the web. For example, a user can collect
+occurrence data from GBIF, iDigBio, and iNaturalist, and easily combine
+them, then use other pacakages to clean and vizualize the data.
+
 \section{Use cases}\label{use-cases}
+
+The following are three use cases for \texttt{rgbif}: niche modelling,
+spatial change in biodiversity, and distribution mapping.
 
@@
 \centering
-\includegraphics{components/figure/manuscript-unnamed-chunk-25-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-26-1.pdf}
 \caption{}
@@
 \centering
-\includegraphics{components/figure/manuscript-unnamed-chunk-30-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
 \caption{}
@@
 \centering
-\includegraphics{components/figure/manuscript-unnamed-chunk-37-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-38-1.pdf}
 \caption{}
@@
 
-The \texttt{rgbif} package is relatively stable now, and should not have
+The \texttt{rgbif} package is relatively stable, and should not have
 many breaking changes unless necessitated due to changes in the GBIF
@@
 This project was supported in part by the Alfred P Sloan Foundation
-(Grant 2013-6-22).
+(Grant 2013-6-22), and the Helmsley Foundation (Grant XXXXXX).
 
@@
 Software can be found at
-\href{https://github.com/ropensci/rgbif}{github.com/ropensci/rgbif},
-under the open and permissive MIT license.
+\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
+under an MIT license.
 

### 2016-03-11 manuscript.Rmd
manuscrpt changes

@@
     footnote: Corresponding author
+  - name: Carl Boettiger
+    affiliation: boettig
+    email: carl(at)ropensci.org
 address:
   - code: cstar
+    address: |
+      University of California, Berkeley, CA, USA
+  - code: boettig
     address: |
@@
 Perhaps the most fundamental element in many fields of ecology is
-the individual. The number of individuals of each species in a given
-location forms the basis for many sub-fields of ecology and evolution.
-Some research questions necessitate collecting new data, while
+the individual organism. The number of individuals of each species in
+a given location forms the basis for many sub-fields of ecology and
+evolution. Some research questions necessitate collecting new data, while
 others can easily take advantage of existing data. In fact, some
@@
 
+# Global Biodiversity Information Facility
+
 There are many online services that collect and maintain specimen records.
 However, Global Biodiversity Information Facility (hereafter, GBIF,
-http://www.gbif.org) is the largest collection of biodiversity
-records globally, currently with 640 million records, 1.6 million
-taxa, 15,000 datasets from 780 publishers (current as of 2015-11-25).
+[http://www.gbif.org](http://www.gbif.org)) is the largest collection of
+biodiversity records globally, currently with 643 million records, 1.6 million
+taxa, 15,450 datasets from 780 publishers (as of 2016-02-09).
 Many large biodiversity warehouses such as iNaturalist
-(http://www.inaturalist.org), VertNet (http://vertnet.org), and USGS's
-Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov)
+([http://www.inaturalist.org](http://www.inaturalist.org)), VertNet
+([http://vertnet.org](http://vertnet.org)), and USGS's Biodiversity Information
+Serving Our Nation (BISON; [http://bison.usgs.ornl.gov](http://bison.usgs.ornl.gov))
 all feed into GBIF.
 
-Herein, we describe the `rgbif` software library [@rgbif] for working with
+The most important data organizational level in GBIF is the individual record. In R
+using `rgbif` you'll recognize this as a row in the data.frame that is returned from
+`occ_search()` or `occ_data()`. Going upstream, each record is part of a dataset,
+where each dataset is submitted by an organization, organizations are organized into
+nodes, datasets are published through institutions (which may be hosted at another
+organization), and a network is a group of datasets (managed by GBIF).
+
+Each record has some taxonomic name associated with it, which itself is linked to
+a lot of other taxonomic data. GBIF maintains their own taxonomic data.
+
+All this organization matters because you can navigate through GBIF data through all
+these organizational levels.
+
+# The rgbif package
+
+Herein, we describe the `rgbif` software package [@rgbif] for working with
 GBIF data in the R programming environment [@R]. R is a widely used
-language in academia, and in non-profit and private sectors.
-Importantly, R makes it easy to do all of the steps of the
+language in academia, as well as non-profit and private sectors.
+Importantly, R makes it easy to execute all steps of the
 research process, including data management, data manipulation
-and cleaning, statistics, and vizualization. Thus, an R client
+and cleaning, statistics, and visualization. Thus, an R client
 for getting GBIF data is a powerful tool to facilitate
@@
 
-# The rgbif package
-
 The `rgbif` package is nearly completely written in R (a small Javascript
-library is included for reading well known text), uses an [MIT license][mit]
+library is included for reading well known text [@wkt]), uses an [MIT license][mit]
 to maximize use everywhere. `rgbif` is developed publicly on GitHub
@@
 
-The GBIF reqistry API services are spread across five sets of functions
+The GBIF registry API services are spread across five sets of functions
 via the main API:
@@
 
-Here, we seaerch for dataset metrics for a single dataset, with uuid
+Here, we search for dataset metrics for a single dataset, with uuid
 `ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
@@
 known to GBIF's database. This serves two purposes: 1) when referring to
-a taxonomic name, you can point to a URI on the internet, and 2) you can
+a taxonomic name, you can point to a URI on the Internet, and 2) you can
 search for metadata on a taxon, and occurrences of that taxon in
@@
 does a very quick search and returns minimal data. Here, we're searching for
-the query tem _Pum_, and we get back many names:
+the query term _Pum_, and we get back many names:
 
@@
 `/occurrence/download` route (many functions, see below).
-`occ_search()` is the main funtion for the search route,
+`occ_search()` is the main function for the search route,
 and is more appropriate for smaller data, while
@@
 
-Large is of course a subjective term. When you hit a "large dataset"
-will depend primarily on the size of the your data request. GBIF imposes
-for any given search a limit of 200,000 records in the search service,
-after which point you can't download any more records for that search.
-However, you can download more records for different searches.
+GBIF imposes for any given search a limit of 200,000 records in the search
+service, after which point you can't download any more records for that
+search. However, you can download more records for different searches.
 
@@
 is defined is different. For example, in the download API you can
-do greater than searches (i.e., `latitude > 50`), whereas you can not
-do that in the occurrence search API. Thus, unfortnately, we couldn't
-make the query interace exactly the same for both search and
+do greater than searches (i.e., `latitude > 50`), whereas you cannot
+do that in the occurrence search API. Thus, unfortunately, we couldn't
+make the query interface exactly the same for both search and
 download functions.
@@
 
-* Get a single numeric count of occurrenes - `occ_count()`
+* Get a single numeric count of occurrences - `occ_count()`
 * Search for occurrences - `occ_search()`
@@
 
-The output of `occ_search()` borrows the tidy `data.frame` idea from the `dplyr`
-R package, so that no matter how large the `data.frame`, the output is easily
-assessed because only a few of the records (rows) are shown, only a few columns
-are shown (with others shown in name only), and metadata is shown on top of
-the `data.frame` to indicate data found and returned, media
-records found, unique taxonomic hierarchies returned, and the query executed.
+The output of `occ_search()` presents a compact `data.frame` so that no matter
+how large the `data.frame`, the output is easily assessed because only a few of
+the records (rows) are shown, only a few columns are shown (with others shown in
+name only), and metadata is shown on top of the `data.frame` to indicate data
+found and returned, media records found, unique taxonomic hierarchies returned,
+and the query executed.
 
@@
 
-Spatial search, based on well known text format, or a bounding box set of
+Spatial search, based on well known text format [@wkt], or a bounding box set of
 four coordinates
@@
 
-Get only those occurrences with spatial issues
+Get only those occurrences with spatial issues. Spatial issues are a set of issues
+that are returned in the `issues` field. They each indicate something different
+about that record. For example, the issue `COUNTRY_COORDINATE_MISMATCH` indicates
+that the interpreted occurrence coordinates fall outside of the indicated country.
+You can see how that might be useful when it comes to cleaning your data prior
+to analysis/visualization.
 
@@
 data. `rgbif` per se is largely not concerned with making this easier,
-although we do have a simple wrappar around `ggplot2` to make it easy to
+although we do have a simple wrapper around `ggplot2` to make it easy to
 get a quick plot of occurrence data. For example, here we plot 100
@@
 
-Another package, [spoccutils][spoccutils]
+Another package, [mapr][mapr], is the perfect mapping companion to `rgbif`.
+It has convenient functions for handling input data from `rgbif`, `spocc`,
+or arbitrary data.frame's, and output plots for base plots, `ggplot2`,
+`ggmap` (`ggplot2` with map layers underneath), and interactive maps on
+GitHub gists or with Leaflet.js.
 
@@
 
-Some of the GBIF taxonomic services are also avialable in [taxize][taxize],
+Some of the GBIF taxonomic services are also available in [taxize][taxize],
 an R package that focuses on getting data from taxonomic data sources on
@@
 from GBIF, iDigBio, and iNaturalist, and easily combine them, then use other
-pacakages to clean and vizualize the data.
+packages to clean and visualize the data.
 
@@
 
-The following are three use cases for `rgbif`: niche modelling,
+The following are three use cases for `rgbif`: niche modeling,
 spatial change in biodiversity, and distribution mapping.
 
-### Ecological niche modelling
+### Ecological niche modeling
 
-```{r niche-modelling, child="use-cases/niche-modelling.Rmd"}
+```{r niche-modeling, child="use-cases/niche-modelling.Rmd"}
 ```
@@
 
-# Acknowledgements
+# Acknowledgments
 
-This project was supported in part by the Alfred P Sloan
-Foundation (Grant 2013-6-22), and the Helmsley Foundation
-(Grant XXXXXX).
+This project was supported in part by the Alfred P Sloan Foundation
+(Grant No. G-2014-13485), and in part by the Helmsley Foundation
+(Grant No. 2016PG-BRI004).
 
@@
 [spocc]: https://github.com/ropensci/spocc
-[spoccutils]: https://github.com/ropensci/spocc
+[mapr]: https://github.com/ropensci/mapr

### 2016-03-11 manuscript.tex
manuscrpt changes

@@
    \cortext[c1]{Corresponding author}
+    \author[boettig]{Carl Boettiger}
+   \ead{carl(at)ropensci.org}
+
       \address[cstar]{University of California, Berkeley, CA, USA}    
+    \address[boettig]{University of California, Berkeley, CA, USA}
 
@@
 of the research process, including data management, data manipulation
-and cleaning, statistics, and vizualization. Thus, an R client for
+and cleaning, statistics, and visualization. Thus, an R client for
 getting GBIF data is a powerful tool to facilitate reproducible
@@
 The \texttt{rgbif} package is nearly completely written in R (a small
-Javascript library is included for reading well known text), uses an
-\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
-use everywhere. \texttt{rgbif} is developed publicly on GitHub at
+Javascript library is included for reading well known text (Herring,
+2011)), uses an \href{http://choosealicense.com/licenses/mit/}{MIT
+license} to maximize use everywhere. \texttt{rgbif} is developed
+publicly on GitHub at
 \href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2580).
-
-Here, we seaerch for dataset metrics for a single dataset, with uuid
+lot of them (2637).
+
+Here, we search for dataset metrics for a single dataset, with uuid
 \texttt{ec93a739-1681-4b04-b62f-3a687127a17f}, a checklist of the ants
@@
 known to GBIF's database. This serves two purposes: 1) when referring to
-a taxonomic name, you can point to a URI on the internet, and 2) you can
+a taxonomic name, you can point to a URI on the Internet, and 2) you can
 search for metadata on a taxon, and occurrences of that taxon in GBIF.
@@
 \texttt{name\_suggest()}, which does a very quick search and returns
-minimal data. Here, we're searching for the query tem \emph{Pum}, and we
-get back many names:
+minimal data. Here, we're searching for the query term \emph{Pum}, and
+we get back many names:
 
@@
 \texttt{/occurrence/download} route (many functions, see below).
-\texttt{occ\_search()} is the main funtion for the search route, and is
+\texttt{occ\_search()} is the main function for the search route, and is
 more appropriate for smaller data, while \texttt{occ\_download*()}
@@
 searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
-not do that in the occurrence search API. Thus, unfortnately, we
-couldn't make the query interace exactly the same for both search and
+not do that in the occurrence search API. Thus, unfortunately, we
+couldn't make the query interface exactly the same for both search and
 download functions.
@@
 \item
-  Get a single numeric count of occurrenes - \texttt{occ\_count()}
+  Get a single numeric count of occurrences - \texttt{occ\_count()}
 \item
@@
 
-Spatial search, based on well known text format, or a bounding box set
-of four coordinates
+Spatial search, based on well known text format (Herring, 2011), or a
+bounding box set of four coordinates
 
@@
 data. \texttt{rgbif} per se is largely not concerned with making this
-easier, although we do have a simple wrappar around \texttt{ggplot2} to
+easier, although we do have a simple wrapper around \texttt{ggplot2} to
 make it easy to get a quick plot of occurrence data. For example, here
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{components/figure/manuscript-unnamed-chunk-20-1.pdf}
-\caption{}
-\end{figure}
-
-Another package, \href{https://github.com/ropensci/spocc}{spoccutils}
+
+Another package, \href{https://github.com/ropensci/mapr}{mapr}, is the
+perfect mapping companion to \texttt{rgbif}. It has convenient functions
+for handling input data from \texttt{rgbif}, \texttt{spocc}, or
+arbitrary data.frame's, and output plots for base plots,
+\texttt{ggplot2}, \texttt{ggmap} (\texttt{ggplot2} with map layers
+underneath), and interactive maps on GitHub gists or with Leaflet.js.
 
@@
 
-Some of the GBIF taxonomic services are also avialable in
+Some of the GBIF taxonomic services are also available in
 \href{https://github.com/ropensci/taxize}{taxize}, an R package that
@@
 occurrence data from GBIF, iDigBio, and iNaturalist, and easily combine
-them, then use other pacakages to clean and vizualize the data.
+them, then use other packages to clean and visualize the data.
 
@@
 
-The following are three use cases for \texttt{rgbif}: niche modelling,
+The following are three use cases for \texttt{rgbif}: niche modeling,
 spatial change in biodiversity, and distribution mapping.
@@
 \subsubsection{Ecological niche
-modelling}\label{ecological-niche-modelling}
+modeling}\label{ecological-niche-modeling}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{components/figure/manuscript-unnamed-chunk-26-1.pdf}
-\caption{}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
-\caption{}
-\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
 \includegraphics{components/figure/manuscript-unnamed-chunk-38-1.pdf}
-\caption{}
-\end{figure}
 
@@
 
-\section{Acknowledgements}\label{acknowledgements}
+\section{Acknowledgments}\label{acknowledgments}
 
@@
 
+Herring J. 2011. OpenGIS implementation standard for geographic
+information-simple feature access-part 1: Common architecture. \emph{OGC
+Document} 4:122--127.
+
 Mar{í}a Mendoza., Ospina OE., C{á}rdenas-Henao H., Garc{í}a-R JC. 2015.

### 2016-03-11 manuscript.Rmd
converted ms to discuss python in addition to R client

@@
 linenumbers: true
-title: "rgbif: R client for working with GBIF species occurrence data"
+title: "R and Python clients for GBIF species occurrence data"
 author:
@@
 
-# The rgbif package
+# The clients
+
+## The rgbif package
 
@@
 
-## Package interface
+## The pygbif library
 
-`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
+`pygbif` [@pygbif] is a Python library for working with GBIF data in
+the Python programming environment [@python]. Python is a general
+purpose programming language used widely in all sectors, and for
+all parts of software development including server and client
+side use cases. Python is used exclusively in some scientific
+disciplines (e.g., astronomy), and has partial usage in other
+disciplines. A Python client for GBIF data is an important tool
+given the even wider usage of Python than R, though maybe slightly
+less than R for ecology/biology.
+
+```{python eval=FALSE}
+pip install pygbif
+```
+
+```{python eval=FALSE}
+import pygbif
+```
+
+The `pygbif` library is less mature and complete than the R package. It also
+uses an [MIT license][mit] to maximize use everywhere. `pygbif` is developed
+publicly on GitHub at [https://github.com/sckott/pygbif][ghpygbif], where
+development versions of the package can be installed, and bugs and feature
+requests reported. Stable versions of `pygbif` can be installed from
+[pypi][pypipygbif], the distribution network for Python libraries.
+
+### Library interfaces
+
+`rgbif` and `pygbif` are designed following the [GBIF Application Programming Interface][gbifdocs],
 or API. The GBIF API has four major components: registry, taxonomic names,
@@
 GBIF service; only dataset information is available vis this service, however. We
-ignore maps in `rgbif` as it is concerned with generating maps primarily for web
-applications. `rgbif` has a suite of functions dealing with each of registry, taxonomic
+ignore maps in both libraries as it is concerned with generating maps primarily for web
+applications. Both libraries have a suite of functions dealing with each of registry, taxonomic
 names, and occurrences - we'll go through each in turn describing design and example usage.
 
-## GBIF feedback loop
+## GBIF headers
 
-With each request `rgbif` makes to GBIF's API, we send request headers that
-tell GBIF that the request is coming from `rgbif`, including what version of
-the package. This helps GBIF know what proportion of requests are coming from
-this package, and therefore R, as most requests likely will come from `rgbif`;
-this information is helpful for them in thinking about how people are using
-GBIF data.
+With each request `rgbif` and `pygbif` make to GBIF's API, we send request headers that
+tell GBIF what client the request is coming from, including what version of
+the library. This helps GBIF know what proportion of requests are coming from
+which client, and therefore from R vs. Python; this information is helpful for GBIF
+in thinking about how people are using GBIF data.
 
@@
 The GBIF registry API services are spread across five sets of functions
-via the main API:
+via the main GBIF API:
 
@@
 
-And dataset information in general is available via the OAI-PMH service,
-functions in `rgbif` prefixed with `gbif_oai_`.
+Dataset information in general is available via the OAI-PMH service,
+functions in `rgbif` prefixed with `gbif_oai_`, but not available in `pygbif`
+yet.
 
@@
 
-See also `datasets()` and `dataset_suggest()` for searching for datasets.
+See also `datasets()` and `dataset_suggest()` in `rgbif` for searching for datasets.
+
+In Python, we can similary search for datasets. Here, search for datasets of type
+`OCCURRENCE`:
+
+```{python eval=FALSE}
+from pygbif import registry
+registry.datasets(type="OCCURRENCE")
+```
 
@@
 
+And in Python, get metrics for the same dataset as above:
+
+```{python}
+from pygbif import registry
+registry.dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
+```
+
 ### Networks, nodes, and installations
@@
 above datasets, but can be useful if you want to explore data from given
-organizations. Here, we search for the first 10 GBIF networks, returning
+organizations. Here, in R we search for the first 10 GBIF networks, returning
 just the title field.
@@
 ```{r}
-networks(limit=10)$data$title
+networks(limit = 10)$data$title
+```
+
+And in Python:
+
+```{python}
+from pygbif import registry
+registry.networks(limit = 10)
 ```
@@
 
-The GBIF taxonomic names API services are spread across five functions:
+The GBIF taxonomic names API services are spread across five functions in `rgbif`:
 
@@
 * GBIF name parser - `parsenames()`
+
+`pygbif` only has `name_backbone()` and `name_suggest()` at this time.
 
@@
 
-We can search the backbone taxonomy with the function `name_backbone()`. Here,
-we're searching for the name _Poa_, restricting to genera, and the family
-_Poaceae_.
+We can search the backbone taxonomy with the function `name_backbone()` in both
+R and Python clients. Here, we're searching for the name _Poa_, restricting to
+genera, and the family _Poaceae_, in R
 
@@
 res[c('usageKey', 'kingdom')]
+```
+
+and in Python
+
+```{python eval=FALSE}
+from pygbif import species
+res = species.name_backbone(name='Poa', rank='genus', family='Poaceae')
+[ res[x] for x in ['usageKey', 'kingdom'] ]
 ```
@@
 
+The same in Python
+
+```{python eval=FALSE}
+from pygbif import species
+species.name_suggest(q='Pum', limit = 6)
+```
+
 With these results, you can then proceed to search for occurrences with the
@@
 GBIF provides two ways to get occurrence data: through the
-`/occurrence/search` route (see `occ_search`), or via the
+`/occurrence/search` route (see `occ_search` in `rgbif`, or
+`occurrences.search` in `pygbif`), or via the
 `/occurrence/download` route (many functions, see below).
-`occ_search()` is the main function for the search route,
+`occ_search()`/`occurrences.search` is the main function for the search route,
 and is more appropriate for smaller data, while
-`occ_download*()` functions are more appropriate for larger
+the download functions are more appropriate for larger
 data requests.
@@
 even given the 200,000 limit. For those that need more data, we have
-created a similar interface in the `download_*()` functions, that should
+created a similar interface in the download functions, that should
 be easy to use. Users should take note that using the download service
@@
 
-```{r}
+```{r eval=FALSE}
 library("dplyr")
@@
   select(gbifID, decimalLatitude, decimalLongitude)
+#>       gbifID abstract accessRights accrualMethod accrualPeriodicity accrualPolicy alternative audience
+#> 1  725767384       NA                         NA                 NA            NA          NA       NA
+#> 2  725767447       NA                         NA                 NA            NA          NA       NA
+#> 3  725767450       NA                         NA                 NA            NA          NA       NA
+#> 4  725767513       NA                         NA                 NA            NA          NA       NA
+#> 5  725767546       NA                         NA                 NA            NA          NA       NA
+#> 6  725767579       NA                         NA                 NA            NA          NA       NA
+#> 7  725767609       NA                         NA                 NA            NA          NA       NA
+#> 8  725767645       NA                         NA                 NA            NA          NA       NA
+#> 9  725767678       NA                         NA                 NA            NA          NA       NA
+#> 10 725767681       NA                         NA                 NA            NA          NA       NA
+#> ..       ...      ...          ...           ...                ...           ...         ...      ...
+#> Variables not shown: available (lgl), bibliographicCitation (chr), conformsTo (lgl), contributor (lgl),
+#>      coverage (lgl), created (chr), creator (lgl), date (lgl), dateAccepted (lgl), dateCopyrighted
+#>      (lgl), dateSubmitted (lgl), description (lgl), educationLevel (lgl), extent (lgl), format (lgl),
+#>      hasFormat (lgl), hasPart (lgl), hasVersion (lgl), identifier (chr), instructionalMethod (lgl),
 ```
+
+In Python
+
+```{python eval = FALSE}
+from pygbif import occurrences as occ
+occ.download('taxonKey = 3119195')
+(res = occ.download_get("0000840-150615163101818", overwrite = True))
+```
+
+We don't have `pygbif` functionality at the moment for importing data, but
+it's coming soon.
 
@@
 
-The downloaded dataset from GBIF is actually a Darwin Core Archive (DwC-A), an
+The downloaded dataset from GBIF is a Darwin Core Archive (DwC-A), an
 internationally recognized biodiversity informatics standard
-(http://rs.tdwg.org/dwc/). The DwC-A downloaded is a compressed folder with
-a number of files, including metadata, citations for each of the datasets included
-in the download, and the data itself, in separate files for each dataset as well
-as one single `.txt` file. In `occ_download_import()`, we simply fetch data from the
-`.txt` file. If you want to dig into the metadata, citations, etc., it is easily
-accessible from the folder on your computer.
+([http://rs.tdwg.org/dwc/](http://rs.tdwg.org/dwc/)). The DwC-A downloaded is a
+compressed folder with a number of files, including metadata, citations for each
+of the datasets included in the download, and the data itself, in separate files for
+each dataset as well as one single `.txt` file. In `rgbif::occ_download_import()`, we
+simply fetch data from the `.txt` file. If you want to dig into the metadata,
+citations, etc., it is easily accessible from the folder on your computer.
 
@@
 
-* Get a single numeric count of occurrences - `occ_count()`
-* Search for occurrences - `occ_search()`
-* A simplified and optimized version of `occ_search()` - `occ_data()`
-* Get occurrences by occurrence identifier - `occ_get()`
-* Get occurrence metadata - `occ_metadata()`
+* Get a single numeric count of occurrences - `rgbif`: `occ_count()` / `pygbif`: `occurrences.count`
+* Search for occurrences - `rgbif`: `occ_search()` / `pygbif`: `occurrences.count`
+* A simplified and optimized version of `rgbif`: `occ_search()` or `occ_data()` / `pygbif`: `occurrences.count`
+* Get occurrences by occurrence identifier - `rgbif`: `occ_get()` / `pygbif`: `occurrences.count`
+* Get occurrence metadata - `rgbif`: `occ_metadata()` / `pygbif`: `occurrences.count`
 
 #### Search for occurrences
+
+__R__
 
@@
 
+__Python__
+
+The equivalent occurrence search workhorse in `pygibf` is `occurrences.search`.
+
+Search by species name, using `name_backbone()` first to get key
+
+```{python eval=FALSE}
+from pygbif import species
+from pygbif import occurrences as occ
+key = species.name_suggest(q = 'Helianthus annuus', rank = 'species')['data'][0]['key']
+occ.search(taxonKey = key, limit = 2)
+```
+
+Instead of getting a taxon key first, you can search for a name directly
+
+```{python eval=FALSE}
+occ.search(scientificName = 'Ursus americanus')
+```
+
+Search for many species
+
+```{python eval=FALSE}
+from pygbif import species
+from pygbif import occurrences as occ
+splist = ['Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa']
+keys = [ species.name_suggest(x)['data'][0]['key'] for x in splist ]
+occ.search(taxonKey = keys, limit = 5)
+```
+
+Spatial search, based on well known text format [@wkt], or a bounding box set of
+four coordinates
+
+```{python eval=FALSE}
+from pygbif import occurrences as occ
+# well known text
+occ.search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
+# bounding box
+occ.search(geometry = '-125.0,38.4,-121.8,40.9', limit = 20)
+```
+
+Get only occurrences with lat/long data
+
+```{python eval=FALSE}
+from pygbif import occurrences as occ
+occ.search(hasCoordinate = True, limit = 5)
+```
+
+Get only those occurrences with spatial issues. Spatial issues are a set of issues
+that are returned in the `issues` field. They each indicate something different
+about that record. For example, the issue `COUNTRY_COORDINATE_MISMATCH` indicates
+that the interpreted occurrence coordinates fall outside of the indicated country.
+You can see how that might be useful when it comes to cleaning your data prior
+to analysis/visualization.
+
+```{python eval=FALSE}
+from pygbif import occurrences as occ
+occ.search(hasGeospatialIssue = True, limit = 5)
+```
+
 #### Data cleaning
@@
 ```
+
+There's no equivalent interface in `pygbif`.
 
@@
 GitHub gists or with Leaflet.js.
+
+There's no equivalent interface in `pygbif`.
 
@@
 
-The `rgbif` R package provides a programmatic interface to GBIF's
+The `rgbif` and `pygbif` libraries provide programmatic interfaces to GBIF's
 application programming interface (API) - a powerful tool for making
@@
 GBIF API.
+
+The `pygbif` library is still in early development, and will
+greatly benefit from any feedback and use cases.
 
@@
 [mapr]: https://github.com/ropensci/mapr
+[ghpygbif]: https://github.com/sckott/pygbif
+[pypipygbif]: https://pypi.python.org/pypi/pygbif

### 2016-03-11 manuscript.tex
converted ms to discuss python in addition to R client

@@
             pdfauthor={},
-            pdftitle={rgbif: R client for working with GBIF species occurrence data},
+            pdftitle={R and Python clients for GBIF species occurrence data},
             colorlinks=true,
@@
 
-  \title{rgbif: R client for working with GBIF species occurrence data}
+  \title{R and Python clients for GBIF species occurrence data}
     \author[cstar]{Scott Chamberlain\corref{c1}}
@@
 Perhaps the most fundamental element in many fields of ecology is the
-individual. The number of individuals of each species in a given
-location forms the basis for many sub-fields of ecology and evolution.
-Some research questions necessitate collecting new data, while others
-can easily take advantage of existing data. In fact, some ecology fields
-are built largely on existing data, e.g., macro-ecology (Brown, 1995;
-Beck et al., 2012).
+individual organism. The number of individuals of each species in a
+given location forms the basis for many sub-fields of ecology and
+evolution. Some research questions necessitate collecting new data,
+while others can easily take advantage of existing data. In fact, some
+ecology fields are built largely on existing data, e.g., macro-ecology
+(Brown, 1995; Beck et al., 2012).
 
@@
 
+\section{Global Biodiversity Information
+Facility}\label{global-biodiversity-information-facility}
+
 There are many online services that collect and maintain specimen
 records. However, Global Biodiversity Information Facility (hereafter,
-GBIF, \url{http://www.gbif.org}) is the largest collection of
-biodiversity records globally, currently with 640 million records, 1.6
-million taxa, 15,000 datasets from 780 publishers (current as of
-2015-11-25). Many large biodiversity warehouses such as iNaturalist
-(\url{http://www.inaturalist.org}), VertNet (\url{http://vertnet.org}),
-and USGS's Biodiversity Information Serving Our Nation (BISON;
-\url{http://bison.usgs.ornl.gov}) all feed into GBIF.
-
-Herein, we describe the \texttt{rgbif} software library (Chamberlain et
+GBIF, \href{http://www.gbif.org}{\url{http://www.gbif.org}}) is the
+largest collection of biodiversity records globally, currently with 643
+million records, 1.6 million taxa, 15,450 datasets from 780 publishers
+(as of 2016-02-09). Many large biodiversity warehouses such as
+iNaturalist
+(\href{http://www.inaturalist.org}{\url{http://www.inaturalist.org}}),
+VertNet (\href{http://vertnet.org}{\url{http://vertnet.org}}), and
+USGS's Biodiversity Information Serving Our Nation (BISON;
+\href{http://bison.usgs.ornl.gov}{\url{http://bison.usgs.ornl.gov}}) all
+feed into GBIF.
+
+The most important data organizational level in GBIF is the individual
+record. In R using \texttt{rgbif} you'll recognize this as a row in the
+data.frame that is returned from \texttt{occ\_search()} or
+\texttt{occ\_data()}. Going upstream, each record is part of a dataset,
+where each dataset is submitted by an organization, organizations are
+organized into nodes, datasets are published through institutions (which
+may be hosted at another organization), and a network is a group of
+datasets (managed by GBIF).
+
+Each record has some taxonomic name associated with it, which itself is
+linked to a lot of other taxonomic data. GBIF maintains their own
+taxonomic data.
+
+All this organization matters because you can navigate through GBIF data
+through all these organizational levels.
+
+\section{The clients}\label{the-clients}
+
+\subsection{The rgbif package}\label{the-rgbif-package}
+
+Herein, we describe the \texttt{rgbif} software package (Chamberlain et
 al.) for working with GBIF data in the R programming environment (R Core
-Team, 2014). R is a widely used language in academia, and in non-profit
-and private sectors. Importantly, R makes it easy to do all of the steps
-of the research process, including data management, data manipulation
-and cleaning, statistics, and visualization. Thus, an R client for
-getting GBIF data is a powerful tool to facilitate reproducible
-research.
-
-\section{The rgbif package}\label{the-rgbif-package}
+Team, 2014). R is a widely used language in academia, as well as
+non-profit and private sectors. Importantly, R makes it easy to execute
+all steps of the research process, including data management, data
+manipulation and cleaning, statistics, and visualization. Thus, an R
+client for getting GBIF data is a powerful tool to facilitate
+reproducible research.
 
@@
 
-\subsection{Package interface}\label{package-interface}
-
-\texttt{rgbif} is designed following the
+\subsection{The pygbif library}\label{the-pygbif-library}
+
+\texttt{pygbif} (Chamberlain) is a Python library for working with GBIF
+data in the Python programming environment ({\textbf{???}}). Python is a
+general purpose programming language used widely in all sectors, and for
+all parts of software development including server and client side use
+cases. Python is used exclusively in some scientific disciplines (e.g.,
+astronomy), and has partial usage in other disciplines. A Python client
+for GBIF data is an important tool given the even wider usage of Python
+than R, though maybe slightly less than R for ecology/biology.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pip install pygbif}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{import} \NormalTok{pygbif}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{pygbif} library is less mature and complete than the R
+package. It also uses an
+\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
+use everywhere. \texttt{pygbif} is developed publicly on GitHub at
+\href{https://github.com/sckott/pygbif}{\url{https://github.com/sckott/pygbif}},
+where development versions of the package can be installed, and bugs and
+feature requests reported. Stable versions of \texttt{pygbif} can be
+installed from \href{https://pypi.python.org/pypi/pygbif}{pypi}, the
+distribution network for Python libraries.
+
+\subsubsection{Library interfaces}\label{library-interfaces}
+
+\texttt{rgbif} and \texttt{pygbif} are designed following the
 \href{http://www.gbif.org/developer/summary}{GBIF Application
@@
 information is available vis this service, however. We ignore maps in
-\texttt{rgbif} as it is concerned with generating maps primarily for web
-applications. \texttt{rgbif} has a suite of functions dealing with each
+both libraries as it is concerned with generating maps primarily for web
+applications. Both libraries have a suite of functions dealing with each
 of registry, taxonomic names, and occurrences - we'll go through each in
@@
 
-\subsection{GBIF feedback loop}\label{gbif-feedback-loop}
-
-With each request \texttt{rgbif} makes to GBIF's API, we send request
-headers that tell GBIF that the request is coming from \texttt{rgbif},
-including what version of the package. This helps GBIF know what
-proportion of requests are coming from this package, and therefore R, as
-most requests likely will come from \texttt{rgbif}; this information is
-helpful for them in thinking about how people are using GBIF data.
+\subsection{GBIF headers}\label{gbif-headers}
+
+With each request \texttt{rgbif} and \texttt{pygbif} make to GBIF's API,
+we send request headers that tell GBIF what client the request is coming
+from, including what version of the library. This helps GBIF know what
+proportion of requests are coming from which client, and therefore from
+R vs.~Python; this information is helpful for GBIF in thinking about how
+people are using GBIF data.
 
@@
 
-And dataset information in general is available via the OAI-PMH service,
-functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}.
+Dataset information in general is available via the OAI-PMH service,
+functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}, but not
+available in \texttt{pygbif} yet.
 
@@
 \NormalTok{res$data$datasetTitle[}\DecValTok{1}\NormalTok{:}\DecValTok{10}\NormalTok{]}
-\CommentTok{#>  [1] "SDNHM Birds Collection"                                                }
-\CommentTok{#>  [2] "CM Birds Collection"                                                   }
-\CommentTok{#>  [3] "condoncollection"                                                      }
-\CommentTok{#>  [4] "Taxonomy in Flux Checklist"                                            }
+\CommentTok{#>  [1] "UCDavis - Western USA - Monarch Butterflies - 1892-2005"                                                                         }
+\CommentTok{#>  [2] "A geographic distribution database of Mononychellus mites (Acari: Tetranychidae) on cassava (Manihot esculenta)"                 }
+\CommentTok{#>  [3] "Plantas Acuáticas de la Orinoquía Colombiana"                                                                                    }
+\CommentTok{#>  [4] "USBombus, contemporary survey data of North American bumble bees (Hymenoptera,  Apidae, Bombus) distributed in the United States"}
 \CommentTok{#>  [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"                                                               }
-\CommentTok{#>  [6] "Bryophyte Collection - University of Washington Herbarium (WTU)"       }
-\CommentTok{#>  [7] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"}
-\CommentTok{#>  [8] "UWFC Ichthyology Collection"                                           }
-\CommentTok{#>  [9] "Lichen Collection - University of Washington Herbarium (WTU)"          }
-\CommentTok{#> [10] "UWBM Mammalogy Collection"}
-\end{Highlighting}
-\end{Shaded}
-
-See also \texttt{datasets()} and \texttt{dataset\_suggest()} for
-searching for datasets.
+\CommentTok{#>  [6] "CM Birds Collection"                                                                                                             }
+\CommentTok{#>  [7] "SDNHM Birds Collection"                                                                                                          }
+\CommentTok{#>  [8] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"                                                          }
+\CommentTok{#>  [9] "University of British Columbia Herbarium (UBC) - Vascular Plant Collection"                                                      }
+\CommentTok{#> [10] "Bryophyte Collection - University of Washington Herbarium (WTU)"}
+\end{Highlighting}
+\end{Shaded}
+
+See also \texttt{datasets()} and \texttt{dataset\_suggest()} in
+\texttt{rgbif} for searching for datasets.
+
+In Python, we can similary search for datasets. Here, search for
+datasets of type \texttt{OCCURRENCE}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
+\NormalTok{registry.datasets(}\DataTypeTok{type}\NormalTok{=}\StringTok{"OCCURRENCE"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2637).
+lot of them (2737).
 
@@
 
+And in Python, get metrics for the same dataset as above:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
+\NormalTok{registry.dataset_metrics(uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 \subsubsection{Networks, nodes, and
@@
 above datasets, but can be useful if you want to explore data from given
-organizations. Here, we search for the first 10 GBIF networks, returning
-just the title field.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit=}\DecValTok{10}\NormalTok{)$data$title}
+organizations. Here, in R we search for the first 10 GBIF networks,
+returning just the title field.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit =} \DecValTok{10}\NormalTok{)$data$title}
 \CommentTok{#>  [1] "GBIF Backbone Sources"                                      }
@@
 
+And in Python:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
+\NormalTok{registry.networks(limit = }\DecValTok{10}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 \subsection{Taxonomic names}\label{taxonomic-names}
 
-The GBIF taxonomic names API services are spread across five functions:
+The GBIF taxonomic names API services are spread across five functions
+in \texttt{rgbif}:
 
@@
 \end{itemize}
+
+\texttt{pygbif} only has \texttt{name\_backbone()} and
+\texttt{name\_suggest()} at this time.
 
@@
 We can search the backbone taxonomy with the function
-\texttt{name\_backbone()}. Here, we're searching for the name
-\emph{Poa}, restricting to genera, and the family \emph{Poaceae}.
+\texttt{name\_backbone()} in both R and Python clients. Here, we're
+searching for the name \emph{Poa}, restricting to genera, and the family
+\emph{Poaceae}, in R
 
@@
 
+and in Python
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
+\NormalTok{res = species.name_backbone(name=}\StringTok{'Poa'}\NormalTok{, rank=}\StringTok{'genus'}\NormalTok{, family=}\StringTok{'Poaceae'}\NormalTok{)}
+\NormalTok{[ res[x] }\KeywordTok{for} \NormalTok{x in [}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{] ]}
+\end{Highlighting}
+\end{Shaded}
+
 \subsubsection{Name searching}\label{name-searching}
@@
 \endhead
-3269133 & Pumilus & GENUS\tabularnewline
-4407849 & Pumilinura & GENUS\tabularnewline
-4323990 & Pumiliopsis & GENUS\tabularnewline
-4324083 & Pumiliopes & GENUS\tabularnewline
-4161281 & Pumilea & GENUS\tabularnewline
-4312370 & Pumilopagurus & GENUS\tabularnewline
+4848380 & Pumiliornis & GENUS\tabularnewline
+2235094 & Pumilibranchipus & GENUS\tabularnewline
+1795637 & Pumora & GENUS\tabularnewline
+4581745 & Pumicia & GENUS\tabularnewline
+4598790 & Pumilicopta & GENUS\tabularnewline
+1593095 & Pumilomyia & GENUS\tabularnewline
 \bottomrule
 \end{longtable}
+
+The same in Python
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
+\NormalTok{species.name_suggest(q=}\StringTok{'Pum'}\NormalTok{, limit = }\DecValTok{6}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 GBIF provides two ways to get occurrence data: through the
-\texttt{/occurrence/search} route (see \texttt{occ\_search}), or via the
-\texttt{/occurrence/download} route (many functions, see below).
-\texttt{occ\_search()} is the main function for the search route, and is
-more appropriate for smaller data, while \texttt{occ\_download*()}
-functions are more appropriate for larger data requests.
-
-Large is of course a subjective term. When you hit a ``large dataset''
-will depend primarily on the size of the your data request. GBIF imposes
-for any given search a limit of 200,000 records in the search service,
-after which point you can't download any more records for that search.
-However, you can download more records for different searches.
+\texttt{/occurrence/search} route (see \texttt{occ\_search} in
+\texttt{rgbif}, or \texttt{occurrences.search} in \texttt{pygbif}), or
+via the \texttt{/occurrence/download} route (many functions, see below).
+\texttt{occ\_search()}/\texttt{occurrences.search} is the main function
+for the search route, and is more appropriate for smaller data, while
+the download functions are more appropriate for larger data requests.
+
+GBIF imposes for any given search a limit of 200,000 records in the
+search service, after which point you can't download any more records
+for that search. However, you can download more records for different
+searches.
 
@@
 given the 200,000 limit. For those that need more data, we have created
-a similar interface in the \texttt{download\_*()} functions, that should
-be easy to use. Users should take note that using the download service
-has a few extra steps to get data into R, but is straight-forward.
+a similar interface in the download functions, that should be easy to
+use. Users should take note that using the download service has a few
+extra steps to get data into R, but is straight-forward.
 
@@
 is different. For example, in the download API you can do greater than
-searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
-not do that in the occurrence search API. Thus, unfortunately, we
-couldn't make the query interface exactly the same for both search and
-download functions.
+searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you cannot
+do that in the occurrence search API. Thus, unfortunately, we couldn't
+make the query interface exactly the same for both search and download
+functions.
 
@@
 \StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
-\CommentTok{#>       gbifID decimalLatitude decimalLongitude}
-\CommentTok{#> 1  725767384        61.01005         24.41740}
-\CommentTok{#> 2  725767447        59.82923         23.13550}
-\CommentTok{#> 3  725767450        60.38505         25.17449}
-\CommentTok{#> 4  725767513        68.37648         23.51963}
-\CommentTok{#> 5  725767546        67.19203         24.85820}
-\CommentTok{#> 6  725767579        60.21607         24.67412}
-\CommentTok{#> 7  725767609        66.49260         25.70471}
-\CommentTok{#> 8  725767645        61.36634         24.76218}
-\CommentTok{#> 9  725767678        62.29174         27.96500}
-\CommentTok{#> 10 725767681        60.28615         22.38489}
-\CommentTok{#> ..       ...             ...              ...}
-\end{Highlighting}
-\end{Shaded}
+\CommentTok{#>       gbifID abstract accessRights accrualMethod accrualPeriodicity accrualPolicy alternative audience}
+\CommentTok{#> 1  725767384       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 2  725767447       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 3  725767450       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 4  725767513       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 5  725767546       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 6  725767579       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 7  725767609       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 8  725767645       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 9  725767678       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> 10 725767681       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{#> ..       ...      ...          ...           ...                ...           ...         ...      ...}
+\CommentTok{#> Variables not shown: available (lgl), bibliographicCitation (chr), conformsTo (lgl), contributor (lgl),}
+\CommentTok{#>      coverage (lgl), created (chr), creator (lgl), date (lgl), dateAccepted (lgl), dateCopyrighted}
+\CommentTok{#>      (lgl), dateSubmitted (lgl), description (lgl), educationLevel (lgl), extent (lgl), format (lgl),}
+\CommentTok{#>      hasFormat (lgl), hasPart (lgl), hasVersion (lgl), identifier (chr), instructionalMethod (lgl),}
+\end{Highlighting}
+\end{Shaded}
+
+In Python
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\NormalTok{occ.download(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
+\NormalTok{(res = occ.download_get(}\StringTok{"0000840-150615163101818"}\NormalTok{, overwrite = }\OtherTok{True}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+We don't have \texttt{pygbif} functionality at the moment for importing
+data, but it's coming soon.
 
@@
 
-The downloaded dataset from GBIF is actually a Darwin Core Archive
-(DwC-A), an internationally recognized biodiversity informatics standard
-(\url{http://rs.tdwg.org/dwc/}). The DwC-A downloaded is a compressed
-folder with a number of files, including metadata, citations for each of
-the datasets included in the download, and the data itself, in separate
-files for each dataset as well as one single \texttt{.txt} file. In
-\texttt{occ\_download\_import()}, we simply fetch data from the
+The downloaded dataset from GBIF is a Darwin Core Archive (DwC-A), an
+internationally recognized biodiversity informatics standard
+(\href{http://rs.tdwg.org/dwc/}{\url{http://rs.tdwg.org/dwc/}}). The
+DwC-A downloaded is a compressed folder with a number of files,
+including metadata, citations for each of the datasets included in the
+download, and the data itself, in separate files for each dataset as
+well as one single \texttt{.txt} file. In
+\texttt{rgbif::occ\_download\_import()}, we simply fetch data from the
 \texttt{.txt} file. If you want to dig into the metadata, citations,
@@
 \item
-  Get a single numeric count of occurrences - \texttt{occ\_count()}
-\item
-  Search for occurrences - \texttt{occ\_search()}
-\item
-  A simplified and optimized version of \texttt{occ\_search()} -
-  \texttt{occ\_data()}
-\item
-  Get occurrences by occurrence identifier - \texttt{occ\_get()}
-\item
-  Get occurrence metadata - \texttt{occ\_metadata()}
+  Get a single numeric count of occurrences - \texttt{rgbif}:
+  \texttt{occ\_count()} / \texttt{pygbif}: \texttt{occurrences.count}
+\item
+  Search for occurrences - \texttt{rgbif}: \texttt{occ\_search()} /
+  \texttt{pygbif}: \texttt{occurrences.count}
+\item
+  A simplified and optimized version of \texttt{rgbif}:
+  \texttt{occ\_search()} or \texttt{occ\_data()} / \texttt{pygbif}:
+  \texttt{occurrences.count}
+\item
+  Get occurrences by occurrence identifier - \texttt{rgbif}:
+  \texttt{occ\_get()} / \texttt{pygbif}: \texttt{occurrences.count}
+\item
+  Get occurrence metadata - \texttt{rgbif}: \texttt{occ\_metadata()} /
+  \texttt{pygbif}: \texttt{occurrences.count}
 \end{itemize}
@@
 \paragraph{Search for occurrences}\label{search-for-occurrences}
+
+\textbf{R}
 
@@
 
-The output of \texttt{occ\_search()} borrows the tidy
-\texttt{data.frame} idea from the \texttt{dplyr} R package, so that no
-matter how large the \texttt{data.frame}, the output is easily assessed
-because only a few of the records (rows) are shown, only a few columns
-are shown (with others shown in name only), and metadata is shown on top
-of the \texttt{data.frame} to indicate data found and returned, media
-records found, unique taxonomic hierarchies returned, and the query
-executed.
+The output of \texttt{occ\_search()} presents a compact
+\texttt{data.frame} so that no matter how large the \texttt{data.frame},
+the output is easily assessed because only a few of the records (rows)
+are shown, only a few columns are shown (with others shown in name
+only), and metadata is shown on top of the \texttt{data.frame} to
+indicate data found and returned, media records found, unique taxonomic
+hierarchies returned, and the query executed.
 
@@
 \KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
-\CommentTok{#> Records found [21577] }
+\CommentTok{#> Records found [21538] }
 \CommentTok{#> Records returned [2] }
@@
 \CommentTok{#>                name        key decimalLatitude decimalLongitude}
-\CommentTok{#> 1 Helianthus annuus 1143516596        35.42767        -105.0688}
-\CommentTok{#> 2 Helianthus annuus 1095851641         0.00000           0.0000}
+\CommentTok{#> 1 Helianthus annuus 1249279611        34.04810       -117.79884}
+\CommentTok{#> 2 Helianthus annuus 1249286909        32.58747        -97.10081}
 \CommentTok{#> Variables not shown: issues (chr), datasetKey (chr), publishingOrgKey}
@@
 \CommentTok{#>      verbatimLocality (chr), collectionCode (chr), occurrenceID (chr),}
-\CommentTok{#>      taxonID (chr), recordedBy (chr), catalogNumber (chr),}
+\CommentTok{#>      taxonID (chr), license (chr), recordedBy (chr), catalogNumber (chr),}
 \CommentTok{#>      http...unknown.org.occurrenceDetails (chr), institutionCode (chr),}
-\CommentTok{#>      rights (chr), occurrenceRemarks (chr), identificationID (chr),}
-\CommentTok{#>      elevation (dbl), elevationAccuracy (dbl), stateProvince (chr),}
-\CommentTok{#>      recordNumber (chr), locality (chr), municipality (chr), language}
-\CommentTok{#>      (chr), type (chr), ownerInstitutionCode (chr), identifiedBy (chr)}
+\CommentTok{#>      rights (chr), eventTime (chr), identificationID (chr),}
+\CommentTok{#>      coordinateAccuracy (dbl), coordinateAccuracyInMeters (dbl),}
+\CommentTok{#>      occurrenceRemarks (chr)}
 \end{Highlighting}
@@
 
-Get only those occurrences with spatial issues
+Get only those occurrences with spatial issues. Spatial issues are a set
+of issues that are returned in the \texttt{issues} field. They each
+indicate something different about that record. For example, the issue
+\texttt{COUNTRY\_COORDINATE\_MISMATCH} indicates that the interpreted
+occurrence coordinates fall outside of the indicated country. You can
+see how that might be useful when it comes to cleaning your data prior
+to analysis/visualization.
 
@@
 \KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasGeospatialIssue =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Python}
+
+The equivalent occurrence search workhorse in \texttt{pygibf} is
+\texttt{occurrences.search}.
+
+Search by species name, using \texttt{name\_backbone()} first to get key
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\NormalTok{key = species.name_suggest(q = }\StringTok{'Helianthus annuus'}\NormalTok{, rank = }\StringTok{'species'}\NormalTok{)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{]}
+\NormalTok{occ.search(taxonKey = key, limit = }\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Instead of getting a taxon key first, you can search for a name directly
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{occ.search(scientificName = }\StringTok{'Ursus americanus'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Search for many species
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\NormalTok{splist = [}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{]}
+\NormalTok{keys = [ species.name_suggest(x)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{] }\KeywordTok{for} \NormalTok{x in splist ]}
+\NormalTok{occ.search(taxonKey = keys, limit = }\DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Spatial search, based on well known text format (Herring, 2011), or a
+bounding box set of four coordinates
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\CommentTok{# well known text}
+\NormalTok{occ.search(geometry = }\StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, limit = }\DecValTok{20}\NormalTok{)}
+\CommentTok{# bounding box}
+\NormalTok{occ.search(geometry = }\StringTok{'-125.0,38.4,-121.8,40.9'}\NormalTok{, limit = }\DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Get only occurrences with lat/long data
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\NormalTok{occ.search(hasCoordinate = }\OtherTok{True}\NormalTok{, limit = }\DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Get only those occurrences with spatial issues. Spatial issues are a set
+of issues that are returned in the \texttt{issues} field. They each
+indicate something different about that record. For example, the issue
+\texttt{COUNTRY\_COORDINATE\_MISMATCH} indicates that the interpreted
+occurrence coordinates fall outside of the indicated country. You can
+see how that might be useful when it comes to cleaning your data prior
+to analysis/visualization.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\NormalTok{occ.search(hasGeospatialIssue = }\OtherTok{True}\NormalTok{, limit = }\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
-\CommentTok{#> [1] 2}
-\end{Highlighting}
-\end{Shaded}
+\CommentTok{#> [1] 4}
+\end{Highlighting}
+\end{Shaded}
+
+There's no equivalent interface in \texttt{pygbif}.
 
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-20-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-34-1.pdf}
 
@@
 underneath), and interactive maps on GitHub gists or with Leaflet.js.
+
+There's no equivalent interface in \texttt{pygbif}.
 
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-26-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-40-1.pdf}
 
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-45-1.pdf}
 
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-38-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-52-1.pdf}
 
@@
 
-The \texttt{rgbif} R package provides a programmatic interface to GBIF's
-application programming interface (API) - a powerful tool for making
-research using species occurrence data reproducible. In fact, the
-\texttt{rgbif} package has already been used in 22 scholarly
-publications (as of 2015-11-14).
+The \texttt{rgbif} and \texttt{pygbif} libraries provide programmatic
+interfaces to GBIF's application programming interface (API) - a
+powerful tool for making research using species occurrence data
+reproducible. In fact, the \texttt{rgbif} package has already been used
+in 22 scholarly publications (as of 2015-11-14).
 
@@
 API.
+
+The \texttt{pygbif} library is still in early development, and will
+greatly benefit from any feedback and use cases.
 
@@
 This project was supported in part by the Alfred P Sloan Foundation
-(Grant 2013-6-22), and the Helmsley Foundation (Grant XXXXXX).
+(Grant No. G-2014-13485), and in part by the Helmsley Foundation (Grant
+No. 2016PG-BRI004).
 
@@
 
-Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{Rgbif: Interface to
+Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{Pygbif: Interface to
 the global 'biodiversity' information facility 'aPI'}.
+
+Chamberlain S. \emph{Rgbif: Python client for gBIF}.
 

### 2016-08-05 manuscript.Rmd
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

@@
     address: |
-      University of California, Berkeley, CA, USA
+      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
   - code: boettig
     address: |
-      University of California, Berkeley, CA, USA
+      rOpenSci, Department of Enivornmental Science, Policy and Management, University of California, Berkeley, CA, USA
 abstract: |
-      1. xxx
+      Corresponding Author:
 
-      2. xxx
+      Scott Chamberlain
 
-      3. xxx
+      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
 
-      4. xxxx
+      Email address: [scott@ropensci.org](mailto:scott@ropensci.org)
+
+      \newpage
+
+      Background. The number of individuals of each species in
+        a given location forms the basis for many sub-fields of ecology and
+        evolution. Data on individuals, including which species, and
+        where they're found, can be used for a large number of research
+        questions. Global Biodiversity Information Facility (hereafter, GBIF)
+        is the largest of these. Programmatic clients for GBIF would
+        make research dealing with GBIF data much easier.
+
+      Methods. We have developed a client to access GBIF data for each
+        of the R and Python programming languages: `rgbif` and `pygbif`.
+
+      Results. For both clients, we describe their design and utility,
+        and demonstrate many use cases.
+
+      Discussion. Programmatic acccess to GBIF will facilitate more open
+        and reproducible science. We hope that all scientists using GBIF
+        data will do so with one of these two clients.
 
@@
 )
+
 basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
@@
                cache.path = paste("components/cache/", basename, "/", sep=""))
+
+# tibble options
+options(tibble.max_extra_cols = 10)
 ```
+
+\newpage
 
@@
 
-All this organization matters because you can navigate through GBIF data through all
-these organizational levels.
+This organization matters because you can navigate through GBIF data through
+these various organizational levels.
 
@@
 
-## The rgbif package
+## rgbif
 
@@
 
-## The pygbif library
+## pygbif
 
 `pygbif` [@pygbif] is a Python library for working with GBIF data in
-the Python programming environment [@python]. Python is a general
+the Python programming environment. Python is a general
 purpose programming language used widely in all sectors, and for
@@
 given the even wider usage of Python than R, though maybe slightly
-less than R for ecology/biology.
+less than R for ecology/biology disciplines.
 
@@
 uses an [MIT license][mit] to maximize use everywhere. `pygbif` is developed
-publicly on GitHub at [https://github.com/sckott/pygbif][ghpygbif], where
+publicly on GitHub at <https://github.com/sckott/pygbif>, where
 development versions of the package can be installed, and bugs and feature
@@
 
-The output of these examples, except one, aren't shown, but all run correctly.
+The output of these examples, except one, aren't shown.
 
@@
 Spatial search, based on well known text format [@wkt], or a bounding box set of
-four coordinates
+four coordinates. The well known text string and the bounding box in the below
+example specify the same rectangular area in California, centering
+approximately on Sacramento. Whereas the bounding box format requires
+`longitude SW corner, latitude SW corner, longitude NE corner, latitude NE corner`,
+the well known text string requires an extra long/lat pair to close the polygon.
 
@@
 # well known text
-occ_search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
+wkt <- 'POLYGON((-122.6 39.9,-120.0 39.9,-120.0 37.9,-122.6 37.9,-122.6 39.9))'
+occ_search(geometry = wkt, limit = 20)
 # bounding box
-occ_search(geometry = c(-125.0,38.4,-121.8,40.9), limit = 20)
+occ_search(geometry = c(-122.6,37.9,-120.0,39.9), limit = 20)
 ```
 
-Get only occurrences with lat/long data
+Get only occurrences with lat/long data using the `hasCoordinate` parameter
 
@@
 
-There's no equivalent interface in `pygbif`.
+There's no equivalent interface in `pygbif` yet.
 
@@
 
+# R vs. Python
+
+Both R and Python are commonly used in science, and can be used for similar
+tasks. Python, however, is a more general programming lanugae, and can be
+used in more contexts than R can be used in.
+
+The `rgbif` and `pygbif` clients can do a lot of the same tasks. We
+envision `rgbif` as being more common in workflows of academics asking
+research questions, whereas `pygbif` can do that as well, but may be more
+easily used in a website.
+
+The R client `rgbif` has much more development time than `pygbif`,
+but through time `pygbif` will become equally mature.
+
 # Use cases
@@
 The `rgbif` and `pygbif` libraries provide programmatic interfaces to GBIF's
-application programming interface (API) - a powerful tool for making
-research using species occurrence data reproducible. In fact,
+application programming interface (API) - a powerful tool for working with
+species occurrence data, and facilitating reproducible research. In fact,
 the `rgbif` package has already been used in 22 scholarly publications
@@
 
-The `pygbif` library is still in early development, and will
+The `pygbif` library is in early development, and will
 greatly benefit from any feedback and use cases.
@@
 scenarios. Making removal of these occurrences easy would be
-very useful, but is actually quite a hard problem.
+very useful, but is actually quite a hard problem. There are many
+other problems like this.
 
@@
 corresponds to a snapshot of the GitHub repository at
-[github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
-can be found at [https://github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
-under an MIT license.
+<https://github.com/sckott/msrgbif>. Software can be found at
+<https://github.com/ropensci/rgbif> and <https://github.com/sckott/pygbif>,
+both under MIT licenses. We thank all the users that have used `rgbif`
+and `pygbif` and have given feedback and reported bugs. In addition, we
+greatly appreciate the all contributions to the two packages, found
+at <https://github.com/ropensci/rgbif/graphs/contributors> and
+<https://github.com/sckott/pygbif/graphs/contributors>.
 

### 2016-08-05 manuscript.tex
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

@@
 
-      \address[cstar]{University of California, Berkeley, CA, USA}    
-    \address[boettig]{University of California, Berkeley, CA, USA}
+      \address[cstar]{rOpenSci, Museum of Paleontology, University of California, Berkeley,
+CA, USA}
+    \address[boettig]{rOpenSci, Department of Enivornmental Science, Policy and Management,
+University of California, Berkeley, CA, USA}
 
   \begin{abstract}
-  \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
-  \item
-    xxx
-  \item
-    xxx
-  \item
-    xxx
-  \item
-    xxxx
-  \end{enumerate}
+  Corresponding Author:
+
+  Scott Chamberlain
+
+  rOpenSci, Museum of Paleontology, University of California, Berkeley,
+  CA, USA
+
+  Email address:
+  \href{mailto:scott@ropensci.org}{\href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}}
+
+  \newpage
+
+  Background. The number of individuals of each species in a given
+  location forms the basis for many sub-fields of ecology and evolution.
+  Data on individuals, including which species, and where they're found,
+  can be used for a large number of research questions. Global
+  Biodiversity Information Facility (hereafter, GBIF) is the largest of
+  these. Programmatic clients for GBIF would make research dealing with
+  GBIF data much easier.
+
+  Methods. We have developed a client to access GBIF data for each of the
+  R and Python programming languages: \texttt{rgbif} and \texttt{pygbif}.
+
+  Results. For both clients, we describe their design and utility, and
+  demonstrate many use cases.
+
+  Discussion. Programmatic acccess to GBIF will facilitate more open and
+  reproducible science. We hope that all scientists using GBIF data will
+  do so with one of these two clients.
   \end{abstract}
@@
  \end{frontmatter}
+
+\newpage
 
@@
 
-All this organization matters because you can navigate through GBIF data
-through all these organizational levels.
+This organization matters because you can navigate through GBIF data
+through these various organizational levels.
 
@@
 
-\subsection{The rgbif package}\label{the-rgbif-package}
+\subsection{rgbif}\label{rgbif}
 
@@
 
-\subsection{The pygbif library}\label{the-pygbif-library}
+\subsection{pygbif}\label{pygbif}
 
 \texttt{pygbif} (Chamberlain) is a Python library for working with GBIF
-data in the Python programming environment ({\textbf{???}}). Python is a
-general purpose programming language used widely in all sectors, and for
-all parts of software development including server and client side use
-cases. Python is used exclusively in some scientific disciplines (e.g.,
-astronomy), and has partial usage in other disciplines. A Python client
-for GBIF data is an important tool given the even wider usage of Python
-than R, though maybe slightly less than R for ecology/biology.
+data in the Python programming environment. Python is a general purpose
+programming language used widely in all sectors, and for all parts of
+software development including server and client side use cases. Python
+is used exclusively in some scientific disciplines (e.g., astronomy),
+and has partial usage in other disciplines. A Python client for GBIF
+data is an important tool given the even wider usage of Python than R,
+though maybe slightly less than R for ecology/biology disciplines.
 
@@
 use everywhere. \texttt{pygbif} is developed publicly on GitHub at
-\href{https://github.com/sckott/pygbif}{\url{https://github.com/sckott/pygbif}},
-where development versions of the package can be installed, and bugs and
-feature requests reported. Stable versions of \texttt{pygbif} can be
-installed from \href{https://pypi.python.org/pypi/pygbif}{pypi}, the
-distribution network for Python libraries.
+\url{https://github.com/sckott/pygbif}, where development versions of
+the package can be installed, and bugs and feature requests reported.
+Stable versions of \texttt{pygbif} can be installed from
+\href{https://pypi.python.org/pypi/pygbif}{pypi}, the distribution
+network for Python libraries.
 
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (2737).
+lot of them (16057).
 
@@
 refer to taxonomic names throughout their services. The backbone has
-4410899 unique names and 2497114 species names. The backbone taxonomy is
+5194833 unique names and 2420842 species names. The backbone taxonomy is
 also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
@@
 
-The output of these examples, except one, aren't shown, but all run
-correctly.
+The output of these examples, except one, aren't shown.
 
@@
 Spatial search, based on well known text format (Herring, 2011), or a
-bounding box set of four coordinates
+bounding box set of four coordinates. The well known text string and the
+bounding box in the below example specify the same rectangular area in
+California, centering approximately on Sacramento. Whereas the bounding
+box format requires
+\texttt{longitude SW corner, latitude SW corner, longitude NE corner, latitude NE corner},
+the well known text string requires an extra long/lat pair to close the
+polygon.
 
@@
 \CommentTok{# well known text}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\NormalTok{wkt <-}\StringTok{ 'POLYGON((-122.6 39.9,-120.0 39.9,-120.0 37.9,-122.6 37.9,-122.6 39.9))'}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{wkt, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
 \CommentTok{# bounding box}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{125.0}\NormalTok{,}\FloatTok{38.4}\NormalTok{,-}\FloatTok{121.8}\NormalTok{,}\FloatTok{40.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Get only occurrences with lat/long data
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{122.6}\NormalTok{,}\FloatTok{37.9}\NormalTok{,-}\FloatTok{120.0}\NormalTok{,}\FloatTok{39.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Get only occurrences with lat/long data using the \texttt{hasCoordinate}
+parameter
 
@@
 
-There's no equivalent interface in \texttt{pygbif}.
+There's no equivalent interface in \texttt{pygbif} yet.
 
@@
 
+\section{R vs.~Python}\label{r-vs.python}
+
+Both R and Python are commonly used in science, and can be used for
+similar tasks. Python, however, is a more general programming lanugae,
+and can be used in more contexts than R can be used in.
+
+The \texttt{rgbif} and \texttt{pygbif} clients can do a lot of the same
+tasks. We envision \texttt{rgbif} as being more common in workflows of
+academics asking research questions, whereas \texttt{pygbif} can do that
+as well, but may be more easily used in a website.
+
+The R client \texttt{rgbif} has much more development time than
+\texttt{pygbif}, but through time \texttt{pygbif} will become equally
+mature.
+
 \section{Use cases}\label{use-cases}
@@
 \KeywordTok{library}\NormalTok{(}\StringTok{"maptools"}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
-\NormalTok{df <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{class}\NormalTok{(x$data)) %in%}\StringTok{ "data.frame"} \NormalTok{]}
-\NormalTok{df <-}\StringTok{ }\KeywordTok{ldply}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
-\NormalTok{df2 <-}\StringTok{ }\NormalTok{df[,}\KeywordTok{c}\NormalTok{(}\StringTok{'decimalLongitude'}\NormalTok{,}\StringTok{'decimalLatitude'}\NormalTok{)]}
+\NormalTok{df_data <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{any}\NormalTok{(}\KeywordTok{class}\NormalTok{(x$data) %in%}\StringTok{ "tbl_df"}\NormalTok{)) ]}
+\NormalTok{df_data <-}\StringTok{ }\NormalTok{dplyr::}\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df_data, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
+\NormalTok{df2 <-}\StringTok{ }\NormalTok{df_data %>%}\StringTok{ }\NormalTok{dplyr::}\KeywordTok{select}\NormalTok{(decimalLongitude, decimalLatitude)}
 \end{Highlighting}
@@
 \KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{"RCurl"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"httr"}\NormalTok{)}
 \KeywordTok{library}\NormalTok{(}\StringTok{"RColorBrewer"}\NormalTok{)}
@@
 Bounding lat/long data is from
-\href{https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt}{here}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{getURL}\NormalTok{(}\StringTok{'https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt'}\NormalTok{)}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}\DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{sep=}\StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{))}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{dat$city, }\DataTypeTok{minlon=}\NormalTok{dat$minlon, }\DataTypeTok{minlat=}\NormalTok{dat$minlat, }\DataTypeTok{maxlon=}\NormalTok{dat$maxlon, }\DataTypeTok{maxlat=}\NormalTok{dat$maxlat)}
-\end{Highlighting}
-\end{Shaded}
+\url{https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{url <-}\StringTok{ 'https://raw.githubusercontent.com/amyxzhang/}
+\StringTok{boundingbox-cities/master/boundbox.txt'}
+\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{content}\NormalTok{(}\KeywordTok{GET}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\StringTok{""}\NormalTok{, url)), }\DataTypeTok{as =} \StringTok{"text"}\NormalTok{)}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
+  \DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }
+  \DataTypeTok{sep =} \StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{),}
+  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
+  \DataTypeTok{city =} \NormalTok{dat$city, }\DataTypeTok{minlon =} \NormalTok{dat$minlon, }
+  \DataTypeTok{minlat =} \NormalTok{dat$minlat, }\DataTypeTok{maxlon =} \NormalTok{dat$maxlon, }
+  \DataTypeTok{maxlat =} \NormalTok{dat$maxlat,}
+  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+A helper function to get count data. GBIF has a count API, but we can't
+use that with a geometry search as that API doesn't support geospatial
+search. We can however use the search API via \texttt{occ\_search()} and
+set \texttt{limit = 1} so that we
 
@@
   \NormalTok{coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
-  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords)$meta$count}
-  \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{x[}\StringTok{'city'}\NormalTok{], }\DataTypeTok{richness=}\NormalTok{num, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
+  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)$meta$count}
+  \KeywordTok{data.frame}\NormalTok{(}
+    \DataTypeTok{city =} \NormalTok{x[}\StringTok{'city'}\NormalTok{], }
+    \DataTypeTok{richness =} \NormalTok{num, }
+    \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
+  \NormalTok{)}
 \NormalTok{\}}
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by=}\StringTok{"city"}\NormalTok{)}
+\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by =} \StringTok{"city"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat+maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon+maxlon)/}\DecValTok{2}\NormalTok{)}
+\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat +}\StringTok{ }\NormalTok{maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon +}\StringTok{ }\NormalTok{maxlon)/}\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 
-From
-\href{http://esp.cr.usgs.gov/data/little/}{\url{http://esp.cr.usgs.gov/data/little/}}.
-And save shapefile in same directory
+From \url{http://esp.cr.usgs.gov/data/little}. And save shapefile in
+same directory
 
@@
 \begin{Highlighting}[]
-\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{), }\DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
+     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{),}
+     \DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{,}
+     \DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
 \CommentTok{# add shape from Atlas of US Trees}
-\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{),}
+     \DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
 \CommentTok{# add Gbif presence points}
-\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
-\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
-\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{, }\StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E.
-1971)"}\NormalTok{, }\DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
+\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude,}
+       \DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
+       \DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{,}
+       \DataTypeTok{legend =} \StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{,}
+       \DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
 \end{Highlighting}
@@
 interfaces to GBIF's application programming interface (API) - a
-powerful tool for making research using species occurrence data
-reproducible. In fact, the \texttt{rgbif} package has already been used
-in 22 scholarly publications (as of 2015-11-14).
+powerful tool for working with species occurrence data, and facilitating
+reproducible research. In fact, the \texttt{rgbif} package has already
+been used in 22 scholarly publications (as of 2015-11-14).
 
@@
 
-The \texttt{pygbif} library is still in early development, and will
-greatly benefit from any feedback and use cases.
+The \texttt{pygbif} library is in early development, and will greatly
+benefit from any feedback and use cases.
 
@@
 these occurrences easy would be very useful, but is actually quite a
-hard problem.
+hard problem. There are many other problems like this.
 
@@
 corresponds to a snapshot of the GitHub repository at
-\href{https://github.com/sckott/msrgbif}{github.com/sckott/msrgbif}.
-Software can be found at
-\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
-under an MIT license.
+\url{https://github.com/sckott/msrgbif}. Software can be found at
+\url{https://github.com/ropensci/rgbif} and
+\url{https://github.com/sckott/pygbif}, both under MIT licenses. We
+thank all the users that have used \texttt{rgbif} and \texttt{pygbif}
+and have given feedback and reported bugs. In addition, we greatly
+appreciate the all contributions to the two packages, found at
+\url{https://github.com/ropensci/rgbif/graphs/contributors} and
+\url{https://github.com/sckott/pygbif/graphs/contributors}.
 
@@
 
-Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{Pygbif: Interface to
-the global 'biodiversity' information facility 'aPI'}.
-
-Chamberlain S. \emph{Rgbif: Python client for gBIF}.
+Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{rgbif: An r interface
+to the global 'biodiversity' information facility API}.
+
+Chamberlain S. \emph{pygbif: A python interface to the global
+biodiversity information facility API}.
 

### 2016-08-05 use-cases/city-diversity.Rmd
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

@@
 library("plyr")
-library("RCurl")
+library("httr")
 library("RColorBrewer")
@@
 
-Bounding lat/long data is from [here](https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt).
+Bounding lat/long data is from <https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt>.
 
 ```{r box}
-rawdat <- getURL('https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt')
-dat <- read.table(text = rawdat, header = FALSE, sep="\t", col.names=c("city","minlat","maxlon","maxlat","minlon"))
-dat <- data.frame(city=dat$city, minlon=dat$minlon, minlat=dat$minlat, maxlon=dat$maxlon, maxlat=dat$maxlat)
+url <- 'https://raw.githubusercontent.com/amyxzhang/
+boundingbox-cities/master/boundbox.txt'
+rawdat <- content(GET(sub("\n", "", url)), as = "text")
+dat <- read.table(
+  text = rawdat, header = FALSE, 
+  sep = "\t", col.names = c("city","minlat","maxlon","maxlat","minlon"),
+  stringsAsFactors = FALSE)
+dat <- data.frame(
+  city = dat$city, minlon = dat$minlon, 
+  minlat = dat$minlat, maxlon = dat$maxlon, 
+  maxlat = dat$maxlat,
+  stringsAsFactors = FALSE
+)
 ```
+
+A helper function to get count data. GBIF has a count API, but we can't use
+that with a geometry search as that API doesn't support geospatial search.
+We can however use the search API via `occ_search()` and set `limit = 1` so 
+that we
 
@@
   coords <- as.numeric(x[c('minlon','minlat','maxlon','maxlat')])
-  num <- occ_search(geometry = coords)$meta$count
-  data.frame(city=x['city'], richness=num, stringsAsFactors = FALSE)
+  num <- occ_search(geometry = coords, limit = 1)$meta$count
+  data.frame(
+    city = x['city'], 
+    richness = num, 
+    stringsAsFactors = FALSE
+  )
 }
@@
 ```{r}
-out <- merge(dat, ldply(out), by="city")
+out <- merge(dat, ldply(out), by = "city")
 ```
@@
 ```{r}
-out <- transform(out, lat = (minlat+maxlat)/2, lon = (minlon+maxlon)/2)
+out <- transform(out, lat = (minlat + maxlat)/2, lon = (minlon + maxlon)/2)
 ```

### 2016-08-05 use-cases/niche-modelling.Rmd
tried to fix front page for peerj preprint specs still not quite right fixed some wkt stuff improved use cases to not take so much time

@@
 library("maptools")
-library("plyr")
+library("dplyr")
 ```
@@
 df <- occ_search(taxonKey = nn, hasCoordinate = TRUE, limit = 500)
-df <- df[ sapply(df, function(x) class(x$data)) %in% "data.frame" ]
-df <- ldply(lapply(df, "[[", "data"))
-df2 <- df[,c('decimalLongitude','decimalLatitude')]
+df_data <- df[ sapply(df, function(x) any(class(x$data) %in% "tbl_df")) ]
+df_data <- dplyr::bind_rows(lapply(df_data, "[[", "data"))
+df2 <- df_data %>% dplyr::select(decimalLongitude, decimalLatitude)
 ```

### 2017-08-10 manuscript.tex
minor tweaks to ms, added many references

@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (16057).
+lot of them (16084).
 
@@
 refer to taxonomic names throughout their services. The backbone has
-5194833 unique names and 2420842 species names. The backbone taxonomy is
+5296599 unique names and 2525274 species names. The backbone taxonomy is
 also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
@@
 reproducible research. In fact, the \texttt{rgbif} package has already
-been used in 22 scholarly publications (as of 2015-11-14).
+been used in 22 scholarly publications (as of 2016-08-10) (Amano,
+Lamming \& Sutherland, 2016, Bartomeus et al. (2013), Barve (2014), Bone
+et al. (2015), Collins et al. (2015), Drozd \& {Š}ipo{š} (2013), Kong,
+Huang \& Duan (2015), Richardson, Roux \& Wilson (2015), Turner,
+Fr{é}ville \& Rieseberg (2015), Verheijen et al. (2015), Zizka \&
+Antonelli (2015), Butterfield et al. (2016), Dellinger et al. (2015),
+Feitosa et al. (2015), Malhado et al. (2015), Werner et al. (2015),
+Robertson, Visser \& Hui (2016), Davison et al. (2015), Janssens et al.
+(2016)).
 
@@
 
+Amano T., Lamming JDL., Sutherland WJ. 2016. Spatial gaps in global
+biodiversity information and the role of citizen science.
+\emph{BioScience} 66:393--400.
+
+Bartomeus I., Park MG., Gibbs J., Danforth BN., Lakso AN., Winfree R.
+2013. Biodiversity ensures plant-pollinator phenological synchrony
+against climate change. \emph{Ecology Letters} 16:1331--1338.
+
+Barve V. 2014. Discovering and developing primary biodiversity data from
+social networking sites: A novel approach. \emph{Ecological Informatics}
+24:194--199.
+
 Beck J., Ballesteros-Mejia L., Buchmann CM., Dengler J., Fritz SA.,
@@
 
+Bone RE., Smith JAC., Arrigo N., Buerki S. 2015. A macro-ecological
+perspective on crassulacean acid metabolism (CAM) photosynthesis
+evolution in afro-madagascan drylands: Eulophiinae orchids as a case
+study. \emph{New Phytologist} 208:469--481.
+
 Brown JH. 1995. \emph{Macroecology}. University of Chicago Press.
@@
 
+Butterfield BJ., Copeland SM., Munson SM., Roybal CM., Wood TE. 2016.
+Prestoration: Using species in restoration that will persist now and
+into the future. \emph{Restor Ecol}.
+
 Ceballos G., Ehrlich PR., Barnosky AD., Garcia A., Pringle RM., Palmer
@@
 biodiversity information facility API}.
+
+Collins R., Ribeiro ED., Machado VN., Hrbek T., Farias I. 2015. A
+preliminary inventory of the catfishes of the lower rio nhamundá, brazil
+(ostariophysi, siluriformes). \emph{BDJ} 3:e4162.
+
+Davison J., Moora M., Opik M., Adholeya A., Ainsaar L., Ba A., Burla S.,
+Diedhiou AG., Hiiesalu I., Jairus T., Johnson NC., Kane A., Koorem K.,
+Kochar M., Ndiaye C., Partel M., Reier U., Saks U., Singh R., Vasar M.,
+Zobel M. 2015. Global assessment of arbuscular mycorrhizal fungus
+diversity reveals very low endemism. \emph{Science} 349:970--973.
+
+Dellinger AS., Essl F., Hojsgaard D., Kirchheimer B., Klatt S., Dawson
+W., Pergl J., Py{š}ek P., Kleunen M van., Weber E., Winter M., Hörandl
+E., Dullinger S. 2015. Niche dynamics of alien species do not differ
+among sexual and apomictic flowering plants. \emph{New Phytologist}
+209:1313--1323.
+
+Drozd P., {Š}ipo{š} J. 2013. R for all (i): Introduction to the new age
+of biological analyses. \emph{Casopis slezskeho zemskeho muzea (A)} 62.
 
@@
 
+Feitosa YO., Absy ML., Latrubesse EM., Stevaux JC. 2015. Late quaternary
+vegetation dynamics from central parts of the madeira river in brazil.
+\emph{Acta Bot. Bras.} 29:120--128.
+
 Ferretti F., Verd GM., Seret B., {Š}prem JS., Micheli F. 2015. Falling
@@
 Document} 4:122--127.
+
+Janssens SB., Vandelook F., Langhe ED., Verstraete B., Smets E.,
+Vandenhouwe I., Swennen R. 2016. Evolutionary dynamics and biogeography
+of musaceae reveal a correlation between the diversification of the
+banana family and the geological and climatic history of southeast asia.
+\emph{New Phytologist} 210:1453--1465.
+
+Kong X., Huang M., Duan R. 2015. SDMdata: A web-based software tool for
+collecting species occurrence records. \emph{PLOS ONE} 10:e0128295.
+
+Malhado AC., Oliveira-Neto JA., Stropp J., Strona G., Dias LC., Pinto
+LB., Ladle RJ. 2015. Climatological correlates of seed size in amazonian
+forest trees. \emph{J Veg Sci} 26:956--963.
 
@@
 
+Richardson DM., Roux JJL., Wilson JR. 2015. Australian acacias as
+invasive species: Lessons to be learnt from regions with long planting
+histories. \emph{Southern Forests: a Journal of Forest Science}
+77:31--39.
+
+Robertson MP., Visser V., Hui C. 2016. Biogeo: An r package for
+assessing and improving data quality of occurrence record datasets.
+\emph{Ecography} 39:394--401.
+
+Turner KG., Fr{é}ville H., Rieseberg LH. 2015. Adaptive plasticity and
+niche expansion in an invasive thistle. \emph{Ecol Evol} 5:3183--3197.
+
+Verheijen LM., Aerts R., Bönisch G., Kattge J., Bodegom PMV. 2015.
+Variation in trait trade-offs allows differentiation among predefined
+plant functional types: Implications for predictive ecology. \emph{New
+Phytologist} 209:563--575.
+
+Werner GDA., Cornwell WK., Cornelissen JHC., Kiers ET. 2015.
+Evolutionary signals of symbiotic persistence in the legumerhizobia
+mutualism. \emph{Proceedings of the National Academy of Sciences}
+112:10262--10269.
+
+Zizka A., Antonelli A. 2015. \emph{speciesgeocodeR: An r package for
+linking species occurrences, user-defined regions and phylogenetic trees
+for biogeography, ecology and evolution}. Cold Spring Harbor Laboratory
+Press.
+
 \end{document}

### 2017-08-10 manuscript.Rmd

@@
 occurrences, and maps. We also include functions to interface with the OAI-PMH
-GBIF service; only dataset information is available vis this service, however. We
-ignore maps in both libraries as it is concerned with generating maps primarily for web
-applications. Both libraries have a suite of functions dealing with each of registry, taxonomic
-names, and occurrences - we'll go through each in turn describing design and example usage.
+GBIF service; only dataset information is available vis this service, however. We now include
+an interface to the GBIF map API, but we for now ignore maps in `pygbif`. Both libraries have a
+suite of functions dealing with each of registry, taxonomic, names, and occurrences - we'll go
+through each in turn describing design of the user interface and example usage.
 
@@
 
-`pygbif` only has `name_backbone()` and `name_suggest()` at this time.
+`pygbif` has all the same method except `rgbif::parsenames()` is `name_parser()`
+in `pygbif`.
 
@@
 
+## GBIF Mapping API
+
+<!-- FILL ME IN -->
+
 ## GBIF data in other R packages
@@
 species occurrence data, and facilitating reproducible research. In fact,
-the `rgbif` package has already been used in 22 scholarly publications
-(as of 2015-11-14).
+the `rgbif` package has already been used in 20 scholarly publications
+(as of 2016-08-10) [@Amano_2016, @Bartomeus_2013, @Barve_2014, @Bone_2015, @Collins_2015, @Drozd_2013, @Kong_2015, @Richardson_2015, @Turner_2015, @Verheijen_2015, @Zizka_2015, @Butterfield_2016, @Dellinger_2015, @Feitosa_2015, @Malhado_2015, @Werner_2015, @Robertson_2016, @Davison_2015, @Janssens_2016].
 

### 2017-08-10 use-cases/oak-occurrences.Rmd

@@
 
-From [http://esp.cr.usgs.gov/data/little/](http://esp.cr.usgs.gov/data/little/). And save shapefile in same directory
+From <http://esp.cr.usgs.gov/data/little>. And save shapefile in same directory
 
-```{r}
+```{r cache = TRUE}
 url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
@@
 
-```{r}
+```{r cache = TRUE}
 alt.USA <- getData('alt', country = 'USA')
@@
 
-```{r}
+```{r cache = TRUE}
 alt.USA <- alt.USA[[1]]
@@
 ```{r fig.keep='last'}
-plot(hill.USA, col = grey(0:100/100), legend = FALSE, xlim = c(-125, -116), ylim = c(32, 42), main = 'Distribution of Quercus lobata', xlab = "Long", ylab = 'Lat')
+plot(hill.USA, col = grey(0:100/100), legend = FALSE,
+     xlim = c(-125, -116), ylim = c(32, 42),
+     main = 'Distribution of Quercus lobata',
+     xlab = "Long", ylab = 'Lat')
 # add shape from Atlas of US Trees
-plot(ql, add = TRUE, col = alpha("white", 0.6), border = FALSE)
+plot(ql, add = TRUE, col = alpha("white", 0.6),
+     border = FALSE)
 # add Gbif presence points
-points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
-legend(x = -121, y = 40.5, "GBIF Data", pch = 19, col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
-legend(x = -121, y = 41.5, "Atlas of United States Trees \n (Little, E. 1971)", pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
+points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude,
+       cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
+legend(x = -121, y = 40.5, "GBIF Data", pch = 19,
+       col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
+legend(x = -121, y = 41.5,
+       legend = "Atlas of United States Trees \n (Little, E. 1971)",
+       pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
 ```

### 2017-09-06 manuscript.Rmd
added Ruby client and fixes use case problems, updated circleci badge in readme

--- 2016-08-05 manuscript.Rmd
@@
 linenumbers: true
-title: "R and Python clients for GBIF species occurrence data"
+title: "R Python, and Ruby clients for GBIF species occurrence data"
 author:
@@
 
-      Methods. We have developed a client to access GBIF data for each
-        of the R and Python programming languages: `rgbif` and `pygbif`.
+      Methods. We have developed clients to access GBIF data for each
+        of the R, Python, and Ruby programming languages: `rgbif`, `pygbif`,
+        `gbifrb`.
 
-      Results. For both clients, we describe their design and utility,
+      Results. For all clients we describe their design and utility,
         and demonstrate many use cases.
@@
 However, Global Biodiversity Information Facility (hereafter, GBIF,
-[http://www.gbif.org](http://www.gbif.org)) is the largest collection of
-biodiversity records globally, currently with 643 million records, 1.6 million
-taxa, 15,450 datasets from 780 publishers (as of 2016-02-09).
+<http://www.gbif.org>) is the largest collection of biodiversity records
+globally, currently with 820 million records, roughly 5.9 million
+taxa, 36,000 datasets from 1,300 publishers (as of 2016-02-09).
 Many large biodiversity warehouses such as iNaturalist
-([http://www.inaturalist.org](http://www.inaturalist.org)), VertNet
-([http://vertnet.org](http://vertnet.org)), and USGS's Biodiversity Information
-Serving Our Nation (BISON; [http://bison.usgs.ornl.gov](http://bison.usgs.ornl.gov))
+(<http://www.inaturalist.org>), VertNet (<http://vertnet.org>), and USGS's
+Biodiversity Information Serving Our Nation (BISON; <http://bison.usgs.ornl.gov>)
 all feed into GBIF.
 
-The most important data organizational level in GBIF is the individual record. In R
-using `rgbif` you'll recognize this as a row in the data.frame that is returned from
-`occ_search()` or `occ_data()`. Going upstream, each record is part of a dataset,
-where each dataset is submitted by an organization, organizations are organized into
+The most important organizational level in GBIF occurrence data is the occurrence
+record. The fields in a record vary, but include information about
+taxonomy (kingdom, phylum, genus, species names) and their identifiers,
+dataset metadata, and locality information including geospatial position.
+Going upstream, each record is part of a dataset, where each dataset is
+submitted by an organization, organizations are organized into
 nodes, datasets are published through institutions (which may be hosted at another
@@
 
-Each record has some taxonomic name associated with it, which itself is linked to
-a lot of other taxonomic data. GBIF maintains their own taxonomic data.
+Each occurrence record has some taxonomic name associated with it, which itself is
+linked to a lot of other taxonomic data - including a master taxonomic backbone
+that integrates taxonomies across many taxonomic authorities.
 
-This organization matters because you can navigate through GBIF data through
-these various organizational levels.
+The organization of GBIF matters because you can navigate GBIF data through
+these hierarchical organizational levels - it helps to be familiar with the
+terminology and how each group relates to another.
 
 # The clients
+
+Although we discuss libraries for R, Python, and Ruby here, we focus mostly on
+the R library `rgbif` as it has seen the most developer and user attention,
+and is the most mature.
 
@@
 
+## gbifrb
+
+`gbifrb` [@gbifrb] is a library for working with GBIF data in
+the Ruby programming environment. Like Python, Ruby is a general
+purpose programming language used widely in all sectors. Unlike
+Python, Ruby is not used extensively in scientific disciplines.
+However, a Ruby client for GBIF data can be an important tool
+given how widely Ruby is used for web and web service
+development.
+
+```{ruby eval=FALSE}
+gem install gbifrb
+```
+
+```{ruby eval=FALSE}
+require 'gbifrb'
+```
+
+The `gbifrb` library is less mature and complete than the R and Python libraries.
+It also uses an [MIT license][mit] to maximize use everywhere. `gbifrb` is developed
+publicly on GitHub at <https://github.com/sckott/gbifrb>, where
+development versions of the package can be installed, and bugs and feature
+requests reported. Stable versions of `gbifrb` can be installed from
+[Rubygems][gemgbif], the distribution network for Ruby libraries.
+
 ### Library interfaces
 
-`rgbif` and `pygbif` are designed following the [GBIF Application Programming Interface][gbifdocs],
+`rgbif`, `pygbif`, and `gbifrb` are designed following the [GBIF Application Programming Interface][gbifdocs],
 or API. The GBIF API has four major components: registry, taxonomic names,
 occurrences, and maps. We also include functions to interface with the OAI-PMH
-GBIF service; only dataset information is available vis this service, however. We
-ignore maps in both libraries as it is concerned with generating maps primarily for web
-applications. Both libraries have a suite of functions dealing with each of registry, taxonomic
-names, and occurrences - we'll go through each in turn describing design and example usage.
+GBIF service; only dataset (registry) information is available via this service, however.
+An interface to the GBIF maps API is in development for `rgbif`, but is non-existent for
+both `pygbif` and `gbifrb`. All three libraries have a suite of functions dealing with
+each of registry, taxonomic, names, and occurrences - we'll go through each in turn
+describing design of the user interface and example usage.
 
@@
 
-With each request `rgbif` and `pygbif` make to GBIF's API, we send request headers that
-tell GBIF what client the request is coming from, including what version of
+With each request `rgbif`, `pygbif`, `gbifrb` make to GBIF's API, we send request
+headers that tell GBIF what library the request is coming from, including what version of
 the library. This helps GBIF know what proportion of requests are coming from
-which client, and therefore from R vs. Python; this information is helpful for GBIF
-in thinking about how people are using GBIF data.
+which library, and therefore from R vs. Python vs. Ruby; this information is helpful
+for GBIF in thinking about how people are using GBIF data.
 
@@
 functions in `rgbif` prefixed with `gbif_oai_`, but not available in `pygbif`
-yet.
+or `gbifrb` yet.
 
@@
 search for datasets on the GBIF web portal. However, programmatic searching using
-this package is much more powerful. Identifying datasets appropriate for a
+any of these libraries is more powerful. Identifying datasets appropriate for a
 research question is helpful as you can get metadata for each dataset, and track
@@
 
-The `dataset_search()` function is one way to search for datasets. Here, we search
-for the query term "oregon", which finds any datasets that have terms matching
-that term.
+The `dataset_search()` function in `rgbif` is one way to search for datasets. Here,
+we search for the term "oregon", which finds any datasets that have words
+matching that term.
 
@@
 
+In Ruby, we can do the same. Here, search for datasets of type `OCCURRENCE`:
+
+```{ruby eval=FALSE}
+require 'gbifrb'
+registry = Gbif::Registry
+registry.datasets(type: "OCCURRENCE")
+```
+
 #### Dataset metrics
@@
 
-Here, we search for dataset metrics for a single dataset, with uuid
+Here, in R we search for dataset metrics for a single dataset, with uuid
 `ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
@@
 
+The same in Ruby:
+
+```{ruby eval=FALSE}
+require 'gbifrb'
+registry = Gbif::Registry
+registry.dataset_metrics(uuid: 'ec93a739-1681-4b04-b62f-3a687127a17f')
+```
+
 ### Networks, nodes, and installations
@@
 
+And in Ruby:
+
+```{ruby}
+require 'gbifrb'
+registry = Gbif::Registry
+registry.networks(limit: 10)
+```
+
 ## Taxonomic names
@@
 
-`pygbif` only has `name_backbone()` and `name_suggest()` at this time.
+`pygbif` and `gbifrb` have all the same functions, except the name parser goes by
+`name_parser()` in `pygbif` and `gbifrb`.
 
@@
 to taxonomic names throughout their services. The backbone has `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$distinctNamesCount` unique
-names and `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$countByRank$SPECIES` species names. The backbone taxonomy is also a dataset with key
+names and `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$countByRank$SPECIES`
+species names. The backbone taxonomy is also a dataset with key
 `d7dddbf4-2cf0-4f39-9b2a-bb099caae36c`
-(http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c).
+(<https://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c>).
 
-We can search the backbone taxonomy with the function `name_backbone()` in both
-R and Python clients. Here, we're searching for the name _Poa_, restricting to
+We can search the backbone taxonomy with the function `name_backbone()` in all
+thee clients. Here, we're searching for the name _Poa_, restricting to
 genera, and the family _Poaceae_, in R
@@
 
-and in Python
+And in Python
 
@@
 [ res[x] for x in ['usageKey', 'kingdom'] ]
+```
+
+And in Ruby
+
+```{ruby eval=FALSE}
+require 'gbifrb'
+species = Gbif::Species
+res = species.name_backbone(name: 'Poa', rank: 'genus', family: 'Poaceae')
+res.select { |k,v| k.match(/usageKey|kingdom/) }
 ```
@@
 
+And in Ruby
+
+```{ruby eval=FALSE}
+require 'gbifrb'
+species = Gbif::Species
+species.name_suggest(q: 'Pum', limit: 6)
+```
+
 With these results, you can then proceed to search for occurrences with the
@@
 GBIF provides two ways to get occurrence data: through the
-`/occurrence/search` route (see `occ_search` in `rgbif`, or
-`occurrences.search` in `pygbif`), or via the
+`/occurrence/search` route (see `occ_search` in `rgbif`,
+`occurrences.search` in `pygbif`, `Occurrences.search` in gbifrb), or via the
 `/occurrence/download` route (many functions, see below).
-`occ_search()`/`occurrences.search` is the main function for the search route,
-and is more appropriate for smaller data, while
-the download functions are more appropriate for larger
-data requests.
 
-GBIF imposes for any given search a limit of 200,000 records in the search
+`occ_search()`/`occurrences.search`/`Occurrences.search` are the main functions
+for the search route, and are more appropriate when you want less data, while
+the download functions are more appropriate for larger data requests.
+
+Small vs. large amounts of data of course is all relative. GBIF imposes for any
+given search a limit of 200,000 records in the search
 service, after which point you can't download any more records for that
@@
 even given the 200,000 limit. For those that need more data, we have
-created a similar interface in the download functions, that should
-be easy to use. Users should take note that using the download service
-has a few extra steps to get data into R, but is straight-forward.
+created a similar interface in the download functions that should
+be easy to use with minimal work. Users should take note that using the
+download service has a few extra steps to get data into R, but is
+straight-forward.
 
@@
 your user credentials. However, simply check when a download job is complete,
-then you should be able to start a new download request.
+then you can start a new download request. See "Queueing Download Requests"
+below for help automating many download requests in R.
 
@@
 
+The Ruby library `gbifrb` does not yet have occurrene download functionality.
+
 #### Downloaded data format
@@
 
+#### Queueing Download Requests
+
+FILL IN TEXT HERE ...
+
 ### Search API
@@
 
-* Get a single numeric count of occurrences - `rgbif`: `occ_count()` / `pygbif`: `occurrences.count`
-* Search for occurrences - `rgbif`: `occ_search()` / `pygbif`: `occurrences.count`
-* A simplified and optimized version of `rgbif`: `occ_search()` or `occ_data()` / `pygbif`: `occurrences.count`
-* Get occurrences by occurrence identifier - `rgbif`: `occ_get()` / `pygbif`: `occurrences.count`
-* Get occurrence metadata - `rgbif`: `occ_metadata()` / `pygbif`: `occurrences.count`
+* Get a single numeric count of occurrences - `rgbif`: `occ_count()` / `pygbif`: `occurrences.count` / `gbifrb`: `Occurrences.count`
+* Search for occurrences - `rgbif`: `occ_search()` / `pygbif`: `occurrences.search` / `gbifrb`: `Occurrences.search`
+* A simplified and optimized version of `rgbif`: `occ_search()` or `occ_data()` / none / none
+* Get occurrences by occurrence identifier - `rgbif`: `occ_get()` / `pygbif`: `occurrences.get` / `gbifrb`: `Occurrences.get`
+* Get occurrence metadata - `rgbif`: `occ_metadata()` / `pygbif`: various / `gbifrb`: various
 
 #### Search for occurrences
-
-__R__
 
@@
 
+__R__
+
 ```{r}
+library(rgbif)
 (key <- name_suggest(q = 'Helianthus annuus', rank = 'species')$key[1])
@@
 
+__Python__
+
+```{python eval=FALSE}
+from pygbif import species
+from pygbif import occurrences as occ
+key = species.name_suggest(q = 'Helianthus annuus', rank = 'species')['data'][0]['key']
+occ.search(taxonKey = key, limit = 2)
+```
+
+__Ruby__
+
+```{ruby eval=FALSE}
+require 'gbifrb'
+species = Gbif::Species
+occ = Gbif::Occurrences
+key = species.name_suggest(q: 'Helianthus annuus', rank: 'species')['data'][0]['key']
+occ.search(taxonKey: key, limit: 2)
+```
+
 Instead of getting a taxon key first, you can search for a name directly
+
+__R__
 
@@
 
+__Python__
+
+```{python eval=FALSE}
+occ.search(scientificName = 'Ursus americanus')
+```
+
+__Ruby__
+
+```{ruby eval=FALSE}
+occ.search(scientificName: 'Ursus americanus')
+```
+
 Search for many species
+
+__R__
 
@@
 occ_search(taxonKey = keys, limit = 5, return = 'data')
+```
+
+__Python__
+
+```{python eval=FALSE}
+from pygbif import species
+from pygbif import occurrences as occ
+splist = ['Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa']
+keys = [ species.name_suggest(x)['data'][0]['key'] for x in splist ]
+occ.search(taxonKey = keys, limit = 5)
+```
+
+__Ruby__
+
+```{ruby eval=FALSE}
+species = Gbif::Species
+occ = Gbif::Occurrences
+splist = ['Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa']
+keys = [ species.name_suggest(x)['data'][0]['key'] for x in splist ]
+occ.search(taxonKey: keys, limit: 5)
 ```
@@
 
+__R__
+
 ```{r eval=FALSE}
@@
 
+__Python__
+
+```{python eval=FALSE}
+from pygbif import occurrences as occ
+# well known text
+occ.search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
+# bounding box
+occ.search(geometry = '-125.0,38.4,-121.8,40.9', limit = 20)
+```
+
+__Ruby__
+
+```{ruby eval=FALSE}
+occ = Gbif::Occurrences
+# well known text
+occ.search(geometry: 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit: 20)
+# bounding box
+occ.search(geometry: '-125.0,38.4,-121.8,40.9', limit: 20)
+```
+
 Get only occurrences with lat/long data using the `hasCoordinate` parameter
+
+__R__
 
@@
 occ_search(hasCoordinate = TRUE, limit = 5)
+```
+
+__Python__
+
+```{python eval=FALSE}
+from pygbif import occurrences as occ
+occ.search(hasCoordinate = True, limit = 5)
+```
+
+__Ruby__
+
+```{ruby eval=FALSE}
+occ = Gbif::Occurrences
+occ.search(hasCoordinate: true, limit: 5)
 ```
@@
 
-The equivalent occurrence search workhorse in `pygibf` is `occurrences.search`.
-
-Search by species name, using `name_backbone()` first to get key
-
-```{python eval=FALSE}
-from pygbif import species
-from pygbif import occurrences as occ
-key = species.name_suggest(q = 'Helianthus annuus', rank = 'species')['data'][0]['key']
-occ.search(taxonKey = key, limit = 2)
-```
-
-Instead of getting a taxon key first, you can search for a name directly
-
-```{python eval=FALSE}
-occ.search(scientificName = 'Ursus americanus')
-```
-
-Search for many species
-
-```{python eval=FALSE}
-from pygbif import species
-from pygbif import occurrences as occ
-splist = ['Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa']
-keys = [ species.name_suggest(x)['data'][0]['key'] for x in splist ]
-occ.search(taxonKey = keys, limit = 5)
-```
-
-Spatial search, based on well known text format [@wkt], or a bounding box set of
-four coordinates
-
-```{python eval=FALSE}
-from pygbif import occurrences as occ
-# well known text
-occ.search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
-# bounding box
-occ.search(geometry = '-125.0,38.4,-121.8,40.9', limit = 20)
-```
-
-Get only occurrences with lat/long data
-
-```{python eval=FALSE}
-from pygbif import occurrences as occ
-occ.search(hasCoordinate = True, limit = 5)
-```
-
-Get only those occurrences with spatial issues. Spatial issues are a set of issues
-that are returned in the `issues` field. They each indicate something different
-about that record. For example, the issue `COUNTRY_COORDINATE_MISMATCH` indicates
-that the interpreted occurrence coordinates fall outside of the indicated country.
-You can see how that might be useful when it comes to cleaning your data prior
-to analysis/visualization.
-
 ```{python eval=FALSE}
@@
 occ.search(hasGeospatialIssue = True, limit = 5)
+```
+
+__Ruby__
+
+```{ruby eval=FALSE}
+occ = Gbif::Occurrences
+occ.search(hasGeospatialIssue: true, limit: 5)
 ```
@@
 into many different pre-defined classes, covering issues with taxonomic names,
-geographic data, and more (see `occ_issues_lookup()` to find out more information
+geographic data, and more (see `rgbif::occ_issues_lookup()` to find out more information
 on GBIF issues; and the same data on [GBIF's development site][gbifissues]).
 
-`occ_issues()` provides a way to easily filter data downloaded via `occ_search()`
+`rgbif::occ_issues()` provides a way to easily filter data downloaded via `rgbif::occ_search()`
 based on GBIF issues.
@@
 
-There's no equivalent interface in `pygbif` yet.
+There's no equivalent interface in `pygbif` or `gbifrb` yet.
 
@@
 
-There's no equivalent interface in `pygbif`.
+There's no equivalent interface in `pygbif` or `gbifrb`.
 
@@
 
-# R vs. Python
+# R vs. Python vs. Ruby
 
@@
 
-The following are three use cases for `rgbif`: niche modeling,
+The following are three use cases for the R library `rgbif`: niche modeling,
 spatial change in biodiversity, and distribution mapping.
@@
 
-The `rgbif` and `pygbif` libraries provide programmatic interfaces to GBIF's
-application programming interface (API) - a powerful tool for working with
-species occurrence data, and facilitating reproducible research. In fact,
-the `rgbif` package has already been used in 22 scholarly publications
-(as of 2015-11-14).
+The `rgbif`, `pygbif`, and `gbibrb` libraries provide programmatic interfaces
+to GBIF's application programming interface (API) - a powerful tool for
+working with species occurrence data, and facilitating reproducible research.
+In fact, the `rgbif` package has already been used in more than 20 scholarly
+publications (as of 2016-08-10), including [@Amano_2016, @Bartomeus_2013, @Barve_2014, @Bone_2015, @Collins_2015, @Drozd_2013, @Kong_2015, @Richardson_2015, @Turner_2015, @Verheijen_2015, @Zizka_2015, @Butterfield_2016, @Dellinger_2015, @Feitosa_2015, @Malhado_2015, @Werner_2015, @Robertson_2016, @Davison_2015, @Janssens_2016].
 
@@
 many breaking changes unless necessitated due to changes in the
-GBIF API.
+GBIF API. However, it will gain function(s) to work with the maps
+API in the near future.
 
-The `pygbif` library is in early development, and will
+The `pygbif` and `gbifrb` libraries are in early development, and will
 greatly benefit from any feedback and use cases.
@@
 very useful, but is actually quite a hard problem. There are many
-other problems like this.
+other problems like this, for which these three libraries will help
+in making more efficient and reproducible.
 
@@
 corresponds to a snapshot of the GitHub repository at
-<https://github.com/sckott/msrgbif>. Software can be found at
-<https://github.com/ropensci/rgbif> and <https://github.com/sckott/pygbif>,
-both under MIT licenses. We thank all the users that have used `rgbif`
-and `pygbif` and have given feedback and reported bugs. In addition, we
-greatly appreciate the all contributions to the two packages, found
-at <https://github.com/ropensci/rgbif/graphs/contributors> and
-<https://github.com/sckott/pygbif/graphs/contributors>.
+<https://github.com/sckott/gbifms>. Software can be found at
+<https://github.com/ropensci/rgbif>, <https://github.com/sckott/pygbif>,
+and <https://github.com/sckott/gibfrb>, all under MIT licenses. We thank
+all the users that have used `rgbif`, `pygbif`, and `gbifrb` and have
+given feedback and reported bugs. In addition, we greatly appreciate
+all the contributors to the three libraries, found at
+<https://github.com/ropensci/rgbif/graphs/contributors>,
+<https://github.com/sckott/pygbif/graphs/contributors>, and
+<https://github.com/sckott/gbifrb/graphs/contributors>.
 

### 2017-09-06 manuscript.tex
added Ruby client and fixes use case problems, updated circleci badge in readme

@@
  \let\@evenfoot\@oddfoot}
+ \def\tightlist{}
 \makeatother
@@
 \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
-\newcommand{\NormalTok}[1]{{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
 \usepackage{longtable}
@@
             pdfauthor={},
-            pdftitle={R and Python clients for GBIF species occurrence data},
+            pdftitle={R Python, and Ruby clients for GBIF species occurrence data},
             colorlinks=true,
@@
 
-  \title{R and Python clients for GBIF species occurrence data}
+  \title{R Python, and Ruby clients for GBIF species occurrence data}
     \author[cstar]{Scott Chamberlain\corref{c1}}
@@
   Email address:
-  \href{mailto:scott@ropensci.org}{\href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}}
+  \href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}
 
@@
 
-  Methods. We have developed a client to access GBIF data for each of the
-  R and Python programming languages: \texttt{rgbif} and \texttt{pygbif}.
-
-  Results. For both clients, we describe their design and utility, and
+  Methods. We have developed clients to access GBIF data for each of the
+  R, Python, and Ruby programming languages: \texttt{rgbif},
+  \texttt{pygbif}, \texttt{gbifrb}.
+
+  Results. For all clients we describe their design and utility, and
   demonstrate many use cases.
@@
 habitat suitability models with real occurrence data (Ficetola et al.,
-2014); ancestral range reconstruction (Ferretti et al., 2015; Mar{í}a
+2014); ancestral range reconstruction (Ferretti et al., 2015; María
 Mendoza et al., 2015); development of invasive species watch lists
@@
 records. However, Global Biodiversity Information Facility (hereafter,
-GBIF, \href{http://www.gbif.org}{\url{http://www.gbif.org}}) is the
-largest collection of biodiversity records globally, currently with 643
-million records, 1.6 million taxa, 15,450 datasets from 780 publishers
-(as of 2016-02-09). Many large biodiversity warehouses such as
-iNaturalist
-(\href{http://www.inaturalist.org}{\url{http://www.inaturalist.org}}),
-VertNet (\href{http://vertnet.org}{\url{http://vertnet.org}}), and
-USGS's Biodiversity Information Serving Our Nation (BISON;
-\href{http://bison.usgs.ornl.gov}{\url{http://bison.usgs.ornl.gov}}) all
-feed into GBIF.
-
-The most important data organizational level in GBIF is the individual
-record. In R using \texttt{rgbif} you'll recognize this as a row in the
-data.frame that is returned from \texttt{occ\_search()} or
-\texttt{occ\_data()}. Going upstream, each record is part of a dataset,
+GBIF, \url{http://www.gbif.org}) is the largest collection of
+biodiversity records globally, currently with 820 million records,
+roughly 5.9 million taxa, 36,000 datasets from 1,300 publishers (as of
+2016-02-09). Many large biodiversity warehouses such as iNaturalist
+(\url{http://www.inaturalist.org}), VertNet (\url{http://vertnet.org}),
+and USGS's Biodiversity Information Serving Our Nation (BISON;
+\url{http://bison.usgs.ornl.gov}) all feed into GBIF.
+
+The most important organizational level in GBIF occurrence data is the
+occurrence record. The fields in a record vary, but include information
+about taxonomy (kingdom, phylum, genus, species names) and their
+identifiers, dataset metadata, and locality information including
+geospatial position. Going upstream, each record is part of a dataset,
 where each dataset is submitted by an organization, organizations are
@@
 
-Each record has some taxonomic name associated with it, which itself is
-linked to a lot of other taxonomic data. GBIF maintains their own
-taxonomic data.
-
-This organization matters because you can navigate through GBIF data
-through these various organizational levels.
+Each occurrence record has some taxonomic name associated with it, which
+itself is linked to a lot of other taxonomic data - including a master
+taxonomic backbone that integrates taxonomies across many taxonomic
+authorities.
+
+The organization of GBIF matters because you can navigate GBIF data
+through these hierarchical organizational levels - it helps to be
+familiar with the terminology and how each group relates to another.
 
 \section{The clients}\label{the-clients}
+
+Although we discuss libraries for R, Python, and Ruby here, we focus
+mostly on the R library \texttt{rgbif} as it has seen the most developer
+and user attention, and is the most mature.
 
@@
 license} to maximize use everywhere. \texttt{rgbif} is developed
-publicly on GitHub at
-\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
-where development versions of the package can be installed, and bugs and
+publicly on GitHub at \url{https://github.com/ropensci/rgbif}, where
+development versions of the package can be installed, and bugs and
 feature requests reported. Stable versions of \texttt{rgbif} can be
@@
 \begin{Highlighting}[]
-\CharTok{import} \NormalTok{pygbif}
+\ImportTok{import}\NormalTok{ pygbif}
 \end{Highlighting}
@@
 
+\subsection{gbifrb}\label{gbifrb}
+
+\texttt{gbifrb} (Chamberlain) is a library for working with GBIF data in
+the Ruby programming environment. Like Python, Ruby is a general purpose
+programming language used widely in all sectors. Unlike Python, Ruby is
+not used extensively in scientific disciplines. However, a Ruby client
+for GBIF data can be an important tool given how widely Ruby is used for
+web and web service development.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{gem install gbifrb}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{gbifrb} library is less mature and complete than the R and
+Python libraries. It also uses an
+\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
+use everywhere. \texttt{gbifrb} is developed publicly on GitHub at
+\url{https://github.com/sckott/gbifrb}, where development versions of
+the package can be installed, and bugs and feature requests reported.
+Stable versions of \texttt{gbifrb} can be installed from
+{[}Rubygems{]}{[}gemgbif{]}, the distribution network for Ruby
+libraries.
+
 \subsubsection{Library interfaces}\label{library-interfaces}
 
-\texttt{rgbif} and \texttt{pygbif} are designed following the
-\href{http://www.gbif.org/developer/summary}{GBIF Application
-Programming Interface}, or API. The GBIF API has four major components:
-registry, taxonomic names, occurrences, and maps. We also include
-functions to interface with the OAI-PMH GBIF service; only dataset
-information is available vis this service, however. We ignore maps in
-both libraries as it is concerned with generating maps primarily for web
-applications. Both libraries have a suite of functions dealing with each
-of registry, taxonomic names, and occurrences - we'll go through each in
-turn describing design and example usage.
+\texttt{rgbif}, \texttt{pygbif}, and \texttt{gbifrb} are designed
+following the \href{http://www.gbif.org/developer/summary}{GBIF
+Application Programming Interface}, or API. The GBIF API has four major
+components: registry, taxonomic names, occurrences, and maps. We also
+include functions to interface with the OAI-PMH GBIF service; only
+dataset (registry) information is available via this service, however.
+An interface to the GBIF maps API is in development for \texttt{rgbif},
+but is non-existent for both \texttt{pygbif} and \texttt{gbifrb}. All
+three libraries have a suite of functions dealing with each of registry,
+taxonomic, names, and occurrences - we'll go through each in turn
+describing design of the user interface and example usage.
 
@@
 
-With each request \texttt{rgbif} and \texttt{pygbif} make to GBIF's API,
-we send request headers that tell GBIF what client the request is coming
-from, including what version of the library. This helps GBIF know what
-proportion of requests are coming from which client, and therefore from
-R vs.~Python; this information is helpful for GBIF in thinking about how
-people are using GBIF data.
+With each request \texttt{rgbif}, \texttt{pygbif}, \texttt{gbifrb} make
+to GBIF's API, we send request headers that tell GBIF what library the
+request is coming from, including what version of the library. This
+helps GBIF know what proportion of requests are coming from which
+library, and therefore from R vs.~Python vs.~Ruby; this information is
+helpful for GBIF in thinking about how people are using GBIF data.
 
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+\tightlist
 \item
@@
 functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}, but not
-available in \texttt{pygbif} yet.
+available in \texttt{pygbif} or \texttt{gbifrb} yet.
 
@@
 discovery process. One can search for datasets on the GBIF web portal.
-However, programmatic searching using this package is much more
+However, programmatic searching using any of these libraries is more
 powerful. Identifying datasets appropriate for a research question is
@@
 
-The \texttt{dataset\_search()} function is one way to search for
-datasets. Here, we search for the query term ``oregon'', which finds any
-datasets that have terms matching that term.
+The \texttt{dataset\_search()} function in \texttt{rgbif} is one way to
+search for datasets. Here, we search for the term ``oregon'', which
+finds any datasets that have words matching that term.
 
@@
 \NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"oregon"}\NormalTok{)}
-\NormalTok{res$data$datasetTitle[}\DecValTok{1}\NormalTok{:}\DecValTok{10}\NormalTok{]}
-\CommentTok{#>  [1] "UCDavis - Western USA - Monarch Butterflies - 1892-2005"                                                                         }
-\CommentTok{#>  [2] "A geographic distribution database of Mononychellus mites (Acari: Tetranychidae) on cassava (Manihot esculenta)"                 }
-\CommentTok{#>  [3] "Plantas Acuáticas de la Orinoquía Colombiana"                                                                                    }
-\CommentTok{#>  [4] "USBombus, contemporary survey data of North American bumble bees (Hymenoptera,  Apidae, Bombus) distributed in the United States"}
-\CommentTok{#>  [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"                                                               }
-\CommentTok{#>  [6] "CM Birds Collection"                                                                                                             }
-\CommentTok{#>  [7] "SDNHM Birds Collection"                                                                                                          }
-\CommentTok{#>  [8] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"                                                          }
-\CommentTok{#>  [9] "University of British Columbia Herbarium (UBC) - Vascular Plant Collection"                                                      }
-\CommentTok{#> [10] "Bryophyte Collection - University of Washington Herbarium (WTU)"}
+\NormalTok{res}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{datasetTitle[}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{]}
+\CommentTok{#>  [1] "Oregon State Ichthyology Collection"                                                                                                                                                  }
+\CommentTok{#>  [2] "Oregon State University Herpetological Collection"                                                                                                                                    }
+\CommentTok{#>  [3] "Mygalomorph spiders from southwestern Oregon, USA, with descriptions of four new species"                                                                                             }
+\CommentTok{#>  [4] "A new species of Helobdella (Hirudinida: Glossiphoniidae) from Oregon, USA"                                                                                                           }
+\CommentTok{#>  [5] "Annotated Checklist of the large branchiopod crustaceans of Idaho, Oregon and Washington, USA, with the “ rediscovery ” of a new species of Branchinecta (Anostraca: Branchinectidae)"}
+\CommentTok{#>  [6] "A new species of Chrysobothris Eschscholtz from Oregon and Washington, with notes on other Buprestidae (Coleoptera) occurring in the United States and Canada"                        }
+\CommentTok{#>  [7] "Three new species of Grylloblatta Walker (Insecta: Grylloblattodea: Grylloblattidae), from southern Oregon and northern California"                                                   }
+\CommentTok{#>  [8] "A new species of Cladotanytarsus (Lenziella) from Oregon supports the systematic concept of the subgenus (Diptera: Chironomidae)"                                                     }
+\CommentTok{#>  [9] "A new monster from southwest Oregon forests: Cryptomasterbehemoth sp.
+n.
+(Opiliones, Laniatores, Travunioidea)"                                                                       }
+\CommentTok{#> [10] "Two new species of Fluminicola (Caenogastropoda, Lithoglyphidae) from southwest Oregon, USA, and a range extension for F. multifarius"}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
-\NormalTok{registry.datasets(}\DataTypeTok{type}\NormalTok{=}\StringTok{"OCCURRENCE"}\NormalTok{)}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ registry}
+\NormalTok{registry.datasets(}\BuiltInTok{type}\OperatorTok{=}\StringTok{"OCCURRENCE"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+In Ruby, we can do the same. Here, search for datasets of type
+\texttt{OCCURRENCE}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{registry = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Registry}
+\NormalTok{registry.datasets(}\StringTok{type: "OCCURRENCE"}\NormalTok{)}
 \end{Highlighting}
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (16084).
-
-Here, we search for dataset metrics for a single dataset, with uuid
+lot of them (21502).
+
+Here, in R we search for dataset metrics for a single dataset, with uuid
 \texttt{ec93a739-1681-4b04-b62f-3a687127a17f}, a checklist of the ants
@@
 \NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
-\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
-           \DataTypeTok{count =} \KeywordTok{unname}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(res$countByRank)))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{longtable}[c]{@{}lr@{}}
+\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res}\OperatorTok{$}\NormalTok{countByRank),}
+           \DataTypeTok{count =} \KeywordTok{unname}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(res}\OperatorTok{$}\NormalTok{countByRank)))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}lr@{}}
 \toprule
@@
 \begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
-\NormalTok{registry.dataset_metrics(uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ registry}
+\NormalTok{registry.dataset_metrics(uuid}\OperatorTok{=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The same in Ruby:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{registry = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Registry}
+\NormalTok{registry.dataset_metrics(}\StringTok{uuid: 'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit =} \DecValTok{10}\NormalTok{)$data$title}
+\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit =} \DecValTok{10}\NormalTok{)}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{title}
 \CommentTok{#>  [1] "GBIF Backbone Sources"                                      }
@@
 \begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
-\NormalTok{registry.networks(limit = }\DecValTok{10}\NormalTok{)}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ registry}
+\NormalTok{registry.networks(limit }\OperatorTok{=} \DecValTok{10}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+And in Ruby:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{registry = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Registry}
+\NormalTok{registry.networks(}\StringTok{limit: }\DecValTok{10}\NormalTok{)}
 \end{Highlighting}
@@
 
-\texttt{pygbif} only has \texttt{name\_backbone()} and
-\texttt{name\_suggest()} at this time.
+\texttt{pygbif} and \texttt{gbifrb} have all the same functions, except
+the name parser goes by \texttt{name\_parser()} in \texttt{pygbif} and
+\texttt{gbifrb}.
 
@@
 refer to taxonomic names throughout their services. The backbone has
-5296599 unique names and 2525274 species names. The backbone taxonomy is
+5869207 unique names and 2818534 species names. The backbone taxonomy is
 also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
-(\url{http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).
+(\url{https://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).
 
 We can search the backbone taxonomy with the function
-\texttt{name\_backbone()} in both R and Python clients. Here, we're
-searching for the name \emph{Poa}, restricting to genera, and the family
+\texttt{name\_backbone()} in all thee clients. Here, we're searching for
+the name \emph{Poa}, restricting to genera, and the family
 \emph{Poaceae}, in R
@@
 
-and in Python
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
-\NormalTok{res = species.name_backbone(name=}\StringTok{'Poa'}\NormalTok{, rank=}\StringTok{'genus'}\NormalTok{, family=}\StringTok{'Poaceae'}\NormalTok{)}
-\NormalTok{[ res[x] }\KeywordTok{for} \NormalTok{x in [}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{] ]}
+And in Python
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ species}
+\NormalTok{res }\OperatorTok{=}\NormalTok{ species.name_backbone(name}\OperatorTok{=}\StringTok{'Poa'}\NormalTok{, rank}\OperatorTok{=}\StringTok{'genus'}\NormalTok{, family}\OperatorTok{=}\StringTok{'Poaceae'}\NormalTok{)}
+\NormalTok{[ res[x] }\ControlFlowTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ [}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{] ]}
+\end{Highlighting}
+\end{Shaded}
+
+And in Ruby
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
+\NormalTok{res = species.name_backbone(}\StringTok{name: 'Poa'}\NormalTok{, }\StringTok{rank: 'genus'}\NormalTok{, }\StringTok{family: 'Poaceae'}\NormalTok{)}
+\NormalTok{res.select \{ |k,v| k.match(}\OtherTok{/usageKey|kingdom/}\NormalTok{) \}}
 \end{Highlighting}
@@
 
-\begin{longtable}[c]{@{}rll@{}}
+\begin{longtable}[]{@{}rll@{}}
 \toprule
@@
 \endhead
-4848380 & Pumiliornis & GENUS\tabularnewline
-2235094 & Pumilibranchipus & GENUS\tabularnewline
-1795637 & Pumora & GENUS\tabularnewline
-4581745 & Pumicia & GENUS\tabularnewline
-4598790 & Pumilicopta & GENUS\tabularnewline
-1593095 & Pumilomyia & GENUS\tabularnewline
+2142856 & Althepus pum & SPECIES\tabularnewline
+8589398 & Pumiliopimoidae & FAMILY\tabularnewline
+8783253 & Pumililema & GENUS\tabularnewline
+4823360 & Pumiliopareia & GENUS\tabularnewline
+4635949 & Pumilina & GENUS\tabularnewline
+4648228 & Pumilopaurus & GENUS\tabularnewline
 \bottomrule
@@
 \begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
-\NormalTok{species.name_suggest(q=}\StringTok{'Pum'}\NormalTok{, limit = }\DecValTok{6}\NormalTok{)}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ species}
+\NormalTok{species.name_suggest(q}\OperatorTok{=}\StringTok{'Pum'}\NormalTok{, limit }\OperatorTok{=} \DecValTok{6}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+And in Ruby
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
+\NormalTok{species.name_suggest(}\StringTok{q: 'Pum'}\NormalTok{, }\StringTok{limit: }\DecValTok{6}\NormalTok{)}
 \end{Highlighting}
@@
 \texttt{/occurrence/search} route (see \texttt{occ\_search} in
-\texttt{rgbif}, or \texttt{occurrences.search} in \texttt{pygbif}), or
-via the \texttt{/occurrence/download} route (many functions, see below).
-\texttt{occ\_search()}/\texttt{occurrences.search} is the main function
-for the search route, and is more appropriate for smaller data, while
-the download functions are more appropriate for larger data requests.
-
-GBIF imposes for any given search a limit of 200,000 records in the
-search service, after which point you can't download any more records
-for that search. However, you can download more records for different
-searches.
+\texttt{rgbif}, \texttt{occurrences.search} in \texttt{pygbif},
+\texttt{Occurrences.search} in gbifrb), or via the
+\texttt{/occurrence/download} route (many functions, see below).
+
+\texttt{occ\_search()}/\texttt{occurrences.search}/\texttt{Occurrences.search}
+are the main functions for the search route, and are more appropriate
+when you want less data, while the download functions are more
+appropriate for larger data requests.
+
+Small vs.~large amounts of data of course is all relative. GBIF imposes
+for any given search a limit of 200,000 records in the search service,
+after which point you can't download any more records for that search.
+However, you can download more records for different searches.
 
@@
 given the 200,000 limit. For those that need more data, we have created
-a similar interface in the download functions, that should be easy to
-use. Users should take note that using the download service has a few
-extra steps to get data into R, but is straight-forward.
+a similar interface in the download functions that should be easy to use
+with minimal work. Users should take note that using the download
+service has a few extra steps to get data into R, but is
+straight-forward.
 
@@
 simultaneously for your user credentials. However, simply check when a
-download job is complete, then you should be able to start a new
-download request.
+download job is complete, then you can start a new download request. See
+``Queueing Download Requests'' below for help automating many download
+requests in R.
 
@@
 is different. For example, in the download API you can do greater than
-searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you cannot
-do that in the occurrence search API. Thus, unfortunately, we couldn't
-make the query interface exactly the same for both search and download
-functions.
+searches (i.e., \texttt{latitude\ \textgreater{}\ 50}), whereas you
+cannot do that in the occurrence search API. Thus, unfortunately, we
+couldn't make the query interface exactly the same for both search and
+download functions.
 
@@
 \NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_download_import}\NormalTok{(res)}
-\NormalTok{dat %>%}
+\NormalTok{dat }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
@@
 \begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
 \NormalTok{occ.download(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
-\NormalTok{(res = occ.download_get(}\StringTok{"0000840-150615163101818"}\NormalTok{, overwrite = }\OtherTok{True}\NormalTok{))}
+\NormalTok{(res }\OperatorTok{=}\NormalTok{ occ.download_get(}\StringTok{"0000840-150615163101818"}\NormalTok{, overwrite }\OperatorTok{=} \VariableTok{True}\NormalTok{))}
 \end{Highlighting}
@@
 
+The Ruby library \texttt{gbifrb} does not yet have occurrene download
+functionality.
+
 \paragraph{Downloaded data format}\label{downloaded-data-format}
@@
 internationally recognized biodiversity informatics standard
-(\href{http://rs.tdwg.org/dwc/}{\url{http://rs.tdwg.org/dwc/}}). The
-DwC-A downloaded is a compressed folder with a number of files,
-including metadata, citations for each of the datasets included in the
-download, and the data itself, in separate files for each dataset as
-well as one single \texttt{.txt} file. In
+(\url{http://rs.tdwg.org/dwc/}). The DwC-A downloaded is a compressed
+folder with a number of files, including metadata, citations for each of
+the datasets included in the download, and the data itself, in separate
+files for each dataset as well as one single \texttt{.txt} file. In
 \texttt{rgbif::occ\_download\_import()}, we simply fetch data from the
@@
 
+\paragraph{Queueing Download Requests}\label{queueing-download-requests}
+
+FILL IN TEXT HERE \ldots{}
+
 \subsubsection{Search API}\label{search-api}
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+\tightlist
 \item
   Get a single numeric count of occurrences - \texttt{rgbif}:
-  \texttt{occ\_count()} / \texttt{pygbif}: \texttt{occurrences.count}
+  \texttt{occ\_count()} / \texttt{pygbif}: \texttt{occurrences.count} /
+  \texttt{gbifrb}: \texttt{Occurrences.count}
 \item
   Search for occurrences - \texttt{rgbif}: \texttt{occ\_search()} /
-  \texttt{pygbif}: \texttt{occurrences.count}
+  \texttt{pygbif}: \texttt{occurrences.search} / \texttt{gbifrb}:
+  \texttt{Occurrences.search}
 \item
   A simplified and optimized version of \texttt{rgbif}:
-  \texttt{occ\_search()} or \texttt{occ\_data()} / \texttt{pygbif}:
-  \texttt{occurrences.count}
+  \texttt{occ\_search()} or \texttt{occ\_data()} / none / none
 \item
   Get occurrences by occurrence identifier - \texttt{rgbif}:
-  \texttt{occ\_get()} / \texttt{pygbif}: \texttt{occurrences.count}
+  \texttt{occ\_get()} / \texttt{pygbif}: \texttt{occurrences.get} /
+  \texttt{gbifrb}: \texttt{Occurrences.get}
 \item
   Get occurrence metadata - \texttt{rgbif}: \texttt{occ\_metadata()} /
-  \texttt{pygbif}: \texttt{occurrences.count}
+  \texttt{pygbif}: various / \texttt{gbifrb}: various
 \end{itemize}
@@
 \paragraph{Search for occurrences}\label{search-for-occurrences}
-
-\textbf{R}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{(key <-}\StringTok{ }\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q =} \StringTok{'Helianthus annuus'}\NormalTok{, }\DataTypeTok{rank =} \StringTok{'species'}\NormalTok{)$key[}\DecValTok{1}\NormalTok{])}
-\CommentTok{#> [1] 3119195}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
-\CommentTok{#> Records found [21538] }
+\textbf{R}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(rgbif)}
+\NormalTok{(key <-}\StringTok{ }\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q =} \StringTok{'Helianthus annuus'}\NormalTok{, }\DataTypeTok{rank =} \StringTok{'species'}\NormalTok{)}\OperatorTok{$}\NormalTok{key[}\DecValTok{1}\NormalTok{])}
+\CommentTok{#> [1] 9206251}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
+\CommentTok{#> Records found [17858] }
 \CommentTok{#> Records returned [2] }
 \CommentTok{#> No. unique hierarchies [1] }
-\CommentTok{#> No. media records [2] }
-\CommentTok{#> Args [taxonKey=3119195, limit=2, offset=0, fields=all] }
-\CommentTok{#> First 10 rows of data}
-\CommentTok{#> }
+\CommentTok{#> No. media records [1] }
+\CommentTok{#> No. facets [0] }
+\CommentTok{#> Args [limit=2, offset=0, taxonKey=9206251, fields=all] }
+\CommentTok{#> # A tibble: 2 x 75}
 \CommentTok{#>                name        key decimalLatitude decimalLongitude}
-\CommentTok{#> 1 Helianthus annuus 1249279611        34.04810       -117.79884}
-\CommentTok{#> 2 Helianthus annuus 1249286909        32.58747        -97.10081}
-\CommentTok{#> Variables not shown: issues (chr), datasetKey (chr), publishingOrgKey}
-\CommentTok{#>      (chr), publishingCountry (chr), protocol (chr), lastCrawled (chr),}
-\CommentTok{#>      lastParsed (chr), extensions (chr), basisOfRecord (chr), taxonKey}
-\CommentTok{#>      (int), kingdomKey (int), phylumKey (int), classKey (int), orderKey}
-\CommentTok{#>      (int), familyKey (int), genusKey (int), speciesKey (int),}
-\CommentTok{#>      scientificName (chr), kingdom (chr), phylum (chr), order (chr),}
-\CommentTok{#>      family (chr), genus (chr), species (chr), genericName (chr),}
-\CommentTok{#>      specificEpithet (chr), taxonRank (chr), dateIdentified (chr), year}
-\CommentTok{#>      (int), month (int), day (int), eventDate (chr), modified (chr),}
-\CommentTok{#>      lastInterpreted (chr), references (chr), identifiers (chr), facts}
-\CommentTok{#>      (chr), relations (chr), geodeticDatum (chr), class (chr), countryCode}
-\CommentTok{#>      (chr), country (chr), rightsHolder (chr), identifier (chr),}
-\CommentTok{#>      verbatimEventDate (chr), datasetName (chr), gbifID (chr),}
-\CommentTok{#>      verbatimLocality (chr), collectionCode (chr), occurrenceID (chr),}
-\CommentTok{#>      taxonID (chr), license (chr), recordedBy (chr), catalogNumber (chr),}
-\CommentTok{#>      http...unknown.org.occurrenceDetails (chr), institutionCode (chr),}
-\CommentTok{#>      rights (chr), eventTime (chr), identificationID (chr),}
-\CommentTok{#>      coordinateAccuracy (dbl), coordinateAccuracyInMeters (dbl),}
-\CommentTok{#>      occurrenceRemarks (chr)}
+\CommentTok{#>               <chr>      <int>           <dbl>            <dbl>}
+\CommentTok{#> 1 Helianthus annuus 1433793045        59.66859         16.54257}
+\CommentTok{#> 2 Helianthus annuus 1434024463        63.71622         20.31247}
+\CommentTok{#> # ... with 71 more variables: issues <chr>, datasetKey <chr>,}
+\CommentTok{#> #   publishingOrgKey <chr>, publishingCountry <chr>, protocol <chr>,}
+\CommentTok{#> #   lastCrawled <chr>, lastParsed <chr>, crawlId <int>, extensions <chr>,}
+\CommentTok{#> #   basisOfRecord <chr>, ...}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Python}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ species}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
+\NormalTok{key }\OperatorTok{=}\NormalTok{ species.name_suggest(q }\OperatorTok{=} \StringTok{'Helianthus annuus'}\NormalTok{, rank }\OperatorTok{=} \StringTok{'species'}\NormalTok{)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{]}
+\NormalTok{occ.search(taxonKey }\OperatorTok{=}\NormalTok{ key, limit }\OperatorTok{=} \DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Ruby}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
+\NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
+\NormalTok{key = species.name_suggest(}\StringTok{q: 'Helianthus annuus'}\NormalTok{, }\StringTok{rank: 'species'}\NormalTok{)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{]}
+\NormalTok{occ.search(}\StringTok{taxonKey: }\NormalTok{key, }\StringTok{limit: }\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 
+\textbf{R}
+
 \begin{Shaded}
@@
 
+\textbf{Python}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{occ.search(scientificName }\OperatorTok{=} \StringTok{'Ursus americanus'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Ruby}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{occ.search(}\StringTok{scientificName: 'Ursus americanus'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 Search for many species
 
+\textbf{R}
+
 \begin{Shaded}
@@
 \NormalTok{splist <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{)}
-\NormalTok{keys <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(splist, function(x) }\KeywordTok{name_suggest}\NormalTok{(x)$key[}\DecValTok{1}\NormalTok{], }\DataTypeTok{USE.NAMES =} \OtherTok{FALSE}\NormalTok{)}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{keys, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{)}
+\NormalTok{keys <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(splist, }\ControlFlowTok{function}\NormalTok{(x) }\KeywordTok{name_suggest}\NormalTok{(x)}\OperatorTok{$}\NormalTok{key[}\DecValTok{1}\NormalTok{], }\DataTypeTok{USE.NAMES =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keys, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Python}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ species}
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
+\NormalTok{splist }\OperatorTok{=}\NormalTok{ [}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{]}
+\NormalTok{keys }\OperatorTok{=}\NormalTok{ [ species.name_suggest(x)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{] }\ControlFlowTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ splist ]}
+\NormalTok{occ.search(taxonKey }\OperatorTok{=}\NormalTok{ keys, limit }\OperatorTok{=} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Ruby}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
+\NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
+\NormalTok{splist = [}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{]}
+\NormalTok{keys = [ species.name_suggest(x)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{] }\KeywordTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ splist ]}
+\NormalTok{occ.search(}\StringTok{taxonKey: }\NormalTok{keys, }\StringTok{limit: }\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 box format requires
-\texttt{longitude SW corner, latitude SW corner, longitude NE corner, latitude NE corner},
+\texttt{longitude\ SW\ corner,\ latitude\ SW\ corner,\ longitude\ NE\ corner,\ latitude\ NE\ corner},
 the well known text string requires an extra long/lat pair to close the
@@
 \NormalTok{wkt <-}\StringTok{ 'POLYGON((-122.6 39.9,-120.0 39.9,-120.0 37.9,-122.6 37.9,-122.6 39.9))'}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{wkt, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
 \CommentTok{# bounding box}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{122.6}\NormalTok{,}\FloatTok{37.9}\NormalTok{,-}\FloatTok{120.0}\NormalTok{,}\FloatTok{39.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{122.6}\NormalTok{,}\FloatTok{37.9}\NormalTok{,}\OperatorTok{-}\FloatTok{120.0}\NormalTok{,}\FloatTok{39.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Python}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
+\CommentTok{# well known text}
+\NormalTok{occ.search(geometry }\OperatorTok{=} \StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, limit }\OperatorTok{=} \DecValTok{20}\NormalTok{)}
+\CommentTok{# bounding box}
+\NormalTok{occ.search(geometry }\OperatorTok{=} \StringTok{'-125.0,38.4,-121.8,40.9'}\NormalTok{, limit }\OperatorTok{=} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Ruby}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
+\CommentTok{# well known text}
+\NormalTok{occ.search(}\StringTok{geometry: 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, }\StringTok{limit: }\DecValTok{20}\NormalTok{)}
+\CommentTok{# bounding box}
+\NormalTok{occ.search(}\StringTok{geometry: '-125.0,38.4,-121.8,40.9'}\NormalTok{, }\StringTok{limit: }\DecValTok{20}\NormalTok{)}
 \end{Highlighting}
@@
 \KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Python}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
+\NormalTok{occ.search(hasCoordinate }\OperatorTok{=} \VariableTok{True}\NormalTok{, limit }\OperatorTok{=} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Ruby}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
+\NormalTok{occ.search(}\StringTok{hasCoordinate: }\DecValTok{true}\NormalTok{, }\StringTok{limit: }\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 
-The equivalent occurrence search workhorse in \texttt{pygibf} is
-\texttt{occurrences.search}.
-
-Search by species name, using \texttt{name\_backbone()} first to get key
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
-\NormalTok{key = species.name_suggest(q = }\StringTok{'Helianthus annuus'}\NormalTok{, rank = }\StringTok{'species'}\NormalTok{)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{]}
-\NormalTok{occ.search(taxonKey = key, limit = }\DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Instead of getting a taxon key first, you can search for a name directly
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{occ.search(scientificName = }\StringTok{'Ursus americanus'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Search for many species
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
-\NormalTok{splist = [}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{]}
-\NormalTok{keys = [ species.name_suggest(x)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{] }\KeywordTok{for} \NormalTok{x in splist ]}
-\NormalTok{occ.search(taxonKey = keys, limit = }\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Spatial search, based on well known text format (Herring, 2011), or a
-bounding box set of four coordinates
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
-\CommentTok{# well known text}
-\NormalTok{occ.search(geometry = }\StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, limit = }\DecValTok{20}\NormalTok{)}
-\CommentTok{# bounding box}
-\NormalTok{occ.search(geometry = }\StringTok{'-125.0,38.4,-121.8,40.9'}\NormalTok{, limit = }\DecValTok{20}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Get only occurrences with lat/long data
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
-\NormalTok{occ.search(hasCoordinate = }\OtherTok{True}\NormalTok{, limit = }\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Get only those occurrences with spatial issues. Spatial issues are a set
-of issues that are returned in the \texttt{issues} field. They each
-indicate something different about that record. For example, the issue
-\texttt{COUNTRY\_COORDINATE\_MISMATCH} indicates that the interpreted
-occurrence coordinates fall outside of the indicated country. You can
-see how that might be useful when it comes to cleaning your data prior
-to analysis/visualization.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
-\NormalTok{occ.search(hasGeospatialIssue = }\OtherTok{True}\NormalTok{, limit = }\DecValTok{5}\NormalTok{)}
+\begin{Shaded}
+\begin{Highlighting}[]
+\ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
+\NormalTok{occ.search(hasGeospatialIssue }\OperatorTok{=} \VariableTok{True}\NormalTok{, limit }\OperatorTok{=} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbf{Ruby}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
+\NormalTok{occ.search(}\StringTok{hasGeospatialIssue: }\DecValTok{true}\NormalTok{, }\StringTok{limit: }\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 with taxonomic names, geographic data, and more (see
-\texttt{occ\_issues\_lookup()} to find out more information on GBIF
-issues; and the same data on
+\texttt{rgbif::occ\_issues\_lookup()} to find out more information on
+GBIF issues; and the same data on
 \href{http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html}{GBIF's
@@
 
-\texttt{occ\_issues()} provides a way to easily filter data downloaded
-via \texttt{occ\_search()} based on GBIF issues.
+\texttt{rgbif::occ\_issues()} provides a way to easily filter data
+downloaded via \texttt{rgbif::occ\_search()} based on GBIF issues.
 
@@
 \KeywordTok{NROW}\NormalTok{(out)}
-\CommentTok{#> [1] 4}
-\NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
-\CommentTok{#> [1] 4}
-\end{Highlighting}
-\end{Shaded}
-
-There's no equivalent interface in \texttt{pygbif} yet.
+\CommentTok{#> [1] 5}
+\NormalTok{out }\OperatorTok{%>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(}\OperatorTok{-}\NormalTok{cudc) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.}\OperatorTok{$}\NormalTok{data }\OperatorTok{%>%}\StringTok{ }\NormalTok{NROW}
+\CommentTok{#> [1] 0}
+\end{Highlighting}
+\end{Shaded}
+
+There's no equivalent interface in \texttt{pygbif} or \texttt{gbifrb}
+yet.
 
@@
 \begin{Highlighting}[]
-\NormalTok{key <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Puma concolor'}\NormalTok{)$speciesKey}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{100}\NormalTok{, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{)}
-\KeywordTok{gbifmap}\NormalTok{(dat$data)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{components/figure/manuscript-unnamed-chunk-34-1.pdf}
+\NormalTok{key <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Puma concolor'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ key, }\DataTypeTok{limit =} \DecValTok{100}\NormalTok{, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{)}
+\KeywordTok{gbifmap}\NormalTok{(dat}\OperatorTok{$}\NormalTok{data)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{components/figure/manuscript-unnamed-chunk-47-1.pdf}
 
@@
 
-There's no equivalent interface in \texttt{pygbif}.
+There's no equivalent interface in \texttt{pygbif} or \texttt{gbifrb}.
 
@@
 
-\section{R vs.~Python}\label{r-vs.python}
+\section{R vs.~Python vs.~Ruby}\label{r-vs.python-vs.ruby}
 
@@
 
-The following are three use cases for \texttt{rgbif}: niche modeling,
-spatial change in biodiversity, and distribution mapping.
+The following are three use cases for the R library \texttt{rgbif}:
+niche modeling, spatial change in biodiversity, and distribution
+mapping.
 
@@
 \NormalTok{nn <-}\StringTok{ }\KeywordTok{name_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
-\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn$data$nubKey))}
-\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
-\NormalTok{df_data <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{any}\NormalTok{(}\KeywordTok{class}\NormalTok{(x$data) %in%}\StringTok{ "tbl_df"}\NormalTok{)) ]}
-\NormalTok{df_data <-}\StringTok{ }\NormalTok{dplyr::}\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df_data, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
-\NormalTok{df2 <-}\StringTok{ }\NormalTok{df_data %>%}\StringTok{ }\NormalTok{dplyr::}\KeywordTok{select}\NormalTok{(decimalLongitude, decimalLatitude)}
+\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{nubKey))}
+\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
+\NormalTok{df_data <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, }\ControlFlowTok{function}\NormalTok{(x) }\KeywordTok{any}\NormalTok{(}\KeywordTok{class}\NormalTok{(x}\OperatorTok{$}\NormalTok{data) }\OperatorTok{%in%}\StringTok{ "tbl_df"}\NormalTok{)) ]}
+\NormalTok{df_data <-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df_data, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
+\NormalTok{df2 <-}\StringTok{ }\NormalTok{df_data }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(decimalLongitude, decimalLatitude)}
 \end{Highlighting}
@@
 \def\labelenumi{(\arabic{enumi})}
-\itemsep1pt\parskip0pt\parsep0pt
+\tightlist
 \item
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-40-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-53-1.pdf}
 
@@
 \KeywordTok{library}\NormalTok{(}\StringTok{"RColorBrewer"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"wicket"}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
-  \DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }
+  \DataTypeTok{text =}\NormalTok{ rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }
   \DataTypeTok{sep =} \StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{),}
@@
 \NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{city =} \NormalTok{dat$city, }\DataTypeTok{minlon =} \NormalTok{dat$minlon, }
-  \DataTypeTok{minlat =} \NormalTok{dat$minlat, }\DataTypeTok{maxlon =} \NormalTok{dat$maxlon, }
-  \DataTypeTok{maxlat =} \NormalTok{dat$maxlat,}
+  \DataTypeTok{city =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{city, }\DataTypeTok{minlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlon, }
+  \DataTypeTok{minlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlat, }\DataTypeTok{maxlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlon, }
+  \DataTypeTok{maxlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlat,}
   \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
@@
 search. We can however use the search API via \texttt{occ\_search()} and
-set \texttt{limit = 1} so that we
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{getdata <-}\StringTok{ }\NormalTok{function(x)\{}
-  \NormalTok{coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
-  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)$meta$count}
+set \texttt{limit\ =\ 1} so that we
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{getdata <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x)\{}
+\NormalTok{  coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
+\NormalTok{  wkt <-}\StringTok{ }\NormalTok{wicket}\OperatorTok{::}\KeywordTok{wkt_correct}\NormalTok{(wicket}\OperatorTok{::}\KeywordTok{bounding_wkt}\NormalTok{(}\DataTypeTok{values =}\NormalTok{ coords))}
+\NormalTok{  num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)}\OperatorTok{$}\NormalTok{meta}\OperatorTok{$}\NormalTok{count}
   \KeywordTok{data.frame}\NormalTok{(}
-    \DataTypeTok{city =} \NormalTok{x[}\StringTok{'city'}\NormalTok{], }
-    \DataTypeTok{richness =} \NormalTok{num, }
+    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{'city'}\NormalTok{], }
+    \DataTypeTok{richness =}\NormalTok{ num, }
     \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
-  \NormalTok{)}
+\NormalTok{  )}
 \NormalTok{\}}
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat +}\StringTok{ }\NormalTok{maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon +}\StringTok{ }\NormalTok{maxlon)/}\DecValTok{2}\NormalTok{)}
+\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =}\NormalTok{ (minlat }\OperatorTok{+}\StringTok{ }\NormalTok{maxlat)}\OperatorTok{/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =}\NormalTok{ (minlon }\OperatorTok{+}\StringTok{ }\NormalTok{maxlon)}\OperatorTok{/}\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{mapp <-}\StringTok{ }\KeywordTok{map_data}\NormalTok{(}\StringTok{'world'}\NormalTok{)}
-\KeywordTok{ggplot}\NormalTok{(mapp, }\KeywordTok{aes}\NormalTok{(long, lat)) +}
-\StringTok{  }\KeywordTok{geom_polygon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group=}\NormalTok{group), }\DataTypeTok{fill=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{alpha=}\DecValTok{0}\NormalTok{, }\DataTypeTok{color=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.4}\NormalTok{) +}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{out, }\KeywordTok{aes}\NormalTok{(lon, lat, }\DataTypeTok{color=}\NormalTok{richness), }\DataTypeTok{size=}\DecValTok{5}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) +}
-\StringTok{  }\KeywordTok{scale_color_continuous}\NormalTok{(}\DataTypeTok{low =} \StringTok{"#60E1EE"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#0404C8"}\NormalTok{) +}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{""}\NormalTok{, }\DataTypeTok{y=}\StringTok{""}\NormalTok{) +}
-\StringTok{  }\KeywordTok{theme_grey}\NormalTok{(}\DataTypeTok{base_size=}\DecValTok{14}\NormalTok{) +}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\DataTypeTok{legend.key =} \KeywordTok{element_blank}\NormalTok{()) +}
+\KeywordTok{ggplot}\NormalTok{(mapp, }\KeywordTok{aes}\NormalTok{(long, lat)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_polygon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group=}\NormalTok{group), }\DataTypeTok{fill=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{alpha=}\DecValTok{0}\NormalTok{, }\DataTypeTok{color=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.4}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{out, }\KeywordTok{aes}\NormalTok{(lon, lat, }\DataTypeTok{color=}\NormalTok{richness), }\DataTypeTok{size=}\DecValTok{5}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_color_continuous}\NormalTok{(}\DataTypeTok{low =} \StringTok{"#60E1EE"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#0404C8"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{""}\NormalTok{, }\DataTypeTok{y=}\StringTok{""}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme_grey}\NormalTok{(}\DataTypeTok{base_size=}\DecValTok{14}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\DataTypeTok{legend.key =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{color =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{keywidth =} \DecValTok{2}\NormalTok{))}
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-45-1.pdf}
+\includegraphics{components/figure/manuscript-unnamed-chunk-58-1.pdf}
 
@@
 \KeywordTok{library}\NormalTok{(}\StringTok{'sp'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'maptools'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{'sf'}\NormalTok{)}
 \KeywordTok{library}\NormalTok{(}\StringTok{'rgeos'}\NormalTok{)}
@@
 \begin{Highlighting}[]
-\NormalTok{keyQl <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Quercus lobata'}\NormalTok{, }\DataTypeTok{kingdom=}\StringTok{'plants'}\NormalTok{)$speciesKey}
+\NormalTok{keyQl <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Quercus lobata'}\NormalTok{, }\DataTypeTok{kingdom=}\StringTok{'plants'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
 \NormalTok{dat.Ql <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey=}\NormalTok{keyQl, }\DataTypeTok{return=}\StringTok{'data'}\NormalTok{, }\DataTypeTok{limit=}\DecValTok{50000}\NormalTok{)}
@@
 \begin{Highlighting}[]
-\NormalTok{url <-}\StringTok{ 'http://esp.cr.usgs.gov/data/little/querloba.zip'}
+\NormalTok{url <-}\StringTok{ 'https://github.com/sckott/gbifms/blob/gh-pages/querloba.zip?raw=true'}
 \NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
-\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"~/querloba.zip"}\NormalTok{)}
-\KeywordTok{unzip}\NormalTok{(}\StringTok{"~/querloba.zip"}\NormalTok{, }\DataTypeTok{exdir =} \NormalTok{tmp)}
-\NormalTok{ql <-}\StringTok{ }\KeywordTok{readShapePoly}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"querloba.shp"}\NormalTok{))}
+\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"querloba.zip"}\NormalTok{)}
+\KeywordTok{unzip}\NormalTok{(}\StringTok{"querloba.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
+\NormalTok{ql <-}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"querloba/querloba.shp"}\NormalTok{))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
-     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{),}
+\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\OperatorTok{:}\DecValTok{100}\OperatorTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
+     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{125}\NormalTok{, }\OperatorTok{-}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{),}
      \DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{,}
@@
 \CommentTok{# add Gbif presence points}
-\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude,}
-       \DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
-\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
-       \DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
-\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{,}
+\KeywordTok{points}\NormalTok{(dat.Ql}\OperatorTok{$}\NormalTok{decimalLongitude, dat.Ql}\OperatorTok{$}\NormalTok{decimalLatitude,}
+       \DataTypeTok{cex =}\NormalTok{ .}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \OperatorTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
+       \DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{)}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \OperatorTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{,}
        \DataTypeTok{legend =} \StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{,}
-       \DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{components/figure/manuscript-unnamed-chunk-52-1.pdf}
+       \DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{components/figure/manuscript-unnamed-chunk-65-1.pdf}
 
@@
 
-The \texttt{rgbif} and \texttt{pygbif} libraries provide programmatic
-interfaces to GBIF's application programming interface (API) - a
-powerful tool for working with species occurrence data, and facilitating
-reproducible research. In fact, the \texttt{rgbif} package has already
-been used in 22 scholarly publications (as of 2016-08-10) (Amano,
-Lamming \& Sutherland, 2016, Bartomeus et al. (2013), Barve (2014), Bone
-et al. (2015), Collins et al. (2015), Drozd \& {Š}ipo{š} (2013), Kong,
-Huang \& Duan (2015), Richardson, Roux \& Wilson (2015), Turner,
-Fr{é}ville \& Rieseberg (2015), Verheijen et al. (2015), Zizka \&
-Antonelli (2015), Butterfield et al. (2016), Dellinger et al. (2015),
-Feitosa et al. (2015), Malhado et al. (2015), Werner et al. (2015),
-Robertson, Visser \& Hui (2016), Davison et al. (2015), Janssens et al.
-(2016)).
+The \texttt{rgbif}, \texttt{pygbif}, and \texttt{gbibrb} libraries
+provide programmatic interfaces to GBIF's application programming
+interface (API) - a powerful tool for working with species occurrence
+data, and facilitating reproducible research. In fact, the
+\texttt{rgbif} package has already been used in more than 20 scholarly
+publications (as of 2016-08-10), including (Amano, Lamming \&
+Sutherland, 2016, Bartomeus et al. (2013), Barve (2014), Bone et al.
+(2015), Collins et al. (2015), Drozd \& Šipoš (2013), Kong, Huang \&
+Duan (2015), Richardson, Roux \& Wilson (2015), Turner, Fréville \&
+Rieseberg (2015), Verheijen et al. (2015), Zizka \& Antonelli (2015),
+Butterfield et al. (2016), Dellinger et al. (2015), Feitosa et al.
+(2015), Malhado et al. (2015), Werner et al. (2015), Robertson, Visser
+\& Hui (2016), Davison et al. (2015), Janssens et al. (2016)).
 
@@
 many breaking changes unless necessitated due to changes in the GBIF
-API.
-
-The \texttt{pygbif} library is in early development, and will greatly
-benefit from any feedback and use cases.
+API. However, it will gain function(s) to work with the maps API in the
+near future.
+
+The \texttt{pygbif} and \texttt{gbifrb} libraries are in early
+development, and will greatly benefit from any feedback and use cases.
 
@@
 these occurrences easy would be very useful, but is actually quite a
-hard problem. There are many other problems like this.
+hard problem. There are many other problems like this, for which these
+three libraries will help in making more efficient and reproducible.
 
@@
 corresponds to a snapshot of the GitHub repository at
-\url{https://github.com/sckott/msrgbif}. Software can be found at
-\url{https://github.com/ropensci/rgbif} and
-\url{https://github.com/sckott/pygbif}, both under MIT licenses. We
-thank all the users that have used \texttt{rgbif} and \texttt{pygbif}
-and have given feedback and reported bugs. In addition, we greatly
-appreciate the all contributions to the two packages, found at
-\url{https://github.com/ropensci/rgbif/graphs/contributors} and
-\url{https://github.com/sckott/pygbif/graphs/contributors}.
+\url{https://github.com/sckott/gbifms}. Software can be found at
+\url{https://github.com/ropensci/rgbif},
+\url{https://github.com/sckott/pygbif}, and
+\url{https://github.com/sckott/gibfrb}, all under MIT licenses. We thank
+all the users that have used \texttt{rgbif}, \texttt{pygbif}, and
+\texttt{gbifrb} and have given feedback and reported bugs. In addition,
+we greatly appreciate all the contributors to the three libraries, found
+at \url{https://github.com/ropensci/rgbif/graphs/contributors},
+\url{https://github.com/sckott/pygbif/graphs/contributors}, and
+\url{https://github.com/sckott/gbifrb/graphs/contributors}.
 
@@
 
+\hypertarget{refs}{}
+\hypertarget{ref-Amano_2016}{}
 Amano T., Lamming JDL., Sutherland WJ. 2016. Spatial gaps in global
@@
 
+\hypertarget{ref-Bartomeus_2013}{}
 Bartomeus I., Park MG., Gibbs J., Danforth BN., Lakso AN., Winfree R.
@@
 
+\hypertarget{ref-Barve_2014}{}
 Barve V. 2014. Discovering and developing primary biodiversity data from
@@
 
+\hypertarget{ref-Beck_2012}{}
 Beck J., Ballesteros-Mejia L., Buchmann CM., Dengler J., Fritz SA.,
@@
 
+\hypertarget{ref-Bone_2015}{}
 Bone RE., Smith JAC., Arrigo N., Buerki S. 2015. A macro-ecological
@@
 
+\hypertarget{ref-Brown_1995}{}
 Brown JH. 1995. \emph{Macroecology}. University of Chicago Press.
 
+\hypertarget{ref-Brown_2015}{}
 Brown KA., Parks KE., Bethell CA., Johnson SE., Mulligan M. 2015.
@@
 
+\hypertarget{ref-Butterfield_2016}{}
 Butterfield BJ., Copeland SM., Munson SM., Roybal CM., Wood TE. 2016.
@@
 
+\hypertarget{ref-Ceballos_2015}{}
 Ceballos G., Ehrlich PR., Barnosky AD., Garcia A., Pringle RM., Palmer
@@
 
+\hypertarget{ref-rgbif}{}
 Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{rgbif: An r interface
@@
 
+\hypertarget{ref-pygbif}{}
 Chamberlain S. \emph{pygbif: A python interface to the global
@@
 
+\hypertarget{ref-gbifrb}{}
+Chamberlain S. \emph{gbifrb: A ruby interface to the global biodiversity
+information facility API}.
+
+\hypertarget{ref-Collins_2015}{}
 Collins R., Ribeiro ED., Machado VN., Hrbek T., Farias I. 2015. A
@@
 
+\hypertarget{ref-Davison_2015}{}
 Davison J., Moora M., Opik M., Adholeya A., Ainsaar L., Ba A., Burla S.,
@@
 
+\hypertarget{ref-Dellinger_2015}{}
 Dellinger AS., Essl F., Hojsgaard D., Kirchheimer B., Klatt S., Dawson
-W., Pergl J., Py{š}ek P., Kleunen M van., Weber E., Winter M., Hörandl
-E., Dullinger S. 2015. Niche dynamics of alien species do not differ
-among sexual and apomictic flowering plants. \emph{New Phytologist}
+W., Pergl J., Pyšek P., Kleunen M van., Weber E., Winter M., Hörandl E.,
+Dullinger S. 2015. Niche dynamics of alien species do not differ among
+sexual and apomictic flowering plants. \emph{New Phytologist}
 209:1313--1323.
 
-Drozd P., {Š}ipo{š} J. 2013. R for all (i): Introduction to the new age
-of biological analyses. \emph{Casopis slezskeho zemskeho muzea (A)} 62.
-
+\hypertarget{ref-Drozd_2013}{}
+Drozd P., Šipoš J. 2013. R for all (i): Introduction to the new age of
+biological analyses. \emph{Casopis slezskeho zemskeho muzea (A)} 62.
+
+\hypertarget{ref-Faulkner_2014}{}
 Faulkner KT., Robertson MP., Rouget M., Wilson JR. 2014. A simple, rapid
@@
 
+\hypertarget{ref-Di_Febbraro_2013}{}
 Febbraro MD., Lurz PWW., Genovesi P., Maiorano L., Girardello M.,
@@
 
+\hypertarget{ref-Feitosa_2015}{}
 Feitosa YO., Absy ML., Latrubesse EM., Stevaux JC. 2015. Late quaternary
@@
 
-Ferretti F., Verd GM., Seret B., {Š}prem JS., Micheli F. 2015. Falling
+\hypertarget{ref-Ferretti_2015}{}
+Ferretti F., Verd GM., Seret B., Šprem JS., Micheli F. 2015. Falling
 through the cracks: The fading history of a large iconic predator.
@@
 
+\hypertarget{ref-Ficetola_2014}{}
 Ficetola GF., Rondinini C., Bonardi A., Baisero D., Padoa-Schioppa E.
@@
 
+\hypertarget{ref-wkt}{}
 Herring J. 2011. OpenGIS implementation standard for geographic
@@
 
+\hypertarget{ref-Janssens_2016}{}
 Janssens SB., Vandelook F., Langhe ED., Verstraete B., Smets E.,
@@
 
+\hypertarget{ref-Kong_2015}{}
 Kong X., Huang M., Duan R. 2015. SDMdata: A web-based software tool for
@@
 
+\hypertarget{ref-Malhado_2015}{}
 Malhado AC., Oliveira-Neto JA., Stropp J., Strona G., Dias LC., Pinto
@@
 
-Mar{í}a Mendoza., Ospina OE., C{á}rdenas-Henao H., Garc{í}a-R JC. 2015.
-A likelihood inference of historical biogeography in the world's most
+\hypertarget{ref-Mendoza_2015}{}
+María Mendoza., Ospina OE., Cárdenas-Henao H., García-R JC. 2015. A
+likelihood inference of historical biogeography in the world's most
 diverse terrestrial vertebrate genus: Diversification of
@@
 
+\hypertarget{ref-Pimm_2014}{}
 Pimm SL., Jenkins CN., Abell R., Brooks TM., Gittleman JL., Joppa LN.,
@@
 
+\hypertarget{ref-R}{}
 R Core Team. 2014. \emph{R: A language and environment for statistical
@@
 
+\hypertarget{ref-Richardson_2015}{}
 Richardson DM., Roux JJL., Wilson JR. 2015. Australian acacias as
@@
 
+\hypertarget{ref-Robertson_2016}{}
 Robertson MP., Visser V., Hui C. 2016. Biogeo: An r package for
@@
 
-Turner KG., Fr{é}ville H., Rieseberg LH. 2015. Adaptive plasticity and
+\hypertarget{ref-Turner_2015}{}
+Turner KG., Fréville H., Rieseberg LH. 2015. Adaptive plasticity and
 niche expansion in an invasive thistle. \emph{Ecol Evol} 5:3183--3197.
 
+\hypertarget{ref-Verheijen_2015}{}
 Verheijen LM., Aerts R., Bönisch G., Kattge J., Bodegom PMV. 2015.
@@
 
+\hypertarget{ref-Werner_2015}{}
 Werner GDA., Cornwell WK., Cornelissen JHC., Kiers ET. 2015.
@@
 
+\hypertarget{ref-Zizka_2015}{}
 Zizka A., Antonelli A. 2015. \emph{speciesgeocodeR: An r package for

### 2017-09-06 use-cases/city-diversity.Rmd
added Ruby client and fixes use case problems, updated circleci badge in readme

@@
   collapse = TRUE,
-  fig.width=8,
-  fig.height=6,
+  fig.width = 8,
+  fig.height = 6,
   cache.path = "cache/"
@@
 library("RColorBrewer")
+library("wicket")
 ```
@@
   coords <- as.numeric(x[c('minlon','minlat','maxlon','maxlat')])
-  num <- occ_search(geometry = coords, limit = 1)$meta$count
+  wkt <- wicket::wkt_correct(wicket::bounding_wkt(values = coords))
+  num <- occ_search(geometry = wkt, limit = 1)$meta$count
   data.frame(

### 2017-09-06 use-cases/oak-occurrences.Rmd
added Ruby client and fixes use case problems, updated circleci badge in readme

@@
 library('sp')
-library('maptools')
+library('sf')
 library('rgeos')
@@
 ```{r cache = TRUE}
-url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
+url <- 'https://github.com/sckott/gbifms/blob/gh-pages/querloba.zip?raw=true'
 tmp <- tempdir()
-download.file(url, destfile = "~/querloba.zip")
-unzip("~/querloba.zip", exdir = tmp)
-ql <- readShapePoly(file.path(tmp, "querloba.shp"))
+download.file(url, destfile = "querloba.zip")
+unzip("querloba.zip", exdir = tmp)
+ql <- sf::read_sf(file.path(tmp, "querloba/querloba.shp"))
 ```

### 2017-09-07 manuscript.Rmd
changes

@@
 tasks. Python, however, is a more general programming lanugae, and can be
-used in more contexts than R can be used in.
+used in more contexts than R can be used in. Ruby is used very little in
+science; but, like Python, Ruby is very widely used as a general purpose
+programming language, with heavy use in web development and web services.
 
-The `rgbif` and `pygbif` clients can do a lot of the same tasks. We
-envision `rgbif` as being more common in workflows of academics asking
-research questions, whereas `pygbif` can do that as well, but may be more
-easily used in a website.
+The three clients can do a lot of the same tasks. We envision `rgbif`
+being more common in workflows of academics asking research questions,
+whereas `pygbif` and `gbifrb` can do that as well, but may be more easily
+used in a website.
 
-The R client `rgbif` has much more development time than `pygbif`,
-but through time `pygbif` will become equally mature.
+The R client `rgbif` has had much more development time than `pygbif`
+and `gbifrb`, but with time `pygbif` and `gbifrb` will become equally
+mature.
 

### 2017-09-07 manuscript.Rmd
forget about python and ruby on circleci, eval=false for all python and ruby chunks
one more eval=false for a ruby chunk

@@
 
-```{python}
+```{python eval=FALSE}
 from pygbif import registry
@@
 
-```{ruby}
+```{ruby eval=FALSE}
 require 'gbifrb'

### 2017-09-07 manuscript.tex
remake pdf

@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (21502).
+lot of them (21522).
 
@@
 similar tasks. Python, however, is a more general programming lanugae,
-and can be used in more contexts than R can be used in.
-
-The \texttt{rgbif} and \texttt{pygbif} clients can do a lot of the same
-tasks. We envision \texttt{rgbif} as being more common in workflows of
-academics asking research questions, whereas \texttt{pygbif} can do that
-as well, but may be more easily used in a website.
-
-The R client \texttt{rgbif} has much more development time than
-\texttt{pygbif}, but through time \texttt{pygbif} will become equally
-mature.
+and can be used in more contexts than R can be used in. Ruby is used
+very little in science; but, like Python, Ruby is very widely used as a
+general purpose programming language, with heavy use in web development
+and web services.
+
+The three clients can do a lot of the same tasks. We envision
+\texttt{rgbif} being more common in workflows of academics asking
+research questions, whereas \texttt{pygbif} and \texttt{gbifrb} can do
+that as well, but may be more easily used in a website.
+
+The R client \texttt{rgbif} has had much more development time than
+\texttt{pygbif} and \texttt{gbifrb}, but with time \texttt{pygbif} and
+\texttt{gbifrb} will become equally mature.
 
@@
 \begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"sp"}\NormalTok{)}
 \KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}

### 2017-09-08 manuscript.Rmd
update manuscript.pdf, remove queue download section not going to discuss since not done

@@
         evolution. Data on individuals, including which species, and
-        where they're found, can be used for a large number of research
+        where they're found can be used for a large number of research
         questions. Global Biodiversity Information Facility (hereafter, GBIF)
         is the largest of these. Programmatic clients for GBIF would
-        make research dealing with GBIF data much easier.
+        make research dealing with GBIF data much easier and more reproducible.
 
@@
       Results. For all clients we describe their design and utility,
-        and demonstrate many use cases.
+        and demonstrate some use cases.
 
       Discussion. Programmatic acccess to GBIF will facilitate more open
-        and reproducible science. We hope that all scientists using GBIF
-        data will do so with one of these two clients.
+        and reproducible science - the three GBIF clients described herein
+        are a significant contribution towards this goal.
 
@@
 
-#### Queueing Download Requests
-
-FILL IN TEXT HERE ...
-
 ### Search API

### 2017-09-08 manuscript.tex
update manuscript.pdf, remove queue download section not going to discuss since not done

@@
   location forms the basis for many sub-fields of ecology and evolution.
-  Data on individuals, including which species, and where they're found,
+  Data on individuals, including which species, and where they're found
   can be used for a large number of research questions. Global
@@
   these. Programmatic clients for GBIF would make research dealing with
-  GBIF data much easier.
+  GBIF data much easier and more reproducible.
 
@@
   Results. For all clients we describe their design and utility, and
-  demonstrate many use cases.
+  demonstrate some use cases.
 
   Discussion. Programmatic acccess to GBIF will facilitate more open and
-  reproducible science. We hope that all scientists using GBIF data will
-  do so with one of these two clients.
+  reproducible science - the three GBIF clients described herein are a
+  significant contribution towards this goal.
   \end{abstract}
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (21522).
+lot of them (21534).
 
@@
 etc., it is easily accessible from the folder on your computer.
-
-\paragraph{Queueing Download Requests}\label{queueing-download-requests}
-
-FILL IN TEXT HERE \ldots{}
 

### 2017-09-23 use-cases/niche-modelling.Rmd
need possibly pkg sp in niche modelling use case
install pkg sp first in niche modelling use case
spelling

--- 2015-11-14 use-cases/niche-modelling.Rmd
@@
 In this example, we plot actual occurrence data for _Bradypus_ species against a single predictor variable, BIO1 (annual mean temperature).
-This is only ont step in a species distribution modelling nworkflow.
+This is only one step in a species distribution modelling workflow.
 
@@
 ```{r}
+library("sp")
 library("rgbif")
@@
 
-```{r cache=TRUE}
+```{r cache=FALSE}
 nn <- name_lookup("bradypus*", rank = "species")

### 2017-09-23 manuscript.Rmd

@@
 
-      Discussion. Programmatic acccess to GBIF will facilitate more open
+      Discussion. Programmatic access to GBIF will facilitate more open
         and reproducible science - the three GBIF clients described herein
@@
 
-In Python, we can similary search for datasets. Here, search for datasets of type
+In Python, we can similarly search for datasets. Here, search for datasets of type
 `OCCURRENCE`:
@@
 your user credentials. However, simply check when a download job is complete,
-then you can start a new download request. See "Queueing Download Requests"
+then you can start a new download request. See "Queuing Download Requests"
 below for help automating many download requests in R.
@@
 
-The Ruby library `gbifrb` does not yet have occurrene download functionality.
+The Ruby library `gbifrb` does not yet have occurrence download functionality.
 
@@
 Both R and Python are commonly used in science, and can be used for similar
-tasks. Python, however, is a more general programming lanugae, and can be
+tasks. Python, however, is a more general programming language, and can be
 used in more contexts than R can be used in. Ruby is used very little in

### 2017-09-26 manuscript.tex
hopefully fixed issue #8

@@
 
-  Discussion. Programmatic acccess to GBIF will facilitate more open and
+  Discussion. Programmatic access to GBIF will facilitate more open and
   reproducible science - the three GBIF clients described herein are a
@@
 
-In Python, we can similary search for datasets. Here, search for
+In Python, we can similarly search for datasets. Here, search for
 datasets of type \texttt{OCCURRENCE}:
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (21534).
+lot of them (21687).
 
@@
 download job is complete, then you can start a new download request. See
-``Queueing Download Requests'' below for help automating many download
+``Queuing Download Requests'' below for help automating many download
 requests in R.
@@
 
-The Ruby library \texttt{gbifrb} does not yet have occurrene download
+The Ruby library \texttt{gbifrb} does not yet have occurrence download
 functionality.
@@
 Both R and Python are commonly used in science, and can be used for
-similar tasks. Python, however, is a more general programming lanugae,
+similar tasks. Python, however, is a more general programming language,
 and can be used in more contexts than R can be used in. Ruby is used
@@
 species against a single predictor variable, BIO1 (annual mean
-temperature). This is only ont step in a species distribution modelling
-nworkflow.
+temperature). This is only one step in a species distribution modelling
+workflow.
 
@@
 \begin{Highlighting}[]
+\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ropensci/wicket"}\NormalTok{, }\DataTypeTok{force =} \OtherTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
@@
 \NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
-  \DataTypeTok{text =}\NormalTok{ rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }
+  \DataTypeTok{text =}\NormalTok{ rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{,}
   \DataTypeTok{sep =} \StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{),}
@@
 \NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
-  \DataTypeTok{city =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{city, }\DataTypeTok{minlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlon, }
-  \DataTypeTok{minlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlat, }\DataTypeTok{maxlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlon, }
+  \DataTypeTok{city =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{city, }\DataTypeTok{minlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlon,}
+  \DataTypeTok{minlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlat, }\DataTypeTok{maxlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlon,}
   \DataTypeTok{maxlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlat,}
@@
   \KeywordTok{data.frame}\NormalTok{(}
-    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{'city'}\NormalTok{], }
-    \DataTypeTok{richness =}\NormalTok{ num, }
+    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{'city'}\NormalTok{],}
+    \DataTypeTok{richness =}\NormalTok{ num,}
     \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
@@
 \KeywordTok{library}\NormalTok{(}\StringTok{'scales'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Get GBIF Data for Quercus lobata}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{keyQl <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Quercus lobata'}\NormalTok{, }\DataTypeTok{kingdom=}\StringTok{'plants'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
-\NormalTok{dat.Ql <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey=}\NormalTok{keyQl, }\DataTypeTok{return=}\StringTok{'data'}\NormalTok{, }\DataTypeTok{limit=}\DecValTok{50000}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)}
-
-From \url{http://esp.cr.usgs.gov/data/little}. And save shapefile in
-same directory
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{url <-}\StringTok{ 'https://github.com/sckott/gbifms/blob/gh-pages/querloba.zip?raw=true'}
+\KeywordTok{library}\NormalTok{(}\StringTok{'rnaturalearth'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get GBIF Data for Fraxinus excelsior}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{keyFe <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name =} \StringTok{'Fraxinus excelsior'}\NormalTok{, }\DataTypeTok{kingdom =} \StringTok{'plants'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
+\NormalTok{dat.Fe <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keyFe, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{, }\DataTypeTok{limit =}\NormalTok{ 10000L)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get Distribution map of F. excelsior European Forest Genetic
+Resources Programme}
+
+From \url{http://www.euforgen.org/species/fraxinus-excelsior/}. And save
+shapefile in same directory
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{url <-}\StringTok{ 'http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus_excelsior.zip'}
 \NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
-\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"querloba.zip"}\NormalTok{)}
-\KeywordTok{unzip}\NormalTok{(}\StringTok{"querloba.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
-\NormalTok{ql <-}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"querloba/querloba.shp"}\NormalTok{))}
+\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"fraxinus_excelsior.zip"}\NormalTok{)}
+\KeywordTok{unzip}\NormalTok{(}\StringTok{"fraxinus_excelsior.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
+\NormalTok{fe <-}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"Fraxinus_excelsior_EUFORGEN.shp"}\NormalTok{))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{alt.USA <-}\StringTok{ }\KeywordTok{getData}\NormalTok{(}\StringTok{'alt'}\NormalTok{, }\DataTypeTok{country =} \StringTok{'USA'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Create Hillshade of US}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{alt.USA <-}\StringTok{ }\NormalTok{alt.USA[[}\DecValTok{1}\NormalTok{]]}
-\NormalTok{slope.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'slope'}\NormalTok{)}
-\NormalTok{aspect.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'aspect'}\NormalTok{)}
-\NormalTok{hill.USA <-}\StringTok{ }\KeywordTok{hillShade}\NormalTok{(slope.USA, aspect.USA, }\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{direction =} \DecValTok{315}\NormalTok{)}
+\NormalTok{eur <-}\StringTok{ }\NormalTok{rnaturalearth}\OperatorTok{::}\KeywordTok{ne_countries}\NormalTok{(}\DataTypeTok{continent =} \StringTok{"europe"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"map_units"}\NormalTok{)}
+\NormalTok{eur1 <-}\StringTok{ }\NormalTok{eur[eur}\OperatorTok{$}\NormalTok{sovereignt }\OperatorTok{!=}\StringTok{ "Russia"}\NormalTok{, ]}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\OperatorTok{:}\DecValTok{100}\OperatorTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
-     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{125}\NormalTok{, }\OperatorTok{-}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{),}
-     \DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{,}
-     \DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
-\CommentTok{# add shape from Atlas of US Trees}
-\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{),}
-     \DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(eur1, }\DataTypeTok{col =} \StringTok{"darkgrey"}\NormalTok{, }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
+     \DataTypeTok{main =} \StringTok{'Distribution of Fraxinus excelsior'}\NormalTok{)}
+\CommentTok{# add distribution range layer}
+\KeywordTok{plot}\NormalTok{(fe, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
 \CommentTok{# add Gbif presence points}
-\KeywordTok{points}\NormalTok{(dat.Ql}\OperatorTok{$}\NormalTok{decimalLongitude, dat.Ql}\OperatorTok{$}\NormalTok{decimalLatitude,}
+\KeywordTok{points}\NormalTok{(dat.Fe}\OperatorTok{$}\NormalTok{decimalLongitude, dat.Fe}\OperatorTok{$}\NormalTok{decimalLatitude,}
        \DataTypeTok{cex =}\NormalTok{ .}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
-\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \OperatorTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
-       \DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{)}
-\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \OperatorTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{,}
-       \DataTypeTok{legend =} \StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{,}
-       \DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{components/figure/manuscript-unnamed-chunk-65-1.pdf}
+\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \DecValTok{38}\NormalTok{, }\DataTypeTok{y =} \DecValTok{81}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"GBIF Data"}\NormalTok{, }\StringTok{"Range Layer"}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{bg =} \StringTok{"grey"}\NormalTok{,}
+       \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{'darkgreen'}\NormalTok{, }\KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{)), }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{components/figure/manuscript-unnamed-chunk-64-1.pdf}
 

### 2017-09-26 use-cases/oak-occurrences.Rmd
hopefully fixed issue #8

--- 2015-11-14 use-cases/oak-occurrences.Rmd
@@
 library('sp')
-library('maptools')
+library('sf')
 library('rgeos')
 library('scales')
+library('rnaturalearth')
 ```
 
-_Get GBIF Data for Quercus lobata_
+_Get GBIF Data for Fraxinus excelsior_
 
 ```{r cache=TRUE}
-keyQl <- name_backbone(name='Quercus lobata', kingdom='plants')$speciesKey
-dat.Ql <- occ_search(taxonKey=keyQl, return='data', limit=50000)
+keyFe <- name_backbone(name = 'Fraxinus excelsior', kingdom = 'plants')$speciesKey
+dat.Fe <- occ_search(taxonKey = keyFe, return = 'data', limit = 10000L)
 ```
 
-_Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)_
+_Get Distribution map of F. excelsior European Forest Genetic Resources Programme_
 
-From [http://esp.cr.usgs.gov/data/little/](http://esp.cr.usgs.gov/data/little/). And save shapefile in same directory
+From <http://www.euforgen.org/species/fraxinus-excelsior/>. And save shapefile in same directory
 
-```{r}
-url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
+```{r cache = TRUE}
+url <- 'http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus_excelsior.zip'
 tmp <- tempdir()
-download.file(url, destfile = "~/querloba.zip")
-unzip("~/querloba.zip", exdir = tmp)
-ql <- readShapePoly(file.path(tmp, "querloba.shp"))
+download.file(url, destfile = "fraxinus_excelsior.zip")
+unzip("fraxinus_excelsior.zip", exdir = tmp)
+fe <- sf::read_sf(file.path(tmp, "Fraxinus_excelsior_EUFORGEN.shp"))
 ```
@@
 
-```{r}
-alt.USA <- getData('alt', country = 'USA')
-```
-
-_Create Hillshade of US_
-
-```{r}
-alt.USA <- alt.USA[[1]]
-slope.USA <- terrain(alt.USA, opt = 'slope')
-aspect.USA <- terrain(alt.USA, opt = 'aspect')
-hill.USA <- hillShade(slope.USA, aspect.USA, angle = 45, direction = 315)
+```{r cache = TRUE}
+eur <- rnaturalearth::ne_countries(continent = "europe", type = "map_units")
+eur1 <- eur[eur$sovereignt != "Russia", ]
 ```
@@
 ```{r fig.keep='last'}
-plot(hill.USA, col = grey(0:100/100), legend = FALSE, xlim = c(-125, -116), ylim = c(32, 42), main = 'Distribution of Quercus lobata', xlab = "Long", ylab = 'Lat')
-# add shape from Atlas of US Trees
-plot(ql, add = TRUE, col = alpha("white", 0.6), border = FALSE)
+plot(eur1, col = "darkgrey", legend = FALSE,
+     main = 'Distribution of Fraxinus excelsior')
+# add distribution range layer
+plot(fe, add = TRUE, col = alpha("white", 0.5), border = FALSE)
 # add Gbif presence points
-points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
-legend(x = -121, y = 40.5, "GBIF Data", pch = 19, col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
-legend(x = -121, y = 41.5, "Atlas of United States Trees \n (Little, E. 1971)", pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
+points(dat.Fe$decimalLongitude, dat.Fe$decimalLatitude,
+       cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
+legend(x = 38, y = 81, c("GBIF Data", "Range Layer"), pch = 19, bg = "grey",
+       col = c('darkgreen', alpha("white", 0.5)), pt.cex = 1, cex = .8)
 ```

### 2017-09-26 git_deploy.sh
attempt fix deploy script #9

@@ -2,7 +2,7 @@
 set -e
 
 {
-  git pull
+  git fetch --unshallow
 } > /dev/null 2>&1
 
 git config --global user.email 'myrmecocystus@gmail.com' && git config --global user.name 'Scott Chamberlain'
@@ -10,5 +10,5 @@ echo $(date) > date.txt
 git add date.txt
 git commit -m 'built and deployed from circle [ci skip]'
 {
-  git push https://sckott:$GH_TOKEN@github.com/sckott/msrgbif gh-pages
+  git push https://sckott:$GH_TOKEN@github.com/sckott/gbifms gh-pages
 } &> /dev/null 2>&1

### 2017-09-26 manuscript.Rmd
update ms with zenodo DOI, update readme with zenodo badge

@@
 All scripts and data used in this paper can be found in the permanent
-data archive Zenodo under the digital object identifier (DOI). This DOI
+data archive Zenodo under the digital object identifier
+(<https://doi.org/10.5281/zenodo.997547>). This DOI
 corresponds to a snapshot of the GitHub repository at
-<https://github.com/sckott/gbifms>. Software can be found at
-<https://github.com/ropensci/rgbif>, <https://github.com/sckott/pygbif>,
+<https://github.com/sckott/gbifms> that matches this preprint. Software
+can be found at <https://github.com/ropensci/rgbif>, <https://github.com/sckott/pygbif>,
 and <https://github.com/sckott/gibfrb>, all under MIT licenses. We thank

### 2017-09-26 manuscript.tex
update ms with zenodo DOI, update readme with zenodo badge

@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (21687).
+lot of them (21697).
 
@@
 \begin{Highlighting}[]
-\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ropensci/wicket"}\NormalTok{, }\DataTypeTok{force =} \OtherTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
 \KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
@@
 
-This example comes from \href{https://twitter.com/ajpelu}{Antonio J.
-Perez-Luque} who
+This example is inspired by a tweet from
+\href{https://twitter.com/ajpelu}{Antonio J. Perez-Luque} who
 \href{https://twitter.com/ajpelu/status/473951167567757312}{shared his
@@
 \href{http://esp.cr.usgs.gov/data/little/}{Atlas of US Trees}.
+
+The data in question from the example above is no longer available, so
+below we use a different species.
 
@@
 All scripts and data used in this paper can be found in the permanent
-data archive Zenodo under the digital object identifier (DOI). This DOI
-corresponds to a snapshot of the GitHub repository at
-\url{https://github.com/sckott/gbifms}. Software can be found at
-\url{https://github.com/ropensci/rgbif},
+data archive Zenodo under the digital object identifier
+(\url{https://doi.org/10.5281/zenodo.997547}). This DOI corresponds to a
+snapshot of the GitHub repository at
+\url{https://github.com/sckott/gbifms} that matches this preprint.
+Software can be found at \url{https://github.com/ropensci/rgbif},
 \url{https://github.com/sckott/pygbif}, and

### 2017-09-26 manuscript.Rmd
update manuscript with correct doi

@@
 data archive Zenodo under the digital object identifier
-(<https://doi.org/10.5281/zenodo.997547>). This DOI
+(<https://doi.org/10.5281/zenodo.997554>). This DOI
 corresponds to a snapshot of the GitHub repository at

### 2017-09-26 manuscript.tex
update manuscript with correct doi

@@
 data archive Zenodo under the digital object identifier
-(\url{https://doi.org/10.5281/zenodo.997547}). This DOI corresponds to a
+(\url{https://doi.org/10.5281/zenodo.997554}). This DOI corresponds to a
 snapshot of the GitHub repository at

### 2017-10-04 manuscript.Rmd
comma between R and Python in title needed

@@
 linenumbers: true
-title: "R Python, and Ruby clients for GBIF species occurrence data"
+title: "R, Python, and Ruby clients for GBIF species occurrence data"
 author:

### 2020-02-25 use-cases/city-diversity.Rmd
attempt wicket fix
install stuff
tweaks to use cases

@@
 ```{r include=FALSE}
-opts_chunk$set(
+knitr::opts_chunk$set(
   warning = FALSE,

### 2020-02-25 use-cases/oak-occurrences.Rmd
install rnaturalearth
install stuff
tweaks to use cases

@@
 
-This example comes from [Antonio J.
+This example is inspired by a tweet from [Antonio J.
 Perez-Luque](https://twitter.com/ajpelu) who [shared his plot on Twitter][tweet].
 Antonio compared the occurrences of Valley Oak (_Quercus lobata_) from [GBIF][gbif] to the distribution of the same species from the [Atlas of US Trees][atlas].
+
+The data in question from the example above is no longer available, so below
+we use a different species.
 

### 2020-02-27 manuscript_mee.Rmd
tweaks to mee paper

--- 2017-09-26 manuscript.Rmd
@@
 ---
-layout: review, 11pt
-linenumbers: true
-title: "R Python, and Ruby clients for GBIF species occurrence data"
+preamble: |
+  \journal{Methods in Ecology and Evolution}
+  \linenumbers
+  \usepackage{setspace}
+  \doublespacing
+bibliography: components/references.bib
+csl: components/methods-in-ecology-and-evolution.csl
+output:
+  rticles::elsevier_article:
+    keep_tex: true
+    latex_engine: xelatex
+layout:  3p # review = doublespace, 3p = singlespace, 5p = two-column
+title: "R, Python, and Ruby clients for GBIF species occurrence data"
 author:
@@
     affiliation: cstar
-    email: scott(at)ropensci.org
+    email: myrmecocystus(at)gmail.com
     footnote: Corresponding author
@@
     affiliation: boettig
-    email: carl(at)ropensci.org
+    email: cboettig(at)berkeley.edu
 address:
@@
 
-      Email address: [scott@ropensci.org](mailto:scott@ropensci.org)
+      Email address: [myrmecocystus@gmail.com](mailto:myrmecocystus@gmail.com)
 
@@
         are a significant contribution towards this goal.
-
-bibliography: components/references.bib
-csl: components/peerj.csl
-documentclass: components/elsarticle
-
-output:
-  pdf_document:
-    template: components/elsarticle.latex
-    keep_tex: true
-    fig_caption: true
 ---
@@
 distribution network for R packages. `rgbif` is part of the rOpenSci project
-(http://ropensci.org), a developer network making R software to facilitate
+(https://ropensci.org), a developer network making R software to facilitate
 reproducible research.
@@
 
-### Library interfaces
+# Library interfaces
 
@@
 
-### Datasets
+## Datasets
 
@@
 
-#### Dataset metrics
+## Dataset metrics
 
@@
 
-### Networks, nodes, and installations
+## Networks, nodes, and installations
 
@@
 
-### GBIF Backbone
+## GBIF Backbone
 
@@
 
-### Name searching
+## Name searching
 
@@
 
-### Download API
+## Occurrences Download API
 
@@
 
-#### Downloaded data format
+### Downloaded data format
 
@@
 
-### Search API
+## Search API
 
@@
 
-#### Search for occurrences
+### Search for occurrences
 
@@
 
-#### Data cleaning
+# Data cleaning
 
@@
 
-## Mapping
+# Mapping
 
 An obvious downstream use case for species occurrence data is to map the
-data. `rgbif` per se is largely not concerned with making this easier,
-although we do have a simple wrapper around `ggplot2` to make it easy to
-get a quick plot of occurrence data. For example, here we plot 100
+data. `rgbif` has enough to do without handling mapping. Another package, [mapr][],
+works with `rgbif` to make this easy. For example, here we plot 100
 occurrences for _Puma concolor_.
@@
 ```{r fig.width=8, fig.height=4}
+library(mapr)
 key <- name_backbone(name='Puma concolor')$speciesKey
 dat <- occ_search(taxonKey = key, limit = 100, hasCoordinate = TRUE)
-gbifmap(dat$data)
+map_ggplot(dat, name = "species")
 ```
 
-Another package, [mapr][mapr], is the perfect mapping companion to `rgbif`.
-It has convenient functions for handling input data from `rgbif`, `spocc`,
+`mapr` has convenient functions for handling input data from `rgbif`, `spocc`,
 or arbitrary data.frame's, and output plots for base plots, `ggplot2`,
@@
 
-## GBIF data in other R packages
+# Citing GBIF data
+
+All the data within GBIF was painstakingly collected by people and institutions
+across the globe. They all, including GBIF, deserve to be cited appropropriately
+if you use any GBIF data in your research. There are clear forces making this
+difficult however. Publishers often limit, without reason if the publication is
+online only, the number of citations one can include a manuscript. Even when 
+this happens, you can always include additional citations in an appendix.
+
+To make citing data from GBIF easier `rgbif` has a function `gbif_citation()`,
+which accepts the output of many different `rgbif` functions and gives
+citations for each dataset represented in the data therein. For example,
+suppose you searched for occurrences for `taxonKey` 9206251 (_Helianthus annuus_):
+
+```{r}
+res <- occ_search(taxonKey = 9206251, limit = 20)
+```
+
+We retrieved `r NROW(res$data)` occurrences from `r length(unique(res$data$datasetKey))`
+different datasets.
+
+To get citations for each of the four datasets, one can pass the output of `occ_search()`,
+here the object `res`, to the `gbif_citation()` function:
+
+```{r}
+cites <- gbif_citation(res)
+cites
+```
+
+And we get four citations, one of each of the datasets behind the `r NROW(res$data)`
+occurrence records retrieved.
+
+We don't yet have the same functionality available in `pygbif`, but it will be
+coming soon.
+
+## Downloads vs. searches
+
+The distinction between "downloads" and "searches" is not especially meaningful with
+respect to the science done with GBIF data. However, downloads and searches
+are different with respect to easily being able to cite GBIF data, which 
+matters for GBIF and for data providers.
+
+Downloads refers to the `rgbif` functions that start with `occ_download`, and the
+`pygbif` methods that start with `occurrences.download`. These methods use the GBIF
+"downloads API", which you can interact with from `rgbif` and `pygbif` programatically,
+similar to how you would interact with the GBIF website - each interaction creating
+a download request, which processes in the background. The user has to ask if 
+the request is ready to download. Downloads can request a very large amount of data.
+
+Searches are faster and more dynamic. Search functions in `rgbif` include `occ_search`,
+`occ_data`, and `occ_get`, while `pygbif ` methods include `occurrences.search`, 
+`occurrences.get`. These methods can only request up to 200,000 occurrence records.
+
+The big difference between downloads and searches with respect to citations is that
+each download generates a Digital Object Identifier (or DOI). This DOI can be used
+cite the entire set of data, that can inlude many datasets. There's no good way to
+refer to the entire results of searches, other than the user manually creating a 
+dataset somewhere public and linking to it from Zenodo to get a DOI. However,
+simply using downloads is simpler in terms of getting a single citeable DOI for
+the data used in a research paper/project.
+
+In `rgbif`, the output of `occ_download_meta()` can be passed to `gbif_citation()`
+to get the citation for the entire download.
+
+```{r eval=FALSE}
+key <- "0000122-171020152545675"
+res <- occ_download_meta(key)
+gbif_citation(res)
+#> $download
+#> [1] "GBIF Occurrence Download https://doi.org/10.15468/dl.yghxj7 Accessed from R via rgbif (https://github.com/ropensci/rgbif) on 2017-10-20"
+#> 
+#> $datasets
+#> NULL
+```
+
+If you however download the data first, then `gbif_citation()` can get citations 
+for all the datasets contained within the download.
+
+```{r eval=FALSE}
+d1 <- occ_download_get(key, overwrite = TRUE)
+gbif_citation(d1)
+#> $download
+#> [1] "GBIF Occurrence Download https://doi.org/10.15468/dl.yghxj7 Accessed from R via rgbif (https://github.com/ropensci/rgbif) on 2017-10-20"
+#> 
+#> $datasets
+#> $datasets[[1]]
+#> <<rgbif citation>>
+#>    Citation: Grant S, Jones J (2017). Field Museum of Natural History (Zoology)
+#>         Invertebrate Collection. Version 18.6. Field Museum. Occurrence Dataset
+#>         https://doi.org/10.15468/6q5vuc accessed via GBIF.org on 2017-10-20..
+#>         Accessed from R via rgbif (https://github.com/ropensci/rgbif) on
+#>         2020-02-24
+#>    Rights: To the extent possible under law, the publisher has waived all
+#>         rights to these data and has dedicated them to the Public Domain (CC0
+#>         1.0). Users may copy, modify, distribute and use the work, including
+#>         for commercial purposes, without restriction.
+#> 
+#> $datasets[[2]]
+#> <<rgbif citation>>
+#>    Citation: Creuwels J (2017). Naturalis Biodiversity Center (NL) - Crustacea.
+#>         Naturalis Biodiversity Center. Occurrence Dataset
+#>         https://doi.org/10.15468/vjoltu accessed via GBIF.org on 2017-10-20..
+#>         Accessed from R via rgbif (https://github.com/ropensci/rgbif) on
+#>         2020-02-24
+#>    Rights: To the extent possible under law, the publisher has waived all
+#>         rights to these data and has dedicated them to the Public Domain (CC0
+#>         1.0). Users may copy, modify, distribute and use the work, including
+#>         for commercial purposes, without restriction.
+```
+
+# GBIF data in other R packages
 
@@
 
-### taxize
+## taxize
 
@@
 
-### spocc
+## spocc
 
@@
 
-### Ecological niche modeling
+## Ecological niche modeling
 
@@
 
-### Biodiversity in big cities
+## Biodiversity in big cities
 
@@
 
-### Valley oak occurrence data comparison
+## Valley oak occurrence data comparison
 
@@
 In fact, the `rgbif` package has already been used in more than 20 scholarly
-publications (as of 2016-08-10), including [@Amano_2016, @Bartomeus_2013, @Barve_2014, @Bone_2015, @Collins_2015, @Drozd_2013, @Kong_2015, @Richardson_2015, @Turner_2015, @Verheijen_2015, @Zizka_2015, @Butterfield_2016, @Dellinger_2015, @Feitosa_2015, @Malhado_2015, @Werner_2015, @Robertson_2016, @Davison_2015, @Janssens_2016].
+publications (as of 2020-02-25), for example @Amano_2016, @Bartomeus_2013,
+and @Werner_2015 (see Appendix A for all citations).
 
-The `rgbif` package is relatively stable, and should not have
+The `rgbif` package is stable, and should not have
 many breaking changes unless necessitated due to changes in the
-GBIF API. However, it will gain function(s) to work with the maps
-API in the near future.
-
-The `pygbif` and `gbifrb` libraries are in early development, and will
-greatly benefit from any feedback and use cases.
+GBIF API. The `pygbif` and `gbifrb` libraries are in early development,
+and will greatly benefit from any feedback and use cases.
 
@@
 
-[mit]: http://choosealicense.com/licenses/mit/
+[mit]: https://choosealicense.com/licenses/mit/
 [ghrgbif]: https://github.com/ropensci/rgbif
 [cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
-[gbifdocs]: http://www.gbif.org/developer/summary
-[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
+[gbifdocs]: https://www.gbif.org/developer/summary
+[gbifissues]: https://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
 [taxize]: https://github.com/ropensci/taxize
@@
 [pypipygbif]: https://pypi.python.org/pypi/pygbif
+[gemgbif]: https://rubygems.org/gems/gbifrb

### 2020-02-27 manuscript_mee.tex
tweaks to mee paper

--- 2017-09-26 manuscript.tex
@@
-\documentclass[author-year, review, 11pt]{components/elsarticle} %review=doublespace preprint=single 5p=2 column
+\documentclass[3p]{elsarticle} %review=doublespace preprint=single 5p=2 column
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
 \usepackage[hyphens]{url}
+
 \usepackage{lineno} % add
-  \linenumbers % turns line numbering on
-\bibliographystyle{elsarticle-harv}
-\biboptions{sort&compress} % For natbib
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 \usepackage{graphicx}
 \usepackage{booktabs} % book-quality tables
-%% Redefines the elsarticle footer
-\makeatletter
-\def\ps@pprintTitle{%
- \let\@oddhead\@empty
- \let\@evenhead\@empty
- \def\@oddfoot{\it \hfill\today}%
- \let\@evenfoot\@oddfoot}
- \def\tightlist{}
-\makeatother
-
-% A modified page layout
-\textwidth 6.75in
-\oddsidemargin -0.15in
-\evensidemargin -0.15in
-\textheight 9in
-\topmargin -0.5in
 %%%%%%%%%%%%%%%% end my additions to header
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\bibliographystyle{elsarticle-harv}
 \usepackage{color}
@@
 \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
 \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
 \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
 \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
 \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
 \newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
 \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
 \usepackage{longtable}
@@
             pdfauthor={},
-            pdftitle={R Python, and Ruby clients for GBIF species occurrence data},
-            colorlinks=true,
+            pdftitle={R, Python, and Ruby clients for GBIF species occurrence data},
+            colorlinks=false,
             urlcolor=blue,
@@
 \urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
+
 \setcounter{secnumdepth}{0}
@@
 \setcounter{secnumdepth}{0}
+
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
 % Pandoc header
+
+\journal{Methods in Ecology and Evolution}
+\linenumbers
+\usepackage{setspace}
+\doublespacing
 
@@
 
-  \title{R Python, and Ruby clients for GBIF species occurrence data}
-    \author[cstar]{Scott Chamberlain\corref{c1}}
-   \ead{scott(at)ropensci.org} 
-   \cortext[c1]{Corresponding author}
+  \title{R, Python, and Ruby clients for GBIF species occurrence data}
+    \author[cstar]{Scott Chamberlain\corref{Corresponding author}}
+   \ead{myrmecocystus(at)gmail.com} 
     \author[boettig]{Carl Boettiger}
-   \ead{carl(at)ropensci.org}
-
-      \address[cstar]{rOpenSci, Museum of Paleontology, University of California, Berkeley,
-CA, USA}
-    \address[boettig]{rOpenSci, Department of Enivornmental Science, Policy and Management,
-University of California, Berkeley, CA, USA}
+   \ead{cboettig(at)berkeley.edu} 
+      \address[cstar]{rOpenSci, Museum of Paleontology, University of
+California, Berkeley, CA, USA}
+    \address[boettig]{rOpenSci, Department of Enivornmental Science,
+Policy and Management, University of California, Berkeley, CA, USA}
 
@@
   Email address:
-  \href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}
+  \href{mailto:myrmecocystus@gmail.com}{\nolinkurl{myrmecocystus@gmail.com}}
 
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 ecology fields are built largely on existing data, e.g., macro-ecology
-(Brown, 1995; Beck et al., 2012).
+(Brown 1995; Beck \emph{et al.} 2012).
 
@@
 records have been used for a suite of other use cases: validating
-habitat suitability models with real occurrence data (Ficetola et al.,
-2014); ancestral range reconstruction (Ferretti et al., 2015; María
-Mendoza et al., 2015); development of invasive species watch lists
-(Faulkner et al., 2014); evaluating risk of invasive species spread
-(Febbraro et al., 2013); and effects of climate change on future
-biodiversity (Brown et al., 2015).
+habitat suitability models with real occurrence data (Ficetola \emph{et
+al.} 2014); ancestral range reconstruction (Ferretti \emph{et al.} 2015;
+Marı́a Mendoza \emph{et al.} 2015); development of invasive species watch
+lists (Faulkner \emph{et al.} 2014); evaluating risk of invasive species
+spread (Febbraro \emph{et al.} 2013); and effects of climate change on
+future biodiversity (Brown \emph{et al.} 2015).
 
 In addition to wide utility, this data is important for conservation.
-Biodiversity loss is one of the greatest challenges of our time (Pimm et
-al., 2014), and some have called this the sixth great mass extinction
-(Ceballos et al., 2015). Given this challenge there is a great need for
-data on specimen records, whether collected from live sightings in the
-field or specimens in museums.
-
+Biodiversity loss is one of the greatest challenges of our time (Pimm
+\emph{et al.} 2014), and some have called this the sixth great mass
+extinction (Ceballos \emph{et al.} 2015). Given this challenge there is
+a great need for data on specimen records, whether collected from live
+sightings in the field or specimens in museums.
+
+\hypertarget{global-biodiversity-information-facility}{%
 \section{Global Biodiversity Information
-Facility}\label{global-biodiversity-information-facility}
+Facility}\label{global-biodiversity-information-facility}}
 
@@
 
-\section{The clients}\label{the-clients}
+\hypertarget{the-clients}{%
+\section{The clients}\label{the-clients}}
 
@@
 
-\subsection{rgbif}\label{rgbif}
-
-Herein, we describe the \texttt{rgbif} software package (Chamberlain et
-al.) for working with GBIF data in the R programming environment (R Core
-Team, 2014). R is a widely used language in academia, as well as
-non-profit and private sectors. Importantly, R makes it easy to execute
-all steps of the research process, including data management, data
-manipulation and cleaning, statistics, and visualization. Thus, an R
-client for getting GBIF data is a powerful tool to facilitate
+\hypertarget{rgbif}{%
+\subsection{rgbif}\label{rgbif}}
+
+Herein, we describe the \texttt{rgbif} software package (Chamberlain
+\emph{et al.}) for working with GBIF data in the R programming
+environment (R Core Team 2014). R is a widely used language in academia,
+as well as non-profit and private sectors. Importantly, R makes it easy
+to execute all steps of the research process, including data management,
+data manipulation and cleaning, statistics, and visualization. Thus, an
+R client for getting GBIF data is a powerful tool to facilitate
 reproducible research.
@@
 The \texttt{rgbif} package is nearly completely written in R (a small
-Javascript library is included for reading well known text (Herring,
-2011)), uses an \href{http://choosealicense.com/licenses/mit/}{MIT
+Javascript library is included for reading well known text (Herring
+2011)), uses an \href{https://choosealicense.com/licenses/mit/}{MIT
 license} to maximize use everywhere. \texttt{rgbif} is developed
@@
 distribution network for R packages. \texttt{rgbif} is part of the
-rOpenSci project (\url{http://ropensci.org}), a developer network making
-R software to facilitate reproducible research.
-
-\subsection{pygbif}\label{pygbif}
+rOpenSci project (https://ropensci.org), a developer network making R
+software to facilitate reproducible research.
+
+\hypertarget{pygbif}{%
+\subsection{pygbif}\label{pygbif}}
 
@@
 package. It also uses an
-\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
+\href{https://choosealicense.com/licenses/mit/}{MIT license} to maximize
 use everywhere. \texttt{pygbif} is developed publicly on GitHub at
@@
 
-\subsection{gbifrb}\label{gbifrb}
+\hypertarget{gbifrb}{%
+\subsection{gbifrb}\label{gbifrb}}
 
@@
 \begin{Highlighting}[]
-\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{require }\StringTok{\textquotesingle{}gbifrb\textquotesingle{}}
 \end{Highlighting}
@@
 Python libraries. It also uses an
-\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
+\href{https://choosealicense.com/licenses/mit/}{MIT license} to maximize
 use everywhere. \texttt{gbifrb} is developed publicly on GitHub at
@@
 Stable versions of \texttt{gbifrb} can be installed from
-{[}Rubygems{]}{[}gemgbif{]}, the distribution network for Ruby
-libraries.
-
-\subsubsection{Library interfaces}\label{library-interfaces}
+\href{https://rubygems.org/gems/gbifrb}{Rubygems}, the distribution
+network for Ruby libraries.
+
+\hypertarget{library-interfaces}{%
+\section{Library interfaces}\label{library-interfaces}}
 
 \texttt{rgbif}, \texttt{pygbif}, and \texttt{gbifrb} are designed
-following the \href{http://www.gbif.org/developer/summary}{GBIF
+following the \href{https://www.gbif.org/developer/summary}{GBIF
 Application Programming Interface}, or API. The GBIF API has four major
@@
 
-\subsection{GBIF headers}\label{gbif-headers}
+\hypertarget{gbif-headers}{%
+\subsection{GBIF headers}\label{gbif-headers}}
 
@@
 
-\subsection{Registry}\label{registry}
+\hypertarget{registry}{%
+\subsection{Registry}\label{registry}}
 
@@
 
-\subsubsection{Datasets}\label{datasets}
+\hypertarget{datasets}{%
+\subsection{Datasets}\label{datasets}}
 
@@
 \begin{Highlighting}[]
-\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"oregon"}\NormalTok{)}
+\NormalTok{res <{-}}\StringTok{ }\KeywordTok{dataset\_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"oregon"}\NormalTok{)}
 \NormalTok{res}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{datasetTitle[}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{]}
-\CommentTok{#>  [1] "Oregon State Ichthyology Collection"                                                                                                                                                  }
-\CommentTok{#>  [2] "Oregon State University Herpetological Collection"                                                                                                                                    }
-\CommentTok{#>  [3] "Mygalomorph spiders from southwestern Oregon, USA, with descriptions of four new species"                                                                                             }
-\CommentTok{#>  [4] "A new species of Helobdella (Hirudinida: Glossiphoniidae) from Oregon, USA"                                                                                                           }
-\CommentTok{#>  [5] "Annotated Checklist of the large branchiopod crustaceans of Idaho, Oregon and Washington, USA, with the “ rediscovery ” of a new species of Branchinecta (Anostraca: Branchinectidae)"}
-\CommentTok{#>  [6] "A new species of Chrysobothris Eschscholtz from Oregon and Washington, with notes on other Buprestidae (Coleoptera) occurring in the United States and Canada"                        }
-\CommentTok{#>  [7] "Three new species of Grylloblatta Walker (Insecta: Grylloblattodea: Grylloblattidae), from southern Oregon and northern California"                                                   }
-\CommentTok{#>  [8] "A new species of Cladotanytarsus (Lenziella) from Oregon supports the systematic concept of the subgenus (Diptera: Chironomidae)"                                                     }
-\CommentTok{#>  [9] "A new monster from southwest Oregon forests: Cryptomasterbehemoth sp.
-n.
-(Opiliones, Laniatores, Travunioidea)"                                                                       }
-\CommentTok{#> [10] "Two new species of Fluminicola (Caenogastropoda, Lithoglyphidae) from southwest Oregon, USA, and a range extension for F. multifarius"}
+\CommentTok{\#>  [1] "Oregon State Ichthyology Collection"                                                                                                                                                  }
+\CommentTok{\#>  [2] "Oregon State University Herpetological Collection"                                                                                                                                    }
+\CommentTok{\#>  [3] "Mygalomorph spiders from southwestern Oregon, USA, with descriptions of four new species"                                                                                             }
+\CommentTok{\#>  [4] "A new species of Helobdella (Hirudinida: Glossiphoniidae) from Oregon, USA"                                                                                                           }
+\CommentTok{\#>  [5] "Cave millipedes of the United States.
+XVI.
+Two new species from Oregon Caves National Monument, Oregon (Chordeumatida, Conotylidae and Caseyidae)"                                    }
+\CommentTok{\#>  [6] "Annotated Checklist of the large branchiopod crustaceans of Idaho, Oregon and Washington, USA, with the “ rediscovery ” of a new species of Branchinecta (Anostraca: Branchinectidae)"}
+\CommentTok{\#>  [7] "A new species of Chrysobothris Eschscholtz from Oregon and Washington, with notes on other Buprestidae (Coleoptera) occurring in the United States and Canada"                        }
+\CommentTok{\#>  [8] "Three new species of Grylloblatta Walker (Insecta: Grylloblattodea: Grylloblattidae), from southern Oregon and northern California"                                                   }
+\CommentTok{\#>  [9] "A new species of Cladotanytarsus (Lenziella) from Oregon supports the systematic concept of the subgenus (Diptera: Chironomidae)"                                                     }
+\CommentTok{\#> [10] "Two new species of Fluminicola (Caenogastropoda, Lithoglyphidae) from southwest Oregon, USA, and a range extension for F. multifarius"}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{require }\StringTok{\textquotesingle{}gbifrb\textquotesingle{}}
 \NormalTok{registry = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Registry}
@@
 
-\paragraph{Dataset metrics}\label{dataset-metrics}
+\hypertarget{dataset-metrics}{%
+\subsection{Dataset metrics}\label{dataset-metrics}}
 
@@
 available for datasets of type \emph{checklist}, but there are quite a
-lot of them (21697).
+lot of them (29853).
 
@@
 \begin{Highlighting}[]
-\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
+\NormalTok{res <{-}}\StringTok{ }\KeywordTok{dataset\_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{\textquotesingle{}ec93a739{-}1681{-}4b04{-}b62f{-}3a687127a17f\textquotesingle{}}\NormalTok{)}
 \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res}\OperatorTok{$}\NormalTok{countByRank),}
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ registry}
-\NormalTok{registry.dataset_metrics(uuid}\OperatorTok{=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
+\NormalTok{registry.dataset\_metrics(uuid}\OperatorTok{=}\StringTok{\textquotesingle{}ec93a739{-}1681{-}4b04{-}b62f{-}3a687127a17f\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{require }\StringTok{\textquotesingle{}gbifrb\textquotesingle{}}
 \NormalTok{registry = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Registry}
-\NormalTok{registry.dataset_metrics(}\StringTok{uuid: 'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\subsubsection{Networks, nodes, and
-installations}\label{networks-nodes-and-installations}
+\NormalTok{registry.dataset\_metrics(}\StringTok{uuid: \textquotesingle{}ec93a739{-}1681{-}4b04{-}b62f{-}3a687127a17f\textquotesingle{}}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{networks-nodes-and-installations}{%
+\subsection{Networks, nodes, and
+installations}\label{networks-nodes-and-installations}}
 
@@
 \KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit =} \DecValTok{10}\NormalTok{)}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{title}
-\CommentTok{#>  [1] "GBIF Backbone Sources"                                      }
-\CommentTok{#>  [2] "Canadensys"                                                 }
-\CommentTok{#>  [3] "Southwest Collections of Arthropods Network (SCAN)"         }
-\CommentTok{#>  [4] "VertNet"                                                    }
-\CommentTok{#>  [5] "Dryad"                                                      }
-\CommentTok{#>  [6] "GBIF Network"                                               }
-\CommentTok{#>  [7] "The Knowledge Network for Biocomplexity (KNB) "             }
-\CommentTok{#>  [8] "Online Zoological Collections of Australian Museums (OZCAM)"}
-\CommentTok{#>  [9] "Catalogue of Life"                                          }
-\CommentTok{#> [10] "Ocean Biogeographic Information System (OBIS)"}
+\CommentTok{\#> [1] "TrIAS"                                        }
+\CommentTok{\#> [2] "Arctos"                                       }
+\CommentTok{\#> [3] "Freshwater Network"                           }
+\CommentTok{\#> [4] "Ocean Biogeographic Information System (OBIS)"}
 \end{Highlighting}
@@
 
-\subsection{Taxonomic names}\label{taxonomic-names}
+\hypertarget{taxonomic-names}{%
+\subsection{Taxonomic names}\label{taxonomic-names}}
 
@@
 
-\subsubsection{GBIF Backbone}\label{gbif-backbone}
+\hypertarget{gbif-backbone}{%
+\subsection{GBIF Backbone}\label{gbif-backbone}}
 
@@
 refer to taxonomic names throughout their services. The backbone has
-5869207 unique names and 2818534 species names. The backbone taxonomy is
+6575748 unique names and 2730631 species names. The backbone taxonomy is
 also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
@@
 \begin{Highlighting}[]
-\NormalTok{res <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Poa'}\NormalTok{, }\DataTypeTok{rank=}\StringTok{'genus'}\NormalTok{, }\DataTypeTok{family=}\StringTok{'Poaceae'}\NormalTok{)}
-\NormalTok{res[}\KeywordTok{c}\NormalTok{(}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{)]}
-\CommentTok{#> $usageKey}
-\CommentTok{#> [1] 2704173}
-\CommentTok{#> }
-\CommentTok{#> $kingdom}
-\CommentTok{#> [1] "Plantae"}
+\NormalTok{res <{-}}\StringTok{ }\KeywordTok{name\_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{\textquotesingle{}Poa\textquotesingle{}}\NormalTok{, }\DataTypeTok{rank=}\StringTok{\textquotesingle{}genus\textquotesingle{}}\NormalTok{, }\DataTypeTok{family=}\StringTok{\textquotesingle{}Poaceae\textquotesingle{}}\NormalTok{)}
+\NormalTok{res[}\KeywordTok{c}\NormalTok{(}\StringTok{\textquotesingle{}usageKey\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}kingdom\textquotesingle{}}\NormalTok{)]}
+\CommentTok{\#> \# A tibble: 1 x 2}
+\CommentTok{\#>   usageKey kingdom}
+\CommentTok{\#>      <int> <chr>  }
+\CommentTok{\#> 1  2704173 Plantae}
 \end{Highlighting}
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ species}
-\NormalTok{res }\OperatorTok{=}\NormalTok{ species.name_backbone(name}\OperatorTok{=}\StringTok{'Poa'}\NormalTok{, rank}\OperatorTok{=}\StringTok{'genus'}\NormalTok{, family}\OperatorTok{=}\StringTok{'Poaceae'}\NormalTok{)}
-\NormalTok{[ res[x] }\ControlFlowTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ [}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{] ]}
+\NormalTok{res }\OperatorTok{=}\NormalTok{ species.name\_backbone(name}\OperatorTok{=}\StringTok{\textquotesingle{}Poa\textquotesingle{}}\NormalTok{, rank}\OperatorTok{=}\StringTok{\textquotesingle{}genus\textquotesingle{}}\NormalTok{, family}\OperatorTok{=}\StringTok{\textquotesingle{}Poaceae\textquotesingle{}}\NormalTok{)}
+\NormalTok{[ res[x] }\ControlFlowTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ [}\StringTok{\textquotesingle{}usageKey\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}kingdom\textquotesingle{}}\NormalTok{] ]}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{require }\StringTok{\textquotesingle{}gbifrb\textquotesingle{}}
 \NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
-\NormalTok{res = species.name_backbone(}\StringTok{name: 'Poa'}\NormalTok{, }\StringTok{rank: 'genus'}\NormalTok{, }\StringTok{family: 'Poaceae'}\NormalTok{)}
+\NormalTok{res = species.name\_backbone(}\StringTok{name: \textquotesingle{}Poa\textquotesingle{}}\NormalTok{, }\StringTok{rank: \textquotesingle{}genus\textquotesingle{}}\NormalTok{, }\StringTok{family: \textquotesingle{}Poaceae\textquotesingle{}}\NormalTok{)}
 \NormalTok{res.select \{ |k,v| k.match(}\OtherTok{/usageKey|kingdom/}\NormalTok{) \}}
@@
 
-\subsubsection{Name searching}\label{name-searching}
+\hypertarget{name-searching}{%
+\subsection{Name searching}\label{name-searching}}
 
@@
 \begin{Highlighting}[]
-\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q=}\StringTok{'Pum'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{6}\NormalTok{)}
+\KeywordTok{name\_suggest}\NormalTok{(}\DataTypeTok{q=}\StringTok{\textquotesingle{}Pum\textquotesingle{}}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{6}\NormalTok{)}
 \end{Highlighting}
@@
 8589398 & Pumiliopimoidae & FAMILY\tabularnewline
+8608394 & Pumiliopimoa & GENUS\tabularnewline
+4579844 & Pumilibaxa & GENUS\tabularnewline
+4581745 & Pumicia & GENUS\tabularnewline
 8783253 & Pumililema & GENUS\tabularnewline
-4823360 & Pumiliopareia & GENUS\tabularnewline
-4635949 & Pumilina & GENUS\tabularnewline
-4648228 & Pumilopaurus & GENUS\tabularnewline
 \bottomrule
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ species}
-\NormalTok{species.name_suggest(q}\OperatorTok{=}\StringTok{'Pum'}\NormalTok{, limit }\OperatorTok{=} \DecValTok{6}\NormalTok{)}
+\NormalTok{species.name\_suggest(q}\OperatorTok{=}\StringTok{\textquotesingle{}Pum\textquotesingle{}}\NormalTok{, limit }\OperatorTok{=} \DecValTok{6}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{require }\StringTok{\textquotesingle{}gbifrb\textquotesingle{}}
 \NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
-\NormalTok{species.name_suggest(}\StringTok{q: 'Pum'}\NormalTok{, }\StringTok{limit: }\DecValTok{6}\NormalTok{)}
+\NormalTok{species.name\_suggest(}\StringTok{q: \textquotesingle{}Pum\textquotesingle{}}\NormalTok{, }\StringTok{limit: }\DecValTok{6}\NormalTok{)}
 \end{Highlighting}
@@
 
-\subsection{Occurrences}\label{occurrences}
+\hypertarget{occurrences}{%
+\subsection{Occurrences}\label{occurrences}}
 
@@
 
-\subsubsection{Download API}\label{download-api}
+\hypertarget{occurrences-download-api}{%
+\subsection{Occurrences Download API}\label{occurrences-download-api}}
 
@@
 \texttt{3119195}, which is the key for \emph{Helianthus annuus}
-(\url{http://www.gbif.org/species/3119195}).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{occ_download}\NormalTok{(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
-\CommentTok{#> <<gbif download>>}
-\CommentTok{#>   Username: xxxx}
-\CommentTok{#>   E-mail: xxxx}
-\CommentTok{#>   Download key: 0000840-150615163101818}
+(http://www.gbif.org/species/3119195).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{occ\_download}\NormalTok{(}\StringTok{\textquotesingle{}taxonKey = 3119195\textquotesingle{}}\NormalTok{)}
+\CommentTok{\#> <<gbif download>>}
+\CommentTok{\#>   Username: xxxx}
+\CommentTok{\#>   E{-}mail: xxxx}
+\CommentTok{\#>   Download key: 0000840{-}150615163101818}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{(res <-}\StringTok{ }\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000840-150615163101818"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
-\CommentTok{#> <<gbif downloaded get>>}
-\CommentTok{#>   Path: ./0000840-150615163101818.zip}
-\CommentTok{#>   File size: 3.19 MB}
+\NormalTok{(res <{-}}\StringTok{ }\KeywordTok{occ\_download\_get}\NormalTok{(}\StringTok{"0000840{-}150615163101818"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
+\CommentTok{\#> <<gbif downloaded get>>}
+\CommentTok{\#>   Path: ./0000840{-}150615163101818.zip}
+\CommentTok{\#>   File size: 3.19 MB}
 \end{Highlighting}
@@
 \KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_download_import}\NormalTok{(res)}
-\NormalTok{dat }\OperatorTok{%>%}
+\NormalTok{dat <{-}}\StringTok{ }\KeywordTok{occ\_download\_import}\NormalTok{(res)}
+\NormalTok{dat }\OperatorTok{\%>\%}
 \StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
-\CommentTok{#>       gbifID abstract accessRights accrualMethod accrualPeriodicity accrualPolicy alternative audience}
-\CommentTok{#> 1  725767384       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 2  725767447       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 3  725767450       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 4  725767513       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 5  725767546       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 6  725767579       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 7  725767609       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 8  725767645       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 9  725767678       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> 10 725767681       NA                         NA                 NA            NA          NA       NA}
-\CommentTok{#> ..       ...      ...          ...           ...                ...           ...         ...      ...}
-\CommentTok{#> Variables not shown: available (lgl), bibliographicCitation (chr), conformsTo (lgl), contributor (lgl),}
-\CommentTok{#>      coverage (lgl), created (chr), creator (lgl), date (lgl), dateAccepted (lgl), dateCopyrighted}
-\CommentTok{#>      (lgl), dateSubmitted (lgl), description (lgl), educationLevel (lgl), extent (lgl), format (lgl),}
-\CommentTok{#>      hasFormat (lgl), hasPart (lgl), hasVersion (lgl), identifier (chr), instructionalMethod (lgl),}
+\CommentTok{\#>       gbifID abstract accessRights accrualMethod accrualPeriodicity accrualPolicy alternative audience}
+\CommentTok{\#> 1  725767384       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 2  725767447       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 3  725767450       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 4  725767513       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 5  725767546       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 6  725767579       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 7  725767609       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 8  725767645       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 9  725767678       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> 10 725767681       NA                         NA                 NA            NA          NA       NA}
+\CommentTok{\#> ..       ...      ...          ...           ...                ...           ...         ...      ...}
+\CommentTok{\#> Variables not shown: available (lgl), bibliographicCitation (chr), conformsTo (lgl), contributor (lgl),}
+\CommentTok{\#>      coverage (lgl), created (chr), creator (lgl), date (lgl), dateAccepted (lgl), dateCopyrighted}
+\CommentTok{\#>      (lgl), dateSubmitted (lgl), description (lgl), educationLevel (lgl), extent (lgl), format (lgl),}
+\CommentTok{\#>      hasFormat (lgl), hasPart (lgl), hasVersion (lgl), identifier (chr), instructionalMethod (lgl),}
 \end{Highlighting}
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
-\NormalTok{occ.download(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
-\NormalTok{(res }\OperatorTok{=}\NormalTok{ occ.download_get(}\StringTok{"0000840-150615163101818"}\NormalTok{, overwrite }\OperatorTok{=} \VariableTok{True}\NormalTok{))}
+\NormalTok{occ.download(}\StringTok{\textquotesingle{}taxonKey = 3119195\textquotesingle{}}\NormalTok{)}
+\NormalTok{(res }\OperatorTok{=}\NormalTok{ occ.download\_get(}\StringTok{"0000840{-}150615163101818"}\NormalTok{, overwrite }\OperatorTok{=} \VariableTok{True}\NormalTok{))}
 \end{Highlighting}
@@
 
-\paragraph{Downloaded data format}\label{downloaded-data-format}
+\hypertarget{downloaded-data-format}{%
+\subsubsection{Downloaded data format}\label{downloaded-data-format}}
 
@@
 
-\subsubsection{Search API}\label{search-api}
+\hypertarget{search-api}{%
+\subsection{Search API}\label{search-api}}
 
@@
 
-\paragraph{Search for occurrences}\label{search-for-occurrences}
+\hypertarget{search-for-occurrences}{%
+\subsubsection{Search for occurrences}\label{search-for-occurrences}}
 
@@
 \KeywordTok{library}\NormalTok{(rgbif)}
-\NormalTok{(key <-}\StringTok{ }\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q =} \StringTok{'Helianthus annuus'}\NormalTok{, }\DataTypeTok{rank =} \StringTok{'species'}\NormalTok{)}\OperatorTok{$}\NormalTok{key[}\DecValTok{1}\NormalTok{])}
-\CommentTok{#> [1] 9206251}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
-\CommentTok{#> Records found [17858] }
-\CommentTok{#> Records returned [2] }
-\CommentTok{#> No. unique hierarchies [1] }
-\CommentTok{#> No. media records [1] }
-\CommentTok{#> No. facets [0] }
-\CommentTok{#> Args [limit=2, offset=0, taxonKey=9206251, fields=all] }
-\CommentTok{#> # A tibble: 2 x 75}
-\CommentTok{#>                name        key decimalLatitude decimalLongitude}
-\CommentTok{#>               <chr>      <int>           <dbl>            <dbl>}
-\CommentTok{#> 1 Helianthus annuus 1433793045        59.66859         16.54257}
-\CommentTok{#> 2 Helianthus annuus 1434024463        63.71622         20.31247}
-\CommentTok{#> # ... with 71 more variables: issues <chr>, datasetKey <chr>,}
-\CommentTok{#> #   publishingOrgKey <chr>, publishingCountry <chr>, protocol <chr>,}
-\CommentTok{#> #   lastCrawled <chr>, lastParsed <chr>, crawlId <int>, extensions <chr>,}
-\CommentTok{#> #   basisOfRecord <chr>, ...}
+\NormalTok{(key <{-}}\StringTok{ }\KeywordTok{name\_suggest}\NormalTok{(}\DataTypeTok{q =} \StringTok{\textquotesingle{}Helianthus annuus\textquotesingle{}}\NormalTok{, }\DataTypeTok{rank =} \StringTok{\textquotesingle{}species\textquotesingle{}}\NormalTok{)}\OperatorTok{$}\NormalTok{key[}\DecValTok{1}\NormalTok{])}
+\CommentTok{\#> [1] 9206251}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
+\CommentTok{\#> Records found [46490] }
+\CommentTok{\#> Records returned [2] }
+\CommentTok{\#> No. unique hierarchies [1] }
+\CommentTok{\#> No. media records [2] }
+\CommentTok{\#> No. facets [0] }
+\CommentTok{\#> Args [limit=2, offset=0, taxonKey=9206251, fields=all] }
+\CommentTok{\#> \# A tibble: 2 x 71}
+\CommentTok{\#>   key   scientificName decimalLatitude decimalLongitude issues datasetKey}
+\CommentTok{\#>   <chr> <chr>                    <dbl>            <dbl> <chr>  <chr>     }
+\CommentTok{\#> 1 2542\textasciitilde{} Helianthus an\textasciitilde{}            25.8            {-}100. cdrou\textasciitilde{} 50c9509d{-}\textasciitilde{}}
+\CommentTok{\#> 2 2550\textasciitilde{} Helianthus an\textasciitilde{}            25.6            {-}100. cdrou\textasciitilde{} 50c9509d{-}\textasciitilde{}}
+\CommentTok{\#> \# ... with 65 more variables: publishingOrgKey <chr>, installationKey <chr>,}
+\CommentTok{\#> \#   publishingCountry <chr>, protocol <chr>, lastCrawled <chr>,}
+\CommentTok{\#> \#   lastParsed <chr>, crawlId <int>, extensions <chr>, basisOfRecord <chr>,}
+\CommentTok{\#> \#   taxonKey <int>, ...}
 \end{Highlighting}
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
-\NormalTok{key }\OperatorTok{=}\NormalTok{ species.name_suggest(q }\OperatorTok{=} \StringTok{'Helianthus annuus'}\NormalTok{, rank }\OperatorTok{=} \StringTok{'species'}\NormalTok{)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{]}
+\NormalTok{key }\OperatorTok{=}\NormalTok{ species.name\_suggest(q }\OperatorTok{=} \StringTok{\textquotesingle{}Helianthus annuus\textquotesingle{}}\NormalTok{, rank }\OperatorTok{=} \StringTok{\textquotesingle{}species\textquotesingle{}}\NormalTok{)[}\StringTok{\textquotesingle{}data\textquotesingle{}}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{\textquotesingle{}key\textquotesingle{}}\NormalTok{]}
 \NormalTok{occ.search(taxonKey }\OperatorTok{=}\NormalTok{ key, limit }\OperatorTok{=} \DecValTok{2}\NormalTok{)}
@@
 \begin{Highlighting}[]
-\NormalTok{require }\StringTok{'gbifrb'}
+\NormalTok{require }\StringTok{\textquotesingle{}gbifrb\textquotesingle{}}
 \NormalTok{species = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Species}
 \NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
-\NormalTok{key = species.name_suggest(}\StringTok{q: 'Helianthus annuus'}\NormalTok{, }\StringTok{rank: 'species'}\NormalTok{)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{]}
+\NormalTok{key = species.name\_suggest(}\StringTok{q: \textquotesingle{}Helianthus annuus\textquotesingle{}}\NormalTok{, }\StringTok{rank: \textquotesingle{}species\textquotesingle{}}\NormalTok{)[}\StringTok{\textquotesingle{}data\textquotesingle{}}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{\textquotesingle{}key\textquotesingle{}}\NormalTok{]}
 \NormalTok{occ.search(}\StringTok{taxonKey: }\NormalTok{key, }\StringTok{limit: }\DecValTok{2}\NormalTok{)}
@@
 \begin{Highlighting}[]
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{scientificName =} \StringTok{'Ursus americanus'}\NormalTok{)}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{scientificName =} \StringTok{\textquotesingle{}Ursus americanus\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{occ.search(scientificName }\OperatorTok{=} \StringTok{'Ursus americanus'}\NormalTok{)}
+\NormalTok{occ.search(scientificName }\OperatorTok{=} \StringTok{\textquotesingle{}Ursus americanus\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{occ.search(}\StringTok{scientificName: 'Ursus americanus'}\NormalTok{)}
+\NormalTok{occ.search(}\StringTok{scientificName: \textquotesingle{}Ursus americanus\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{splist <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{)}
-\NormalTok{keys <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(splist, }\ControlFlowTok{function}\NormalTok{(x) }\KeywordTok{name_suggest}\NormalTok{(x)}\OperatorTok{$}\NormalTok{key[}\DecValTok{1}\NormalTok{], }\DataTypeTok{USE.NAMES =} \OtherTok{FALSE}\NormalTok{)}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keys, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{)}
+\NormalTok{splist <{-}}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{\textquotesingle{}Cyanocitta stelleri\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Junco hyemalis\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Aix sponsa\textquotesingle{}}\NormalTok{)}
+\NormalTok{keys <{-}}\StringTok{ }\KeywordTok{sapply}\NormalTok{(splist, }\ControlFlowTok{function}\NormalTok{(x) }\KeywordTok{name\_suggest}\NormalTok{(x)}\OperatorTok{$}\NormalTok{key[}\DecValTok{1}\NormalTok{], }\DataTypeTok{USE.NAMES =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keys, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{, }\DataTypeTok{return =} \StringTok{\textquotesingle{}data\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
-\NormalTok{splist }\OperatorTok{=}\NormalTok{ [}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{]}
-\NormalTok{keys }\OperatorTok{=}\NormalTok{ [ species.name_suggest(x)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{] }\ControlFlowTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ splist ]}
+\NormalTok{splist }\OperatorTok{=}\NormalTok{ [}\StringTok{\textquotesingle{}Cyanocitta stelleri\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Junco hyemalis\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Aix sponsa\textquotesingle{}}\NormalTok{]}
+\NormalTok{keys }\OperatorTok{=}\NormalTok{ [ species.name\_suggest(x)[}\StringTok{\textquotesingle{}data\textquotesingle{}}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{\textquotesingle{}key\textquotesingle{}}\NormalTok{] }\ControlFlowTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ splist ]}
 \NormalTok{occ.search(taxonKey }\OperatorTok{=}\NormalTok{ keys, limit }\OperatorTok{=} \DecValTok{5}\NormalTok{)}
@@
 \NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
-\NormalTok{splist = [}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{]}
-\NormalTok{keys = [ species.name_suggest(x)[}\StringTok{'data'}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{'key'}\NormalTok{] }\KeywordTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ splist ]}
+\NormalTok{splist = [}\StringTok{\textquotesingle{}Cyanocitta stelleri\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Junco hyemalis\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Aix sponsa\textquotesingle{}}\NormalTok{]}
+\NormalTok{keys = [ species.name\_suggest(x)[}\StringTok{\textquotesingle{}data\textquotesingle{}}\NormalTok{][}\DecValTok{0}\NormalTok{][}\StringTok{\textquotesingle{}key\textquotesingle{}}\NormalTok{] }\KeywordTok{for}\NormalTok{ x }\KeywordTok{in}\NormalTok{ splist ]}
 \NormalTok{occ.search(}\StringTok{taxonKey: }\NormalTok{keys, }\StringTok{limit: }\DecValTok{5}\NormalTok{)}
@@
 
-Spatial search, based on well known text format (Herring, 2011), or a
+Spatial search, based on well known text format (Herring 2011), or a
 bounding box set of four coordinates. The well known text string and the
@@
 \begin{Highlighting}[]
-\CommentTok{# well known text}
-\NormalTok{wkt <-}\StringTok{ 'POLYGON((-122.6 39.9,-120.0 39.9,-120.0 37.9,-122.6 37.9,-122.6 39.9))'}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
-\CommentTok{# bounding box}
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{122.6}\NormalTok{,}\FloatTok{37.9}\NormalTok{,}\OperatorTok{-}\FloatTok{120.0}\NormalTok{,}\FloatTok{39.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\CommentTok{\# well known text}
+\NormalTok{wkt <{-}}\StringTok{ \textquotesingle{}POLYGON(({-}122.6 39.9,{-}120.0 39.9,{-}120.0 37.9,{-}122.6 37.9,{-}122.6 39.9))\textquotesingle{}}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\CommentTok{\# bounding box}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(}\OperatorTok{{-}}\FloatTok{122.6}\NormalTok{,}\FloatTok{37.9}\NormalTok{,}\OperatorTok{{-}}\FloatTok{120.0}\NormalTok{,}\FloatTok{39.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
 \end{Highlighting}
@@
 \ImportTok{from}\NormalTok{ pygbif }\ImportTok{import}\NormalTok{ occurrences }\ImportTok{as}\NormalTok{ occ}
-\CommentTok{# well known text}
-\NormalTok{occ.search(geometry }\OperatorTok{=} \StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, limit }\OperatorTok{=} \DecValTok{20}\NormalTok{)}
-\CommentTok{# bounding box}
-\NormalTok{occ.search(geometry }\OperatorTok{=} \StringTok{'-125.0,38.4,-121.8,40.9'}\NormalTok{, limit }\OperatorTok{=} \DecValTok{20}\NormalTok{)}
+\CommentTok{\# well known text}
+\NormalTok{occ.search(geometry }\OperatorTok{=} \StringTok{\textquotesingle{}POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))\textquotesingle{}}\NormalTok{, limit }\OperatorTok{=} \DecValTok{20}\NormalTok{)}
+\CommentTok{\# bounding box}
+\NormalTok{occ.search(geometry }\OperatorTok{=} \StringTok{\textquotesingle{}{-}125.0,38.4,{-}121.8,40.9\textquotesingle{}}\NormalTok{, limit }\OperatorTok{=} \DecValTok{20}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{occ = }\DataTypeTok{Gbif}\NormalTok{::}\DataTypeTok{Occurrences}
-\CommentTok{# well known text}
-\NormalTok{occ.search(}\StringTok{geometry: 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, }\StringTok{limit: }\DecValTok{20}\NormalTok{)}
-\CommentTok{# bounding box}
-\NormalTok{occ.search(}\StringTok{geometry: '-125.0,38.4,-121.8,40.9'}\NormalTok{, }\StringTok{limit: }\DecValTok{20}\NormalTok{)}
+\CommentTok{\# well known text}
+\NormalTok{occ.search(}\StringTok{geometry: \textquotesingle{}POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))\textquotesingle{}}\NormalTok{, }\StringTok{limit: }\DecValTok{20}\NormalTok{)}
+\CommentTok{\# bounding box}
+\NormalTok{occ.search(}\StringTok{geometry: \textquotesingle{}{-}125.0,38.4,{-}121.8,40.9\textquotesingle{}}\NormalTok{, }\StringTok{limit: }\DecValTok{20}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasGeospatialIssue =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
+\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{hasGeospatialIssue =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 
-\paragraph{Data cleaning}\label{data-cleaning}
+\hypertarget{data-cleaning}{%
+\section{Data cleaning}\label{data-cleaning}}
 
@@
 GBIF issues; and the same data on
-\href{http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html}{GBIF's
+\href{https://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html}{GBIF's
 development site}).
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{issue =} \StringTok{'DEPTH_UNLIKELY'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
+\NormalTok{out <{-}}\StringTok{ }\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{issue =} \StringTok{\textquotesingle{}DEPTH\_UNLIKELY\textquotesingle{}}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
 \KeywordTok{NROW}\NormalTok{(out)}
-\CommentTok{#> [1] 5}
-\NormalTok{out }\OperatorTok{%>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(}\OperatorTok{-}\NormalTok{cudc) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.}\OperatorTok{$}\NormalTok{data }\OperatorTok{%>%}\StringTok{ }\NormalTok{NROW}
-\CommentTok{#> [1] 0}
+\CommentTok{\#> [1] 5}
+\NormalTok{out }\OperatorTok{\%>\%}\StringTok{ }\KeywordTok{occ\_issues}\NormalTok{(}\OperatorTok{{-}}\NormalTok{cudc) }\OperatorTok{\%>\%}\StringTok{ }\NormalTok{.}\OperatorTok{$}\NormalTok{data }\OperatorTok{\%>\%}\StringTok{ }\NormalTok{NROW}
+\CommentTok{\#> [1] 500}
 \end{Highlighting}
@@
 
-\subsection{Mapping}\label{mapping}
+\hypertarget{mapping}{%
+\section{Mapping}\label{mapping}}
 
 An obvious downstream use case for species occurrence data is to map the
-data. \texttt{rgbif} per se is largely not concerned with making this
-easier, although we do have a simple wrapper around \texttt{ggplot2} to
-make it easy to get a quick plot of occurrence data. For example, here
-we plot 100 occurrences for \emph{Puma concolor}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{key <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Puma concolor'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ key, }\DataTypeTok{limit =} \DecValTok{100}\NormalTok{, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{)}
-\KeywordTok{gbifmap}\NormalTok{(dat}\OperatorTok{$}\NormalTok{data)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{components/figure/manuscript-unnamed-chunk-47-1.pdf}
-
-Another package, \href{https://github.com/ropensci/mapr}{mapr}, is the
-perfect mapping companion to \texttt{rgbif}. It has convenient functions
-for handling input data from \texttt{rgbif}, \texttt{spocc}, or
-arbitrary data.frame's, and output plots for base plots,
-\texttt{ggplot2}, \texttt{ggmap} (\texttt{ggplot2} with map layers
-underneath), and interactive maps on GitHub gists or with Leaflet.js.
+data. \texttt{rgbif} has enough to do without handling mapping. Another
+package, \href{https://github.com/ropensci/mapr}{mapr}, works with
+\texttt{rgbif} to make this easy. For example, here we plot 100
+occurrences for \emph{Puma concolor}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(mapr)}
+\NormalTok{key <{-}}\StringTok{ }\KeywordTok{name\_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{\textquotesingle{}Puma concolor\textquotesingle{}}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
+\NormalTok{dat <{-}}\StringTok{ }\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ key, }\DataTypeTok{limit =} \DecValTok{100}\NormalTok{, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{)}
+\KeywordTok{map\_ggplot}\NormalTok{(dat, }\DataTypeTok{name =} \StringTok{"species"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{components/figure/manuscript_mee-unnamed-chunk-47-1.pdf}
+
+\texttt{mapr} has convenient functions for handling input data from
+\texttt{rgbif}, \texttt{spocc}, or arbitrary data.frame's, and output
+plots for base plots, \texttt{ggplot2}, \texttt{ggmap} (\texttt{ggplot2}
+with map layers underneath), and interactive maps on GitHub gists or
+with Leaflet.js.
 
@@
 
-\subsection{GBIF data in other R
-packages}\label{gbif-data-in-other-r-packages}
+\hypertarget{citing-gbif-data}{%
+\section{Citing GBIF data}\label{citing-gbif-data}}
+
+All the data within GBIF was painstakingly collected by people and
+institutions across the globe. They all, including GBIF, deserve to be
+cited appropropriately if you use any GBIF data in your research. There
+are clear forces making this difficult however. Publishers often limit,
+without reason if the publication is online only, the number of
+citations one can include a manuscript. Even when this happens, you can
+always include additional citations in an appendix.
+
+To make citing data from GBIF easier \texttt{rgbif} has a function
+\texttt{gbif\_citation()}, which accepts the output of many different
+\texttt{rgbif} functions and gives citations for each dataset
+represented in the data therein. For example, suppose you searched for
+occurrences for \texttt{taxonKey} 9206251 (\emph{Helianthus annuus}):
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res <{-}}\StringTok{ }\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{taxonKey =} \DecValTok{9206251}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+We retrieved 20 occurrences from 4 different datasets.
+
+To get citations for each of the four datasets, one can pass the output
+of \texttt{occ\_search()}, here the object \texttt{res}, to the
+\texttt{gbif\_citation()} function:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cites <{-}}\StringTok{ }\KeywordTok{gbif\_citation}\NormalTok{(res)}
+\NormalTok{cites}
+\CommentTok{\#> [[1]]}
+\CommentTok{\#> <<rgbif citation>>}
+\CommentTok{\#>    Citation: iNaturalist.org (2020). iNaturalist Research{-}grade Observations.}
+\CommentTok{\#>         Occurrence dataset https://doi.org/10.15468/ab3s5x accessed via}
+\CommentTok{\#>         GBIF.org on 2020{-}02{-}25.. Accessed from R via rgbif}
+\CommentTok{\#>         (https://github.com/ropensci/rgbif) on 2020{-}02{-}25}
+\CommentTok{\#>    Rights:}
+\CommentTok{\#> }
+\CommentTok{\#> [[2]]}
+\CommentTok{\#> <<rgbif citation>>}
+\CommentTok{\#>    Citation: Shah M, Coulson S (2020). Artportalen (Swedish Species Observation}
+\CommentTok{\#>         System). Version 92.180. ArtDatabanken. Occurrence dataset}
+\CommentTok{\#>         https://doi.org/10.15468/kllkyl accessed via GBIF.org on 2020{-}02{-}25..}
+\CommentTok{\#>         Accessed from R via rgbif (https://github.com/ropensci/rgbif) on}
+\CommentTok{\#>         2020{-}02{-}25}
+\CommentTok{\#>    Rights:}
+\CommentTok{\#> }
+\CommentTok{\#> [[3]]}
+\CommentTok{\#> <<rgbif citation>>}
+\CommentTok{\#>    Citation: naturgucker.de. naturgucker. Occurrence dataset}
+\CommentTok{\#>         https://doi.org/10.15468/uc1apo accessed via GBIF.org on 2020{-}02{-}25..}
+\CommentTok{\#>         Accessed from R via rgbif (https://github.com/ropensci/rgbif) on}
+\CommentTok{\#>         2020{-}02{-}25}
+\CommentTok{\#>    Rights:}
+\CommentTok{\#> }
+\CommentTok{\#> [[4]]}
+\CommentTok{\#> <<rgbif citation>>}
+\CommentTok{\#>    Citation: Vanreusel W, Barendse R, Steeman R, Gielen K, Swinnen K, Desmet P,}
+\CommentTok{\#>         Herremans M (2020). Waarnemingen.be {-} Non{-}native plant occurrences in}
+\CommentTok{\#>         Flanders and the Brussels Capital Region, Belgium. Version 1.14.}
+\CommentTok{\#>         Natuurpunt. Occurrence dataset https://doi.org/10.15468/smdvdo accessed}
+\CommentTok{\#>         via GBIF.org on 2020{-}02{-}25.. Accessed from R via rgbif}
+\CommentTok{\#>         (https://github.com/ropensci/rgbif) on 2020{-}02{-}25}
+\CommentTok{\#>    Rights:}
+\end{Highlighting}
+\end{Shaded}
+
+And we get four citations, one of each of the datasets behind the 20
+occurrence records retrieved.
+
+We don't yet have the same functionality available in \texttt{pygbif},
+but it will be coming soon.
+
+\hypertarget{downloads-vs.-searches}{%
+\subsection{Downloads vs.~searches}\label{downloads-vs.-searches}}
+
+The distinction between ``downloads'' and ``searches'' is not especially
+meaningful with respect to the science done with GBIF data. However,
+downloads and searches are different with respect to easily being able
+to cite GBIF data, which matters for GBIF and for data providers.
+
+Downloads refers to the \texttt{rgbif} functions that start with
+\texttt{occ\_download}, and the \texttt{pygbif} methods that start with
+\texttt{occurrences.download}. These methods use the GBIF ``downloads
+API'', which you can interact with from \texttt{rgbif} and
+\texttt{pygbif} programatically, similar to how you would interact with
+the GBIF website - each interaction creating a download request, which
+processes in the background. The user has to ask if the request is ready
+to download. Downloads can request a very large amount of data.
+
+Searches are faster and more dynamic. Search functions in \texttt{rgbif}
+include \texttt{occ\_search}, \texttt{occ\_data}, and \texttt{occ\_get},
+while \texttt{pygbif} methods include \texttt{occurrences.search},
+\texttt{occurrences.get}. These methods can only request up to 200,000
+occurrence records.
+
+The big difference between downloads and searches with respect to
+citations is that each download generates a Digital Object Identifier
+(or DOI). This DOI can be used cite the entire set of data, that can
+inlude many datasets. There's no good way to refer to the entire results
+of searches, other than the user manually creating a dataset somewhere
+public and linking to it from Zenodo to get a DOI. However, simply using
+downloads is simpler in terms of getting a single citeable DOI for the
+data used in a research paper/project.
+
+In \texttt{rgbif}, the output of \texttt{occ\_download\_meta()} can be
+passed to \texttt{gbif\_citation()} to get the citation for the entire
+download.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{key <{-}}\StringTok{ "0000122{-}171020152545675"}
+\NormalTok{res <{-}}\StringTok{ }\KeywordTok{occ\_download\_meta}\NormalTok{(key)}
+\KeywordTok{gbif\_citation}\NormalTok{(res)}
+\CommentTok{\#> $download}
+\CommentTok{\#> [1] "GBIF Occurrence Download https://doi.org/10.15468/dl.yghxj7 Accessed from R via rgbif (https://github.com/ropensci/rgbif) on 2017{-}10{-}20"}
+\CommentTok{\#> }
+\CommentTok{\#> $datasets}
+\CommentTok{\#> NULL}
+\end{Highlighting}
+\end{Shaded}
+
+If you however download the data first, then \texttt{gbif\_citation()}
+can get citations for all the datasets contained within the download.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{d1 <{-}}\StringTok{ }\KeywordTok{occ\_download\_get}\NormalTok{(key, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{)}
+\KeywordTok{gbif\_citation}\NormalTok{(d1)}
+\CommentTok{\#> $download}
+\CommentTok{\#> [1] "GBIF Occurrence Download https://doi.org/10.15468/dl.yghxj7 Accessed from R via rgbif (https://github.com/ropensci/rgbif) on 2017{-}10{-}20"}
+\CommentTok{\#> }
+\CommentTok{\#> $datasets}
+\CommentTok{\#> $datasets[[1]]}
+\CommentTok{\#> <<rgbif citation>>}
+\CommentTok{\#>    Citation: Grant S, Jones J (2017). Field Museum of Natural History (Zoology)}
+\CommentTok{\#>         Invertebrate Collection. Version 18.6. Field Museum. Occurrence Dataset}
+\CommentTok{\#>         https://doi.org/10.15468/6q5vuc accessed via GBIF.org on 2017{-}10{-}20..}
+\CommentTok{\#>         Accessed from R via rgbif (https://github.com/ropensci/rgbif) on}
+\CommentTok{\#>         2020{-}02{-}24}
+\CommentTok{\#>    Rights: To the extent possible under law, the publisher has waived all}
+\CommentTok{\#>         rights to these data and has dedicated them to the Public Domain (CC0}
+\CommentTok{\#>         1.0). Users may copy, modify, distribute and use the work, including}
+\CommentTok{\#>         for commercial purposes, without restriction.}
+\CommentTok{\#> }
+\CommentTok{\#> $datasets[[2]]}
+\CommentTok{\#> <<rgbif citation>>}
+\CommentTok{\#>    Citation: Creuwels J (2017). Naturalis Biodiversity Center (NL) {-} Crustacea.}
+\CommentTok{\#>         Naturalis Biodiversity Center. Occurrence Dataset}
+\CommentTok{\#>         https://doi.org/10.15468/vjoltu accessed via GBIF.org on 2017{-}10{-}20..}
+\CommentTok{\#>         Accessed from R via rgbif (https://github.com/ropensci/rgbif) on}
+\CommentTok{\#>         2020{-}02{-}24}
+\CommentTok{\#>    Rights: To the extent possible under law, the publisher has waived all}
+\CommentTok{\#>         rights to these data and has dedicated them to the Public Domain (CC0}
+\CommentTok{\#>         1.0). Users may copy, modify, distribute and use the work, including}
+\CommentTok{\#>         for commercial purposes, without restriction.}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{gbif-data-in-other-r-packages}{%
+\section{GBIF data in other R
+packages}\label{gbif-data-in-other-r-packages}}
 
@@
 
-\subsubsection{taxize}\label{taxize}
+\hypertarget{taxize}{%
+\subsection{taxize}\label{taxize}}
 
@@
 
-\subsubsection{spocc}\label{spocc}
+\hypertarget{spocc}{%
+\subsection{spocc}\label{spocc}}
 
@@
 
-\section{R vs.~Python vs.~Ruby}\label{r-vs.python-vs.ruby}
+\hypertarget{r-vs.-python-vs.-ruby}{%
+\section{R vs.~Python vs.~Ruby}\label{r-vs.-python-vs.-ruby}}
 
@@
 
-\section{Use cases}\label{use-cases}
+\hypertarget{use-cases}{%
+\section{Use cases}\label{use-cases}}
 
@@
 
-\subsubsection{Ecological niche
-modeling}\label{ecological-niche-modeling}
+\hypertarget{ecological-niche-modeling}{%
+\subsection{Ecological niche modeling}\label{ecological-niche-modeling}}
 
@@
 \begin{Highlighting}[]
-\NormalTok{files <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\KeywordTok{system.file}\NormalTok{(}\DataTypeTok{package =} \StringTok{"dismo"}\NormalTok{), }\StringTok{"/ex"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{),}
+\NormalTok{files <{-}}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\KeywordTok{system.file}\NormalTok{(}\DataTypeTok{package =} \StringTok{"dismo"}\NormalTok{), }\StringTok{"/ex"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{),}
                     \StringTok{"grd"}\NormalTok{, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{predictors <-}\StringTok{ }\KeywordTok{stack}\NormalTok{(files)}
+\NormalTok{predictors <{-}}\StringTok{ }\KeywordTok{stack}\NormalTok{(files)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{data}\NormalTok{(wrld_simpl)}
+\KeywordTok{data}\NormalTok{(wrld\_simpl)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{nn <-}\StringTok{ }\KeywordTok{name_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
-\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{nubKey))}
-\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
-\NormalTok{df_data <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, }\ControlFlowTok{function}\NormalTok{(x) }\KeywordTok{any}\NormalTok{(}\KeywordTok{class}\NormalTok{(x}\OperatorTok{$}\NormalTok{data) }\OperatorTok{%in%}\StringTok{ "tbl_df"}\NormalTok{)) ]}
-\NormalTok{df_data <-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df_data, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
-\NormalTok{df2 <-}\StringTok{ }\NormalTok{df_data }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(decimalLongitude, decimalLatitude)}
+\NormalTok{nn <{-}}\StringTok{ }\KeywordTok{name\_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
+\NormalTok{nn <{-}}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{nubKey))}
+\NormalTok{df <{-}}\StringTok{ }\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
+\NormalTok{df\_data <{-}}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, }\ControlFlowTok{function}\NormalTok{(x) }\KeywordTok{any}\NormalTok{(}\KeywordTok{class}\NormalTok{(x}\OperatorTok{$}\NormalTok{data) }\OperatorTok{\%in\%}\StringTok{ "tbl\_df"}\NormalTok{)) ]}
+\NormalTok{df\_data <{-}}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{bind\_rows}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df\_data, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
+\NormalTok{df2 <{-}}\StringTok{ }\NormalTok{df\_data }\OperatorTok{\%>\%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(decimalLongitude, decimalLatitude)}
 \end{Highlighting}
@@
 \KeywordTok{plot}\NormalTok{(predictors, }\DecValTok{1}\NormalTok{)}
-\KeywordTok{plot}\NormalTok{(wrld_simpl, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(wrld\_simpl, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{)}
 \KeywordTok{points}\NormalTok{(df2, }\DataTypeTok{col =} \StringTok{"blue"}\NormalTok{)}
@@
 
-\includegraphics{components/figure/manuscript-unnamed-chunk-53-1.pdf}
-
-\subsubsection{Biodiversity in big
-cities}\label{biodiversity-in-big-cities}
+\includegraphics{components/figure/manuscript_mee-unnamed-chunk-57-1.pdf}
+
+\hypertarget{biodiversity-in-big-cities}{%
+\subsection{Biodiversity in big
+cities}\label{biodiversity-in-big-cities}}
 
@@
 \begin{Highlighting}[]
-\NormalTok{url <-}\StringTok{ 'https://raw.githubusercontent.com/amyxzhang/}
-\StringTok{boundingbox-cities/master/boundbox.txt'}
-\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{content}\NormalTok{(}\KeywordTok{GET}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\StringTok{""}\NormalTok{, url)), }\DataTypeTok{as =} \StringTok{"text"}\NormalTok{)}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
+\NormalTok{url <{-}}\StringTok{ \textquotesingle{}https://raw.githubusercontent.com/amyxzhang/}
+\StringTok{boundingbox{-}cities/master/boundbox.txt\textquotesingle{}}
+\NormalTok{rawdat <{-}}\StringTok{ }\KeywordTok{content}\NormalTok{(}\KeywordTok{GET}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\StringTok{""}\NormalTok{, url)), }\DataTypeTok{as =} \StringTok{"text"}\NormalTok{)}
+\NormalTok{dat <{-}}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
   \DataTypeTok{text =}\NormalTok{ rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{,}
@@
   \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
+\NormalTok{dat <{-}}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
   \DataTypeTok{city =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{city, }\DataTypeTok{minlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlon,}
@@
 \begin{Highlighting}[]
-\NormalTok{getdata <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x)\{}
-\NormalTok{  coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
-\NormalTok{  wkt <-}\StringTok{ }\NormalTok{wicket}\OperatorTok{::}\KeywordTok{wkt_correct}\NormalTok{(wicket}\OperatorTok{::}\KeywordTok{bounding_wkt}\NormalTok{(}\DataTypeTok{values =}\NormalTok{ coords))}
-\NormalTok{  num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)}\OperatorTok{$}\NormalTok{meta}\OperatorTok{$}\NormalTok{count}
+\NormalTok{getdata <{-}}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x)\{}
+\NormalTok{  coords <{-}}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{\textquotesingle{}minlon\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}minlat\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}maxlon\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}maxlat\textquotesingle{}}\NormalTok{)])}
+\NormalTok{  wkt <{-}}\StringTok{ }\NormalTok{wicket}\OperatorTok{::}\KeywordTok{wkt\_correct}\NormalTok{(wicket}\OperatorTok{::}\KeywordTok{bounding\_wkt}\NormalTok{(}\DataTypeTok{values =}\NormalTok{ coords))}
+\NormalTok{  num <{-}}\StringTok{ }\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)}\OperatorTok{$}\NormalTok{meta}\OperatorTok{$}\NormalTok{count}
   \KeywordTok{data.frame}\NormalTok{(}
-    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{'city'}\NormalTok{],}
+    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{\textquotesingle{}city\textquotesingle{}}\NormalTok{],}
     \DataTypeTok{richness =}\NormalTok{ num,}
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{apply}\NormalTok{(dat, }\DecValTok{1}\NormalTok{, getdata)}
+\NormalTok{out <{-}}\StringTok{ }\KeywordTok{apply}\NormalTok{(dat, }\DecValTok{1}\NormalTok{, getdata)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by =} \StringTok{"city"}\NormalTok{)}
+\NormalTok{out <{-}}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by =} \StringTok{"city"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =}\NormalTok{ (minlat }\OperatorTok{+}\StringTok{ }\NormalTok{maxlat)}\OperatorTok{/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =}\NormalTok{ (minlon }\OperatorTok{+}\StringTok{ }\NormalTok{maxlon)}\OperatorTok{/}\DecValTok{2}\NormalTok{)}
+\NormalTok{out <{-}}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =}\NormalTok{ (minlat }\OperatorTok{+}\StringTok{ }\NormalTok{maxlat)}\OperatorTok{/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =}\NormalTok{ (minlon }\OperatorTok{+}\StringTok{ }\NormalTok{maxlon)}\OperatorTok{/}\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{mapp <-}\StringTok{ }\KeywordTok{map_data}\NormalTok{(}\StringTok{'world'}\NormalTok{)}
+\NormalTok{mapp <{-}}\StringTok{ }\KeywordTok{map\_data}\NormalTok{(}\StringTok{\textquotesingle{}world\textquotesingle{}}\NormalTok{)}
 \KeywordTok{ggplot}\NormalTok{(mapp, }\KeywordTok{aes}\NormalTok{(long, lat)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_polygon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group=}\NormalTok{group), }\DataTypeTok{fill=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{alpha=}\DecValTok{0}\NormalTok{, }\DataTypeTok{color=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.4}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{out, }\KeywordTok{aes}\NormalTok{(lon, lat, }\DataTypeTok{color=}\NormalTok{richness), }\DataTypeTok{size=}\DecValTok{5}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_color_continuous}\NormalTok{(}\DataTypeTok{low =} \StringTok{"#60E1EE"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#0404C8"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom\_polygon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group=}\NormalTok{group), }\DataTypeTok{fill=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{alpha=}\DecValTok{0}\NormalTok{, }\DataTypeTok{color=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.4}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom\_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{out, }\KeywordTok{aes}\NormalTok{(lon, lat, }\DataTypeTok{color=}\NormalTok{richness), }\DataTypeTok{size=}\DecValTok{5}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale\_color\_continuous}\NormalTok{(}\DataTypeTok{low =} \StringTok{"\#60E1EE"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"\#0404C8"}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{""}\NormalTok{, }\DataTypeTok{y=}\StringTok{""}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_grey}\NormalTok{(}\DataTypeTok{base_size=}\DecValTok{14}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\DataTypeTok{legend.key =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{color =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{keywidth =} \DecValTok{2}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{components/figure/manuscript-unnamed-chunk-58-1.pdf}
-
-\subsubsection{Valley oak occurrence data
-comparison}\label{valley-oak-occurrence-data-comparison}
+\StringTok{  }\KeywordTok{theme\_grey}\NormalTok{(}\DataTypeTok{base\_size=}\DecValTok{14}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\DataTypeTok{legend.key =} \KeywordTok{element\_blank}\NormalTok{()) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{color =} \KeywordTok{guide\_legend}\NormalTok{(}\DataTypeTok{keywidth =} \DecValTok{2}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{components/figure/manuscript_mee-unnamed-chunk-62-1.pdf}
+
+\hypertarget{valley-oak-occurrence-data-comparison}{%
+\subsection{Valley oak occurrence data
+comparison}\label{valley-oak-occurrence-data-comparison}}
 
@@
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{'rgbif'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'raster'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'sp'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'sf'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'rgeos'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'scales'}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{'rnaturalearth'}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}rgbif\textquotesingle{}}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}raster\textquotesingle{}}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}sp\textquotesingle{}}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}sf\textquotesingle{}}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}rgeos\textquotesingle{}}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}scales\textquotesingle{}}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{\textquotesingle{}rnaturalearth\textquotesingle{}}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{keyFe <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name =} \StringTok{'Fraxinus excelsior'}\NormalTok{, }\DataTypeTok{kingdom =} \StringTok{'plants'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
-\NormalTok{dat.Fe <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keyFe, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{, }\DataTypeTok{limit =}\NormalTok{ 10000L)}
+\NormalTok{keyFe <{-}}\StringTok{ }\KeywordTok{name\_backbone}\NormalTok{(}\DataTypeTok{name =} \StringTok{\textquotesingle{}Fraxinus excelsior\textquotesingle{}}\NormalTok{, }\DataTypeTok{kingdom =} \StringTok{\textquotesingle{}plants\textquotesingle{}}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
+\NormalTok{dat.Fe <{-}}\StringTok{ }\KeywordTok{occ\_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keyFe, }\DataTypeTok{return =} \StringTok{\textquotesingle{}data\textquotesingle{}}\NormalTok{, }\DataTypeTok{limit =}\NormalTok{ 10000L)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{url <-}\StringTok{ 'http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus_excelsior.zip'}
-\NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
-\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"fraxinus_excelsior.zip"}\NormalTok{)}
-\KeywordTok{unzip}\NormalTok{(}\StringTok{"fraxinus_excelsior.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
-\NormalTok{fe <-}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"Fraxinus_excelsior_EUFORGEN.shp"}\NormalTok{))}
+\NormalTok{url <{-}}\StringTok{ \textquotesingle{}http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus\_excelsior.zip\textquotesingle{}}
+\NormalTok{tmp <{-}}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
+\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"fraxinus\_excelsior.zip"}\NormalTok{)}
+\KeywordTok{unzip}\NormalTok{(}\StringTok{"fraxinus\_excelsior.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
+\NormalTok{fe <{-}}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read\_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"Fraxinus\_excelsior\_EUFORGEN.shp"}\NormalTok{))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{eur <-}\StringTok{ }\NormalTok{rnaturalearth}\OperatorTok{::}\KeywordTok{ne_countries}\NormalTok{(}\DataTypeTok{continent =} \StringTok{"europe"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"map_units"}\NormalTok{)}
-\NormalTok{eur1 <-}\StringTok{ }\NormalTok{eur[eur}\OperatorTok{$}\NormalTok{sovereignt }\OperatorTok{!=}\StringTok{ "Russia"}\NormalTok{, ]}
+\NormalTok{eur <{-}}\StringTok{ }\NormalTok{rnaturalearth}\OperatorTok{::}\KeywordTok{ne\_countries}\NormalTok{(}\DataTypeTok{continent =} \StringTok{"europe"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"map\_units"}\NormalTok{)}
+\NormalTok{eur1 <{-}}\StringTok{ }\NormalTok{eur[eur}\OperatorTok{$}\NormalTok{sovereignt }\OperatorTok{!=}\StringTok{ "Russia"}\NormalTok{, ]}
 \end{Highlighting}
@@
 \KeywordTok{plot}\NormalTok{(eur1, }\DataTypeTok{col =} \StringTok{"darkgrey"}\NormalTok{, }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
-     \DataTypeTok{main =} \StringTok{'Distribution of Fraxinus excelsior'}\NormalTok{)}
-\CommentTok{# add distribution range layer}
+     \DataTypeTok{main =} \StringTok{\textquotesingle{}Distribution of Fraxinus excelsior\textquotesingle{}}\NormalTok{)}
+\CommentTok{\# add distribution range layer}
 \KeywordTok{plot}\NormalTok{(fe, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
-\CommentTok{# add Gbif presence points}
+\CommentTok{\# add Gbif presence points}
 \KeywordTok{points}\NormalTok{(dat.Fe}\OperatorTok{$}\NormalTok{decimalLongitude, dat.Fe}\OperatorTok{$}\NormalTok{decimalLatitude,}
-       \DataTypeTok{cex =}\NormalTok{ .}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+       \DataTypeTok{cex =} \FloatTok{.7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
 \KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \DecValTok{38}\NormalTok{, }\DataTypeTok{y =} \DecValTok{81}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"GBIF Data"}\NormalTok{, }\StringTok{"Range Layer"}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{bg =} \StringTok{"grey"}\NormalTok{,}
-       \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{'darkgreen'}\NormalTok{, }\KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{)), }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{components/figure/manuscript-unnamed-chunk-64-1.pdf}
-
+       \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{\textquotesingle{}darkgreen\textquotesingle{}}\NormalTok{, }\KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{)), }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{.8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{components/figure/manuscript_mee-unnamed-chunk-68-1.pdf}
+
+\hypertarget{conclusions-and-future-directions}{%
 \section{Conclusions and future
-directions}\label{conclusions-and-future-directions}
+directions}\label{conclusions-and-future-directions}}
 
@@
 \texttt{rgbif} package has already been used in more than 20 scholarly
-publications (as of 2016-08-10), including (Amano, Lamming \&
-Sutherland, 2016, Bartomeus et al. (2013), Barve (2014), Bone et al.
-(2015), Collins et al. (2015), Drozd \& Šipoš (2013), Kong, Huang \&
-Duan (2015), Richardson, Roux \& Wilson (2015), Turner, Fréville \&
-Rieseberg (2015), Verheijen et al. (2015), Zizka \& Antonelli (2015),
-Butterfield et al. (2016), Dellinger et al. (2015), Feitosa et al.
-(2015), Malhado et al. (2015), Werner et al. (2015), Robertson, Visser
-\& Hui (2016), Davison et al. (2015), Janssens et al. (2016)).
-
-The \texttt{rgbif} package is relatively stable, and should not have
-many breaking changes unless necessitated due to changes in the GBIF
-API. However, it will gain function(s) to work with the maps API in the
-near future.
-
-The \texttt{pygbif} and \texttt{gbifrb} libraries are in early
-development, and will greatly benefit from any feedback and use cases.
+publications (as of 2020-02-25), including (Drozd \& Šipoš 2013;
+Bartomeus \emph{et al.} 2013; Barve 2014; Richardson \emph{et al.} 2015;
+Feitosa \emph{et al.} 2015; Collins \emph{et al.} 2015; Malhado \emph{et
+al.} 2015; Kong \emph{et al.} 2015; Werner \emph{et al.} 2015; Bone
+\emph{et al.} 2015; Turner \emph{et al.} 2015; Davison \emph{et al.}
+2015; Verheijen \emph{et al.} 2015; Dellinger \emph{et al.} 2015; Zizka
+\& Antonelli 2015; Janssens \emph{et al.} 2016; Robertson \emph{et al.}
+2016; Amano \emph{et al.} 2016; Butterfield \emph{et al.} 2016).
+
+The \texttt{rgbif} package is stable, and should not have many breaking
+changes unless necessitated due to changes in the GBIF API. The
+\texttt{pygbif} and \texttt{gbifrb} libraries are in early development,
+and will greatly benefit from any feedback and use cases.
 
@@
 
-\section{Acknowledgments}\label{acknowledgments}
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
 
 This project was supported in part by the Alfred P Sloan Foundation
-(Grant No. G-2014-13485), and in part by the Helmsley Foundation (Grant
-No. 2016PG-BRI004).
-
-\section{Data Accessibility}\label{data-accessibility}
+(Grant No.~G-2014-13485), and in part by the Helmsley Foundation (Grant
+No.~2016PG-BRI004).
+
+\hypertarget{data-accessibility}{%
+\section{Data Accessibility}\label{data-accessibility}}
 
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Amano_2016}{}
-Amano T., Lamming JDL., Sutherland WJ. 2016. Spatial gaps in global
-biodiversity information and the role of citizen science.
-\emph{BioScience} 66:393--400.
-
-\hypertarget{ref-Bartomeus_2013}{}
-Bartomeus I., Park MG., Gibbs J., Danforth BN., Lakso AN., Winfree R.
-2013. Biodiversity ensures plant-pollinator phenological synchrony
-against climate change. \emph{Ecology Letters} 16:1331--1338.
-
-\hypertarget{ref-Barve_2014}{}
-Barve V. 2014. Discovering and developing primary biodiversity data from
-social networking sites: A novel approach. \emph{Ecological Informatics}
-24:194--199.
-
-\hypertarget{ref-Beck_2012}{}
-Beck J., Ballesteros-Mejia L., Buchmann CM., Dengler J., Fritz SA.,
-Gruber B., Hof C., Jansen F., Knapp S., Kreft H., Schneider A-K., Winter
-M., Dormann CF. 2012. Whats on the horizon for macroecology?
-\emph{Ecography} 35:673--683.
-
-\hypertarget{ref-Bone_2015}{}
-Bone RE., Smith JAC., Arrigo N., Buerki S. 2015. A macro-ecological
-perspective on crassulacean acid metabolism (CAM) photosynthesis
-evolution in afro-madagascan drylands: Eulophiinae orchids as a case
-study. \emph{New Phytologist} 208:469--481.
-
-\hypertarget{ref-Brown_1995}{}
-Brown JH. 1995. \emph{Macroecology}. University of Chicago Press.
-
-\hypertarget{ref-Brown_2015}{}
-Brown KA., Parks KE., Bethell CA., Johnson SE., Mulligan M. 2015.
-Predicting plant diversity patterns in madagascar: Understanding the
-effects of climate and land cover change in a biodiversity hotspot.
-\emph{PLOS ONE} 10:e0122721.
-
-\hypertarget{ref-Butterfield_2016}{}
-Butterfield BJ., Copeland SM., Munson SM., Roybal CM., Wood TE. 2016.
-Prestoration: Using species in restoration that will persist now and
-into the future. \emph{Restor Ecol}.
-
-\hypertarget{ref-Ceballos_2015}{}
-Ceballos G., Ehrlich PR., Barnosky AD., Garcia A., Pringle RM., Palmer
-TM. 2015. Accelerated modern human-induced species losses: Entering the
-sixth mass extinction. \emph{Science Advances} 1:e1400253--e1400253.
-
-\hypertarget{ref-rgbif}{}
-Chamberlain S., Ram K., Barve V., Mcglinn D. \emph{rgbif: An r interface
-to the global 'biodiversity' information facility API}.
-
-\hypertarget{ref-pygbif}{}
-Chamberlain S. \emph{pygbif: A python interface to the global
-biodiversity information facility API}.
-
-\hypertarget{ref-gbifrb}{}
-Chamberlain S. \emph{gbifrb: A ruby interface to the global biodiversity
-information facility API}.
-
-\hypertarget{ref-Collins_2015}{}
-Collins R., Ribeiro ED., Machado VN., Hrbek T., Farias I. 2015. A
-preliminary inventory of the catfishes of the lower rio nhamundá, brazil
-(ostariophysi, siluriformes). \emph{BDJ} 3:e4162.
-
-\hypertarget{ref-Davison_2015}{}
-Davison J., Moora M., Opik M., Adholeya A., Ainsaar L., Ba A., Burla S.,
-Diedhiou AG., Hiiesalu I., Jairus T., Johnson NC., Kane A., Koorem K.,
-Kochar M., Ndiaye C., Partel M., Reier U., Saks U., Singh R., Vasar M.,
-Zobel M. 2015. Global assessment of arbuscular mycorrhizal fungus
-diversity reveals very low endemism. \emph{Science} 349:970--973.
-
-\hypertarget{ref-Dellinger_2015}{}
-Dellinger AS., Essl F., Hojsgaard D., Kirchheimer B., Klatt S., Dawson
-W., Pergl J., Pyšek P., Kleunen M van., Weber E., Winter M., Hörandl E.,
-Dullinger S. 2015. Niche dynamics of alien species do not differ among
-sexual and apomictic flowering plants. \emph{New Phytologist}
-209:1313--1323.
-
-\hypertarget{ref-Drozd_2013}{}
-Drozd P., Šipoš J. 2013. R for all (i): Introduction to the new age of
-biological analyses. \emph{Casopis slezskeho zemskeho muzea (A)} 62.
-
-\hypertarget{ref-Faulkner_2014}{}
-Faulkner KT., Robertson MP., Rouget M., Wilson JR. 2014. A simple, rapid
-methodology for developing invasive species watch lists.
-\emph{Biological Conservation} 179:25--32.
-
-\hypertarget{ref-Di_Febbraro_2013}{}
-Febbraro MD., Lurz PWW., Genovesi P., Maiorano L., Girardello M.,
-Bertolino S. 2013. The use of climatic niches in screening procedures
-for introduced species to evaluate risk of spread: A case with the
-american eastern grey squirrel. \emph{PLoS ONE} 8:e66559.
-
-\hypertarget{ref-Feitosa_2015}{}
-Feitosa YO., Absy ML., Latrubesse EM., Stevaux JC. 2015. Late quaternary
-vegetation dynamics from central parts of the madeira river in brazil.
-\emph{Acta Bot. Bras.} 29:120--128.
-
-\hypertarget{ref-Ferretti_2015}{}
-Ferretti F., Verd GM., Seret B., Šprem JS., Micheli F. 2015. Falling
-through the cracks: The fading history of a large iconic predator.
-\emph{Fish and Fisheries}:n/a--n/a.
-
-\hypertarget{ref-Ficetola_2014}{}
-Ficetola GF., Rondinini C., Bonardi A., Baisero D., Padoa-Schioppa E.
-2014. Habitat availability for amphibians and extinction threat: A
-global analysis. \emph{Diversity and Distributions} 21:302--311.
-
-\hypertarget{ref-wkt}{}
-Herring J. 2011. OpenGIS implementation standard for geographic
+\begin{cslreferences}
+\leavevmode\hypertarget{ref-Amano_2016}{}%
+Amano, T., Lamming, J.D.L. \& Sutherland, W.J. (2016). Spatial gaps in
+global biodiversity information and the role of citizen science.
+\emph{BioScience}, \textbf{66}, 393--400. Retrieved from
+\url{http://dx.doi.org/10.1093/biosci/biw022}
+
+\leavevmode\hypertarget{ref-Bartomeus_2013}{}%
+Bartomeus, I., Park, M.G., Gibbs, J., Danforth, B.N., Lakso, A.N. \&
+Winfree, R. (2013). Biodiversity ensures plant-pollinator phenological
+synchrony against climate change (M. Eubanks, Ed.). \emph{Ecology
+Letters}, \textbf{16}, 1331--1338. Retrieved from
+\url{http://dx.doi.org/10.1111/ele.12170}
+
+\leavevmode\hypertarget{ref-Barve_2014}{}%
+Barve, V. (2014). Discovering and developing primary biodiversity data
+from social networking sites: A novel approach. \emph{Ecological
+Informatics}, \textbf{24}, 194--199. Retrieved from
+\url{http://dx.doi.org/10.1016/j.ecoinf.2014.08.008}
+
+\leavevmode\hypertarget{ref-Beck_2012}{}%
+Beck, J., Ballesteros-Mejia, L., Buchmann, C.M., Dengler, J., Fritz,
+S.A., Gruber, B., Hof, C., Jansen, F., Knapp, S., Kreft, H., Schneider,
+A.-K., Winter, M. \& Dormann, C.F. (2012). Whats on the horizon for
+macroecology? \emph{Ecography}, \textbf{35}, 673--683. Retrieved from
+\url{http://dx.doi.org/10.1111/j.1600-0587.2012.07364.x}
+
+\leavevmode\hypertarget{ref-Bone_2015}{}%
+Bone, R.E., Smith, J.A.C., Arrigo, N. \& Buerki, S. (2015). A
+macro-ecological perspective on crassulacean acid metabolism (CAM)
+photosynthesis evolution in afro-madagascan drylands: Eulophiinae
+orchids as a case study. \emph{New Phytologist}, \textbf{208}, 469--481.
+Retrieved from \url{http://dx.doi.org/10.1111/nph.13572}
+
+\leavevmode\hypertarget{ref-Brown_1995}{}%
+Brown, J.H. (1995). \emph{Macroecology}. University of Chicago Press.
+
+\leavevmode\hypertarget{ref-Brown_2015}{}%
+Brown, K.A., Parks, K.E., Bethell, C.A., Johnson, S.E. \& Mulligan, M.
+(2015). Predicting plant diversity patterns in madagascar: Understanding
+the effects of climate and land cover change in a biodiversity hotspot
+(L. Kumar, Ed.). \emph{PLOS ONE}, \textbf{10}, e0122721. Retrieved from
+\url{http://dx.doi.org/10.1371/journal.pone.0122721}
+
+\leavevmode\hypertarget{ref-Butterfield_2016}{}%
+Butterfield, B.J., Copeland, S.M., Munson, S.M., Roybal, C.M. \& Wood,
+T.E. (2016). Prestoration: Using species in restoration that will
+persist now and into the future. \emph{Restor Ecol}. Retrieved from
+\url{http://dx.doi.org/10.1111/rec.12381}
+
+\leavevmode\hypertarget{ref-Ceballos_2015}{}%
+Ceballos, G., Ehrlich, P.R., Barnosky, A.D., Garcia, A., Pringle, R.M.
+\& Palmer, T.M. (2015). Accelerated modern human-induced species losses:
+Entering the sixth mass extinction. \emph{Science Advances}, \textbf{1},
+e1400253--e1400253. Retrieved from
+\url{http://dx.doi.org/10.1126/sciadv.1400253}
+
+\leavevmode\hypertarget{ref-gbifrb}{}%
+Chamberlain, S. \emph{gbifrb: A ruby interface to the global
+biodiversity information facility API}. Retrieved from
+\url{https://github.com/sckott/gbifrb}
+
+\leavevmode\hypertarget{ref-pygbif}{}%
+Chamberlain, S. \emph{pygbif: A python interface to the global
+biodiversity information facility API}. Retrieved from
+\url{https://github.com/sckott/pygbif}
+
+\leavevmode\hypertarget{ref-rgbif}{}%
+Chamberlain, S., Ram, K., Barve, V. \& Mcglinn, D. \emph{rgbif: An r
+interface to the global 'biodiversity' information facility API}.
+Retrieved from \url{https://github.com/ropensci/rgbif}
+
+\leavevmode\hypertarget{ref-Collins_2015}{}%
+Collins, R., Ribeiro, E.D., Machado, V.N., Hrbek, T. \& Farias, I.
+(2015). A preliminary inventory of the catfishes of the lower rio
+nhamundá, brazil (ostariophysi, siluriformes). \emph{BDJ}, \textbf{3},
+e4162. Retrieved from \url{http://dx.doi.org/10.3897/bdj.3.e4162}
+
+\leavevmode\hypertarget{ref-Davison_2015}{}%
+Davison, J., Moora, M., Opik, M., Adholeya, A., Ainsaar, L., Ba, A.,
+Burla, S., Diedhiou, A.G., Hiiesalu, I., Jairus, T., Johnson, N.C.,
+Kane, A., Koorem, K., Kochar, M., Ndiaye, C., Partel, M., Reier, U.,
+Saks, U., Singh, R., Vasar, M. \& Zobel, M. (2015). Global assessment of
+arbuscular mycorrhizal fungus diversity reveals very low endemism.
+\emph{Science}, \textbf{349}, 970--973. Retrieved from
+\url{http://dx.doi.org/10.1126/science.aab1161}
+
+\leavevmode\hypertarget{ref-Dellinger_2015}{}%
+Dellinger, A.S., Essl, F., Hojsgaard, D., Kirchheimer, B., Klatt, S.,
+Dawson, W., Pergl, J., Pyšek, P., Kleunen, M. van, Weber, E., Winter,
+M., Hörandl, E. \& Dullinger, S. (2015). Niche dynamics of alien species
+do not differ among sexual and apomictic flowering plants. \emph{New
+Phytologist}, \textbf{209}, 1313--1323. Retrieved from
+\url{http://dx.doi.org/10.1111/nph.13694}
+
+\leavevmode\hypertarget{ref-Drozd_2013}{}%
+Drozd, P. \& Šipoš, J. (2013). R for all (i): Introduction to the new
+age of biological analyses. \emph{Casopis slezskeho zemskeho muzea (A)},
+\textbf{62}. Retrieved from
+\url{http://dx.doi.org/10.2478/cszma-2013-0004}
+
+\leavevmode\hypertarget{ref-Faulkner_2014}{}%
+Faulkner, K.T., Robertson, M.P., Rouget, M. \& Wilson, J.R.U. (2014). A
+simple, rapid methodology for developing invasive species watch lists.
+\emph{Biological Conservation}, \textbf{179}, 25--32. Retrieved from
+\url{http://dx.doi.org/10.1016/j.biocon.2014.08.014}
+
+\leavevmode\hypertarget{ref-Di_Febbraro_2013}{}%
+Febbraro, M.D., Lurz, P.W.W., Genovesi, P., Maiorano, L., Girardello, M.
+\& Bertolino, S. (2013). The use of climatic niches in screening
+procedures for introduced species to evaluate risk of spread: A case
+with the american eastern grey squirrel (H. Verbruggen, Ed.). \emph{PLoS
+ONE}, \textbf{8}, e66559. Retrieved from
+\url{http://dx.doi.org/10.1371/journal.pone.0066559}
+
+\leavevmode\hypertarget{ref-Feitosa_2015}{}%
+Feitosa, Y.O., Absy, M.L., Latrubesse, E.M. \& Stevaux, J.C. (2015).
+Late quaternary vegetation dynamics from central parts of the madeira
+river in brazil. \emph{Acta Bot. Bras.}, \textbf{29}, 120--128.
+Retrieved from \url{http://dx.doi.org/10.1590/0102-33062014abb3711}
+
+\leavevmode\hypertarget{ref-Ferretti_2015}{}%
+Ferretti, F., Verd, G.M., Seret, B., Šprem, J.S. \& Micheli, F. (2015).
+Falling through the cracks: The fading history of a large iconic
+predator. \emph{Fish and Fisheries}, n/a--n/a. Retrieved from
+\url{http://dx.doi.org/10.1111/faf.12108}
+
+\leavevmode\hypertarget{ref-Ficetola_2014}{}%
+Ficetola, G.F., Rondinini, C., Bonardi, A., Baisero, D. \&
+Padoa-Schioppa, E. (2014). Habitat availability for amphibians and
+extinction threat: A global analysis (D. Richardson, Ed.).
+\emph{Diversity and Distributions}, \textbf{21}, 302--311. Retrieved
+from \url{http://dx.doi.org/10.1111/ddi.12296}
+
+\leavevmode\hypertarget{ref-wkt}{}%
+Herring, J. (2011). OpenGIS implementation standard for geographic
 information-simple feature access-part 1: Common architecture. \emph{OGC
-Document} 4:122--127.
-
-\hypertarget{ref-Janssens_2016}{}
-Janssens SB., Vandelook F., Langhe ED., Verstraete B., Smets E.,
-Vandenhouwe I., Swennen R. 2016. Evolutionary dynamics and biogeography
-of musaceae reveal a correlation between the diversification of the
-banana family and the geological and climatic history of southeast asia.
-\emph{New Phytologist} 210:1453--1465.
-
-\hypertarget{ref-Kong_2015}{}
-Kong X., Huang M., Duan R. 2015. SDMdata: A web-based software tool for
-collecting species occurrence records. \emph{PLOS ONE} 10:e0128295.
-
-\hypertarget{ref-Malhado_2015}{}
-Malhado AC., Oliveira-Neto JA., Stropp J., Strona G., Dias LC., Pinto
-LB., Ladle RJ. 2015. Climatological correlates of seed size in amazonian
-forest trees. \emph{J Veg Sci} 26:956--963.
-
-\hypertarget{ref-Mendoza_2015}{}
-María Mendoza., Ospina OE., Cárdenas-Henao H., García-R JC. 2015. A
-likelihood inference of historical biogeography in the world's most
-diverse terrestrial vertebrate genus: Diversification of
+Document}, \textbf{4}, 122--127. Retrieved from
+\url{http://www.opengeospatial.org/standards/sfa}
+
+\leavevmode\hypertarget{ref-Janssens_2016}{}%
+Janssens, S.B., Vandelook, F., Langhe, E.D., Verstraete, B., Smets, E.,
+Vandenhouwe, I. \& Swennen, R. (2016). Evolutionary dynamics and
+biogeography of musaceae reveal a correlation between the
+diversification of the banana family and the geological and climatic
+history of southeast asia. \emph{New Phytologist}, \textbf{210},
+1453--1465. Retrieved from \url{http://dx.doi.org/10.1111/nph.13856}
+
+\leavevmode\hypertarget{ref-Kong_2015}{}%
+Kong, X., Huang, M. \& Duan, R. (2015). SDMdata: A web-based software
+tool for collecting species occurrence records (J.H. Badger, Ed.).
+\emph{PLOS ONE}, \textbf{10}, e0128295. Retrieved from
+\url{http://dx.doi.org/10.1371/journal.pone.0128295}
+
+\leavevmode\hypertarget{ref-Malhado_2015}{}%
+Malhado, A.C.M., Oliveira-Neto, J.A., Stropp, J., Strona, G., Dias,
+L.C.P., Pinto, L.B. \& Ladle, R.J. (2015). Climatological correlates of
+seed size in amazonian forest trees (T. Nakashizuka, Ed.). \emph{J Veg
+Sci}, \textbf{26}, 956--963. Retrieved from
+\url{http://dx.doi.org/10.1111/jvs.12301}
+
+\leavevmode\hypertarget{ref-Mendoza_2015}{}%
+Marı́a Mendoza, Ospina, O.E., Cárdenas-Henao, H. \& Garcı́a-R, J.C.
+(2015). A likelihood inference of historical biogeography in the world's
+most diverse terrestrial vertebrate genus: Diversification of
 direct-developing frogs (craugastoridae: Pristimantis) across the
-neotropics. \emph{Molecular Phylogenetics and Evolution} 85:50--58.
-
-\hypertarget{ref-Pimm_2014}{}
-Pimm SL., Jenkins CN., Abell R., Brooks TM., Gittleman JL., Joppa LN.,
-Raven PH., Roberts CM., Sexton JO. 2014. The biodiversity of species and
-their rates of extinction, distribution, and protection. \emph{Science}
-344:1246752--1246752.
-
-\hypertarget{ref-R}{}
-R Core Team. 2014. \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing.
-
-\hypertarget{ref-Richardson_2015}{}
-Richardson DM., Roux JJL., Wilson JR. 2015. Australian acacias as
-invasive species: Lessons to be learnt from regions with long planting
-histories. \emph{Southern Forests: a Journal of Forest Science}
-77:31--39.
-
-\hypertarget{ref-Robertson_2016}{}
-Robertson MP., Visser V., Hui C. 2016. Biogeo: An r package for
+neotropics. \emph{Molecular Phylogenetics and Evolution}, \textbf{85},
+50--58. Retrieved from
+\url{http://dx.doi.org/10.1016/j.ympev.2015.02.001}
+
+\leavevmode\hypertarget{ref-Pimm_2014}{}%
+Pimm, S.L., Jenkins, C.N., Abell, R., Brooks, T.M., Gittleman, J.L.,
+Joppa, L.N., Raven, P.H., Roberts, C.M. \& Sexton, J.O. (2014). The
+biodiversity of species and their rates of extinction, distribution, and
+protection. \emph{Science}, \textbf{344}, 1246752--1246752. Retrieved
+from \url{http://dx.doi.org/10.1126/science.1246752}
+
+\leavevmode\hypertarget{ref-R}{}%
+R Core Team. (2014). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing, Vienna, Austria.
+Retrieved from \url{http://www.R-project.org/}
+
+\leavevmode\hypertarget{ref-Richardson_2015}{}%
+Richardson, D.M., Roux, J.J.L. \& Wilson, J.R. (2015). Australian
+acacias as invasive species: Lessons to be learnt from regions with long
+planting histories. \emph{Southern Forests: a Journal of Forest
+Science}, \textbf{77}, 31--39. Retrieved from
+\url{http://dx.doi.org/10.2989/20702620.2014.999305}
+
+\leavevmode\hypertarget{ref-Robertson_2016}{}%
+Robertson, M.P., Visser, V. \& Hui, C. (2016). Biogeo: An r package for
 assessing and improving data quality of occurrence record datasets.
-\emph{Ecography} 39:394--401.
-
-\hypertarget{ref-Turner_2015}{}
-Turner KG., Fréville H., Rieseberg LH. 2015. Adaptive plasticity and
-niche expansion in an invasive thistle. \emph{Ecol Evol} 5:3183--3197.
-
-\hypertarget{ref-Verheijen_2015}{}
-Verheijen LM., Aerts R., Bönisch G., Kattge J., Bodegom PMV. 2015.
-Variation in trait trade-offs allows differentiation among predefined
-plant functional types: Implications for predictive ecology. \emph{New
-Phytologist} 209:563--575.
-
-\hypertarget{ref-Werner_2015}{}
-Werner GDA., Cornwell WK., Cornelissen JHC., Kiers ET. 2015.
-Evolutionary signals of symbiotic persistence in the legumerhizobia
-mutualism. \emph{Proceedings of the National Academy of Sciences}
-112:10262--10269.
-
-\hypertarget{ref-Zizka_2015}{}
-Zizka A., Antonelli A. 2015. \emph{speciesgeocodeR: An r package for
-linking species occurrences, user-defined regions and phylogenetic trees
-for biogeography, ecology and evolution}. Cold Spring Harbor Laboratory
-Press.
+\emph{Ecography}, \textbf{39}, 394--401. Retrieved from
+\url{http://dx.doi.org/10.1111/ecog.02118}
+
+\leavevmode\hypertarget{ref-Turner_2015}{}%
+Turner, K.G., Fréville, H. \& Rieseberg, L.H. (2015). Adaptive
+plasticity and niche expansion in an invasive thistle. \emph{Ecol Evol},
+\textbf{5}, 3183--3197. Retrieved from
+\url{http://dx.doi.org/10.1002/ece3.1599}
+
+\leavevmode\hypertarget{ref-Verheijen_2015}{}%
+Verheijen, L.M., Aerts, R., Bönisch, G., Kattge, J. \& Bodegom, P.M.V.
+(2015). Variation in trait trade-offs allows differentiation among
+predefined plant functional types: Implications for predictive ecology.
+\emph{New Phytologist}, \textbf{209}, 563--575. Retrieved from
+\url{http://dx.doi.org/10.1111/nph.13623}
+
+\leavevmode\hypertarget{ref-Werner_2015}{}%
+Werner, G.D.A., Cornwell, W.K., Cornelissen, J.H.C. \& Kiers, E.T.
+(2015). Evolutionary signals of symbiotic persistence in the
+legumerhizobia mutualism. \emph{Proceedings of the National Academy of
+Sciences}, \textbf{112}, 10262--10269. Retrieved from
+\url{http://dx.doi.org/10.1073/pnas.1424030112}
+
+\leavevmode\hypertarget{ref-Zizka_2015}{}%
+Zizka, A. \& Antonelli, A. (2015). \emph{speciesgeocodeR: An r package
+for linking species occurrences, user-defined regions and phylogenetic
+trees for biogeography, ecology and evolution}. Cold Spring Harbor
+Laboratory Press. Retrieved from \url{http://dx.doi.org/10.1101/032755}
+\end{cslreferences}
 

### 2020-02-27 use-cases/city-diversity.Rmd
tweaks to mee paper

@@
 
-```{r box}
+```{r box, cache=TRUE}
 url <- 'https://raw.githubusercontent.com/amyxzhang/
